### 2002-01-25 docs/paper/figures.tex
Initial revision
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@2 75dff457-8751-0410-ab1a-cd701654ce6b

\clearpage

\setcounter{thefigs}{0}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{radec.sample8b.ps}
\caption{\label{radec.sample8b} Angular distribution, in an Aitoff
projection of equatorial coordinates, of the SDSS sample used in the
current analysis. This spectroscopic sample covers around 1,169 square
degrees and contains about 100,000 galaxy spectra.}
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{response_sdss.ps}
\caption{\label{response_sdss} Estimated filter response for all five
bands in the SDSS, as a function of rest frame wavelength for galaxies
at different redshifts. A 4 gigayear old instantaneous burst using the
models of \citet{bruzual93a} is shown for reference. {\it Top panel:}
The system consisting of \band{0.0}{u}, \band{0.0}{g}, \band{0.0}{r},
\band{0.0}{i}, and \band{0.0}{z}.  {\it Bottom panel:} The system
consisting of \band{0.0}{u}, \band{0.1}{g}, \band{0.1}{r},
\band{0.1}{i}, and \band{0.2}{z}. We use this second system rather
than the first because it requires less interpolation to determine
\band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}, and no extrapolation to
determine \band{0.0}{u} and \band{0.1}{z}. }
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{plotaij.sample8b15.ps}
\caption{\label{plotaij.sample8b15} {\it Top panel}: Distribution of
the components of the three-parameter fit to the five-band SDSS
photometry for a random subsample consisting of 16,000 of the SDSS
galaxies. $a_0$ is linearly proportional to the flux between $3500\AA$
and $7000\AA$, while $a_1$ and $a_2$ contribute no flux in this
range. Thus, the ratios $a_1/a_0$ and $a_2/a_0$ describe the spectral
type of the galaxy. $a_2/a_0$ is the more variable parameter and thus
is the better separator of galaxy type. {\it Bottom panel}: At fixed
$a_1/a_0=-0.2$, the inferred spectra corresponding to various values
of $a_2/a_0$. Near $a_2/a_0=0.5$, the spectrum is similar to that of
an elliptical galaxy. For higher values, the spectrum becomes bluer. }
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{model.ps}
\caption{\label{model} Galaxy fluxes determined from the inferred
galaxy spectra (see Figure \ref{plotaij.sample8b15}) relative to the
observed galaxy fluxes, for all five SDSS bands, shown for a random
subsample consisting of 16,000 of the SDSS galaxies. The
residuals are shown against redshift and absolute magnitude, the two
most important systematic variables in the analysis of this paper. The
solid lines show the median values. The dashed lines show the 10\% and
90\% levels (corresponding to $1.3\sigma$ if the distribution were
Gaussian). The systematics are very small in $g$, $r$, $i$ and $z$,
but more significant in $u$. The scatter is largest in $u$ and
$z$. Nevertheless, the inferred fluxes are a good representation of
the observed fluxes.  }
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{colors.sample8b15.ps}
\caption{\label{colors.sample8b15} For galaxies in the absolute
magnitude range $-21.5<M_r<-21$, the color distribution calculated
from the inferred SEDs as a function of redshift. The left column
shows the four colors relative to \band{0.1}{r} for each galaxy versus
redshift. The right column shows the distribution of colors for
galaxies in the range $0.04<z<0.12$ {\it (dotted)} compared to the
distribution of colors of galaxies in the range $0.12<z<0.19$ {\it
(solid)}. The resulting color distribution is clearly fairly constant
with redshift.}
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{kcorrect.sample8b15.ps}
\caption{\label{kcorrect.sample8b15} $K$-corrections as a function of
redshift in all five bands for a random subsample consisting of 8,000
of the SDSS galaxies. The $K$-corrections are largest, and therefore
the most uncertain, for the \band{0.0}{u} and \band{0.1}{g} bands. The
range of $K$-corrections at each redshift reflects the range of galaxy
types at each redshift.}
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{ciCompare.sample8b15.ps}
\caption{\label{ciCompare.sample8b15} Difference in the
$K$-corrections in each band between the method used in Figure
\ref{kcorrect.sample8b15} and the method of simply interpolating
between adjacent bands fitting a power-law SED. The differences are
small in \band{0.1}{r}, \band{0.1}{i}, and \band{0.2}{z}, where galaxy
SEDs have simple shapes. There are large differences in \band{0.0}{u}
and \band{0.1}{g}, for which the 4000 \AA\ break is important. Note
particularly the systematic differences in \band{0.1}{g} with
redshift.}
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{cibreakCompare.sample8b15.ps}
\caption{\label{cibreakCompare.sample8b15} Same as Figure
\ref{ciCompare.sample8b15}, now comparing the $K$-corrections of Figure
\ref{kcorrect.sample8b15} with the ``interpolation with a break''
method. This method fits a power law between adjacent bands, except at
4000 \AA, where we fit for the size of the 4000\AA\ break (assuming
that $f(\lambda)\propto \lambda^2$ for $\lambda<4000$ \AA). This
greatly improves the agreement in \band{0.1}{g} while making the
disagreement in \band{0.0}{u} only slightly worse. These results
indicate that it important to account for the structure in the blue
region of the spectrum when performing $K$-corrections.}
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{specK.ps}
\caption{\label{specK} Similar to Figure \ref{ciCompare.sample8b15},
now comparing the $K$-corrections of Figure \ref{kcorrect.sample8b15}
with $K$-corrections determined from the spectra (which can only be
calculated for the \band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}
bands). In all bands, the $K$-corrections are very similar, giving us
confidence in our results. This result is actually remarkable more for
what it says about the high quality of the spectrophotometry in the
SDSS survey.}
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{aperturevsz.M.sample8b15.ps}
\caption{\label{aperturevsz.M.sample8b15} }
\end{figure}

### 2002-01-25 docs/paper/paper.tex
Initial revision
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@2 75dff457-8751-0410-ab1a-cd701654ce6b

\documentclass[10pt,preprint]{aastex}

\newcommand{\vv}[1]{{\bf #1}}
\newcommand{\df}{\delta}
\newcommand{\dfft}{{\tilde{\delta}}}
\newcommand{\betaft}{{\tilde{\beta}}}
\newcommand{\erf}{{\mathrm{erf}}}
\newcommand{\erfc}{{\mathrm{erfc}}}
\newcommand{\Step}{{\mathrm{Step}}}
\newcommand{\ee}[1]{\times 10^{#1}}
\newcommand{\avg}[1]{{\langle{#1}\rangle}}
\newcommand{\Avg}[1]{{\left\langle{#1}\right\rangle}}
\def\simless{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'074$}}\mathchar"7218\,$}}} % < or of order
\def\simgreat{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'076$}}\mathchar"7218\,$}}} % > or of order
\newcommand{\iras}{{\sl IRAS\/}}
\newcommand{\petroratio}{{{\mathcal{R}}_P}}
\newcommand{\petroradius}{{{r}_P}}
\newcommand{\petronumber}{{{N}_P}}
\newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
\newcommand{\band}[2]{\ensuremath{^{#1}{#2}}}

\setlength{\footnotesep}{9.6pt}

\newcounter{thefigs}
\newcommand{\fignum}{\arabic{thefigs}}

\newcounter{thetabs}
\newcommand{\tabnum}{\arabic{thetabs}}

\newcounter{address}

%% You can insert a short comment on the title page using the command below.

\slugcomment{Submitted to \aj}

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.  Running heads
%% will not print in the manuscript style.

\shortauthors{Blanton {\it et al.} (2000)}
\shorttitle{Low Redshift Galaxy Evolution}

%% This paper uses runs 752/756, rerun 4
%% It uses redshifts from version 3 of spectro1d

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Modeling Galaxy Spectral Energy Distributions in the SDSS
Spectroscopic Survey}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, you can use \\ to force line breaks.

%\author{Michael Blanton}
%\affil{NASA/Fermilab Astrophysics Center\\
%Fermi National Accelerator Laboratory, Batavia, IL 60510-0500}
%\author{\and lots and lots of other, probably more important people}
%\affil{Some Institution\\
%Somewhere, Some City, Some State or Province}
%\email{blanton@fnal.gov}

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}}, and
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}
%David W. Hogg\altaffilmark{\ref{NYU}}, and
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

\altaffiltext{1}{Based on observations obtained with the
Sloan Digital Sky Survey} 
\setcounter{address}{2}
\altaffiltext{\theaddress}{
\stepcounter{address}
New York University, Department of Physics, 4 Washington Place, New
York, NY 10003
\label{NYU}}
\altaffiltext{\theaddress}{
\stepcounter{address}
Department of Physics and Astronomy, The Johns Hopkins University,
Baltimore, MD 21218
\label{JHU}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%University of Pittsburgh,
%Department of Physics and Astronomy,
%3941 O'Hara Street,
%Pittsburgh, PA 15260
%\label{Pitt}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%rinceton University Observatory, Princeton,
%NJ 08544
%\label{Princeton}}
%\addtocounter{address}{1}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Fermi National Accelerator Laboratory, P.O. Box 500,
%Batavia, IL 60510
%\label{Fermilab}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Astronomy, University of Washington,
%Box 351580,
%Seattle, WA 98195 
%\label{UW}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%University of Chicago, Astronomy \&
%Astrophysics Center, 5640 S. Ellis Ave., Chicago, IL 60637
%\label{Chicago}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Hubble Fellow 
%\label{Hubble}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Sussex Astronomy Centre,
%University of Sussex,
%Falmer, Brighton BN1 9QJ, UK
%\label{Sussex}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Ohio State University,
%Department of Astronomy,
%Columbus, OH 43210
%\label{Ohio}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Apache Point Observatory,
%2001 Apache Point Road,
%P.O. Box 59, Sunspot, NM 88349-0059
%\label{APO}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Astronomy, California Institute of Technology,
%Pasadena, CA 91125
%\label{Caltech}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Observatoire Midi-Pyr\'en\'ees, 
%14 ave Edouard Belin, Toulouse, F-31400, France
%\label{Pyrenees}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Astronomy and Research Center for 
%the Early Universe,
%School of Science, University of Tokyo,
%Tokyo 113-0033, Japan
%\label{Tokyo}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%UC Berkeley, Dept. of Astronomy, 601 Campbell Hall, Berkeley, CA  94720-3411
%\label{Berkeley}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Institute for Cosmic Ray Research, University of
%Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
%\label{CosmicRay}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Institute for Advanced Study, Olden Lane,
%Princeton, NJ 08540
%\label{IAS}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%U.S. Naval Observatory,
%3450 Massachusetts Ave., NW,
%Washington, DC  20392-5420
%\label{USNO}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%University of Michigan, Department of Physics,
%500 East University, Ann Arbor, MI 48109
%\label{Michigan}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Physics \& Astronomy,
%The University of Edinburgh,
%James Clerk Maxwell Building,
%The King's Buildings,
%Mayfield Road,
%Edinburgh EH9 3JZ, UK
%\label{Edinburgh}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Physics, Carnegie Mellon University, 
%5000 Forbes Avenue, Pittsburgh, PA 15213-3890 
%\label{CarnegieMellon}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Space Telescope Science Institute, Baltimore, MD 21218
%\label{STScI}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Astronomy and Astrophysics,
%The Pennsylvania State University,
%University Park, PA 16802
%\label{PennState}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Astronomy,
%University of Maryland,
%College Park, MD 20742-2421 
%\label{Maryland}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Astrophysikalisches Institut Potsdam,
%An der Sternwarte 16, D-14482 Potsdam, Germany
%\label{Potsdam}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%Department of Physics, Drexel University, Philadelphia, PA 19104
%\label{Drexel}}
%\altaffiltext{\theaddress}{
%\stepcounter{address}
%National Astronomical Observatory, Mitaka, Tokyo 181-8588, Japan
%\label{NAOJ}}
%\addtocounter{address}{1}
%\altaffiltext{\theaddress}{Physics Dept., University of California, Davis, CA 95616
%\label{UCDavis}}
%\addtocounter{address}{1}
%\altaffiltext{\theaddress}{IGPP/Lawrence Livermore National Laboratory
%\label{IGPP}}
%\addtocounter{address}{1}
%\altaffiltext{\theaddress}{Department of Astronomy, University of California, Berkeley, C
%A 94720-3411
%\label{Berkeley}}
%\stepcounter{address}
%\altaffiltext{\theaddress}{Remote Sensing Division, Code 7215, Naval
%Research Laboratory, Washington, DC 20375
%\label{NRL}}
%\addtocounter{address}{1}
%\altaffiltext{\theaddress}{U.S. Naval Observatory, Flagstaff Station,
%P.O. Box 1149,
%Flagstaff, AZ  86002-1149
%\label{Flagstaff}}

\clearpage

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
We present a method of inferring galaxy spectral energy distributions
(SEDs) from photometric observations of galaxies in broad band
filters. Using the Sloan Digital Sky Survey (SDSS) observations in the
optical wavelength regime, we demonstrate that the method yields
robust results. We show how the method can be used to recover galaxy
magnitudes in fixed frame bandpasses over a range of redshifts. We
compare our method to others, including galaxy spectrophotometry. For
the SDSS, fixed frame magnitudes around the SDSS $r$, $i$, and $z$
bands are highly reliable, at the level of a few percent; $g$-band
fixed frame magnitudes are somewhat more uncertain, and $u$-band fixed
frame magnitudes are highly uncertain, at the level of 10--20\%. All
code and templates from this paper are public.
\end{abstract}

\keywords{galaxies: fundamental parameters --- galaxies: photometry
--- galaxies: statistics}

%
% Introduction and motivation
%

\section{Motivation}
\label{motivation}

Photometric surveys of the sky observe galaxies in bandpasses which
are fixed in the rest-frame of the observer. However, spectra of
distant galaxies are redshifted due to the expansion of the
universe. Thus, the more distant the galaxy, the further the observed
bandpass is blueshifted relative to the rest-frame spectral energy
distribution (SED) of the object observed. To compare observations of
galaxies at different redshifts, it is important to account for the
fact that one is observing different parts of the galaxy SED.

In the past, this accounting has in general come in the form of
``$K$-corrections'' applied to the observed magnitudes. That is, a
function $K(z)$ is added to the standard cosmological bolometric
distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
been applied regardless of the galaxy type, though more recently it
has become standard to use a discrete set of $K(z)$ functions
depending on galaxy type (based either on morphology or spectral
features). As described in \cite{oke68a} and later papers, the
function is based on the projection of an assumed galaxy SED
redshifted to $z$ onto the filter response.

However, galaxies do not all have the same SED, nor are they selected
from some discrete set of SEDs. For this reason, schemes for applying
the $K$-corrections can fail to be self-consistent: the SED assumed
for the $K$-corrections can be significantly inconsistent with the
observed galaxy colors! As we demand more precision from our
astronomical data analysis in new, large multi-band surveys, and in
particular as we try to quantify the evolution of galaxies, we must
take a more sophisticated approach to approximating fixed-frame
observations of galaxies.

Our approach here is to develop a method for inferring the underlying
SEDs of a set of galaxies at a range of redshifts by requiring that
their SEDs all be drawn from a similar population. For each galaxy, we
will recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. In addition, this method provides a natural
way to infer an estimate of the flux within the whole optical range.
The approach is equivalent to the photometric metric redshift
estimation methods of \citet{csabai99a} and \citet{budavari00a}, 
except we use slightly different coordinate systems.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad-band photometry. An implementation of the method in C and IDL is
provided at a URL on the World-Wide Web, through a public CVS
repository, or by request.  Section \ref{data} applies the method to
galaxies in the SDSS, showing that the fits are robust. Section
\ref{kcorrection} shows how one derives $K$-corrections from the
results, and discusses how best to use the results of such
fits. Section \ref{conclusion} concludes and discusses future
development of the method described here.

\section{Fitting SEDs to Galaxy Broad-band Colors}
\label{sedfit}

Our task is to recover a model for the galaxy SED from broad-band
photometric measurements. Since a set of broad band galaxy fluxes does
not correspond uniquely to a particular SED, and because galaxy SEDs
are known to have significant structure over wavelength ranges small
compared to our bandpass, this task is an ill-posed, inverse
problem. However, we are not completely ignorant about the forms which
galaxy SEDs take, so we can attempt to use what we know about galaxy
SEDs to simplify our task.

The method described here for doing so follows closely the methods of
estimating photometric redshifts used by \citet{csabai00a} and
\citet{budavari00a}. It is designed to take advantage of what we
already understand about galaxy spectral energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.
Beginning with a space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), one can find an
orthonormal set of spectra spanning that space. To be more precise, we
define the dot-product in the space of galaxy spectra to be:
\begin{equation}
\vv{b}_i \cdot \vv{b}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
\lambda b_i(\lambda) b_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range corresponding to the range
which is within the observed wavelength range over most of the
redshift range of the sample. Otherwise, orthogonality is defined in
terms of regions of the spectra which are poorly constrained, which
make interpretation of the fits inconvenient.  This choice does not
affect the results for the reconstructed SED fit quantitatively
whatsoever.\footnote{If the template spectra were identical in the
subrange chosen and only differed outside of it, this choice would
cause computational problems, but in fact the spectra are all
independent in the restricted, as well as the full, wavelength range.}

We then further define the basis of the SED space such that the $N_b$
template SEDs, $\vv{v}_i$, are linear combinations of $N_b$ basis
spectra $\vv{b}_i$ and that $\vv{b}_i \cdot \vv{b}_j =
\delta_{ij}$. These conditions do not fully specify the $\vv{b}_i$,
which naturally may be rotated arbitrarily within our
$N_b$-dimensional subspace of spectral space.  In our implementation,
we use the subspace defined by linear combinations of ten
Bruzual-Charlot instantaneous burst models (our $\vv{v}_i(\lambda)$)
with ages ranging from $3 \times 10^7$ to $2\times 10^{10}$ years,
five with metallicity $Z=0.02$ and six with $Z=0.004$, all assuming a
Salpeter Initial Mass Function and smoothed using a Gaussian with a
standard deviation of $\sigma = 300$\AA. The smoothing is necessary
to prevent the method from overfitting; not all the small-scale
information about the templates can be recovered using this method,
and we find that this level of smoothing gives sensible results.

We cannot determine an individual galaxy's position in this
ten-dimensional space from its broad-band colors alone, because there
are only five observed bands. However, as outlined by
\citet{csabai00a}, one can find a lower dimensional space defined by
$N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
galaxies. The SED of galaxy $i$ may be approximated as a linear
combination of the basis spectra in this lower dimensional space:
\begin{equation}
\label{lincomb}
f_{\mathrm{model},i}(\lambda) d\lambda = 
\sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
\end{equation}
where $a_k$ are the components of the galaxy projected in the
low-dimensional space. From this SED it is easy to determine the flux
$F_{\mathrm{model},il}$ measured in each survey bandpass by projecting
the model SED onto the bandpass $l$. We define
\begin{equation}
\chi^2 = \sum_i \sum_l
\frac{(F_{\mathrm{obs},il}-F_{\mathrm{model},il})^2}{\sigma_{il}^2},
\end{equation}
where $\sigma_{il}$ is the estimated error in the observed flux
$F_{\mathrm{obs},il}$ in bandpass $l$ for galaxy $i$. Taking
derivatives of $\chi^2$ with respect to $a_{i,k}$ and $e_{k,j}$
results in a set of equations bilinear in $a_{i,k}$ and $e_{k,j}$.
\citet{csabai00a} describe how to iteratively solve this bilinear
equation by first fixing $e_{k,j}$ and fitting for $a_{i,k}$, then
fixing $a_{i,k}$ and fitting for $e_{k,j}$, and iterating that
procedure.

%We put one additional constraint on our SEDs. To regularize the edges
%of the SEDs, we define two additional bands blueward and redward of
%$u$ and $z$, respectively, which we will call $a_1$ and $a_2$. The
%bandpasses are simply Gaussians centered at 2700\AA\ and 10200\AA\
%with standard deviations of 250\AA\ and 350\AA, respectively. From
%studying synthetic galaxy SEDs redshifted between $z=0$ and $z=0.2$,
%it is clear that the $u-a_1$ color varies not much more than 1.5
%magnitudes around the value 1.5, and that the $z-a_2$ color usually
%varies only by about 0.2 magnitudes around 0.1. Thus, given the $u$
%and $z$ magnitude, we can put a very weak constraint on the $a_1$ and
%$a_2$ magnitudes.  Similarly, when one of the SDSS bands is
%unobserved, we replace it based on an adjacent band and the mean color
%for galaxies in the sample, with a large error bar. These weak
%constraints prevent the inferred galaxy SEDs from becoming too
%unreasonable in the regions of the spectrum without data.

We define the ``flux'' $f_k$ of each template as the flux in the range
$\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
estimate of the flux in this range is $F_{{v}}=\Sigma_k a_k f_k$. The
locus of points of a fixed flux in the component space $a_k$ lie on a
plane. Therefore, a natural choice for the orientation of the axes
$\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
constant flux; in this manner, the coefficient $a_0$ in Equation
(\ref{lincomb}) is directly proportional to $F$ and is thus purely a
measure of the object's flux (in the wavelength range considered),
while the parameters $a_i/a_0$ (where $i>0$) primarily measure the
shape of the galaxy SED in the wavelength range considered. $F$ can be
trivially related to $L$, the luminosity in this range, by the inverse
square of the luminosity distance.  As it turns out, the parameters
$a_i/a_0$ tend to be distributed very approximately in an ellipsoid,
with not much curvature within the space defined by the template
SEDs. Thus, we rotate the axes $\vv{b}_i$ (for $i>0$) such that they
are aligned with the principal axes of the ellipsoid. This choice of
coordinate system is arbitrary, however, and does not affect our
quantitative results.

In this way, we can reconstruct SEDs based on the broad-band
magnitudes of galaxies. From these reconstructed SEDs one can estimate
$K$-corrections, develop a measure of galaxy spectral type, or to
synthesize a very broad-band measure of the galaxy flux.  The
templates determined during the procedure can, of course, be used to
estimate photometric redshifts, as \citet{csabai00a} and
\citet{budavari00a} describe.

The version of the $K$-correction code (v1\_0) implementing the
calculations described here, along with eigentemplates and filter
curves is publicly available on the World Wide Web at {\tt
http://physics.nyu.edu/\~mb144/kcorrect}. In addition, a public CVS
repository can be accessed at {\tt xxx} (instructions for exporting
the appropriate version of the code from this repository are provided
on the web page noted above. The whole of the code can be used through
the Research Systems, Incorporated, IDL language; everything except
for the template-fitting also exists in a stand-alone C version (which
calls the same routines, guaranteeing consistency).

\section{Application to SDSS Data}
\label{data}

In this section we describe the SDSS data set, we describe how we fit
the templates and the coefficients in detail for this data set, and we
describe our distribution of code to do this.

\subsection{SDSS Spectroscopic Data}

The SDSS (\citealt{york00a}) is producing imaging and spectroscopic
surveys over $\pi$ steradians in the Northern Galactic Cap. A
dedicated 2.5m telescope (Siegmund {\it et al.}, in preparation) at
Apache Point Observatory, Sunspot, New Mexico, images the sky in five
bands between 3000 and 10000 \AA\ ($u$, $g$, $r$, $i$, $z$;
\citealt{fukugita96a}) using a drift-scanning, mosaic CCD camera
(\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
22.5$. The ultimate goal is to spectroscopically observe 900,000
galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Bright
Red Galaxies (BRGs; Eisenstein {\it et al.}, in preparation), and
100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation)
will be targeted for spectroscopic follow up using two digital
spectrographs on the same telescope. Many of the details of the survey
are described in a paper accompanying the Early Data Release
(\citealt{stoughton01a}). The survey has begun in earnest, and has
obtained about 20\% of its intended data.

As of October 2001, the SDSS had imaged and targeted around 1,739
square degrees of sky and taken spectra of approximately 170,000
objects over 1,169 square degrees of that area. As described below,
only a subset of these galaxies are used here to develop the
templates, though we calculate the SED fits for all of them. The
results of the photometric pipeline for all of these objects were
extracted from the SDSS Operational Database {\bf ref munn?}. The
photometry used here for the bulk of these objects was the same as
that used when the objects were targeted.  However, for those objects
which were in the EDR photometry, we used the better calibrations and
photometry from the EDR.

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& ZZZZ - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& ZZZZ - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons which are detected as a
function of wavelength (a unitless quantity).

The response functions $R(\lambda)$ have been measured using a
monochrometer by {\bf how ref Doi}.  Using a model for the atmospheric
transmission and the reflectivity of the primary and secondary
mirrors, one can then model the response of the entire system. For
each filter in the SDSS, there are six different CCDs; it has been
shown that the differences between these six camera columns are
negligible. The resulting set of filter curves is shown in Figure
\ref{response_sdss}, in comparison to a model of a galaxy spectral
energy distribution observed at $z=0$ (a 4 Gyr old instantaneous burst
from the models of \citealt{bruzual93a}).

We extract one-dimensional spectra from the two-dimension images using
a pipeline created specifically for the SDSS instrumentation
(\citealt{schlegel02a}). We also use redshifts determined by the same
pipeline; the official SDSS redshifts are obtained from a different
pipeline (\citealt{subbarao02a}). The two independent versions provide
a consistency check on the redshift determination. They are consistent
(for galaxies) at over the 99\% level.\footnote{The official pipeline
tends to work better for objects with unusual spectra, such as certain
types of stars and QSOs.}.

We use throughout the magnitudes determined by the SDSS using the
modified form of the magnitude described by \citet{petrosian76a}, as
described in \citet{strauss02a}, reddening-corrected using the dust
maps of \citet{schlegel98a}. These magnitudes have a nearly constant
metric aperture as a function redshift; in addition, all the bands use
the same aperture, so the measured SED corresponds (to within the
effects of seeing) to the SED of an identifiable region of the galaxy.

\subsection{Fitting SEDs to SDSS Data}

For the purposes of applying the method described in Section
\ref{sedfit} to SDSS data, we take a subsample of the data consisting
of around 25,000 objects in the range $0.0<z<0.5$. The sample includes
both the main sample and the LRGs; we sample these spectra such that
we get an approximately even distribution within our range of
redshifts. In addition, we add results from galaxies in several plates
(totalling about 1,000 objects) which were selected by a photometric
redshift algorithm ({\bf ref who?}) to be at around $z\sim 0.3$. These
objects are invaluable for tying down the blue end of the templates.

Some of the objects have missing or poorly constrained data. For
example, the $u$- or $z$-band fluxes for some objects are swamped by
the photon noise of the sky. We identify such cases as magnitude
errors greater than 0.8 or magnitudes fainter than 22.5 in any
band. We ignore these objects entirely when fitting for the
templates. However, we still want to determine a best-fit SED for each
object. For this purpose, we replace a ``bad'' magnitude based on an
adjacent (usually the redder) band and the average color in those
bands of objects at that redshift. We give these measurements a large
error bar, of 0.8 magnitudes. Thus, we account for the missing
information by simply requiring that the object SED have
``reasonable'' properties.

We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
and still fit allow freedom to fit for the templates themselves. Since
galaxies live in a low-dimensional space, we find it is worth being
able to fit for the templates. Simply using five templates (which
obviously reproduces all the magnitudes exactly) tends to yield
unphysical trends of galaxy SED versus redshift ({\it cf.}  Figure
\ref{constantz} below). Using three templates does nearly as well as
four templates in the sense that the resulting templates reproduce the
magnitudes nearly as well. However, since one of the applications of
these SED determinations is the distribution of galaxy colors in
fixed-frame magnitudes, we don't want to artificially reduce the
dimensionality of the color space to only two.

One more choice needs to be made, the wavelength regime over which to
orthogonalize the templates and over which to calculate the flux. We
choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA and
$\lambda_{\mathrm{max}}=7500$\AA, since this range is included for
almost all galaxies in the sample. We refer here and in other papers
to the flux and luminosity in this range as the ``visual flux'' $f_v$
and the ``visual luminosity'' $L_v$.

\subsection{Results}

The resulting four templates are shown in Figure \ref{templates} in
the range $2000<\lambda<12000$\AA, the region which is constrained by
our observations. Figure \ref{coeffs} is a more illuminating
figure. It shows the pairwise joint distributions of $a_1/a_0$,
$a_2/a_0$, and $a_3/a_0$ in the top three panels. In the bottom panel,
three spectra taken from a one-dimensional sequence along the galay
locus are shown, showing that the spectra become progressively bluer
along that sequence.

These reconstructed spectra do an excellent job of reproducing the
observed galaxy fluxes. Figure \ref{model} shows the differences
between the observed and reconstructed fluxes as a function of
redshift and as a function of absolute magnitude (in $\band{0.1}{r}$).
The solid lines show the median differences. The estimates of the
fluxes in $g$, $r$, $i$, and $z$ do not correlate strongly with any of
these parameters, while for $u$ there is a 5--10\% discrepancy for the
most distant galaxies, in the sense that the model $u$-band flux is
too bright. The dotted lines show the boundaries containing 80\% of
the objects surrounding the median (roughly the $1.3\sigma$ limits if
the distribution were Gaussian). For $g$, $r$, and $i$, the scatter is
very small --- certainly comparable to the errors in calibration ---
and roughly independent of redshift and absolute magnitude. For $z$,
the scatter is somewhat larger, perhaps reflecting the poor
signal-to-noise in $z$, but remains constant with redshift. For $u$,
the scatter is larger, perhaps reflecting the fact that four templates
simply is not enough to characterize galaxy SEDs simultaneously in the
optical and near ultra-violet. After all, galaxies are known to have
large variations in their near-ultraviolet properties (because it
depends strongly on dust content and recent star-formation). This
effect may become more important with redshift as $u$ probes bluer and
bluer regions of the SED. All told, these spectra seem to be fair
estimates of the true underlying spectra.

An important test of the reasonableness of the fits is to check that
for a fixed absolute magnitude, the distribution of the SED fits
depends little on redshift. Figures \ref{coeffs1}--\ref{coeffs4}
demonstrates this to be the case for galaxies in several luminosity
ranges. The left column of Figure \ref{coeffs1} shows the three
coefficient ratios $a_i/a_0$ for each galaxy as a function of
redshift. The right column shows the distribution of these coefficient
ratios for two redshift ranges, showing that they are nearly
consistent. The resulting change in the color distribution inferred
from these coefficients is similarly small, at the level of what one
would expect intrinsic evolution to cause.

In short, these fits to galaxy SEDs provide estimates which reproduce
the galaxy photometry nearly to the level of the errors in the
photometry itself, seem physically reasonable, and are consistent over
the range of redshifts we consider ($0.0. < z < 0.5$).

\section{$K$-corrections}
\label{kcorrection}

\subsection{Calculating $K$-corrections}

With a model of the rest-frame galaxy SED in hand, it is 
straightforward to project that SED onto a filter corresponding to 
an observation at any redshift, and predict the so-called
$K$-correction to the galaxy magnitude in order to compare galaxies 
at different redshifts. One simply calculates the AB magnitude at 
any redshift as:
\begin{equation}
m_{AB} = ZZZZ - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
f(\frac{\lambda}{1+z}) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].
\end{equation}
The simple way to think about this equation is that simply moving away
from the Earth doing does not change the bolometric flux of the
galaxy. So when the spectrum is stretched by the $(1+z)$ redshift
factor, as in $f(\lambda/(1+z))$, the bolometric flux should be
constant; that is, you must apply the extra multiplicative factor
$1/(1+z)$ to counteract the stretching of the spectrum. Once you have
have accounted for the fact that you are observing a particular
bandpass given the redshift of the object, you can then go ahead and
apply to the calculated magnitude the cosmological distance modulus
(that is, the luminosity distance-squared law, as tabulated by, {\it
e.g.} \citealt{hogg97a}).

We want to emphasize here that, while $K$-correcting to a fixed frame
bandpass is sometimes necessary in order to achieve a scientific
objective, it is not always necessary or appropriate. Because
$K$-corrections are an inherently uncertain business (the broad-band
magnitudes just do not uniquely determine the SED) they should be
avoid or minimized when possible. For example, if one was calculating
the evolution of clustering of red and blue galaxies separately, it
would perhaps be wise to perform the separation not on the
$K$-corrected colors of the galaxies but on the median color of, say,
$M_\ast$ galaxies as a function of redshift. Similarly, in situation
where $K$-corrections cannot be ignored, such as the calculation of
the evolution of the luminosity function, their effect should be
minimized by, for example, correcting to the median redshift of
galaxies in the sample.

For this reason, we here define a notation which we will use in
subsequent papers for the effective rest-frame bandpass at any
redshift in band $b$, namely:
\begin{equation}
\band{z}{b}
\end{equation}
read, ``$b$-band at redshift $z$.'' Now, typically one's measurements
in this band will be difficult to connect to, say, solar luminosities
in $\band{z}{b}$, simply because people have not projected the appropriate
stellar spectrophotometry onto these bandpasses. However, it is our
position that stars are {\it better} understood than galaxies, and
that it therefore is in the end simpler to stay as close as possible
to the system in which the galaxies are observed. In any case,
astronomy is quickly reaching a level of precision for which the exact
nature of the bandpasses used has to be known and considered in most
analyses of observational data.

The reader may ask why, if we are basing our $K$-corrections on a full
model of the galaxy SED, we do not correct to bandpasses with simpler
shapes (say, Gaussians). The reason is that we want the effective
bandpasses to actually have been observed for some galaxies in the
sample, so that the $K$-correction for those galaxies is formally and
actually zero. This property is desirable because, as noted above,
there are uncertainties in the $K$-corrections.

\subsection{Testing the $K$-corrections}

We show, in Figure \ref{kcorrect}, the $K$-corrections to $z=0.1$
inferred from this method, for all five SDSS bands. Note that the
$K$-corrections are largest (and thus most uncertain) in
$\band{0.1}{u}$ and $\band{0.1}{g}$.

To test how robust these $K$-correction results are to our model
assumptions, we compare them to other methods of calculating
fixed-frame galaxy magnitudes.  An extremely simple method is to
calculate the flux in any desired bandpass by fitting a power-law
slope and amplitude to the fluxes in the two adjacent bandpasses
(extrapolating when necessary). Figure \ref{ciCompare.sample8b15}
shows the differences in the $K$-corrections inferred from this method
and those inferred from the method of Figure
\ref{kcorrect.sample8b15}, as a function of redshift. $r$, $i$, and
$z$ are all reasonably similar in either method; $u$ and $g$, however,
have distinct trends with redshift, mostly due to the sharper
gradients in the spectra in this regime. To show this fact, we perform
a similar power-law fit, only this time including a break in the
spectrum at 4000\AA. We use the $u$-$g$ color to fit the break,
assuming that the slope blueward of 4000\AA\ is $f(\lambda)\propto
\lambda^{2}$.  Figure \ref{cibreakCompare.sample8b15} shows the
results of this fit; the redshift trend in $g$ is greatly reduced, as
is the trend in $u$, but a large amount of scatter remains in the $u$
band. This results from the fact that you can {\it either} fit the
slope of the SED below the 4000 \AA\ break {\it or} the size of the
4000 \AA\ break itself; it is not possible with this data to constrain
both in an individual spectrum, which is a fundamental limitation of
our analysis of a fixed frame $u$-band luminosity function.

Finally, it is possible to use the galaxy spectra obtained with the
spectrograph to estimate the $K$-correction for each object (in
\band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}). However, this
procedure requires trusting the spectrophotometry over a wide
wavelength range. In addition, the spectra only probe the inner parts
of galaxies, and because the fiber aperture is fixed in angular space,
a different region of the galaxy is probed as a function of
redshift. Nevertheless, in Figure \ref{specK}, we compare the
$K$-corrections calculated based on the spectra to those of Figure
\ref{kcorrect}, finding that they are extremely similar. The fact that
they are so similar might prove useful, in the sense that one could
combine the photometric and spectroscopic results, using the spectra
to constrain small-scale features such as the 4000 \AA\ break and the
photometry to cover a large wavelength range and to probe the full
extent of the galaxy. Such a procedure might resolve the difficulties
in handling the $u$-band discussed in the previous paragraph. We will
not pursue this issue further in this paper and leave it for future
investigation.

\section{Conclusions and Future Work}
\label{conclusions}

We have presented a method and an implementation for estimating galaxy
SEDs for the purpose of calculating fixed-frame galaxy magnitudes over
a range of redshifts. We have demonstrated that it gives sensible and
consistent results. We will be using this method in future papers
which will describe the joint distribution of luminosities and colors
of galaxies, as well as the evolution of the luminosity function of
galaxies. Furthermore, we plan to incorporate observations of objects
in bands other than the SDSS bands to further describe the nature of
galaxy SEDs.

Some of the scatter in the three-dimension space describing the shape
of the galaxy SED is probably due to dust. It may be possible to
evaluate the effects of reddening in this space and, by assuming that
galaxies corrected for internal reddening live in an even lower
dimensional space, perform a reddening correction. We will be
investigating this question in the near future.

Furthermore, one may want to include the effects of evolution in the
$K$-corrections. One approach to this calculation is to define a
one-dimensional locus, and record which position on this locus is
closest to each galaxy. Then, the one-dimensional locus provides a set
of spectra, for which estimates of evolution can be calculated. This
procedure avoids fitting star-formation histories to a huge number of
different positions in galaxy SED space. Since their will be
degeneracies between age and metallicity in these fits in any case, it
may be best to simplify the problem to one dimension.

The distribution of galaxies in the three-dimensional space of galaxy
SEDs described here may be useful for other purposes, as well. As one
example, which has already been explored in a separate project, one
can calculate photometric redshifts (if you are willing to restrict
your galaxy models to a one-dimensional, though possibly curved, locus
within the three-dimensional space). However, there is a fairly severe
redshift-type degeneracy, in the sense that blue galaxies at high
redshift can be confused with red galaxies at lower redshift. These
problems can be hazardous to someone trying to estimate the evolution
of the luminosity function from a photometric redshift sample (such
studies rarely include these correlated errors in their estimate of
their uncertainties). However, one can make an alternative set of
assumptions by modeling the distribution of the galaxies in the
three-dimensional spectral space as a pair of gaussians. For such a
model, and a model for the position and luminosity evolution of each
gaussian, one can calculate the probability of observing any given
galaxy. Thus, one could use the galaxy photometry to constrain the
luminosity evolution of galaxies in each gaussian, as well as the
evolution of the position of the gaussians. This method has its own
systematic errors and its own restrictive set of assumptions, but
provides an alternative path to calculating the luminosity function
directly from photometric redshifts.

\acknowledgments

MB was supported at the beginning of this work by the DOE and NASA
grant NAG 5-7092 at Fermilab. He is also grateful for the hospitality
of the Department of Physics and Astronomy at the State University of
New York at Stony Brook, who kindly provided computing facilities on
his frequent visits there. 
%MAS acknowledges the support of NSF grant AST-0071091.

The Sloan Digital Sky Survey (SDSS) is a joint project of The
University of Chicago, Fermilab, the Institute for Advanced Study, the
Japan Participation Group, the Johns Hopkins University, the
Max-Planck-Institute for Astronomy, New Mexico State University,
Princeton University, the United States Naval Observatory, and the
University of Washington. Apache Point Observatory, site of the SDSS
telescopes, is operated by the Astrophysical Research Consortium
(ARC).  Funding for the project has been provided by the Alfred
P.~Sloan Foundation, the SDSS member institutions, the National
Aeronautics and Space Administration, the National Science Foundation,
the U.~S.~Department of Energy, Monbusho, and the Max Planck
Society. The SDSS Web site is http://www.sdss.org/.

\begin{thebibliography}{DUM}
\input{bib_paper.tex}
\end{thebibliography}

\newpage

\include{tables}

\include{figures}

\end{document}

### 2002-01-25 src/export.h

@@ -1,0 +1,2143 @@
+/* export.h */
+sdsfadsfdfadsf dsaf
+
+/*
+ *  Copyright (c) 1992-1998, Research Systems Inc.  All rights reserved.
+ *  Reproduction by any means whatsoever is prohibited without express
+ *  written permission.
+*/
+
+/*
+ * NOTICE OF INTENT. PLEASE READ:
+ *
+ *     This file is intended to supply the external definitions
+ *     required to write and link code against IDL, either to extend
+ *     IDL (CALL_EXTERNAL, Dynamically loadable modules, LINKIMAGE) or
+ *     to use IDL to extend your own program (Callable IDL). Our goal is
+ *     that including this file in your program will enable your program
+ *     to access the public interfaces described in the IDL documentation,
+ *     especially the External Development Guide.
+ *
+ *     NOT EVERYTHING IN THIS FILE IS A STABLE PUBLIC INTERFACE.
+ *     For various technical reasons, there are things in this file that
+ *     are considered to be private to Research Systems and subject to
+ *     immediate change without notice. Anything in this file that is not
+ *     explicitly documented elsewhere falls into this category. You should
+ *     not use such interfaces, as they are not publically supported and
+ *     may change or disappear without warning.
+ *
+ *     Examples of such interfaces are:
+ *         - Types and constants that are not discussed in the documentation.
+ *         - Functions that are not discussed elsewhere.
+ *         - Fields of structures that are not discussed in the documentation
+ *           even though the structure itself *is* documented.
+ */
+
+#ifndef export_IDL_DEF
+#define export_IDL_DEF
+
+#ifdef __cplusplus
+    extern "C" {
+#endif
+
+
+/***** Definitions from msg_code *****/
+
+#ifndef msg_code_IDL_DEF
+#define msg_code_IDL_DEF
+
+/* Warning: These codes can change between releases. */
+
+#define IDL_M_GENERIC                    -1
+#define IDL_M_NAMED_GENERIC              -2
+#define IDL_M_SYSERR                     -4
+#define IDL_M_NOSUPPORT                  -5
[…]

### 2002-01-25 src/fit_coeffs.c

@@ -1,0 +1,126 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * fit_coeffs.c
+ *
+ * Given the locations of the eigentemplate matrix, the basis set matrix,
+ * and galaxy flux, errors, and redshift, output the best fit coefficients.
+ *
+ * Mike Blanton
+ * 2/2001
+ */
+
+static IDL_LONG nz=1000;
+static double zmin=1.e-4, zmax=1.e-0;
+
+static double *bmatrix=NULL;
+static double *lambda=NULL;
+static double *rmatrix=NULL;
+static double *ematrix=NULL;
+static double *zvals=NULL;
+static IDL_LONG nk,nb,nl,nt;
+static double *filter_lambda=NULL;
+static double *filter_pass=NULL;
+static IDL_LONG *filter_n=NULL;
+static IDL_LONG maxn;
+static double *galaxy_z=NULL;
+static double *galaxy_flux=NULL;
+static double *galaxy_invvar=NULL;
+static double *coeffs=NULL;
+
+int main(int argc,
+				 char **argv)
+{
+	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk;
+	char ematrixfile[2000],bmatrixfile[2000],filterlist[2000],lambdafile[2000];
+
+	/* read arguments */
+	if(argc<5) {
+		fprintf(stderr,"Usage: cat <galaxy file> | fit_coeffs <ematrix file> <bmatrix file> <lambdafile> <filterlist> [zmin] [zmax] [nz]\n");
+		exit(1);
+	} /* end if */
+	i=1;
+	strcpy(ematrixfile,argv[i]); i++;
+	strcpy(bmatrixfile,argv[i]); i++;
+	strcpy(lambdafile,argv[i]); i++;
+	strcpy(filterlist,argv[i]); i++;
+	/* have not implemented optional parameters yet */
+
+	/* load the ematrix; this gives the eigentemplates in terms of the
+	 * basis set */
+	k_load_ascii_table(&ematrix,&ndim,&sizes,ematrixfile);
+	nt=sizes[0];
+	nb=sizes[1];
+	FREEVEC(sizes);
+
+	/* input the bmatrix; this is the spectrum for each basis element */
+	k_load_ascii_table(&bmatrix,&ndim,&sizes,bmatrixfile);
[…]

### 2002-01-25 src/fit_photoz.c

@@ -1,0 +1,116 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * fit_photoz.c
+ *
+ * Given the locations of the eigentemplate matrix, the basis set matrix,
+ * and galaxy fluxes and errors, output the best-fit redshift.
+ *
+ * Mike Blanton
+ * 2/2001
+ */
+
+static IDL_LONG nz=1000;
+static double zmin=1.e-4, zmax=1.e-0;
+
+static double *bmatrix=NULL;
+static double *lambda=NULL;
+static double *rmatrix=NULL;
+static double *ematrix=NULL;
+static double *zvals=NULL;
+static IDL_LONG nk,nb,nl,nt;
+static double *filter_lambda=NULL;
+static double *filter_pass=NULL;
+static IDL_LONG *filter_n=NULL;
+static IDL_LONG maxn;
+static double *galaxy_z=NULL;
+static double *galaxy_flux=NULL;
+static double *galaxy_invvar=NULL;
+static double *coeffs=NULL;
+
+int main(int argc,
+				 char **argv)
+{
+	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk;
+	char ematrixfile[2000],bmatrixfile[2000],filterlist[2000],lambdafile[2000];
+
+	if(argc<5) {
+		fprintf(stderr,"Usage: cat <galaxy file> | fit_photoz <ematrix file> <bmatrix file> <lambdafile> <filterlist> [zmin] [zmax] [nz]\n");
+		exit(1);
+	} /* end if */
+	i=1;
+	strcpy(ematrixfile,argv[i]); i++;
+	strcpy(bmatrixfile,argv[i]); i++;
+	strcpy(lambdafile,argv[i]); i++;
+	strcpy(filterlist,argv[i]); i++;
+
+	/* load the ematrix */
+	k_load_ascii_table(&ematrix,&ndim,&sizes,ematrixfile);
+	nt=sizes[0];
+	nb=sizes[1];
+	FREEVEC(sizes);
+
+	/* input the bmatrix */
+	k_load_ascii_table(&bmatrix,&ndim,&sizes,bmatrixfile);
+	if(sizes[0]!=nb) {
[…]

### 2002-01-25 src/idl_k_create_r.c

@@ -1,0 +1,47 @@
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include "kcorrect.h"
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+static void free_memory()
+{
+}
+
+/********************************************************************/
+IDL_LONG idl_k_create_r
+  (int      argc,
+   void *   argv[])
+{
+	IDL_LONG nk,nb,nl,nz,*filter_n,maxn;
+	double *rmatrix,*bmatrix,*lambda,*zvals,*filter_lambda,*filter_pass;
+	
+	IDL_LONG i,j;
+	IDL_LONG retval=1;
+
+	/* 0. allocate pointers from IDL */
+	i=0;
+	rmatrix=(double *)argv[i]; i++;
+	nk=*((IDL_LONG *)argv[i]); i++;
+	nb=*((IDL_LONG *)argv[i]); i++;
+	bmatrix=(double *)argv[i]; i++;
+	lambda=(double *)argv[i]; i++;
+	nl=*((IDL_LONG *)argv[i]); i++;
+	zvals=(double *)argv[i]; i++;
+	nz=*((IDL_LONG *)argv[i]); i++;
+	filter_n=(IDL_LONG *)argv[i]; i++;
+	filter_lambda=(double *)argv[i]; i++;
+	filter_pass=(double *)argv[i]; i++;
+	maxn=*((IDL_LONG *)argv[i]); i++;
+	
+	/* 1. run the fitting routine */
+	retval=(IDL_LONG) k_create_r(rmatrix,nk,nb,bmatrix,lambda,nl,zvals,nz,
+															 filter_n,filter_lambda,filter_pass,maxn);
+	
+	/* 2. free memory and leave */
+	free_memory();
+	return retval;
+}
+
+/***************************************************************************/
+

### 2002-01-25 src/idl_k_fit_coeffs.c

@@ -1,0 +1,49 @@
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include "kcorrect.h"
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+static void free_memory()
+{
+}
+
+/********************************************************************/
+IDL_LONG idl_k_fit_coeffs
+  (int      argc,
+   void *   argv[])
+{
+   IDL_LONG nt,nz,nk,nb,ngalaxy;
+   double *ematrix,*zvals,*rmatrix,*amatrix;
+	 double *galaxy_flux,*galaxy_invvar,*galaxy_z;
+
+	 IDL_LONG i;
+	 IDL_LONG retval=1;
+
+   /* 0. allocate pointers from IDL */
+	 i=0;
+	 ematrix=(double *)argv[i]; i++;
+   nt=*((IDL_LONG *)argv[i]); i++;
+	 zvals=(double *)argv[i]; i++;
+   nz=*((IDL_LONG *)argv[i]); i++;
+	 rmatrix=(double *)argv[i]; i++;
+   nk=*((IDL_LONG *)argv[i]); i++;
+   nb=*((IDL_LONG *)argv[i]); i++;
+	 amatrix=(double *)argv[i]; i++;
+	 galaxy_flux=(double *)argv[i]; i++;
+	 galaxy_invvar=(double *)argv[i]; i++;
+	 galaxy_z=(double *)argv[i]; i++;
+   ngalaxy=*((IDL_LONG *)argv[i]); i++;
+
+	 /* 1. run the fitting routine */
+	 retval=(IDL_LONG) k_fit_coeffs(ematrix,nt,zvals,nz,rmatrix,nk,nb,
+																	amatrix,galaxy_flux,galaxy_invvar,
+																	galaxy_z,ngalaxy);
+
+	 /* 2. free memory and leave */
+	 free_memory();
+   return retval;
+}
+
+/***************************************************************************/
+

### 2002-01-25 src/idl_k_fit_photoz.c

@@ -1,0 +1,49 @@
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include "kcorrect.h"
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+static void free_memory()
+{
+}
+
+/********************************************************************/
+IDL_LONG idl_k_fit_photoz
+  (int      argc,
+   void *   argv[])
+{
+   IDL_LONG nt,nz,nk,nb,ngalaxy;
+   double *ematrix,*zvals,*rmatrix,*amatrix;
+	 double *galaxy_flux,*galaxy_invvar,*galaxy_photoz;
+
+	 IDL_LONG i;
+	 IDL_LONG retval=1;
+
+   /* 0. allocate pointers from IDL */
+	 i=0;
+	 ematrix=(double *)argv[i]; i++;
+   nt=*((IDL_LONG *)argv[i]); i++;
+	 zvals=(double *)argv[i]; i++;
+   nz=*((IDL_LONG *)argv[i]); i++;
+	 rmatrix=(double *)argv[i]; i++;
+   nk=*((IDL_LONG *)argv[i]); i++;
+   nb=*((IDL_LONG *)argv[i]); i++;
+	 amatrix=(double *)argv[i]; i++;
+	 galaxy_flux=(double *)argv[i]; i++;
+	 galaxy_invvar=(double *)argv[i]; i++;
+	 galaxy_photoz=(double *)argv[i]; i++;
+   ngalaxy=*((IDL_LONG *)argv[i]); i++;
+
+	 /* 1. run the fitting routine */
+	 retval=(IDL_LONG) k_fit_photoz(ematrix,nt,zvals,nz,rmatrix,nk,nb,
+																	amatrix,galaxy_flux,galaxy_invvar,
+																	galaxy_photoz,ngalaxy);
+
+	 /* 2. free memory and leave */
+	 free_memory();
+   return retval;
+}
+
+/***************************************************************************/
+

### 2002-01-25 src/idl_k_fit_templates.c

@@ -1,0 +1,53 @@
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include "kcorrect.h"
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+static void free_memory()
+{
+}
+
+/********************************************************************/
+IDL_LONG idl_k_fit_templates
+  (int      argc,
+   void *   argv[])
+{
+   IDL_LONG nt,nz,nk,nb,ngalaxy,*galaxy_clip,initialized_dmatrix;
+   double *ematrix,*zvals,*rmatrix,*amatrix,*dmatrix;
+	 double *galaxy_flux,*galaxy_invvar,*galaxy_z;
+
+	 IDL_LONG i;
+	 IDL_LONG retval=1;
+
+   /* 0. allocate pointers from IDL */
+	 i=0;
+	 ematrix=(double *)argv[i]; i++;
+   nt=*((IDL_LONG *)argv[i]); i++;
+	 zvals=(double *)argv[i]; i++;
+   nz=*((IDL_LONG *)argv[i]); i++;
+	 rmatrix=(double *)argv[i]; i++;
+   nk=*((IDL_LONG *)argv[i]); i++;
+   nb=*((IDL_LONG *)argv[i]); i++;
+	 amatrix=(double *)argv[i]; i++;
+	 galaxy_flux=(double *)argv[i]; i++;
+	 galaxy_invvar=(double *)argv[i]; i++;
+	 galaxy_z=(double *)argv[i]; i++;
+	 galaxy_clip=(IDL_LONG *)argv[i]; i++;
+   ngalaxy=*((IDL_LONG *)argv[i]); i++;
+	 dmatrix=(double *)argv[i]; i++;
+   initialized_dmatrix=*((IDL_LONG *)argv[i]); i++;
+
+	 /* 1. run the fitting routine */
+	 retval=(IDL_LONG) k_fit_templates(ematrix,nt,zvals,nz,rmatrix,nk,nb,
+																		 amatrix,galaxy_flux,galaxy_invvar,
+																		 galaxy_z,galaxy_clip,ngalaxy,dmatrix,
+																		 initialized_dmatrix);
+
+	 /* 2. free memory and leave */
+	 free_memory();
+   return retval;
+}
+
+/***************************************************************************/
+

### 2002-01-25 src/k_brent.c

@@ -1,0 +1,76 @@
+#include <string.h>
+#include <stdio.h>
+#include <math.h>
+#define NRANSI
+#define ITMAX 100
+#define CGOLD 0.3819660
+#define ZEPS 1.0e-10
+#define SHFT(a,b,c,d) (a)=(b);(b)=(c);(c)=(d);
+#define SIGN(a,b) ((b) >= 0.0 ? fabs(a) : -fabs(a))
+
+double k_brent(double ax, double bx, double cx, double (*f)(double), 
+							 double tol, double *xmin)
+{
+	int iter;
+	double a,b,d,etemp,fu,fv,fw,fx,p,q,r,tol1,tol2,u,v,w,x,xm;
+	double e=0.0;
+
+	a=(ax < cx ? ax : cx);
+	b=(ax > cx ? ax : cx);
+	x=w=v=bx;
+	fw=fv=fx=(*f)(x);
+	for (iter=1;iter<=ITMAX;iter++) {
+		xm=0.5*(a+b);
+		tol2=2.0*(tol1=tol*fabs(x)+ZEPS);
+		if (fabs(x-xm) <= (tol2-0.5*(b-a))) {
+			*xmin=x;
+			return fx;
+		}
+		if (fabs(e) > tol1) {
+			r=(x-w)*(fx-fv);
+			q=(x-v)*(fx-fw);
+			p=(x-v)*q-(x-w)*r;
+			q=2.0*(q-r);
+			if (q > 0.0) p = -p;
+			q=fabs(q);
+			etemp=e;
+			e=d;
+			if (fabs(p) >= fabs(0.5*q*etemp) || p <= q*(a-x) || p >= q*(b-x))
+				d=CGOLD*(e=(x >= xm ? a-x : b-x));
+			else {
+				d=p/q;
+				u=x+d;
+				if (u-a < tol2 || b-u < tol2)
+					d=SIGN(tol1,xm-x);
+			}
+		} else {
+			d=CGOLD*(e=(x >= xm ? a-x : b-x));
+		}
+		u=(fabs(d) >= tol1 ? x+d : x+SIGN(tol1,d));
+		fu=(*f)(u);
+		if (fu <= fx) {
+			if (u >= x) a=x; else b=x;
+			SHFT(v,w,x,u)
+			SHFT(fv,fw,fx,fu)
+		} else {
+			if (u < x) a=u; else b=u;
+			if (fu <= fw || w == x) {
+				v=w;
+				w=u;
[…]

### 2002-01-25 src/k_create_r.c

@@ -1,0 +1,134 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_create_r.c
+ *
+ * Makes the rmatrix. This matrix is the impact of each filter curve
+ * on each template basis element. 
+ *
+ * Amazingly, this is the only part of the code in which the 
+ * filter curves are necessary.
+ *
+ * Mike Blanton
+ * 6/2001 */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+static double *cr_bmatrix=NULL;
+static double *cr_lambda=NULL;
+static IDL_LONG cr_nb;
+static IDL_LONG cr_nl;
+static double cr_z;
+static IDL_LONG cr_b;
+static IDL_LONG cr_filter_n;
+static double *cr_filter_lambda=NULL;
+static double *cr_filter_pass=NULL;
+
+double filter(double lambda) 
+{
+	double sl,rflambda,filt,spectrum;
+	unsigned long i,ip1,indxoff;
+
+	k_locate(cr_filter_lambda, cr_filter_n, lambda, &i);
+	if(i>=cr_filter_n-1 || i<0) return(0.);
+	ip1=i+1;
+	sl=(lambda-cr_filter_lambda[i])/(cr_filter_lambda[ip1]-cr_filter_lambda[i]);
+	filt=cr_filter_pass[i]+sl*(cr_filter_pass[ip1]-cr_filter_pass[i]);
+
+	rflambda=lambda/(1.+cr_z);
+	k_locate(cr_lambda, cr_nl, rflambda, &i);
+	if(i>=cr_nl-1 || i<0) return(0.);
+	ip1=i+1;
+	sl=(rflambda-cr_lambda[i])/(cr_lambda[ip1]-cr_lambda[i]);
+	indxoff=cr_b*cr_nl;
+	spectrum=cr_bmatrix[indxoff+i]
+		+sl*(cr_bmatrix[indxoff+ip1]-cr_bmatrix[indxoff+i]);
+	
+	filt=filt*lambda*spectrum;
+	return(filt);
+} /* end filter */
+
+double scalefilter(double lambda) 
+{
+	double sl,filt;
+	unsigned long i,ip1;
+
+	k_locate(cr_filter_lambda, cr_filter_n, lambda, &i);
+	if(i>=cr_filter_n-1 || i<0) return(0.);
+	ip1=i+1;
[…]

### 2002-01-25 src/k_ddot.f

@@ -1,0 +1,49 @@
+      double precision function k_ddot(n,dx,incx,dy,incy)
+c
+c     forms the dot product of two vectors.
+c     uses unrolled loops for increments equal to one.
+c     jack dongarra, linpack, 3/11/78.
+c     modified 12/3/93, array(1) declarations changed to array(*)
+c
+      double precision dx(*),dy(*),dtemp
+      integer i,incx,incy,ix,iy,m,mp1,n
+c
+      k_ddot = 0.0d0
+      dtemp = 0.0d0
+      if(n.le.0)return
+      if(incx.eq.1.and.incy.eq.1)go to 20
+c
+c        code for unequal increments or equal increments
+c          not equal to 1
+c
+      ix = 1
+      iy = 1
+      if(incx.lt.0)ix = (-n+1)*incx + 1
+      if(incy.lt.0)iy = (-n+1)*incy + 1
+      do 10 i = 1,n
+        dtemp = dtemp + dx(ix)*dy(iy)
+        ix = ix + incx
+        iy = iy + incy
+   10 continue
+      k_ddot = dtemp
+      return
+c
+c        code for both increments equal to 1
+c
+c
+c        clean-up loop
+c
+   20 m = mod(n,5)
+      if( m .eq. 0 ) go to 40
+      do 30 i = 1,m
+        dtemp = dtemp + dx(i)*dy(i)
+   30 continue
+      if( n .lt. 5 ) go to 60
+   40 mp1 = m + 1
+      do 50 i = mp1,n,5
+        dtemp = dtemp + dx(i)*dy(i) + dx(i + 1)*dy(i + 1) +
+     *   dx(i + 2)*dy(i + 2) + dx(i + 3)*dy(i + 3) + dx(i + 4)*dy(i + 4)
+   50 continue
+   60 k_ddot = dtemp
+      return
+      end

### 2002-01-25 src/k_dgemm.f

@@ -1,0 +1,313 @@
+      SUBROUTINE K_DGEMM ( TRANSA, TRANSB, M, N, K, ALPHA, A, LDA, B,
+     $     LDB,BETA, C, LDC )
+*     .. Scalar Arguments ..
+      CHARACTER*1        TRANSA, TRANSB
+      INTEGER            M, N, K, LDA, LDB, LDC
+      DOUBLE PRECISION   ALPHA, BETA
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * ), B( LDB, * ), C( LDC, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DGEMM  performs one of the matrix-matrix operations
+*
+*     C := alpha*op( A )*op( B ) + beta*C,
+*
+*  where  op( X ) is one of
+*
+*     op( X ) = X   or   op( X ) = X',
+*
+*  alpha and beta are scalars, and A, B and C are matrices, with op( A )
+*  an m by k matrix,  op( B )  a  k by n matrix and  C an m by n matrix.
+*
+*  Parameters
+*  ==========
+*
+*  TRANSA - CHARACTER*1.
+*           On entry, TRANSA specifies the form of op( A ) to be used in
+*           the matrix multiplication as follows:
+*
+*              TRANSA = 'N' or 'n',  op( A ) = A.
+*
+*              TRANSA = 'T' or 't',  op( A ) = A'.
+*
+*              TRANSA = 'C' or 'c',  op( A ) = A'.
+*
+*           Unchanged on exit.
+*
+*  TRANSB - CHARACTER*1.
+*           On entry, TRANSB specifies the form of op( B ) to be used in
+*           the matrix multiplication as follows:
+*
+*              TRANSB = 'N' or 'n',  op( B ) = B.
+*
+*              TRANSB = 'T' or 't',  op( B ) = B'.
+*
+*              TRANSB = 'C' or 'c',  op( B ) = B'.
+*
+*           Unchanged on exit.
+*
+*  M      - INTEGER.
+*           On entry,  M  specifies  the number  of rows  of the  matrix
+*           op( A )  and of the  matrix  C.  M  must  be at least  zero.
+*           Unchanged on exit.
+*
+*  N      - INTEGER.
+*           On entry,  N  specifies the number  of columns of the matrix
+*           op( B ) and the number of columns of the matrix C. N must be
+*           at least zero.
+*           Unchanged on exit.
+*
+*  K      - INTEGER.
[…]

### 2002-01-25 src/k_dgemv.f

@@ -1,0 +1,261 @@
+      SUBROUTINE K_DGEMV ( TRANS, M, N, ALPHA, A, LDA, X, INCX,
+     $                   BETA, Y, INCY )
+*     .. Scalar Arguments ..
+      DOUBLE PRECISION   ALPHA, BETA
+      INTEGER            INCX, INCY, LDA, M, N
+      CHARACTER*1        TRANS
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * ), X( * ), Y( * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DGEMV  performs one of the matrix-vector operations
+*
+*     y := alpha*A*x + beta*y,   or   y := alpha*A'*x + beta*y,
+*
+*  where alpha and beta are scalars, x and y are vectors and A is an
+*  m by n matrix.
+*
+*  Parameters
+*  ==========
+*
+*  TRANS  - CHARACTER*1.
+*           On entry, TRANS specifies the operation to be performed as
+*           follows:
+*
+*              TRANS = 'N' or 'n'   y := alpha*A*x + beta*y.
+*
+*              TRANS = 'T' or 't'   y := alpha*A'*x + beta*y.
+*
+*              TRANS = 'C' or 'c'   y := alpha*A'*x + beta*y.
+*
+*           Unchanged on exit.
+*
+*  M      - INTEGER.
+*           On entry, M specifies the number of rows of the matrix A.
+*           M must be at least zero.
+*           Unchanged on exit.
+*
+*  N      - INTEGER.
+*           On entry, N specifies the number of columns of the matrix A.
+*           N must be at least zero.
+*           Unchanged on exit.
+*
+*  ALPHA  - DOUBLE PRECISION.
+*           On entry, ALPHA specifies the scalar alpha.
+*           Unchanged on exit.
+*
+*  A      - DOUBLE PRECISION array of DIMENSION ( LDA, n ).
+*           Before entry, the leading m by n part of the array A must
+*           contain the matrix of coefficients.
+*           Unchanged on exit.
+*
+*  LDA    - INTEGER.
+*           On entry, LDA specifies the first dimension of A as declared
+*           in the calling (sub) program. LDA must be at least
+*           max( 1, m ).
+*           Unchanged on exit.
+*
+*  X      - DOUBLE PRECISION array of DIMENSION at least
[…]

### 2002-01-25 src/k_dposv.f

@@ -1,0 +1,122 @@
+      SUBROUTINE K_DPOSV( UPLO, N, NRHS, A, LDA, B, LDB, INFO )
+*
+*  -- LAPACK driver routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     March 31, 1993
+*
+*     .. Scalar Arguments ..
+      CHARACTER          UPLO
+      INTEGER            INFO, LDA, LDB, N, NRHS
+*     ..
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * ), B( LDB, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DPOSV computes the solution to a real system of linear equations
+*     A * X = B,
+*  where A is an N-by-N symmetric positive definite matrix and X and B
+*  are N-by-NRHS matrices.
+*
+*  The Cholesky decomposition is used to factor A as
+*     A = U**T* U,  if UPLO = 'U', or
+*     A = L * L**T,  if UPLO = 'L',
+*  where U is an upper triangular matrix and L is a lower triangular
+*  matrix.  The factored form of A is then used to solve the system of
+*  equations A * X = B.
+*
+*  Arguments
+*  =========
+*
+*  UPLO    (input) CHARACTER*1
+*          = 'U':  Upper triangle of A is stored;
+*          = 'L':  Lower triangle of A is stored.
+*
+*  N       (input) INTEGER
+*          The number of linear equations, i.e., the order of the
+*          matrix A.  N >= 0.
+*
+*  NRHS    (input) INTEGER
+*          The number of right hand sides, i.e., the number of columns
+*          of the matrix B.  NRHS >= 0.
+*
+*  A       (input/output) DOUBLE PRECISION array, dimension (LDA,N)
+*          On entry, the symmetric matrix A.  If UPLO = 'U', the leading
+*          N-by-N upper triangular part of A contains the upper
+*          triangular part of the matrix A, and the strictly lower
+*          triangular part of A is not referenced.  If UPLO = 'L', the
+*          leading N-by-N lower triangular part of A contains the lower
+*          triangular part of the matrix A, and the strictly upper
+*          triangular part of A is not referenced.
+*
[…]

### 2002-01-25 src/k_dpotf2.f

@@ -1,0 +1,168 @@
+      SUBROUTINE K_DPOTF2( UPLO, N, A, LDA, INFO )
+*
+*  -- LAPACK routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     February 29, 1992
+*
+*     .. Scalar Arguments ..
+      CHARACTER          UPLO
+      INTEGER            INFO, LDA, N
+*     ..
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DPOTF2 computes the Cholesky factorization of a real symmetric
+*  positive definite matrix A.
+*
+*  The factorization has the form
+*     A = U' * U ,  if UPLO = 'U', or
+*     A = L  * L',  if UPLO = 'L',
+*  where U is an upper triangular matrix and L is lower triangular.
+*
+*  This is the unblocked version of the algorithm, calling Level 2 BLAS.
+*
+*  Arguments
+*  =========
+*
+*  UPLO    (input) CHARACTER*1
+*          Specifies whether the upper or lower triangular part of the
+*          symmetric matrix A is stored.
+*          = 'U':  Upper triangular
+*          = 'L':  Lower triangular
+*
+*  N       (input) INTEGER
+*          The order of the matrix A.  N >= 0.
+*
+*  A       (input/output) DOUBLE PRECISION array, dimension (LDA,N)
+*          On entry, the symmetric matrix A.  If UPLO = 'U', the leading
+*          n by n upper triangular part of A contains the upper
+*          triangular part of the matrix A, and the strictly lower
+*          triangular part of A is not referenced.  If UPLO = 'L', the
+*          leading n by n lower triangular part of A contains the lower
+*          triangular part of the matrix A, and the strictly upper
+*          triangular part of A is not referenced.
+*
+*          On exit, if INFO = 0, the factor U or L from the Cholesky
+*          factorization A = U'*U  or A = L*L'.
+*
+*  LDA     (input) INTEGER
+*          The leading dimension of the array A.  LDA >= max(1,N).
+*
[…]

### 2002-01-25 src/k_dpotrf.f

@@ -1,0 +1,184 @@
+      SUBROUTINE K_DPOTRF( UPLO, N, A, LDA, INFO )
+*
+*  -- LAPACK routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     March 31, 1993
+*
+*     .. Scalar Arguments ..
+      CHARACTER          UPLO
+      INTEGER            INFO, LDA, N
+*     ..
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DPOTRF computes the Cholesky factorization of a real symmetric
+*  positive definite matrix A.
+*
+*  The factorization has the form
+*     A = U**T * U,  if UPLO = 'U', or
+*     A = L  * L**T,  if UPLO = 'L',
+*  where U is an upper triangular matrix and L is lower triangular.
+*
+*  This is the block version of the algorithm, calling Level 3 BLAS.
+*
+*  Arguments
+*  =========
+*
+*  UPLO    (input) CHARACTER*1
+*          = 'U':  Upper triangle of A is stored;
+*          = 'L':  Lower triangle of A is stored.
+*
+*  N       (input) INTEGER
+*          The order of the matrix A.  N >= 0.
+*
+*  A       (input/output) DOUBLE PRECISION array, dimension (LDA,N)
+*          On entry, the symmetric matrix A.  If UPLO = 'U', the leading
+*          N-by-N upper triangular part of A contains the upper
+*          triangular part of the matrix A, and the strictly lower
+*          triangular part of A is not referenced.  If UPLO = 'L', the
+*          leading N-by-N lower triangular part of A contains the lower
+*          triangular part of the matrix A, and the strictly upper
+*          triangular part of A is not referenced.
+*
+*          On exit, if INFO = 0, the factor U or L from the Cholesky
+*          factorization A = U**T*U or A = L*L**T.
+*
+*  LDA     (input) INTEGER
+*          The leading dimension of the array A.  LDA >= max(1,N).
+*
+*  INFO    (output) INTEGER
+*          = 0:  successful exit
+*          < 0:  if INFO = -i, the i-th argument had an illegal value
[…]

### 2002-01-25 src/k_dpotrs.f

@@ -1,0 +1,133 @@
+      SUBROUTINE K_DPOTRS( UPLO, N, NRHS, A, LDA, B, LDB, INFO )
+*
+*  -- LAPACK routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     March 31, 1993
+*
+*     .. Scalar Arguments ..
+      CHARACTER          UPLO
+      INTEGER            INFO, LDA, LDB, N, NRHS
+*     ..
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * ), B( LDB, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DPOTRS solves a system of linear equations A*X = B with a symmetric
+*  positive definite matrix A using the Cholesky factorization
+*  A = U**T*U or A = L*L**T computed by DPOTRF.
+*
+*  Arguments
+*  =========
+*
+*  UPLO    (input) CHARACTER*1
+*          = 'U':  Upper triangle of A is stored;
+*          = 'L':  Lower triangle of A is stored.
+*
+*  N       (input) INTEGER
+*          The order of the matrix A.  N >= 0.
+*
+*  NRHS    (input) INTEGER
+*          The number of right hand sides, i.e., the number of columns
+*          of the matrix B.  NRHS >= 0.
+*
+*  A       (input) DOUBLE PRECISION array, dimension (LDA,N)
+*          The triangular factor U or L from the Cholesky factorization
+*          A = U**T*U or A = L*L**T, as computed by DPOTRF.
+*
+*  LDA     (input) INTEGER
+*          The leading dimension of the array A.  LDA >= max(1,N).
+*
+*  B       (input/output) DOUBLE PRECISION array, dimension (LDB,NRHS)
+*          On entry, the right hand side matrix B.
+*          On exit, the solution matrix X.
+*
+*  LDB     (input) INTEGER
+*          The leading dimension of the array B.  LDB >= max(1,N).
+*
+*  INFO    (output) INTEGER
+*          = 0:  successful exit
+*          < 0:  if INFO = -i, the i-th argument had an illegal value
+*
[…]

### 2002-01-25 src/k_dscal.f

@@ -1,0 +1,43 @@
+      subroutine  k_dscal(n,da,dx,incx)
+c
+c     scales a vector by a constant.
+c     uses unrolled loops for increment equal to one.
+c     jack dongarra, linpack, 3/11/78.
+c     modified 3/93 to return if incx .le. 0.
+c     modified 12/3/93, array(1) declarations changed to array(*)
+c
+      double precision da,dx(*)
+      integer i,incx,m,mp1,n,nincx
+c
+      if( n.le.0 .or. incx.le.0 )return
+      if(incx.eq.1)go to 20
+c
+c        code for increment not equal to 1
+c
+      nincx = n*incx
+      do 10 i = 1,nincx,incx
+        dx(i) = da*dx(i)
+   10 continue
+      return
+c
+c        code for increment equal to 1
+c
+c
+c        clean-up loop
+c
+   20 m = mod(n,5)
+      if( m .eq. 0 ) go to 40
+      do 30 i = 1,m
+        dx(i) = da*dx(i)
+   30 continue
+      if( n .lt. 5 ) return
+   40 mp1 = m + 1
+      do 50 i = mp1,n,5
+        dx(i) = da*dx(i)
+        dx(i + 1) = da*dx(i + 1)
+        dx(i + 2) = da*dx(i + 2)
+        dx(i + 3) = da*dx(i + 3)
+        dx(i + 4) = da*dx(i + 4)
+   50 continue
+      return
+      end

### 2002-01-25 src/k_dsyrk.f

@@ -1,0 +1,294 @@
+      SUBROUTINE K_DSYRK ( UPLO, TRANS, N, K, ALPHA, A, LDA,
+     $                   BETA, C, LDC )
+*     .. Scalar Arguments ..
+      CHARACTER*1        UPLO, TRANS
+      INTEGER            N, K, LDA, LDC
+      DOUBLE PRECISION   ALPHA, BETA
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * ), C( LDC, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DSYRK  performs one of the symmetric rank k operations
+*
+*     C := alpha*A*A' + beta*C,
+*
+*  or
+*
+*     C := alpha*A'*A + beta*C,
+*
+*  where  alpha and beta  are scalars, C is an  n by n  symmetric matrix
+*  and  A  is an  n by k  matrix in the first case and a  k by n  matrix
+*  in the second case.
+*
+*  Parameters
+*  ==========
+*
+*  UPLO   - CHARACTER*1.
+*           On  entry,   UPLO  specifies  whether  the  upper  or  lower
+*           triangular  part  of the  array  C  is to be  referenced  as
+*           follows:
+*
+*              UPLO = 'U' or 'u'   Only the  upper triangular part of  C
+*                                  is to be referenced.
+*
+*              UPLO = 'L' or 'l'   Only the  lower triangular part of  C
+*                                  is to be referenced.
+*
+*           Unchanged on exit.
+*
+*  TRANS  - CHARACTER*1.
+*           On entry,  TRANS  specifies the operation to be performed as
+*           follows:
+*
+*              TRANS = 'N' or 'n'   C := alpha*A*A' + beta*C.
+*
+*              TRANS = 'T' or 't'   C := alpha*A'*A + beta*C.
+*
+*              TRANS = 'C' or 'c'   C := alpha*A'*A + beta*C.
+*
+*           Unchanged on exit.
+*
+*  N      - INTEGER.
+*           On entry,  N specifies the order of the matrix C.  N must be
+*           at least zero.
+*           Unchanged on exit.
+*
+*  K      - INTEGER.
+*           On entry with  TRANS = 'N' or 'n',  K  specifies  the number
+*           of  columns   of  the   matrix   A,   and  on   entry   with
+*           TRANS = 'T' or 't' or 'C' or 'c',  K  specifies  the  number
[…]

### 2002-01-25 src/k_dtrsm.f

@@ -1,0 +1,378 @@
+      SUBROUTINE K_DTRSM ( SIDE, UPLO, TRANSA, DIAG, M, N, ALPHA, A, LDA,
+     $                   B, LDB )
+*     .. Scalar Arguments ..
+      CHARACTER*1        SIDE, UPLO, TRANSA, DIAG
+      INTEGER            M, N, LDA, LDB
+      DOUBLE PRECISION   ALPHA
+*     .. Array Arguments ..
+      DOUBLE PRECISION   A( LDA, * ), B( LDB, * )
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  DTRSM  solves one of the matrix equations
+*
+*     op( A )*X = alpha*B,   or   X*op( A ) = alpha*B,
+*
+*  where alpha is a scalar, X and B are m by n matrices, A is a unit, or
+*  non-unit,  upper or lower triangular matrix  and  op( A )  is one  of
+*
+*     op( A ) = A   or   op( A ) = A'.
+*
+*  The matrix X is overwritten on B.
+*
+*  Parameters
+*  ==========
+*
+*  SIDE   - CHARACTER*1.
+*           On entry, SIDE specifies whether op( A ) appears on the left
+*           or right of X as follows:
+*
+*              SIDE = 'L' or 'l'   op( A )*X = alpha*B.
+*
+*              SIDE = 'R' or 'r'   X*op( A ) = alpha*B.
+*
+*           Unchanged on exit.
+*
+*  UPLO   - CHARACTER*1.
+*           On entry, UPLO specifies whether the matrix A is an upper or
+*           lower triangular matrix as follows:
+*
+*              UPLO = 'U' or 'u'   A is an upper triangular matrix.
+*
+*              UPLO = 'L' or 'l'   A is a lower triangular matrix.
+*
+*           Unchanged on exit.
+*
+*  TRANSA - CHARACTER*1.
+*           On entry, TRANSA specifies the form of op( A ) to be used in
+*           the matrix multiplication as follows:
+*
+*              TRANSA = 'N' or 'n'   op( A ) = A.
+*
+*              TRANSA = 'T' or 't'   op( A ) = A'.
+*
+*              TRANSA = 'C' or 'c'   op( A ) = A'.
+*
+*           Unchanged on exit.
+*
+*  DIAG   - CHARACTER*1.
+*           On entry, DIAG specifies whether or not A is unit triangular
+*           as follows:
+*
[…]

### 2002-01-25 src/k_fileopen.c

@@ -1,0 +1,21 @@
+#include <string.h>
+#include <stdio.h>
+
+/*
+ * k_fileopen.c : contains routine to open a file 
+ *              Created 9/28/95 MB
+ */
+
+/* open a file and return file pointer */
+FILE *k_fileopen(const char *Filename,
+								 const char *Mode)
+{
+	FILE *fp;
+	
+	if (!(fp=fopen(Filename,Mode))) {
+		fprintf(stderr,"Could not open file for %s: %s",Mode,Filename);
+		exit(1);
+	} /* end if */
+
+	return(fp);
+} /* end fileopen() */

### 2002-01-25 src/k_fit_coeffs.c

@@ -1,0 +1,131 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_fit_coeffs.c
+ *
+ * Finds the best-fit contribution of each template to each galaxy,
+ * given the templates.
+ *
+ * Mike Blanton
+ * 2/2001
+ */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+static double *cmatrix=NULL;
+static double *covar=NULL;
+static double *rhs=NULL;
+
+/* fit coefficients, given information about the templates and 
+ * the filters (in ematrix and rmatrix) and a set of galaxies with
+ * redshifts */
+IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
+											IDL_LONG nt,             /* number of eigentemplates */
+											double *zvals,      /* z interpolation */
+											IDL_LONG nz,
+											double *rmatrix,    /* r matrix */
+											IDL_LONG nk,             /* number of bandpasses */
+											IDL_LONG nb,             /* number of templates */
+											double *amatrix, /* coefficients */
+											double *galaxy_flux, /* galaxy fluxes [i][k], 
+																							redshifts */
+											double *galaxy_invvar,
+											double *galaxy_z,
+											IDL_LONG ngalaxy)
+{
+	double currz;
+	char uplo;
+	IDL_LONG i,j,jp,k,b,indx,unity,info;
+	
+	/*
+	 * 1. Make cmatrix
+	 */
+#if 0
+	printf("    (making cmatrix)\n");
+	fflush(stdout);
+#endif
+	cmatrix=(double *) malloc(ngalaxy*nk*nt*sizeof(double));
+	for(i=0;i<ngalaxy;i++) {
+		currz=(galaxy_z[i]<0.5*(zvals[0]+zvals[1])) ? 0.5*(zvals[0]+zvals[1]) :
+			galaxy_z[i];
+		for(k=0;k<nk;k++)
+			for(j=0;j<nt;j++) {
+				indx=i*nk*nt+k*nt+j;
[…]

### 2002-01-25 src/k_fit_photoz.c

@@ -1,0 +1,334 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_fit_photoz.c
+ *
+ * Finds the best-fit contribution of each template, and 
+ * the best redshift estimate, to each galaxy, given the templates.
+ *
+ * Mike Blanton
+ * 1/2002
+ */
+
+#define ZRES 0.10
+#define TOL 0.005
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+static double *pz_ematrix=NULL;
+static double *pz_rmatrix=NULL;
+static double *pz_zvals=NULL;
+static double *pz_coeffs=NULL;
+static double *pz_model_flux=NULL;
+static double *pz_galaxy_flux=NULL;
+static double *pz_galaxy_invvar=NULL;
+static double *pz_ematrix1=NULL;
+static double *pz_coeffspnorm=NULL;
+static double *pz_coeffs1=NULL;
+static double pz_z,pz_linepos;
+static IDL_LONG pz_nt,pz_nz,pz_nk,pz_nb,pz_puse,pz_np,pz_p;
+
+#if 0
+static double pz_coeffsps[12]={
+	0.07,0.08,-0.19,
+	0.011,0.085,-0.17,
+	0.009,0.09,-0.15,
+	0.03,0.09,-0.1
+};
+static double pz_coeffspe[12]={
+	0.011,0.085,-0.17,
+	0.009,0.09,-0.15,
+	0.03,0.09,-0.1,
+	0.12,0.08,0.14
+};
+#endif
+
+static double pz_coeffsps[12]={
+	0.04,0.082,-0.185,
+	0.04,0.082,-0.175,
+	0.04,0.082,-0.15,
+	0.04,0.082,-0.1
+};
+static double pz_coeffspe[12]={
+	0.04,0.082,-0.175,
+	0.04,0.082,-0.15,
+	0.04,0.082,-0.1,
+	0.04,0.082,0.0
+};
+
+double pz_calc_chi2(double z) 
+{
+	double *model_flux,chi2;
+	IDL_LONG k;
+	
[…]

### 2002-01-25 src/k_fit_templates.c

@@ -1,0 +1,130 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * fitTemplates.c
+ *
+ * Updates the eigentemplates given the new coefficients.
+ *
+ * Mike Blanton
+ * 1/2002 */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+static double *covar=NULL;
+static double *rhs=NULL;
+
+/* finds best fit eigentemplates, given the galaxy properties and 
+ * a previous guess at the best coefficients to the templates (in
+ * amatrix) */
+IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
+												 IDL_LONG nt,       /* number of eigentemplates */
+												 double *zvals,      /* z interpolation */
+												 IDL_LONG nz,
+												 double *rmatrix,    /* r matrix */
+												 IDL_LONG nk,             /* number of bandpasses */
+												 IDL_LONG nb,             /* number of templates */
+												 double *amatrix, /* current coefficients */
+												 double *galaxy_flux, /* galaxy fluxes [i][k] redshifts */
+												 double *galaxy_invvar,
+												 double *galaxy_z,
+												 IDL_LONG *galaxy_clip,
+												 IDL_LONG ngalaxy,
+												 double *dmatrix,    /* d matrix */
+												 IDL_LONG initialized_dmatrix)
+{
+	char uplo;
+	IDL_LONG i,j,jp,k,b,bp,unity,indx,info,order,use;
+	
+	/* 
+	 * 1. Create system for each galaxy and solve for the appropriate
+	 * parameters, and solve it. (Ax=b)
+	 */
+
+	/* 1a. create A */
+	printf("     (creating A matrix)\n");
+	fflush(stdout);
+	covar=(double *) malloc(nt*nb*nt*nb*sizeof(double));
+	if(!initialized_dmatrix) {
+		for(i=0;i<ngalaxy;i++) {
+			use=1;
+			for(k=0;k<nk;k++)
[…]

### 2002-01-25 src/k_ilaenv.f

@@ -1,0 +1,506 @@
+      INTEGER          FUNCTION K_ILAENV( ISPEC, NAME, OPTS, N1, N2, N3,
+     $                 N4 )
+*
+*  -- LAPACK auxiliary routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     September 30, 1994
+*
+*     .. Scalar Arguments ..
+      CHARACTER*( * )    NAME, OPTS
+      INTEGER            ISPEC, N1, N2, N3, N4
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  K_ILAENV is called from the LAPACK routines to choose problem-dependent
+*  parameters for the local environment.  See ISPEC for a description of
+*  the parameters.
+*
+*  This version provides a set of parameters which should give good,
+*  but not optimal, performance on many of the currently available
+*  computers.  Users are encouraged to modify this subroutine to set
+*  the tuning parameters for their particular machine using the option
+*  and problem size information in the arguments.
+*
+*  This routine will not function correctly if it is converted to all
+*  lower case.  Converting it to all upper case is allowed.
+*
+*  Arguments
+*  =========
+*
+*  ISPEC   (input) INTEGER
+*          Specifies the parameter to be returned as the value of
+*          K_ILAENV.
+*          = 1: the optimal blocksize; if this value is 1, an unblocked
+*               algorithm will give the best performance.
+*          = 2: the minimum block size for which the block routine
+*               should be used; if the usable block size is less than
+*               this value, an unblocked routine should be used.
+*          = 3: the crossover point (in a block routine, for N less
+*               than this value, an unblocked routine should be used)
+*          = 4: the number of shifts, used in the nonsymmetric
+*               eigenvalue routines
+*          = 5: the minimum column dimension for blocking to be used;
[…]

### 2002-01-25 src/k_interpolate.c

@@ -1,0 +1,49 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_interpolate.c
+ *
+ * Just a linear interpolation of a vector
+ *
+ * Mike Blanton
+ * 2/2001
+ *
+ */
+
+/* linear interpolate in a given vector; out of bounds --> 0 */
+double k_interpolate(double currpos,
+										 double vals[],
+										 double pos[],
+										 IDL_LONG n)
+{
+	unsigned long i,ip1;
+	double sp,currval;
+
+	k_locate(pos, (unsigned long) n, currpos, &i);
+	if(i>=n-1 || i<0) return(0.);
+	ip1=i+1;
+	sp=currpos-pos[i];
+	currval=vals[i]+sp*(vals[ip1]-vals[i])/(pos[ip1]-pos[i]);
+	return(currval);
+} /* end interpolate templates */
+
+/* linear interpolate in a given vector; out of bounds --> 0;
+ * assume x-axis is given equally spaced */
+double k_interpolate_es(double currpos,
+											double vals[],
+											double pos[],
+											IDL_LONG n)
+{
+	unsigned long i,ip1;
+	double sp,currval;
+
+	i=(unsigned long) floor((double) n*(currpos-pos[0])/(2.*pos[n-1]-pos[n-2]));
+	if(i>=n-1 || i<0) return(0.);
+	ip1=i+1;
+	sp=currpos-pos[i];
+	currval=vals[i]+sp*(vals[ip1]-vals[i])/(pos[ip1]-pos[i]);
+	return(currval);
+} /* end interpolate templates */

### 2002-01-25 src/k_load_ascii_table.c

@@ -1,0 +1,62 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_load_ascii_table.c
+ *
+ * Load an ascii table, of the form:
+ *
+ * <ndim> <size_{0}> ... <size_{ndim-1}>
+ * <entry_0>
+ * <entry_1>
+ * ...
+ * <entry_n>
+ * ...
+ * <entry_{size_0*size_1*..*size_{ndim-1}-1>
+ *
+ * where the table element [i][j][k] (for ndim==3) would be the 
+ * entry element n, where n=i*size_2*size1+j*size2+k
+ *
+ * Mike Blanton
+ * 1/2002
+ */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+/* load an ascii table in my standard format */
+IDL_LONG k_load_ascii_table(double **table,
+														IDL_LONG *ndim,
+														IDL_LONG **sizes,
+														char filename[])
+{
+	IDL_LONG i,nelem;
+	FILE *fp;
+
+	FREEVEC((*table));
+	FREEVEC((*sizes));
+
+	fp=k_fileopen(filename,"r");
+	fscanf(fp,"%d",ndim);
+	if((*ndim)<=0) {
+		fprintf(stderr,"k_load_ascii_table found 0-dimensional table in %s\n",
+						filename);
+		exit(1);
+	} /* end if */
+	nelem=1;
+	(*sizes)=(IDL_LONG *) malloc((*ndim)*sizeof(IDL_LONG));
+	for(i=0;i<(*ndim);i++) {
+		fscanf(fp,"%d",&((*sizes)[i]));
+		nelem*=(*sizes)[i];
+	} /* end for i */
+	if(nelem<=0) {
+		fprintf(stderr,"k_load_ascii_table found 0-size table in %s\n",
+						filename);
+		exit(1);
+	} /* end if */
+	(*table)=(double *) malloc(nelem*sizeof(double));
+	for(i=0;i<nelem;i++) 
+		fscanf(fp,"%lf",&((*table)[i]));
+	fclose(fp);
+} /* end k_load_ascii_table */

### 2002-01-25 src/k_load_filters.c

@@ -1,0 +1,59 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_load_filters.c
+ *
+ * Mike Blanton
+ * 1/2002
+ */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+/* load the filters in their standard format */
+IDL_LONG k_load_filters(IDL_LONG **filter_n,
+												double **filter_lambda,
+												double **filter_pass,
+												IDL_LONG *maxn,
+												IDL_LONG *nk,
+												char filterlist[])
+{
+	IDL_LONG i,j;
+	FILE *listfp,*fp;
+	char filename[2000];
+
+	FREEVEC((*filter_n));
+	FREEVEC((*filter_lambda));
+	FREEVEC((*filter_pass));
+
+	listfp=k_fileopen(filterlist,"r");
+	fscanf(listfp,"%d",nk);
+	(*filter_n)=(IDL_LONG *) malloc((*nk)*sizeof(IDL_LONG));
+	(*maxn)=0;
+	for(i=0;i<(*nk);i++) {
+		fscanf(listfp,"%s",filename);
+		fp=k_fileopen(filename,"r");
+		fscanf(fp,"%d",&((*filter_n)[i]));
+		fclose(fp);
+		if((*filter_n)[i]>(*maxn)) (*maxn)=(*filter_n)[i];
+	} /* end for i */
+	fclose(listfp);
+		
+	(*filter_lambda)=(double *) malloc((*nk)*(*maxn)*sizeof(double));
+	(*filter_pass)=(double *) malloc((*nk)*(*maxn)*sizeof(double));
+	listfp=k_fileopen(filterlist,"r");
+	fscanf(listfp,"%d",nk);
+	for(i=0;i<(*nk);i++) {
+		fscanf(listfp,"%s",filename);
+		fp=k_fileopen(filename,"r");
+		fscanf(fp,"%d",&((*filter_n)[i]));
+		for(j=0;j<(*filter_n)[i];j++)
+			fscanf(fp,"%lf %lf",&((*filter_lambda)[i*(*maxn)+j]),
+						 &((*filter_pass)[i*(*maxn)+j]));
+		fclose(fp);
+	} /* end for i */
+	fclose(listfp);
+	
+} /* end k_load_ascii_table */

### 2002-01-25 src/k_locate.c

@@ -1,0 +1,21 @@
+/* 
+ * I have altered this routine to be compatible with zero-offset
+ * arrays
+ */
+void k_locate(double xx[], unsigned long n, double x, unsigned long *j)
+{
+	unsigned long ju,jm,jl;
+	int ascnd;
+
+	jl=-1;
+	ju=n;
+	ascnd=(xx[n-1] > xx[0]);
+	while (ju-jl > 1) {
+		jm=(ju+jl) >> 1;
+		if (x > xx[jm] == ascnd)
+			jl=jm;
+		else
+			ju=jm;
+	}
+	*j=jl;
+}

### 2002-01-25 src/k_lsame.f

@@ -1,0 +1,87 @@
+      LOGICAL          FUNCTION K_LSAME( CA, CB )
+*
+*  -- LAPACK auxiliary routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     September 30, 1994
+*
+*     .. Scalar Arguments ..
+      CHARACTER          CA, CB
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  LSAME returns .TRUE. if CA is the same letter as CB regardless of
+*  case.
+*
+*  Arguments
+*  =========
+*
+*  CA      (input) CHARACTER*1
+*  CB      (input) CHARACTER*1
+*          CA and CB specify the single characters to be compared.
+*
+* =====================================================================
+*
+*     .. Intrinsic Functions ..
+      INTRINSIC          ICHAR
+*     ..
+*     .. Local Scalars ..
+      INTEGER            INTA, INTB, ZCODE
+*     ..
+*     .. Executable Statements ..
+*
+*     Test if the characters are equal
+*
+      K_LSAME = CA.EQ.CB
+      IF( K_LSAME )
+     $   RETURN
+*
+*     Now test for equivalence if both characters are alphabetic.
+*
+      ZCODE = ICHAR( 'Z' )
+*
+*     Use 'Z' rather than 'A' so that ASCII can be detected on Prime
+*     machines, on which ICHAR returns a value with bit 8 set.
+*     ICHAR('A') on Prime machines returns 193 which is the same as
+*     ICHAR('A') on an EBCDIC machine.
+*
+      INTA = ICHAR( CA )
+      INTB = ICHAR( CB )
+*
+      IF( ZCODE.EQ.90 .OR. ZCODE.EQ.122 ) THEN
+*
+*        ASCII is assumed - ZCODE is the ASCII code of either lower or
+*        upper case 'Z'.
+*
+         IF( INTA.GE.97 .AND. INTA.LE.122 ) INTA = INTA - 32
+         IF( INTB.GE.97 .AND. INTB.LE.122 ) INTB = INTB - 32
+*
+      ELSE IF( ZCODE.EQ.233 .OR. ZCODE.EQ.169 ) THEN
+*
[…]

### 2002-01-25 src/k_midpnt.c

@@ -1,0 +1,32 @@
+#include <string.h>
+#include <stdio.h>
+#include <math.h>
+#include "kcorrect.h"
+#define FUNC(x) ((*func)(x))
+
+double k_midpnt(double (*func)(double), double a, double b, IDL_LONG n)
+{
+	double x,tnm,sum,del,ddel;
+	static double s;
+	IDL_LONG it,j;
+
+	if (n == 1) {
+		return (s=(b-a)*FUNC(0.5*(a+b)));
+	} else {
+		for(it=1,j=1;j<n-1;j++) it *= 3;
+		tnm=it;
+		del=(b-a)/(3.0*tnm);
+		ddel=del+del;
+		x=a+0.5*del;
+		sum=0.0;
+		for (j=1;j<=it;j++) {
+			sum += FUNC(x);
+			x += ddel;
+			sum += FUNC(x);
+			x += del;
+		}
+		s=(s+(b-a)*sum/tnm)/3.0;
+		return s;
+	}
+}
+#undef FUNC

### 2002-01-25 src/k_model_fluxes.c

@@ -1,0 +1,44 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_model_fluxes.c
+ *
+ * Returns model fluxes, given the model
+ *
+ * Mike Blanton
+ * 1/2002
+ */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+/* calculate the model fluxes, given the coeffs of the model */
+IDL_LONG k_model_fluxes(double *ematrix,    /* eigentemplates */
+												IDL_LONG nt,             /* number of eigentemplates */
+												double *zvals,      /* z interpolation */
+												IDL_LONG nz,
+												double *rmatrix,    /* r matrix */
+												IDL_LONG nk,             /* number of bandpasses */
+												IDL_LONG nb,             /* number of templates */
+												double *coeffs, /* coefficients */
+												double *galaxy_z,
+												double *model_flux,
+												IDL_LONG ngalaxy)
+{
+	IDL_LONG i,j,k,b;
+	
+	for(i=0;i<ngalaxy;i++) {
+		for(k=0;k<nk;k++) {
+			model_flux[k]=0.;
+			for(b=0;b<nb;b++) {
+				for(j=0;j<nt;j++) 
+					model_flux[k+i*nk]+=coeffs[i*nt+j]*ematrix[j*nb+b]*
+						k_interpolate(galaxy_z[i],&(rmatrix[k*nb*nz+b*nz]),zvals,nz);
+			} /* end for b */
+		} /* end for k */
+	} /* end for i */
+
+	return(1);
+} /* end k_fit_photoz */

### 2002-01-25 src/k_pca_etemplates.c

@@ -1,0 +1,143 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "utils.h"
+
+/*
+ * k_pca_etemplates.c
+ *
+ * Use PCA on the templates! Assumes they are orthonormal.
+ *
+ * Mike Blanton
+ * 2/2001
+ *
+ */
+
+static double *covar=NULL;
+static double *tmpb=NULL;
+static double *tmpa=NULL;
+static double *amean=NULL;
+static double *ameanold=NULL;
+static double *eval=NULL;
+static double *fv1=NULL;
+static double *f=NULL;
+
+void k_pca_etemplates(double *amatrix,
+											IDL_LONG nt,
+											IDL_LONG ngalaxy,
+											double *ematrix,
+											IDL_LONG nb,
+											double *bflux,
+											IDL_LONG nclip,
+											IDL_LONG niter)
+{
+	double tmp,cliplimit;
+	IDL_LONG clip,i,j,jp,b,ierr,unity,order,threent,use,effn;
+	char jobz,uplo;
+
+	amean=(double *) malloc((nt-1)*sizeof(double));
+	ameanold=(double *) malloc((nt-1)*sizeof(double));
+	covar=(double *) malloc((nt-1)*(nt-1)*sizeof(double));
+	cliplimit=1.e+10;
+	for(j=1;j<nt;j++) 
+		amean[j-1]=0.;
+	for(clip=0;clip<niter;clip++) {
+		
+		for(j=1;j<nt;j++) {
+			ameanold[j-1]=amean[j-1];
+			amean[j-1]=0.;
+		} /* end for j */
+		
+		effn=0;
+		for(i=0;i<ngalaxy;i++) {
+			use=1;
+			for(j=1;j<nt;j++) 
+				if(amatrix[i*nt+j]==0. ||
+					 fabs(amatrix[i*nt+j]/amatrix[i*nt]-ameanold[j-1])>cliplimit)
+					use=0;
+			if(use) {
+				for(j=1;j<nt;j++) 
+					amean[j-1]+=amatrix[i*nt+j]/amatrix[i*nt];
+				effn++;
+			} /* end if */
+		} /* end for i */
+		for(j=1;j<nt;j++)
[…]

### 2002-01-25 src/k_polint.c

@@ -1,0 +1,44 @@
+#include <string.h>
+#include <stdio.h>
+#include <math.h>
+#include "kcorrect.h"
+#define NRANSI
+
+void k_polint(double xa[], double ya[], IDL_LONG n, double x, double *y, 
+							double *dy)
+{
+	IDL_LONG i,m,ns=1;
+	double den,dif,dift,ho,hp,w;
+	double *c,*d;
+
+	dif=fabs(x-xa[1]);
+	c=k_vector(1,n);
+	d=k_vector(1,n);
+	for (i=1;i<=n;i++) {
+		if ( (dift=fabs(x-xa[i])) < dif) {
+			ns=i;
+			dif=dift;
+		}
+		c[i]=ya[i];
+		d[i]=ya[i];
+	}
+	*y=ya[ns--];
+	for (m=1;m<n;m++) {
+		for (i=1;i<=n-m;i++) {
+			ho=xa[i]-x;
+			hp=xa[i+m]-x;
+			w=c[i+1]-d[i];
+			if ( (den=ho-hp) == 0.0) {
+				printf("Error in routine polint");
+				exit(1);
+			}
+			den=w/den;
+			d[i]=hp*den;
+			c[i]=ho*den;
+		}
+		*y += (*dy=(2*ns < (n-m) ? c[ns+1] : d[ns--]));
+	}
+	k_free_vector(d,1,n);
+	k_free_vector(c,1,n);
+}
+#undef NRANSI

### 2002-01-25 src/k_qromo.c

@@ -1,0 +1,34 @@
+#include <string.h>
+#include <stdio.h>
+#include <math.h>
+#include "kcorrect.h"
+#define EPS 1.0e-4
+#define JMAX 14
+#define JMAXP (JMAX+1)
+#define K 5
+
+double k_qromo(double (*func)(double), double a, double b,
+							 double (*choose)(double(*)(double), double, double, IDL_LONG))
+{
+	IDL_LONG j;
+	double ss,dss,h[JMAXP+1],s[JMAXP+1];
+
+	h[1]=1.0;
+	for (j=1;j<=JMAX;j++) {
+		s[j]=(*choose)(func,a,b,j);
+		if (j >= K) {
+			k_polint(&h[j-K],&s[j-K],K,0.0,&ss,&dss);
+			if (fabs(dss) < EPS*fabs(ss) || ss<EPS) return ss;
+		}
+		s[j+1]=s[j];
+		h[j+1]=h[j]/9.0;
+	}
+	printf("%le %le\n",dss,ss);
+	printf("Too many steps in routing qromo");
+	exit(1);
+	return 0.0;
+}
+#undef EPS
+#undef JMAX
+#undef JMAXP
+#undef K

### 2002-01-25 src/k_utils.c

@@ -1,0 +1,24 @@
+#include <stdio.h>
+#include <stddef.h>
+#include <stdlib.h>
+#define NR_END 1
+#define FREE_ARG char*
+
+double *k_vector(long nl, long nh)
+/* allocate a double vector with subscript range v[nl..nh] */
+{
+	double *v;
+
+	v=(double *)malloc((size_t) ((nh-nl+1+NR_END)*sizeof(double)));
+	if (!v) {
+		fprintf(stderr,"allocation failure in vector()");
+		exit(1);
+	}
+	return v-nl+NR_END;
+}
+
+void k_free_vector(double *v, long nl, long nh)
+/* free a double vector allocated with vector() */
+{
+	free((FREE_ARG) (v+nl-NR_END));
+}

### 2002-01-25 src/k_write_ascii_table.c

@@ -1,0 +1,49 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * k_write_ascii_table.c
+ *
+ * Write an ascii table, of the form:
+ *
+ * <ndim> <size_{0}> ... <size_{ndim-1}>
+ * <entry_0>
+ * <entry_1>
+ * ...
+ * <entry_n>
+ * ...
+ * <entry_{size_0*size_1*..*size_{ndim-1}-1>
+ *
+ * where the table element [i][j][k] (for ndim==3) would be the 
+ * entry element n, where n=i*size_2*size1+j*size2+k
+ *
+ * Mike Blanton
+ * 1/2002
+ */
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+/* write an ascii table in my standard format */
+IDL_LONG k_write_ascii_table(double *table,
+														 IDL_LONG ndim,
+														 IDL_LONG *sizes,
+														 char filename[])
+{
+	IDL_LONG i,nelem;
+	FILE *fp;
+
+	fp=k_fileopen(filename,"w");
+	fprintf(fp,"%d ",ndim);
+	for(i=0;i<ndim;i++) 
+		fprintf(fp,"%d ",sizes[i]);
+	fprintf(fp,"\n");
+	nelem=1;
+	for(i=0;i<ndim;i++) 
+		nelem*=sizes[i];
+	for(i=0;i<nelem;i++) 
+		fprintf(fp,"%e\n",table[i]);
+	fclose(fp);
+
+} /* end k_write_ascii_table */

### 2002-01-25 src/k_xerbla.f

@@ -1,0 +1,46 @@
+      SUBROUTINE K_XERBLA( SRNAME, INFO )
+*
+*  -- LAPACK auxiliary routine (version 2.0) --
+*     Univ. of Tennessee, Univ. of California Berkeley, NAG Ltd.,
+*     Courant Institute, Argonne National Lab, and Rice University
+*     September 30, 1994
+*
+*     .. Scalar Arguments ..
+      CHARACTER*6        SRNAME
+      INTEGER            INFO
+*     ..
+*
+*  Purpose
+*  =======
+*
+*  XERBLA  is an error handler for the LAPACK routines.
+*  It is called by an LAPACK routine if an input parameter has an
+*  invalid value.  A message is printed and execution stops.
+*
+*  Installers may consider modifying the STOP statement in order to
+*  call system-specific exception-handling facilities.
+*
+*  Arguments
+*  =========
+*
+*  SRNAME  (input) CHARACTER*6
+*          The name of the routine which called XERBLA.
+*
+*  INFO    (input) INTEGER
+*          The position of the invalid parameter in the parameter list
+*          of the calling routine.
+*
+* =====================================================================
+*
+*     .. Executable Statements ..
+*
+      WRITE( *, FMT = 9999 )SRNAME, INFO
+*
+      STOP
+*
+ 9999 FORMAT( ' ** On entry to ', A6, ' parameter number ', I2, ' had ',
+     $      'an illegal value' )
+*
+*     End of XERBLA
+*
+      END

### 2002-01-25 src/k_zbrent.c

@@ -1,0 +1,80 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#define NRANSI
+#define ITMAX 100
+#define EPS 3.0e-8
+#define SIGN(a,b) ((b) >= 0.0 ? fabs(a) : -fabs(a))
+
+double k_zbrent(double (*func)(double), double x1, double x2, double tol)
+{
+	int iter;
+	double a=x1,b=x2,c=x2,d,e,min1,min2;
+	double fa=(*func)(a),fb=(*func)(b),fc,p,q,r,s,tol1,xm;
+
+	if ((fa > 0.0 && fb > 0.0) || (fa < 0.0 && fb < 0.0)) {
+		fprintf(stderr,"%le %le ; %le %le\n",a,fa,b,fb);
+		fprintf(stderr,"Root must be bracketed in zbrent\n");
+		if(fa<0.)
+			return(-1.);
+		else 
+			return(-2.);
+	}
+	fc=fb;
+	for (iter=1;iter<=ITMAX;iter++) {
+		if ((fb > 0.0 && fc > 0.0) || (fb < 0.0 && fc < 0.0)) {
+			c=a;
+			fc=fa;
+			e=d=b-a;
+		}
+		if (fabs(fc) < fabs(fb)) {
+			a=b;
+			b=c;
+			c=a;
+			fa=fb;
+			fb=fc;
+			fc=fa;
+		}
+		tol1=2.0*EPS*fabs(b)+0.5*tol;
+		xm=0.5*(c-b);
+		if (fabs(xm) <= tol1 || fb == 0.0) return b;
+		if (fabs(e) >= tol1 && fabs(fa) > fabs(fb)) {
+			s=fb/fa;
+			if (a == c) {
+				p=2.0*xm*s;
+				q=1.0-s;
+			} else {
+				q=fa/fc;
+				r=fb/fc;
+				p=s*(2.0*xm*q*(q-r)-(b-a)*(r-1.0));
+				q=(q-1.0)*(r-1.0)*(s-1.0);
+			}
+			if (p > 0.0) q = -q;
+			p=fabs(p);
+			min1=3.0*xm*q-fabs(tol1*q);
+			min2=fabs(e*q);
+			if (2.0*p < (min1 < min2 ? min1 : min2)) {
+				e=d;
+				d=p/q;
+			} else {
+				d=xm;
+				e=d;
+			}
+		} else {
+			d=xm;
+			e=d;
+		}
[…]

### 2002-01-25 src/kcorrect.h

@@ -1,0 +1,147 @@
+/*
+ * If you do not need or want to link your code to IDL, unset LINKTOIDL
+ * in the Makefile
+ */
+#ifdef LINKTOIDL
+#include "export.h"
+#else
+typedef int IDL_LONG;
+#endif
+
+#define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+/* fit coefficients, given information about the templates and 
+ * the filters (in ematrix and rmatrix) and a set of galaxies with
+ * redshifts */
+IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
+											IDL_LONG nt,             /* number of eigentemplates */
+											double *zvals,      /* z interpolation */
+											IDL_LONG nz,
+											double *rmatrix,    /* r matrix */
+											IDL_LONG nk,             /* number of bandpasses */
+											IDL_LONG nb,             /* number of templates */
+											double *amatrix, /* coefficients */
+											double *galaxy_flux, /* galaxy fluxes [i][k] and redshifts */
+											double *galaxy_invvar,
+											double *galaxy_z,
+											IDL_LONG ngalaxy);
+
+/* finds best fit eigentemplates, given the galaxy properties and 
+ * a previous guess at the best coefficients to the templates (in
+ * amatrix) */
+IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
+												 IDL_LONG nt,             /* number of eigentemplates */
+												 double *zvals,      /* z interpolation */
+												 IDL_LONG nz,
+												 double *rmatrix,    /* r matrix */
+												 IDL_LONG nk,             /* number of bandpasses */
+												 IDL_LONG nb,             /* number of templates */
+												 double *amatrix, /* current coefficients */
+												 double *galaxy_flux, /* galaxy fluxes [i][k] redshifts */
+												 double *galaxy_invvar,
+												 double *galaxy_z,
+												 IDL_LONG *galaxy_clip,
+												 IDL_LONG ngalaxy,
+												 double *dmatrix,    /* d matrix */
+												 IDL_LONG initialized_dmatrix);
+
+/* calculate the model fluxes, given the coeffs of the model */
+IDL_LONG k_model_fluxes(double *ematrix,    /* eigentemplates */
+												IDL_LONG nt,             /* number of eigentemplates */
+												double *zvals,      /* z interpolation */
+												IDL_LONG nz,
+												double *rmatrix,    /* r matrix */
[…]

### 2002-01-25 src/model_fluxes.c

@@ -1,0 +1,123 @@
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+#include "kcorrect.h"
+
+/*
+ * model_fluxes.c
+ *
+ * Calculates the model flux, given the coefficients and the template
+ * information
+ *
+ * Mike Blanton
+ * 1/2002 */
+
+static IDL_LONG nz=1000;
+static double zmin=1.e-4, zmax=1.e-0;
+
+static double *model_fluxes=NULL;
+static double *bmatrix=NULL;
+static double *lambda=NULL;
+static double *rmatrix=NULL;
+static double *ematrix=NULL;
+static double *zvals=NULL;
+static IDL_LONG nk,nb,nl,nt;
+static double *filter_lambda=NULL;
+static double *filter_pass=NULL;
+static IDL_LONG *filter_n=NULL;
+static IDL_LONG maxn;
+static double *galaxy_z=NULL;
+static double *coeffs=NULL;
+
+int main(int argc,
+				 char **argv)
+{
+	double to_z=-1.;
+	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk;
+	char ematrixfile[2000],bmatrixfile[2000],filterlist[2000],lambdafile[2000];
+
+	if(argc<5) {
+		fprintf(stderr,"Usage: cat <coeff file> | model_fluxes <ematrix file> <bmatrix file> <lambdafile> <filterlist> [to_z] [zmin] [zmax] [nz]\n");
+		exit(1);
+	} /* end if */
+	i=1;
+	strcpy(ematrixfile,argv[i]); i++;
+	strcpy(bmatrixfile,argv[i]); i++;
+	strcpy(lambdafile,argv[i]); i++;
+	strcpy(filterlist,argv[i]); i++;
+	if(argc>=6) { to_z=atof(argv[i]); i++; }
+
+	/* load the ematrix; this gives the eigentemplates in terms of the
+	 * basis set */
+	k_load_ascii_table(&ematrix,&ndim,&sizes,ematrixfile);
+	nt=sizes[0];
+	nb=sizes[1];
+	FREEVEC(sizes);
+
+	/* input the bmatrix; this is the spectrum for each basis element */
+	k_load_ascii_table(&bmatrix,&ndim,&sizes,bmatrixfile);
+	if(sizes[0]!=nb) {
+		fprintf(stderr,"ematrix and bmatrix files incompatible.\n");
+		exit(1);
+	} /* end if */
+	nl=sizes[1];
[…]

### 2002-01-29 docs/paper/figures.tex
fix figures
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@6 75dff457-8751-0410-ab1a-cd701654ce6b
fixing stuff
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@8 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \figurenum{\fignum}
-\plotone{radec.sample8b.ps}
-\caption{\label{radec.sample8b} Angular distribution, in an Aitoff
-projection of equatorial coordinates, of the SDSS sample used in the
-current analysis. This spectroscopic sample covers around 1,169 square
-degrees and contains about 100,000 galaxy spectra.}
+\plotone{response_sdss.ps}
+\caption{\label{response_sdss} Estimated filter response for all five
+bands in the SDSS, as a function of observed wavelength. A 4 gigayear
+old instantaneous burst using the models of \citet{bruzual93a} (and
+observed at $z=0$) is shown for reference.}
+%at different redshifts. A 4 gigayear old instantaneous burst using the
+%models of \citet{bruzual93a} is shown for reference. {\it Top panel:}
+%The system consisting of \band{0.0}{u}, \band{0.0}{g}, \band{0.0}{r},
+%\band{0.0}{i}, and \band{0.0}{z}.  {\it Bottom panel:} The system
+%consisting of \band{0.0}{u}, \band{0.1}{g}, \band{0.1}{r},
+%\band{0.1}{i}, and \band{0.2}{z}. We use this second system rather
+%than the first because it requires less interpolation to determine
+%\band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}, and no extrapolation to
+%determine \band{0.0}{u} and \band{0.1}{z}. 
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{response_sdss.ps}
-\caption{\label{response_sdss} Estimated filter response for all five
-bands in the SDSS, as a function of rest frame wavelength for galaxies
-at different redshifts. A 4 gigayear old instantaneous burst using the
-models of \citet{bruzual93a} is shown for reference. {\it Top panel:}
-The system consisting of \band{0.0}{u}, \band{0.0}{g}, \band{0.0}{r},
-\band{0.0}{i}, and \band{0.0}{z}.  {\it Bottom panel:} The system
-consisting of \band{0.0}{u}, \band{0.1}{g}, \band{0.1}{r},
-\band{0.1}{i}, and \band{0.2}{z}. We use this second system rather
-than the first because it requires less interpolation to determine
-\band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}, and no extrapolation to
-determine \band{0.0}{u} and \band{0.1}{z}. }
+\plotone{k_espec_plot.ps}
+\caption{\label{k_espec_plot} The four derived eigenspectra. Note that
+eigenspectra \#1, \#2, and \#3 are constrained to have zero total flux in the
+range between 3500\AA and 7500\AA. Eigenspectrum \#0 is not in any
+sense the ``average'' spectrum. }
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{plotaij.sample8b15.ps}
-\caption{\label{plotaij.sample8b15} {\it Top panel}: Distribution of
-the components of the three-parameter fit to the five-band SDSS
-photometry for a random subsample consisting of 16,000 of the SDSS
-galaxies. $a_0$ is linearly proportional to the flux between $3500\AA$
-and $7000\AA$, while $a_1$ and $a_2$ contribute no flux in this
-range. Thus, the ratios $a_1/a_0$ and $a_2/a_0$ describe the spectral
-type of the galaxy. $a_2/a_0$ is the more variable parameter and thus
-is the better separator of galaxy type. {\it Bottom panel}: At fixed
-$a_1/a_0=-0.2$, the inferred spectra corresponding to various values
-of $a_2/a_0$. Near $a_2/a_0=0.5$, the spectrum is similar to that of
-an elliptical galaxy. For higher values, the spectrum becomes bluer. }
+\plotone{k_model_plot.ps}
+\caption{\label{k_model_plot} Reconstructed galaxy fluxes relative to
+the observed galaxy fluxes, for all five SDSS bands, shown for a
+random subsample consisting of 16,000 of the SDSS galaxies. The
+residuals are shown against redshift.  There is no systematic trend
+with redshift in any band. The 5-$\sigma$ clipped estimate of the
+scatter around the observed fluxes is listed for each band. In $g$,
+$r$, and $i$ the scatter is consistent with the expected photometric
+errors in the survey. In $z$ and $u$ there is more scatter. In the
+case of $z$ this is attributable to Poisson noise in the
+observations. In the case of $u$, it is not, but it may simply be that
+there are more intrinic differences between galaxies in the $u$-band
+than elsewhere.}
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{model.ps}
-\caption{\label{model} Galaxy fluxes determined from the inferred
-galaxy spectra (see Figure \ref{plotaij.sample8b15}) relative to the
-observed galaxy fluxes, for all five SDSS bands, shown for a random
-subsample consisting of 16,000 of the SDSS galaxies. The
-residuals are shown against redshift and absolute magnitude, the two
-most important systematic variables in the analysis of this paper. The
-solid lines show the median values. The dashed lines show the 10\% and
-90\% levels (corresponding to $1.3\sigma$ if the distribution were
-Gaussian). The systematics are very small in $g$, $r$, $i$ and $z$,
-but more significant in $u$. The scatter is largest in $u$ and
-$z$. Nevertheless, the inferred fluxes are a good representation of
-the observed fluxes.  }
+\plotone{k_coeffdist_plot.ps}
+\caption{\label{k_coeffdist_plot} {\it Top panels}: Distribution of
+the components of the four-parameter fit to the five-band SDSS
+photometry for a random subsample consisting of 16,000 of the SDSS
+galaxies. $a_0$ is linearly proportional to the flux between $3500\AA$
+and $7000\AA$, while $a_1$, $a_2$, and $a_3$ contribute no flux in
+this range. Thus, the ratios $a_1/a_0$ and $a_2/a_0$ describe the
+spectral type of the galaxy. $a_2/a_0$ is the most variable parameter
+and thus is the better separator of galaxy type. {\it Bottom panel}:
+At fixed $a_1/a_0$ and $a_2/a_0$, the inferred spectra corresponding
+to various values of $a_3/a_0$. Near $a_3/a_0=-0.16$, the spectrum is
+similar to that of an elliptical galaxy. For higher values, the
+spectrum becomes bluer. }
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{colors.sample8b15.ps}
-\caption{\label{colors.sample8b15} For galaxies in the absolute
-magnitude range $-21.5<M_r<-21$, the color distribution calculated
-from the inferred SEDs as a function of redshift. The left column
-shows the four colors relative to \band{0.1}{r} for each galaxy versus
-redshift. The right column shows the distribution of colors for
-galaxies in the range $0.04<z<0.12$ {\it (dotted)} compared to the
-distribution of colors of galaxies in the range $0.12<z<0.19$ {\it
-(solid)}. The resulting color distribution is clearly fairly constant
-with redshift.}
+\plotone{k_coeff_plot2.ps}
+\caption{\label{k_coeff_plot1} }
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{kcorrect.sample8b15.ps}
+\plotone{k_coeff_plot1.ps}
+\caption{\label{k_coeff_plot1} }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+%\plotone{kcorrect.sample8b15.ps}
 \caption{\label{kcorrect.sample8b15} $K$-corrections as a function of
@@
 \figurenum{\fignum}
-\plotone{ciCompare.sample8b15.ps}
+%\plotone{ciCompare.sample8b15.ps}
 \caption{\label{ciCompare.sample8b15} Difference in the
@@
 \figurenum{\fignum}
-\plotone{cibreakCompare.sample8b15.ps}
+%\plotone{cibreakCompare.sample8b15.ps}
 \caption{\label{cibreakCompare.sample8b15} Same as Figure
@@
 \figurenum{\fignum}
-\plotone{specK.ps}
+%\plotone{specK.ps}
 \caption{\label{specK} Similar to Figure \ref{ciCompare.sample8b15},
@@
 \figurenum{\fignum}
-\plotone{aperturevsz.M.sample8b15.ps}
+%\plotone{aperturevsz.M.sample8b15.ps}
 \caption{\label{aperturevsz.M.sample8b15} }

### 2002-01-29 docs/paper/paper.tex
fixing stuff
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@8 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-The resulting four templates are shown in Figure \ref{templates} in
+The resulting four templates are shown in Figure \ref{k_espec_plot} in
 the range $2000<\lambda<12000$\AA, the region which is constrained by
-our observations. Figure \ref{coeffs} is a more illuminating
+our observations. Figure \ref{k_coeffdist_plot} is a more illuminating
 figure. It shows the pairwise joint distributions of $a_1/a_0$,
-$a_2/a_0$, and $a_3/a_0$ in the top three panels. In the bottom panel,
+$a_2/a_0$, and $a_3/a_0$ in the top two panels. In the bottom panel,
 three spectra taken from a one-dimensional sequence along the galay

### 2002-01-31 docs/paper/paper.tex
running
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@10 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2002-01-25 docs/paper/paper.tex
@@
 
-The resulting four templates are shown in Figure \ref{templates} in
-the range $2000<\lambda<12000$\AA, the region which is constrained by
-our observations. Figure \ref{coeffs} is a more illuminating
-figure. It shows the pairwise joint distributions of $a_1/a_0$,
-$a_2/a_0$, and $a_3/a_0$ in the top three panels. In the bottom panel,
-three spectra taken from a one-dimensional sequence along the galay
-locus are shown, showing that the spectra become progressively bluer
-along that sequence.
+%The resulting four templates are shown in Figure \ref{k_espec_plot} in
+%the range $2000<\lambda<12000$\AA, the region which is constrained by
+%our observations. 
+The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
+joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$. In the
+bottom panel, three spectra taken from a one-dimensional sequence
+along the galay locus are shown, showing that the spectra become
+progressively bluer along that sequence. Note that constraints on the
+SED become poor at the bluest and reddest edges of this diagram (the
+peak of the $z$-band response is at 8700\AA). Therefore, the odd
+behavior near the edges (particularly at the red edge) should not be
+taken too seriously.
 

### 2002-02-04 src/fit_coeffs.c
lots of docs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@14 75dff457-8751-0410-ab1a-cd701654ce6b

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * fit_coeffs.c

### 2002-02-04 src/fit_photoz.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * fit_photoz.c

### 2002-02-04 src/idl_k_model_fluxes.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <math.h>
 #include <stdlib.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 #define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
 static void free_memory()

### 2002-02-04 src/k_create_r.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_create_r.c

### 2002-02-04 src/k_fit_coeffs.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_fit_coeffs.c

### 2002-02-04 src/k_fit_photoz.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_fit_photoz.c

### 2002-02-04 src/k_fit_templates.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * fitTemplates.c

### 2002-02-04 src/k_interpolate.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_interpolate.c

### 2002-02-04 src/k_load_ascii_table.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_load_ascii_table.c

### 2002-02-04 src/k_load_filters.c

@@ -1,7 +1,8 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_load_filters.c
@@ -22,11 +23,17 @@ IDL_LONG k_load_filters(IDL_LONG **filter_n,
 {
 	IDL_LONG i,j;
 	FILE *listfp,*fp;
-	char filename[2000];
+	char filename[2000],fullfilename[2000],filterdir[1000];
 
 	FREEVEC((*filter_n));
 	FREEVEC((*filter_lambda));
 	FREEVEC((*filter_pass));
+	
+	if(getenv("KCORRECT_DIR")!=NULL) {
+		sprintf(filterdir,"%s/data/filters",getenv("KCORRECT_DIR"));
+	} else {
+		strcpy(filterdir,".");
+	} /* end if */
 
 	listfp=k_fileopen(filterlist,"r");
 	fscanf(listfp,"%d",nk);
@@ -34,7 +41,8 @@ IDL_LONG k_load_filters(IDL_LONG **filter_n,
 	(*maxn)=0;
 	for(i=0;i<(*nk);i++) {
 		fscanf(listfp,"%s",filename);
-		fp=k_fileopen(filename,"r");
+		sprintf(fullfilename,"%s/%s",filterdir,filename);
+		fp=k_fileopen(fullfilename,"r");
 		fscanf(fp,"%d",&((*filter_n)[i]));
 		fclose(fp);
 		if((*filter_n)[i]>(*maxn)) (*maxn)=(*filter_n)[i];
@@ -47,7 +55,8 @@ IDL_LONG k_load_filters(IDL_LONG **filter_n,
 	fscanf(listfp,"%d",nk);
 	for(i=0;i<(*nk);i++) {
 		fscanf(listfp,"%s",filename);
-		fp=k_fileopen(filename,"r");
+		sprintf(fullfilename,"%s/%s",filterdir,filename);
+		fp=k_fileopen(fullfilename,"r");
 		fscanf(fp,"%d",&((*filter_n)[i]));
 		for(j=0;j<(*filter_n)[i];j++)
 			fscanf(fp,"%lf %lf",&((*filter_lambda)[i*(*maxn)+j]),
@@ -56,4 +65,5 @@ IDL_LONG k_load_filters(IDL_LONG **filter_n,
 	} /* end for i */
[…]

### 2002-02-04 src/k_midpnt.c

@@ -1,7 +1,7 @@
 #include <string.h>
 #include <stdio.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 #define FUNC(x) ((*func)(x))
 
 double k_midpnt(double (*func)(double), double a, double b, IDL_LONG n)

### 2002-02-04 src/k_polint.c

@@ -1,7 +1,7 @@
 #include <string.h>
 #include <stdio.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 #define NRANSI
 
 void k_polint(double xa[], double ya[], IDL_LONG n, double x, double *y, 

### 2002-02-04 src/k_qromo.c

@@ -1,7 +1,7 @@
 #include <string.h>
 #include <stdio.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 #define EPS 1.0e-4
 #define JMAX 14
 #define JMAXP (JMAX+1)

### 2002-02-04 src/k_write_ascii_table.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * k_write_ascii_table.c

### 2002-02-04 src/model_fluxes.c

@@ -1,7 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "kcorrect.h"
+#include <kcorrect.h>
 
 /*
  * model_fluxes.c
@@ -19,7 +19,7 @@ static double *model_fluxes=NULL;
 static double *bmatrix=NULL;
 static double *lambda=NULL;
 static double *rmatrix=NULL;
-static double *ematrix=NULL;
+static double *ematrix=NULL;  
 static double *zvals=NULL;
 static IDL_LONG nk,nb,nl,nt;
 static double *filter_lambda=NULL;
@@ -96,9 +96,9 @@ int main(int argc,
 	model_fluxes=(double *) malloc(nk*nchunk*sizeof(double));
 	galaxy_z=(double *) malloc(nchunk*sizeof(double));
 	coeffs=(double *) malloc(nchunk*nt*sizeof(double));
-	fscanf(stdin,"%lf",&(galaxy_z[0]));
+	fscanf(stdin,"%lf",&(coeffs[0]));
 	while(!feof(stdin)) {
-		for(j=0;j<nt;j++)
+		for(j=1;j<nt;j++)
 			fscanf(stdin,"%lf",&(coeffs[j]));
 		fscanf(stdin,"%lf",&(galaxy_z[0]));
 		if(to_z!=-1.) galaxy_z[0]=to_z;
@@ -107,6 +107,7 @@ int main(int argc,
 		for(k=0;k<nk;k++)
 			fprintf(stdout,"%e ",model_fluxes[k]);
 		fprintf(stdout,"\n");
+		fscanf(stdin,"%lf",&(coeffs[0]));
 	}
 	
 	FREEVEC(zvals);

### 2002-02-05 docs/paper/paper.tex
fix flux units
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@17 75dff457-8751-0410-ab1a-cd701654ce6b
units
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@18 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 Istv\'an Csabai\altaffilmark{\ref{JHU}},
-Mamoru Doi\altaffilmark{\ref{Tokyo}}, and
+Mamoru Doi\altaffilmark{\ref{Tokyo}},
 %Daniel Eisenstein\altaffilmark{\ref{Arizona}},
-James E. Gunn\altaffilmark{\ref{Princeton}}
-%David W. Hogg\altaffilmark{\ref{NYU}}, and
+James E. Gunn\altaffilmark{\ref{Princeton}}, and
+David W. Hogg\altaffilmark{\ref{NYU}}
 %David J. Schlegel\altaffilmark{\ref{Princeton}}
@@
 
+In the future, all observations of galaxies (and indeed of the sky in
+general) will be performed using integral field spectrographs or
+equivalent devices which combine high spatial resolution, high
+spectral resolution, and high signal-to-noise. By that time, the
+phenomenon of broad-band filters and the quaint nomenclature
+surrounding their usage --- magnitudes, $K$-corrections, color terms,
+{\it etc.} --- will have long since been forgotten. However, today
+virtually all observations of galaxies are made through such filters,
+and special care has to be taken to recover knowledge of galaxy
+spectral energy distributions (SEDs) from these observations. This
+paper describes how to do so in a way well-suited for recovering (for
+example) fixed-frame magnitudes of objects over a large range of
+redshifts.
+
 Photometric surveys of the sky observe galaxies in bandpasses which
@@
 \int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
-\lambda b_i(\lambda) b_j(\lambda),
+b_i(\lambda) b_j(\lambda),
 \end{equation}
@@
 \begin{eqnarray}
-m_{AB} &=& ZZZZ - 2.5 \log_{10}\left[
+m_{AB} &=& -48.60 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
@@
 \cr
-&=& ZZZZ - 2.5 \log_{10}\left[
+&=& -48.60 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
@@
 \begin{equation}
-m_{AB} = ZZZZ - 2.5 \log_{10}\left[
+m_{AB} = -48.60 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}

### 2002-02-05 src/k_create_r.c

@@ -7,15 +7,22 @@
  * k_create_r.c
  *
  * Makes the rmatrix. This matrix is the impact of each filter curve
- * on each template basis element. 
+ * on each template basis element.
  *
  * Amazingly, this is the only part of the code in which the 
  * filter curves are necessary.
+ *
+ * AB magnitudes are assumed. The input spectra should be in units of
+ * ergs cm^{-2} s^{-1} A^{-1}
  *
  * Mike Blanton
  * 6/2001 */
 
 #define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
+
+/* factor of 10^(0.4*48.60) which converts flux in based on 
+ * f_lambda=erg/cm^2/s/A into maggies */
+#define ABSCALE 2.75423e+19
 
 static double *cr_bmatrix=NULL;
 static double *cr_lambda=NULL;
@@ -107,8 +114,8 @@ IDL_LONG k_create_r(double *rmatrix,
 		} /* end for l */
 
 		/* create scale factor for filter */
-		scale=1./k_qromo(scalefilter,cr_filter_lambda[0],
-										 cr_filter_lambda[cr_filter_n-1],k_midpnt);
+		scale=ABSCALE/k_qromo(scalefilter,cr_filter_lambda[0],
+													cr_filter_lambda[cr_filter_n-1],k_midpnt);
 
 		/* loop over redshifts */
 		for(i=0;i<nz;i++) {

### 2002-02-05 docs/paper/paper.tex
Hogg tweaks
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@19 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-In the future, all observations of galaxies (and indeed of the sky in
-general) will be performed using integral field spectrographs or
-equivalent devices which combine high spatial resolution, high
-spectral resolution, and high signal-to-noise. By that time, the
-phenomenon of broad-band filters and the quaint nomenclature
-surrounding their usage --- magnitudes, $K$-corrections, color terms,
-{\it etc.} --- will have long since been forgotten. However, today
-virtually all observations of galaxies are made through such filters,
-and special care has to be taken to recover knowledge of galaxy
-spectral energy distributions (SEDs) from these observations. This
-paper describes how to do so in a way well-suited for recovering (for
-example) fixed-frame magnitudes of objects over a large range of
-redshifts.
+In the future, observations of galaxies (and indeed of any
+astronomical sources) will be performed using integral field
+spectrographs or equivalent devices which combine high spatial
+resolution, high spectral resolution, and high signal-to-noise. By
+that time, the phenomena of broad-band filters and the quaint
+terminology surrounding their usage --- magnitudes, $K$-corrections,
+color terms, {\it etc.} --- will have long since been
+forgotten. However, today, virtually all observations of galaxies are
+made through such filters, and special care has to be taken to recover
+knowledge of galaxy spectral energy distributions (SEDs) from these
+observations. This paper describes how to do so in a way well-suited
+for recovering (for example) fixed-frame magnitudes of objects over a
+large range of redshifts.
 

### 2002-02-05 include/kcorrect.h
call maggies maggies
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@21 75dff457-8751-0410-ab1a-cd701654ce6b

@@ -21,7 +21,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 											IDL_LONG nk,             /* number of bandpasses */
 											IDL_LONG nb,             /* number of templates */
 											double *amatrix, /* coefficients */
-											double *galaxy_flux, /* galaxy fluxes [i][k] and redshifts */
+											double *galaxy_maggies, /* galaxy maggies */
 											double *galaxy_invvar,
 											double *galaxy_z,
 											IDL_LONG ngalaxy);
@@ -37,7 +37,7 @@ IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
 												 IDL_LONG nk,             /* number of bandpasses */
 												 IDL_LONG nb,             /* number of templates */
 												 double *amatrix, /* current coefficients */
-												 double *galaxy_flux, /* galaxy fluxes [i][k] redshifts */
+												 double *galaxy_maggies, /* galaxy maggies */
 												 double *galaxy_invvar,
 												 double *galaxy_z,
 												 IDL_LONG *galaxy_clip,
@@ -45,18 +45,18 @@ IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
 												 double *dmatrix,    /* d matrix */
 												 IDL_LONG initialized_dmatrix);
 
-/* calculate the model fluxes, given the coeffs of the model */
-IDL_LONG k_model_fluxes(double *ematrix,    /* eigentemplates */
-												IDL_LONG nt,             /* number of eigentemplates */
-												double *zvals,      /* z interpolation */
-												IDL_LONG nz,
-												double *rmatrix,    /* r matrix */
-												IDL_LONG nk,             /* number of bandpasses */
-												IDL_LONG nb,             /* number of templates */
-												double *coeffs, /* coefficients */
-												double *galaxy_z,
-												double *model_flux,
-												IDL_LONG ngalaxy);
+/* calculate the reconstructed fluxes, given the coeffs of the model */
+IDL_LONG k_reconstructed_fluxes(double *ematrix,    /* eigentemplates */
+																IDL_LONG nt, /* number of eigentemplates */
+																double *zvals,      /* z interpolation */
+																IDL_LONG nz,
+																double *rmatrix,    /* r matrix */
+																IDL_LONG nk,      /* number of bandpasses */
+																IDL_LONG nb,      /* number of templates */
[…]

### 2002-02-05 src/fit_coeffs.c

@@ -7,7 +7,7 @@
  * fit_coeffs.c
  *
  * Given the locations of the eigentemplate matrix, the basis set matrix,
- * and galaxy flux, errors, and redshift, output the best fit coefficients.
+ * and galaxy maggies, errors, and redshift, output the best fit coefficients.
  *
  * Mike Blanton
  * 2/2001
@@ -27,7 +27,7 @@ static double *filter_pass=NULL;
 static IDL_LONG *filter_n=NULL;
 static IDL_LONG maxn;
 static double *galaxy_z=NULL;
-static double *galaxy_flux=NULL;
+static double *galaxy_maggies=NULL;
 static double *galaxy_invvar=NULL;
 static double *coeffs=NULL;
 
@@ -94,16 +94,16 @@ int main(int argc,
      coefficients */
 	nchunk=1;
 	galaxy_z=(double *) malloc(nchunk*sizeof(double));
-	galaxy_flux=(double *) malloc(nchunk*nk*sizeof(double));
+	galaxy_maggies=(double *) malloc(nchunk*nk*sizeof(double));
 	galaxy_invvar=(double *) malloc(nchunk*nk*sizeof(double));
 	coeffs=(double *) malloc(nchunk*nt*sizeof(double));
 	fscanf(stdin,"%lf",&(galaxy_z[0]));
 	while(!feof(stdin)) {
 		for(k=0;k<nk;k++)
-			fscanf(stdin,"%lf",&(galaxy_flux[k]));
+			fscanf(stdin,"%lf",&(galaxy_maggies[k]));
 		for(k=0;k<nk;k++)
 			fscanf(stdin,"%lf",&(galaxy_invvar[k]));
-		k_fit_coeffs(ematrix,nt,zvals,nz,rmatrix,nk,nb,coeffs,galaxy_flux,
+		k_fit_coeffs(ematrix,nt,zvals,nz,rmatrix,nk,nb,coeffs,galaxy_maggies,
 								 galaxy_invvar,galaxy_z,nchunk);
 		for(j=0;j<nt;j++)
 			fprintf(stdout,"%e ",coeffs[j]);
@@ -120,7 +120,7 @@ int main(int argc,
 	FREEVEC(filter_lambda);
 	FREEVEC(filter_pass);
 	FREEVEC(galaxy_z);
-	FREEVEC(galaxy_flux);
+	FREEVEC(galaxy_maggies);
 	FREEVEC(galaxy_invvar);
 	FREEVEC(coeffs);
 } /* end main */

### 2002-02-05 src/fit_photoz.c

@@ -27,7 +27,7 @@ static double *filter_pass=NULL;
 static IDL_LONG *filter_n=NULL;
 static IDL_LONG maxn;
 static double *galaxy_z=NULL;
-static double *galaxy_flux=NULL;
+static double *galaxy_maggies=NULL;
 static double *galaxy_invvar=NULL;
 static double *coeffs=NULL;
 
@@ -84,21 +84,21 @@ int main(int argc,
 	/* get the redshift estimates */
 	nchunk=1;
 	galaxy_z=(double *) malloc(nchunk*sizeof(double));
-	galaxy_flux=(double *) malloc(nchunk*nk*sizeof(double));
+	galaxy_maggies=(double *) malloc(nchunk*nk*sizeof(double));
 	galaxy_invvar=(double *) malloc(nchunk*nk*sizeof(double));
 	coeffs=(double *) malloc(nchunk*nt*sizeof(double));
-	fscanf(stdin,"%lf",&(galaxy_flux[0]));
+	fscanf(stdin,"%lf",&(galaxy_maggies[0]));
 	while(!feof(stdin)) {
 		for(k=1;k<nk;k++)
-			fscanf(stdin,"%lf",&(galaxy_flux[k]));
+			fscanf(stdin,"%lf",&(galaxy_maggies[k]));
 		for(k=0;k<nk;k++)
 			fscanf(stdin,"%lf",&(galaxy_invvar[k]));
-		k_fit_photoz(ematrix,nt,zvals,nz,rmatrix,nk,nb,coeffs,galaxy_flux,
+		k_fit_photoz(ematrix,nt,zvals,nz,rmatrix,nk,nb,coeffs,galaxy_maggies,
 								 galaxy_invvar,galaxy_z,nchunk);
 		for(j=0;j<nt;j++)
 			fprintf(stdout,"%e ",coeffs[j]);
 		fprintf(stdout,"%e\n",galaxy_z[0]);
-		fscanf(stdin,"%lf",&(galaxy_flux[0]));
+		fscanf(stdin,"%lf",&(galaxy_maggies[0]));
 	} /* end for while */
 	
 	FREEVEC(zvals);
@@ -110,7 +110,7 @@ int main(int argc,
 	FREEVEC(filter_lambda);
 	FREEVEC(filter_pass);
 	FREEVEC(galaxy_z);
-	FREEVEC(galaxy_flux);
+	FREEVEC(galaxy_maggies);
[…]

### 2002-02-05 src/idl_k_fit_coeffs.c

@@ -15,7 +15,7 @@ IDL_LONG idl_k_fit_coeffs
 {
    IDL_LONG nt,nz,nk,nb,ngalaxy;
    double *ematrix,*zvals,*rmatrix,*amatrix;
-	 double *galaxy_flux,*galaxy_invvar,*galaxy_z;
+	 double *galaxy_maggies,*galaxy_invvar,*galaxy_z;
 
 	 IDL_LONG i;
 	 IDL_LONG retval=1;
@@ -30,14 +30,14 @@ IDL_LONG idl_k_fit_coeffs
    nk=*((IDL_LONG *)argv[i]); i++;
    nb=*((IDL_LONG *)argv[i]); i++;
 	 amatrix=(double *)argv[i]; i++;
-	 galaxy_flux=(double *)argv[i]; i++;
+	 galaxy_maggies=(double *)argv[i]; i++;
 	 galaxy_invvar=(double *)argv[i]; i++;
 	 galaxy_z=(double *)argv[i]; i++;
    ngalaxy=*((IDL_LONG *)argv[i]); i++;
 
 	 /* 1. run the fitting routine */
 	 retval=(IDL_LONG) k_fit_coeffs(ematrix,nt,zvals,nz,rmatrix,nk,nb,
-																	amatrix,galaxy_flux,galaxy_invvar,
+																	amatrix,galaxy_maggies,galaxy_invvar,
 																	galaxy_z,ngalaxy);
 
 	 /* 2. free memory and leave */

### 2002-02-05 src/idl_k_fit_photoz.c

@@ -15,7 +15,7 @@ IDL_LONG idl_k_fit_photoz
 {
    IDL_LONG nt,nz,nk,nb,ngalaxy;
    double *ematrix,*zvals,*rmatrix,*amatrix;
-	 double *galaxy_flux,*galaxy_invvar,*galaxy_photoz;
+	 double *galaxy_maggies,*galaxy_invvar,*galaxy_photoz;
 
 	 IDL_LONG i;
 	 IDL_LONG retval=1;
@@ -30,14 +30,14 @@ IDL_LONG idl_k_fit_photoz
    nk=*((IDL_LONG *)argv[i]); i++;
    nb=*((IDL_LONG *)argv[i]); i++;
 	 amatrix=(double *)argv[i]; i++;
-	 galaxy_flux=(double *)argv[i]; i++;
+	 galaxy_maggies=(double *)argv[i]; i++;
 	 galaxy_invvar=(double *)argv[i]; i++;
 	 galaxy_photoz=(double *)argv[i]; i++;
    ngalaxy=*((IDL_LONG *)argv[i]); i++;
 
 	 /* 1. run the fitting routine */
 	 retval=(IDL_LONG) k_fit_photoz(ematrix,nt,zvals,nz,rmatrix,nk,nb,
-																	amatrix,galaxy_flux,galaxy_invvar,
+																	amatrix,galaxy_maggies,galaxy_invvar,
 																	galaxy_photoz,ngalaxy);
 
 	 /* 2. free memory and leave */

### 2002-02-05 src/idl_k_fit_templates.c

@@ -15,7 +15,7 @@ IDL_LONG idl_k_fit_templates
 {
    IDL_LONG nt,nz,nk,nb,ngalaxy,*galaxy_clip,initialized_dmatrix;
    double *ematrix,*zvals,*rmatrix,*amatrix,*dmatrix;
-	 double *galaxy_flux,*galaxy_invvar,*galaxy_z;
+	 double *galaxy_maggies,*galaxy_invvar,*galaxy_z;
 
 	 IDL_LONG i;
 	 IDL_LONG retval=1;
@@ -30,7 +30,7 @@ IDL_LONG idl_k_fit_templates
    nk=*((IDL_LONG *)argv[i]); i++;
    nb=*((IDL_LONG *)argv[i]); i++;
 	 amatrix=(double *)argv[i]; i++;
-	 galaxy_flux=(double *)argv[i]; i++;
+	 galaxy_maggies=(double *)argv[i]; i++;
 	 galaxy_invvar=(double *)argv[i]; i++;
 	 galaxy_z=(double *)argv[i]; i++;
 	 galaxy_clip=(IDL_LONG *)argv[i]; i++;
@@ -40,7 +40,7 @@ IDL_LONG idl_k_fit_templates
 
 	 /* 1. run the fitting routine */
 	 retval=(IDL_LONG) k_fit_templates(ematrix,nt,zvals,nz,rmatrix,nk,nb,
-																		 amatrix,galaxy_flux,galaxy_invvar,
+																		 amatrix,galaxy_maggies,galaxy_invvar,
 																		 galaxy_z,galaxy_clip,ngalaxy,dmatrix,
 																		 initialized_dmatrix);
 

### 2002-02-05 src/k_fit_coeffs.c

@@ -30,7 +30,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 											IDL_LONG nk,             /* number of bandpasses */
 											IDL_LONG nb,             /* number of templates */
 											double *amatrix, /* coefficients */
-											double *galaxy_flux, /* galaxy fluxes [i][k], 
+											double *galaxy_maggies, /* galaxy fluxes [i][k], 
 																							redshifts */
 											double *galaxy_invvar,
 											double *galaxy_z,
@@ -89,7 +89,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 		for(j=0;j<nt;j++) {
 			rhs[j]=0.;
 			for(k=0;k<nk;k++) 
-				rhs[j]+=galaxy_flux[k+i*nk]*cmatrix[i*nk*nt+k*nt+j]
+				rhs[j]+=galaxy_maggies[k+i*nk]*cmatrix[i*nk*nt+k*nt+j]
 						*galaxy_invvar[k+i*nk];
 		} /* end for j */
 
@@ -113,7 +113,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 #endif
 				fprintf(stderr,"at galaxy %d, info=%d\n",i,info);
 				for(k=0;k<nk;k++) 
-					fprintf(stderr,"%e\n",galaxy_flux[k+i*nk]);
+					fprintf(stderr,"%e\n",galaxy_maggies[k+i*nk]);
 				for(k=0;k<nk;k++) 
 					fprintf(stderr,"%e\n",galaxy_invvar[k+i*nk]);
 				for(j=0;j<nt;j++) 

### 2002-02-05 src/k_fit_photoz.c

@@ -22,7 +22,7 @@ static double *pz_rmatrix=NULL;
 static double *pz_zvals=NULL;
 static double *pz_coeffs=NULL;
 static double *pz_model_flux=NULL;
-static double *pz_galaxy_flux=NULL;
+static double *pz_galaxy_maggies=NULL;
 static double *pz_galaxy_invvar=NULL;
 static double *pz_coeffspnorm=NULL;
 static double pz_z,pz_linepos;
@@ -37,8 +37,8 @@ double pz_calc_chi2(double z)
 								 pz_nk,pz_nb,pz_coeffs,&z,pz_model_flux,1);
 	chi2=0.;
 	for(k=0;k<pz_nk;k++)
-		chi2+=(pz_model_flux[k]-pz_galaxy_flux[k])
-			*(pz_model_flux[k]-pz_galaxy_flux[k])*pz_galaxy_invvar[k];
+		chi2+=(pz_model_flux[k]-pz_galaxy_maggies[k])
+			*(pz_model_flux[k]-pz_galaxy_maggies[k])*pz_galaxy_invvar[k];
 	
 	return(chi2);
 } /* end pz_calc_chi2 */
@@ -51,7 +51,7 @@ double pz_fit_coeffs(double z)
 	
 	/* fit coeffs */
 	k_fit_coeffs(pz_ematrix,pz_nt,pz_zvals,pz_nz,pz_rmatrix,
-							 pz_nk,pz_nb,pz_coeffs,pz_galaxy_flux,pz_galaxy_invvar,
+							 pz_nk,pz_nb,pz_coeffs,pz_galaxy_maggies,pz_galaxy_invvar,
 							 &z,1);
 	
 	chi2=pz_calc_chi2(z);
@@ -68,7 +68,7 @@ IDL_LONG k_fit_photoz(double *ematrix,    /* eigentemplates */
 											IDL_LONG nk,             /* number of bandpasses */
 											IDL_LONG nb,             /* number of templates */
 											double *coeffs, /* coefficients */
-											double *galaxy_flux, /* galaxy fluxes [i][k], 
+											double *galaxy_maggies, /* galaxy fluxes [i][k], 
 																							redshifts */
 											double *galaxy_invvar,
 											double *galaxy_z,
[…]

### 2002-02-05 src/k_fit_templates.c

@@ -27,7 +27,7 @@ IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
 												 IDL_LONG nk,             /* number of bandpasses */
 												 IDL_LONG nb,             /* number of templates */
 												 double *amatrix, /* current coefficients */
-												 double *galaxy_flux, /* galaxy fluxes [i][k] redshifts */
+												 double *galaxy_maggies, /* galaxy maggies */
 												 double *galaxy_invvar,
 												 double *galaxy_z,
 												 IDL_LONG *galaxy_clip,
@@ -93,7 +93,7 @@ IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
 						use=0;
 				if(!galaxy_clip[i] && use) {
 					for(k=0;k<nk;k++) 
-						rhs[indx]+=galaxy_flux[k+i*nk]*amatrix[i*nt+j]*
+						rhs[indx]+=galaxy_maggies[k+i*nk]*amatrix[i*nt+j]*
 							k_interpolate(galaxy_z[i],&(rmatrix[k*nb*nz+b*nz]),zvals,nz)
 							*galaxy_invvar[k+i*nk];
 				} /* end if */

### 2002-02-05 src/k_interpolate.c

@@ -40,7 +40,7 @@ double k_interpolate_es(double currpos,
 	unsigned long i,ip1;
 	double sp,currval;
 
-	i=(unsigned long) floor((double) n*(currpos-pos[0])/(2.*pos[n-1]-pos[n-2]));
+	i=(unsigned long) floor((double) n*(currpos-pos[0])/(pos[n-1]-pos[0]));
 	if(i>=n-1 || i<0) return(0.);
 	ip1=i+1;
 	sp=currpos-pos[i];

### 2002-02-06 src/fit_coeffs.c
add testing capabilities
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@26 75dff457-8751-0410-ab1a-cd701654ce6b

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -34,20 +35,32 @@ static double *coeffs=NULL;
 int main(int argc,
 				 char **argv)
 {
-	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk;
+	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk,ncurrchunk;
 	char ematrixfile[2000],bmatrixfile[2000],filterlist[2000],lambdafile[2000];
+	char version[1000],versionpath[2000];
+
+	/* set defaults */
+	strcpy(version,"default");
+	strcpy(versionpath,".");
+	if(getenv("KCORRECT_DIR")!=NULL) {
+		sprintf(versionpath,"%s/data/etemplates",getenv("KCORRECT_DIR"));
+	} /* end if */
 
 	/* read arguments */
-	if(argc<5) {
-		fprintf(stderr,"Usage: cat <galaxy file> | fit_coeffs <ematrix file> <bmatrix file> <lambdafile> <filterlist> [zmin] [zmax] [nz]\n");
+	if(argc<0) {
+		fprintf(stderr,"Usage: cat <galaxy file> | fit_coeffs [version [version path]]\n");
 		exit(1);
 	} /* end if */
 	i=1;
-	strcpy(ematrixfile,argv[i]); i++;
-	strcpy(bmatrixfile,argv[i]); i++;
-	strcpy(lambdafile,argv[i]); i++;
-	strcpy(filterlist,argv[i]); i++;
-	/* have not implemented optional parameters yet */
+	if(argc>=2) 
+		strcpy(version,argv[i]); i++;
+	if(argc>=3) 
+		strcpy(versionpath,argv[i]); i++;
+
+	sprintf(ematrixfile,"%s/ematrix.%s.dat",versionpath,version);
+	sprintf(bmatrixfile,"%s/bmatrix.%s.dat",versionpath,version);
+	sprintf(lambdafile,"%s/lambda.%s.dat",versionpath,version);
+	sprintf(filterlist,"%s/filterlist.%s.dat",versionpath,version);
 
 	/* load the ematrix; this gives the eigentemplates in terms of the
[…]

### 2002-02-06 src/fit_photoz.c

@@ -34,18 +34,32 @@ static double *coeffs=NULL;
 int main(int argc,
 				 char **argv)
 {
-	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk;
+	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk,ncurrchunk;
 	char ematrixfile[2000],bmatrixfile[2000],filterlist[2000],lambdafile[2000];
+	char version[1000],versionpath[2000];
 
-	if(argc<5) {
-		fprintf(stderr,"Usage: cat <galaxy file> | fit_photoz <ematrix file> <bmatrix file> <lambdafile> <filterlist> [zmin] [zmax] [nz]\n");
+	/* set defaults */
+	strcpy(version,"defaultpz");
+	strcpy(versionpath,".");
+	if(getenv("KCORRECT_DIR")!=NULL) {
+		sprintf(versionpath,"%s/data/etemplates",getenv("KCORRECT_DIR"));
+	} /* end if */
+
+	/* read arguments */
+	if(argc<0) {
+		fprintf(stderr,"Usage: cat <galaxy file> | fit_photoz [version [version path]]\n");
 		exit(1);
 	} /* end if */
 	i=1;
-	strcpy(ematrixfile,argv[i]); i++;
-	strcpy(bmatrixfile,argv[i]); i++;
-	strcpy(lambdafile,argv[i]); i++;
-	strcpy(filterlist,argv[i]); i++;
+	if(argc>=2) 
+		strcpy(version,argv[i]); i++;
+	if(argc>=3) 
+		strcpy(versionpath,argv[i]); i++;
+
+	sprintf(ematrixfile,"%s/ematrix.%s.dat",versionpath,version);
+	sprintf(bmatrixfile,"%s/bmatrix.%s.dat",versionpath,version);
+	sprintf(lambdafile,"%s/lambda.%s.dat",versionpath,version);
+	sprintf(filterlist,"%s/filterlist.%s.dat",versionpath,version);
 
 	/* load the ematrix */
 	k_load_ascii_table(&ematrix,&ndim,&sizes,ematrixfile);
@@ -82,23 +96,29 @@ int main(int argc,
 						 filter_lambda,filter_pass,maxn);
 
 	/* get the redshift estimates */
-	nchunk=1;
[…]

### 2002-02-06 src/k_fit_coeffs.c

@@ -111,7 +111,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 #else
 				if(info!=0) {
 #endif
-				fprintf(stderr,"at galaxy %d, info=%d\n",i,info);
+				fprintf(stderr,"at galaxy %d, info=%d, %le\n",i,info,galaxy_z[i]);
 				for(k=0;k<nk;k++) 
 					fprintf(stderr,"%e\n",galaxy_maggies[k+i*nk]);
 				for(k=0;k<nk;k++) 

### 2002-02-06 src/k_interpolate.c

@@ -40,8 +40,10 @@ double k_interpolate_es(double currpos,
 	unsigned long i,ip1;
 	double sp,currval;
 
-	i=(unsigned long) floor((double) n*(currpos-pos[0])/(pos[n-1]-pos[0]));
-	if(i>=n-1 || i<0) return(0.);
+	i=(unsigned long) floor((double) n*(currpos-pos[0])/
+													(2*pos[n-1]-pos[n-2]-pos[0]));
+	if(i>=n || i<0) return(0.);
+	if(i==n-1) i=n-2;
 	ip1=i+1;
 	sp=currpos-pos[i];
 	currval=vals[i]+sp*(vals[ip1]-vals[i])/(pos[ip1]-pos[i]);

### 2002-02-06 src/k_reconstruct_maggies.c

@@ -31,7 +31,7 @@ IDL_LONG k_reconstruct_maggies(double *ematrix,    /* eigentemplates */
 	
 	for(i=0;i<ngalaxy;i++) {
 		for(k=0;k<nk;k++) {
-			reconstruct_maggies[k]=0.;
+			reconstruct_maggies[k+i*nk]=0.;
 			for(b=0;b<nb;b++) {
 				for(j=0;j<nt;j++) 
 					reconstruct_maggies[k+i*nk]+=coeffs[i*nt+j]*ematrix[j*nb+b]*

### 2002-02-06 src/reconstruct_maggies.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -33,19 +34,34 @@ int main(int argc,
 				 char **argv)
 {
 	double to_z=-1.;
-	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk;
+	IDL_LONG i,j,k,*sizes=NULL,ndim,nchunk,ncurrchunk;
 	char ematrixfile[2000],bmatrixfile[2000],filterlist[2000],lambdafile[2000];
+	char version[1000],versionpath[2000];
 
-	if(argc<5) {
-		fprintf(stderr,"Usage: cat <coeff file> | reconstruct_maggies <ematrix file> <bmatrix file> <lambdafile> <filterlist> [to_z] [zmin] [zmax] [nz]\n");
+	/* set defaults */
+	strcpy(version,"default");
+	strcpy(versionpath,".");
+	if(getenv("KCORRECT_DIR")!=NULL) {
+		sprintf(versionpath,"%s/data/etemplates",getenv("KCORRECT_DIR"));
+	} /* end if */
+
+	/* read arguments */
+	if(argc<0) {
+		fprintf(stderr,"Usage: cat <coeff file> | reconstruct_maggies [to_z] [version [version path]]\n");
 		exit(1);
 	} /* end if */
 	i=1;
-	strcpy(ematrixfile,argv[i]); i++;
-	strcpy(bmatrixfile,argv[i]); i++;
-	strcpy(lambdafile,argv[i]); i++;
-	strcpy(filterlist,argv[i]); i++;
-	if(argc>=6) { to_z=atof(argv[i]); i++; }
+	if(argc>=2) 
+		to_z=atof(argv[i]); i++; 
+	if(argc>=3) 
+		strcpy(version,argv[i]); i++;
+	if(argc>=4) 
+		strcpy(versionpath,argv[i]); i++;
+
+	sprintf(ematrixfile,"%s/ematrix.%s.dat",versionpath,version);
+	sprintf(bmatrixfile,"%s/bmatrix.%s.dat",versionpath,version);
+	sprintf(lambdafile,"%s/lambda.%s.dat",versionpath,version);
+	sprintf(filterlist,"%s/filterlist.%s.dat",versionpath,version);
 
[…]

### 2002-02-08 docs/paper/paper.tex
call maggies maggies
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@21 75dff457-8751-0410-ab1a-cd701654ce6b
docs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@28 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \shortauthors{Blanton {\it et al.} (2000)}
-\shorttitle{Low Redshift Galaxy Evolution}
+\shorttitle{Modeling Galaxy SEDs}
 
@@
 \begin{abstract}
-We present a method of inferring galaxy spectral energy distributions
-(SEDs) from photometric observations of galaxies in broad band
-filters. Using the Sloan Digital Sky Survey (SDSS) observations in the
-optical wavelength regime, we demonstrate that the method yields
-robust results. We show how the method can be used to recover galaxy
-magnitudes in fixed frame bandpasses over a range of redshifts. We
-compare our method to others, including galaxy spectrophotometry. For
-the SDSS, fixed frame magnitudes around the SDSS $r$, $i$, and $z$
-bands are highly reliable, at the level of a few percent; $g$-band
-fixed frame magnitudes are somewhat more uncertain, and $u$-band fixed
-frame magnitudes are highly uncertain, at the level of 10--20\%. All
-code and templates from this paper are public.
+Certain astronomical questions, such as the evolution of the
+luminosity function of galaxies and the relationship between galaxy
+luminosities and galaxy colors, require transforming inherently
+redshift-dependent measurements of galaxy magnitudes into measures of
+galaxy properties which are consistent over a range of redshifts.  As
+part of the task of addressing these questions, which we will
+investigate in subsequent papers, we implement here the method of
+\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
+broad-band photometric observations, in the context of the optical
+observations of the Sloan Digital Sky Survey (SDSS). Linear
+combinations of four spectral templates can reproduce the five SDSS
+magnitudes of galaxies to the precision of the photometry. Expressed
+in the correct coordinate system, the distribution of the coefficients
+of the templates is planar, and in fact nearly linear. The resulting
+reconstructed SEDs (the sum of the templates) can be used to recover
+fixed-frame magnitudes over a range of redshifts. We show that the
+results of doing so yield consistent results within the SDSS Main
+Galaxy Sample and the SDSS Luminous Red Galaxy Sample, in the sense
+that within each sample the trend with redshift of intrinsic colors
+for similar type galaxies are small. A potential problem in the
+measurement of the SDSS filter response curves is noted. We also
+validate the results using galaxy spectrophotometry.  All code and
+templates from this paper are publicly available and well-documented;
+they are designed such that they can easily be incorporated into other
+applications. The code is easily adapted to handle galaxy observations
+in any AB magnitude system in the rest-frame optical regime.
 \end{abstract}
@@
 
-In the future, observations of galaxies (and indeed of any
-astronomical sources) will be performed using integral field
+In the future, we expect that observations of galaxies (and indeed of
+any astronomical sources) will be performed using integral field
 spectrographs or equivalent devices which combine high spatial
-resolution, high spectral resolution, and high signal-to-noise. By
-that time, the phenomena of broad-band filters and the quaint
-terminology surrounding their usage --- magnitudes, $K$-corrections,
-color terms, {\it etc.} --- will have long since been
-forgotten. However, today, virtually all observations of galaxies are
-made through such filters, and special care has to be taken to recover
-knowledge of galaxy spectral energy distributions (SEDs) from these
-observations. This paper describes how to do so in a way well-suited
-for recovering (for example) fixed-frame magnitudes of objects over a
-large range of redshifts.
+resolution and high spectral resolution. By that time, the phenomena
+of broad-band filters and the quaint terminology surrounding their
+usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
+will have long since been forgotten. However, today, virtually all
+observations of galaxies are made through such filters, and special
+care has to be taken to recover knowledge of galaxy spectral energy
+distributions (SEDs) from these observations. This paper focuses on a
+method of facing this problem which (in the optical wavelength regime)
+is fast, robust, and consistent over a large range of redshifts.
 
-Photometric surveys of the sky observe galaxies in bandpasses which
-are fixed in the rest-frame of the observer. However, spectra of
-distant galaxies are redshifted due to the expansion of the
+This task is nontrivial because SEDs of galaxies contain important
+information on scales considerably smaller than the width of typical
+broad-band filters. It is further complicated by the fact that SEDs of
+distant galaxies are redshifts due to the expansion of the
 universe. Thus, the more distant the galaxy, the further the observed
 bandpass is blueshifted relative to the rest-frame spectral energy
-distribution (SED) of the object observed. To compare observations of
-galaxies at different redshifts, it is important to account for the
-fact that one is observing different parts of the galaxy SED.
+distribution of the object observed.
 
-In the past, this accounting has in general come in the form of
-``$K$-corrections'' applied to the observed magnitudes. That is, a
-function $K(z)$ is added to the standard cosmological bolometric
+In the past, accounting for these effects has in general come in the
+form of ``$K$-corrections'' applied to the observed magnitudes. That
+is, a function $K(z)$ is added to the standard cosmological bolometric
 distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
@@
 function is based on the projection of an assumed galaxy SED
-redshifted to $z$ onto the filter response.
+redshifted to $z$ onto the measured instrument response as a function
+of wavelength (including the effects of the atmosphere, the
+reflectivity of the mirrors, the filter transmission, and the response
+of the CCD device or photographic emulsion).
 
@@
 
-Our approach here is to develop a method for inferring the underlying
-SEDs of a set of galaxies at a range of redshifts by requiring that
-their SEDs all be drawn from a similar population. For each galaxy, we
-will recover a model SED, which can be used to synthesize the galaxy's
+Our approach here is to use a method for inferring the underlying SEDs
+of a set of galaxies at a range of redshifts by requiring that their
+SEDs all be drawn from a similar population. For each galaxy, we will
+recover a model SED, which can be used to synthesize the galaxy's
 magnitude in any bandpass. In addition, this method provides a natural
 way to infer an estimate of the flux within the whole optical range.
-The approach is equivalent to the photometric metric redshift
-estimation methods of \citet{csabai99a} and \citet{budavari00a}, 
-except we use slightly different coordinate systems.
+The approach is equivalent (nearly identical) to the photometric
+metric redshift estimation methods of \citet{csabai99a} and
+\citet{budavari00a}, except we use slightly different coordinate
+systems.
+
+We implemented this system and are publishing it in order to prepare
+the reader for upcoming papers which will rely heavily on the
+reliability of the SED fits described here.  However, having
+implemented the fits, there is no reason not to share them with the
+community; thus, this paper serves the secondary purpose of describing
+a piece of software being released to the astronomical community. The
+computer software created is a C shared object library, around which
+we have written both stand-alone C programs and IDL routines. We have
+made the source code available publicly, through a URL on the World
+Wide Web and through a public CVS repository. Improvements or ports to
+other languages implemented by users may be incorporated into the code
+upon request. The conditions of use for the code are that this paper
+is cited in any resulting refereed journal article and that the
+version of the code used is specified in any such paper. The version
+of the code used to make the figures for this paper is {\tt kcorrect
+v1_2}.
 
 Section \ref{sedfit} describes our method of fitting galaxy SEDs to
-broad-band photometry. An implementation of the method in C and IDL is
-provided at a URL on the World-Wide Web, through a public CVS
-repository, or by request.  Section \ref{data} applies the method to
+broad-band photometry. Section \ref{data} applies the method to
 galaxies in the SDSS, showing that the fits are robust. Section
@@
 \cr
-&=& -48.60 - 2.5 \log_{10}\left[
+&=& -2.41 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
@@
 where $R(\lambda)$ is the fraction of photons which are detected as a
-function of wavelength (a unitless quantity).
+function of wavelength (a unitless quantity). The difference in the
+zeropoint is simply the factor of speed of light $c$ (expressed in
+\AA\ s$^{-1}$) between the frequency and wavelength scales. {\bf
+clearer on units}
 

### 2002-02-08 include/kcorrect.h

@@ -30,7 +30,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
  * a previous guess at the best coefficients to the templates (in
  * amatrix) */
 IDL_LONG k_fit_templates(double *ematrix,    /* eigentemplates */
-												 IDL_LONG nt,             /* number of eigentemplates */
+												 IDL_LONG nt,           /* number of eigentemplates */
 												 double *zvals,      /* z interpolation */
 												 IDL_LONG nz,
 												 double *rmatrix,    /* r matrix */
@@ -96,6 +96,12 @@ IDL_LONG k_load_ascii_table(double **table,
 														IDL_LONG *ndim,
 														IDL_LONG **sizes,
 														char filename[]);
+
+/* write an ascii table in my standard format */
+IDL_LONG k_write_ascii_table(double *table,
+														 IDL_LONG ndim,
+														 IDL_LONG *sizes,
+														 char filename[]);
 
 /* load the filters in their standard format */
 IDL_LONG k_load_filters(IDL_LONG **filter_n,

### 2002-02-10 docs/paper/paper.tex
fixed stuff
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@29 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 redshift-dependent measurements of galaxy magnitudes into measures of
-galaxy properties which are consistent over a range of redshifts.  As
-part of the task of addressing these questions, which we will
-investigate in subsequent papers, we implement here the method of
-\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
-broad-band photometric observations, in the context of the optical
-observations of the Sloan Digital Sky Survey (SDSS). Linear
-combinations of four spectral templates can reproduce the five SDSS
-magnitudes of galaxies to the precision of the photometry. Expressed
-in the correct coordinate system, the distribution of the coefficients
-of the templates is planar, and in fact nearly linear. The resulting
+galaxy properties which are consistent over a range of redshifts.  To
+address these astronomical questions, investigated in detail in
+subsequent papers, we implement here the method of \citet{csabai00a}
+for fitting spectral energy distributions (SEDs) to broad-band
+photometric observations, in the context of the optical observations
+of the Sloan Digital Sky Survey (SDSS). Linear combinations of four
+spectral templates can reproduce the five SDSS magnitudes of galaxies
+to the precision of the photometry. Expressed in the correct
+coordinate system, the distribution of the coefficients of the
+templates is planar, and in fact nearly linear. The resulting
 reconstructed SEDs (the sum of the templates) can be used to recover

### 2002-02-10 src/k_fit_coeffs.c

@@ -123,7 +123,6 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 					fprintf(stderr,"%e\n",rhs[j]);
 				for(j=0;j<nt;j++) 
 					amatrix[i*nt+j]=0.;
-				return(0);
 			} /* end if */
 			
 			/* 2e. Store result */

### 2002-02-11 docs/paper/figures.tex
lots of docs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@14 75dff457-8751-0410-ab1a-cd701654ce6b
added
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@33 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
+%\clearpage
+%\stepcounter{thefigs}
+%\begin{figure}
+%\figurenum{\fignum}
+%%\plotone{k_espec_plot.ps}
+%\caption{\label{k_espec_plot} The four derived eigenspectra. Note that
+%eigenspectra \#1, \#2, and \#3 are constrained to have zero total flux in the
+%range between 3500\AA and 7500\AA. Eigenspectrum \#0 is not in any
+%sense the ``average'' spectrum. }
+%\end{figure}
+
 \clearpage
@@
 \figurenum{\fignum}
-\plotone{k_espec_plot.ps}
-\caption{\label{k_espec_plot} The four derived eigenspectra. Note that
-eigenspectra \#1, \#2, and \#3 are constrained to have zero total flux in the
-range between 3500\AA and 7500\AA. Eigenspectrum \#0 is not in any
-sense the ``average'' spectrum. }
+\plotone{k_coeffdist_plot.ps}
+\caption{\label{k_coeffdist_plot} {\it Top panels}: Distribution of
+the components of the four-parameter fit to the five-band SDSS
+photometry for a random subsample consisting of 10,000 of the SDSS
+galaxies. $a_0$ is linearly proportional to the flux between $3500\AA$
+and $7000\AA$, while $a_1$, $a_2$, and $a_3$ contribute no flux in
+this range. Thus, the ratios $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$
+describe the spectral type of the galaxy. $a_1/a_0$ is the most
+variable parameter and thus is the better separator of galaxy
+type. {\it Bottom panel}: At fixed $a_2/a_0$ and $a_3/a_0$, the
+inferred spectra corresponding to various values of $a_3/a_0$. Near
+$a_3/a_0=-0.20$, the spectrum is similar to that of an elliptical
+galaxy. For higher values, the spectrum becomes bluer. }
 \end{figure}
@@
 the observed galaxy fluxes, for all five SDSS bands, shown for a
-random subsample consisting of 16,000 of the SDSS galaxies. The
+random subsample consisting of around 10,000 of the SDSS galaxies. The
 residuals are shown against redshift.  There is no systematic trend
@@
 than elsewhere.}
-\end{figure}
-
-\clearpage
-\stepcounter{thefigs}
-\begin{figure}
-\figurenum{\fignum}
-\plotone{k_coeffdist_plot.ps}
-\caption{\label{k_coeffdist_plot} {\it Top panels}: Distribution of
-the components of the four-parameter fit to the five-band SDSS
-photometry for a random subsample consisting of 16,000 of the SDSS
-galaxies. $a_0$ is linearly proportional to the flux between $3500\AA$
-and $7000\AA$, while $a_1$, $a_2$, and $a_3$ contribute no flux in
-this range. Thus, the ratios $a_1/a_0$ and $a_2/a_0$ describe the
-spectral type of the galaxy. $a_2/a_0$ is the most variable parameter
-and thus is the better separator of galaxy type. {\it Bottom panel}:
-At fixed $a_1/a_0$ and $a_2/a_0$, the inferred spectra corresponding
-to various values of $a_3/a_0$. Near $a_3/a_0=-0.16$, the spectrum is
-similar to that of an elliptical galaxy. For higher values, the
-spectrum becomes bluer. }
 \end{figure}

### 2002-02-11 docs/paper/paper.tex
added
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@33 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 of the code used to make the figures for this paper is {\tt kcorrect
-v1_2}.
+v1\_2}.
 
@@
 galaxy SEDs take, so we can attempt to use what we know about galaxy
-SEDs to simplify our task.
-
-The method described here for doing so follows closely the methods of
-estimating photometric redshifts used by \citet{csabai00a} and
-\citet{budavari00a}. It is designed to take advantage of what we
-already understand about galaxy spectral energy distributions.
+SEDs to simplify our task.  The method described here for doing so
+follows closely the methods of estimating photometric redshifts used
+by \citet{csabai00a} and \citet{budavari00a}. It is designed to take
+advantage of what we already understand about galaxy spectral energy
+distributions.
 %In addition, as implemented here, the method is
@@
 %empirical aspects of the results.
-Beginning with a space defined by $N_b$ template galaxy SEDs
-$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), one can find an
-orthonormal set of spectra spanning that space. To be more precise, we
-define the dot-product in the space of galaxy spectra to be:
+
+Let us begin with an SED space defined by $N_b$ template galaxy SEDs
+$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
+large. In principle, this should be a complete set of galaxy SEDs if
+we want to express {\it any} galaxy SED in this basis set; in
+practice, this property is not necessary.  Given a set of
+$\vv{v}_i(\lambda)$, one can find an orthonormal set of spectra
+$\vv{b}_i(\lambda)$ spanning the space.
+
+To be more precise, we define the dot-product in the space of galaxy
+spectra to be:
 \begin{equation}
-\vv{b}_i \cdot \vv{b}_j =
+\vv{v}_i \cdot \vv{v}_j =
 \int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
-b_i(\lambda) b_j(\lambda),
+v_i(\lambda) v_j(\lambda),
 \end{equation}
@@
 independent in the restricted, as well as the full, wavelength range.}
-
-We then further define the basis of the SED space such that the $N_b$
-template SEDs, $\vv{v}_i$, are linear combinations of $N_b$ basis
-spectra $\vv{b}_i$ and that $\vv{b}_i \cdot \vv{b}_j =
-\delta_{ij}$. These conditions do not fully specify the $\vv{b}_i$,
-which naturally may be rotated arbitrarily within our
-$N_b$-dimensional subspace of spectral space.  In our implementation,
-we use the subspace defined by linear combinations of ten
-Bruzual-Charlot instantaneous burst models (our $\vv{v}_i(\lambda)$)
-with ages ranging from $3 \times 10^7$ to $2\times 10^{10}$ years,
-five with metallicity $Z=0.02$ and six with $Z=0.004$, all assuming a
-Salpeter Initial Mass Function and smoothed using a Gaussian with a
-standard deviation of $\sigma = 300$\AA. The smoothing is necessary
-to prevent the method from overfitting; not all the small-scale
-information about the templates can be recovered using this method,
-and we find that this level of smoothing gives sensible results.
-
-We cannot determine an individual galaxy's position in this
-ten-dimensional space from its broad-band colors alone, because there
-are only five observed bands. However, as outlined by
-\citet{csabai00a}, one can find a lower dimensional space defined by
-$N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
-galaxies. The SED of galaxy $i$ may be approximated as a linear
-combination of the basis spectra in this lower dimensional space:
+We then define the basis of the SED space such that the $N_b$ template
+SEDs, $\vv{v}_i$, are linear combinations of $N_b$ basis spectra
+$\vv{b}_i$ and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These
+conditions do not fully specify the $\vv{b}_i$, which naturally may be
+rotated arbitrarily within our $N_b$-dimensional subspace of spectral
+space.
+
+In our case the number of observed bands is less than $N_b$, so we
+cannot determine an individual galaxy's position in this
+$N_b$-dimensional space from its broad-band colors alone.  However, as
+outlined by \citet{csabai00a}, one can find a lower dimensional space
+defined by $N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
+galaxies, as follows. The SED of galaxy $i$ may be reconstructed from a
+linear combination of the basis spectra in this lower dimensional
+space:
 \begin{equation}
 \label{lincomb}
-f_{\mathrm{model},i}(\lambda) d\lambda = 
+f_{\mathrm{rec},i}(\lambda) d\lambda = 
 \sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
@@
 where $a_k$ are the components of the galaxy projected in the
-low-dimensional space. From this SED it is easy to determine the flux
-$F_{\mathrm{model},il}$ measured in each survey bandpass by projecting
-the model SED onto the bandpass $l$. We define
+low-dimensional space. From this SED it is easy to determine the
+reconstructed flux
+$F_{\mathrm{rec},il}$ in each survey bandpass by projecting
+the model SED onto the bandpass $l$: 
+\begin{equation}
+F_{\mathrm{rec},il} = \int_0^\infty d\lambda f(\lambda) S(\lambda),
+\end{equation}
+where $S(\lambda)$ is the response of the instrument.  We define
 \begin{equation}
 \chi^2 = \sum_i \sum_l
-\frac{(F_{\mathrm{obs},il}-F_{\mathrm{model},il})^2}{\sigma_{il}^2},
+\frac{(F_{\mathrm{obs},il}-F_{\mathrm{rec},il})^2}{\sigma_{il}^2},
 \end{equation}
@@
 
-We define the ``flux'' $f_k$ of each template as the flux in the range
-$\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
-estimate of the flux in this range is $F_{{v}}=\Sigma_k a_k f_k$. The
+We define the ``total flux'' $f_k$ of each template as the flux in the
+range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
+estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$. The
 locus of points of a fixed flux in the component space $a_k$ lie on a
@@
 with not much curvature within the space defined by the template
-SEDs. Thus, we rotate the axes $\vv{b}_i$ (for $i>0$) such that they
-are aligned with the principal axes of the ellipsoid. This choice of
-coordinate system is arbitrary, however, and does not affect our
-quantitative results.
+SEDs. Thus, we rotate the axes $\vv{e}_i$ (for $i>0$) such that they
+are aligned with the principal axes of the ellipsoid. Furthermore, we
+normalize the vectors $\vv{e}_i$ such that the $a_0=F_t$ in units of
+ergs cm$^{-2}$ s$^{-1}$. This choice of coordinate system is
+arbitrary and does not affect our quantitative results, but it is
+convenient.
 
@@
 
-The version of the $K$-correction code (v1\_0) implementing the
+The version of the $K$-correction code (v1\_2) implementing the
 calculations described here, along with eigentemplates and filter
 curves is publicly available on the World Wide Web at {\tt
-http://physics.nyu.edu/\~mb144/kcorrect}. In addition, a public CVS
+http://physics.nyu.edu/\~\ mb144/kcorrect}. In addition, a public CVS
 repository can be accessed at {\tt xxx} (instructions for exporting
@@
 the Research Systems, Incorporated, IDL language; everything except
-for the template-fitting also exists in a stand-alone C version (which
+for the template-fitting also exists in stand-alone C programs (which
 calls the same routines, guaranteeing consistency).
@@
 will be targeted for spectroscopic follow up using two digital
-spectrographs on the same telescope. Many of the details of the survey
-are described in a paper accompanying the Early Data Release
-(\citealt{stoughton01a}). The survey has begun in earnest, and has
-obtained about 20\% of its intended data.
+spectrographs on the same telescope. Many of the details of the galaxy
+survey are described in a description of the galaxy target selection
+paper (\citealt{strauss02a}). Other aspects of the survey are
+described in the Early Data Release (EDR; \citealt{stoughton01a}). The
+survey has begun in earnest, and has obtained about 20\% of its
+intended data.
 
@@
 \begin{eqnarray}
-m_{AB} &=& -48.60 - 2.5 \log_{10}\left[
+m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
@@
 \cr
-&=& -2.41 - 2.5 \log_{10}\left[
+&=& -48.60 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
@@
 \end{eqnarray}
-where $R(\lambda)$ is the fraction of photons which are detected as a
-function of wavelength (a unitless quantity). The difference in the
-zeropoint is simply the factor of speed of light $c$ (expressed in
-\AA\ s$^{-1}$) between the frequency and wavelength scales. {\bf
-clearer on units}
+where $R(\lambda)$ is the fraction of photons entering the Earth's
+atmosphere which are detected as a function of wavelength (a unitless
+quantity). In this equation, $f(\lambda)$ is in units of ergs
+cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
+cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
+that an object with $f(\nu) = 3.631 \times 10^{-20}$ ergs cm$^{-2}$
+s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes equal to zero.  The
+difference in the zeropoint in the two equations is simply the factor
+of speed of light $c$ (expressed in \AA\ s$^{-1}$) between the
+frequency and wavelength scales.
 
@@
 shown that the differences between these six camera columns are
-negligible. The resulting set of filter curves is shown in Figure
+small. The resulting set of filter curves is shown in Figure
 \ref{response_sdss}, in comparison to a model of a galaxy spectral
@@
 
-We use throughout the magnitudes determined by the SDSS using the
+We use two types of magnitudes determined by the SDSS. First, the
 modified form of the magnitude described by \citet{petrosian76a}, as
-described in \citet{strauss02a}, reddening-corrected using the dust
-maps of \citet{schlegel98a}. These magnitudes have a nearly constant
-metric aperture as a function redshift; in addition, all the bands use
-the same aperture, so the measured SED corresponds (to within the
-effects of seeing) to the SED of an identifiable region of the galaxy.
+described in \citet{strauss02a}.  These magnitudes have a nearly
+constant metric aperture as a function redshift; in addition, all the
+bands use the same aperture, so the measured SED corresponds (to
+within the effects of seeing) to the SED of an identifiable region of
+the galaxy. However, for faint objects, the Petrosian magnitudes tend
+to become noisy. Thus, for galaxies with $z>0.25$ we instead use the
+higher signal-to-noise ``model magnitudes.'' Model magnitudes are
+calculated in all bands using a single weighted aperture. The weighted
+aperture is the better fitting model (pure exponential or pure de
+Vaucouleurs) to the galaxy image in the $r$-band. In
+\citet{stoughton01a} we are explicitly warned (in paragraphs written
+by the lead author of this paper!)  not to mix Petrosian and model
+magnitudes in a single analysis, since they measure galaxies in very
+different ways. Nevertheless, for the purposes of contraining galaxy
+template SEDs, it is perfectly acceptable to use {\it any}
+well-defined part of any galaxy.  We reddening-correct both types of
+magnitudes using the dust maps of \citet{schlegel98a}.
 
@@
 (totalling about 1,000 objects) which were selected by a photometric
-redshift algorithm ({\bf ref who?}) to be at around $z\sim 0.3$. These
-objects are invaluable for tying down the blue end of the templates.
+redshift algorithm ({\bf ref who?}) to be at around $z\sim
+0.3$--$0.4$. These objects are invaluable for tying down the blue end
+of the templates.
 
@@
 
+For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
+defined by linear combinations of ten Bruzual-Charlot instantaneous
+burst models with ages ranging from $3 \times 10^7$ to $2\times
+10^{10}$ years, five with metallicity $Z=0.02$ and five with
+$Z=0.004$, all assuming a Salpeter Initial Mass Function. We find that
+we cannot recover information about the SED below a certain scale,
+except perhaps around the 4000\AA\ break. Thus, we smooth most of the
+wavelength regime of the templates using a Gaussian with a standard
+deviation of $\sigma = 300$\AA; around 4000\AA, we smooth only with
+$\sigma = 150$\AA.
+
 We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
@@
 four templates in the sense that the resulting templates reproduce the
-magnitudes nearly as well. However, since one of the applications of
-these SED determinations is the distribution of galaxy colors in
-fixed-frame magnitudes, we don't want to artificially reduce the
-dimensionality of the color space to only two.
+$griz$ magnitudes nearly as well. However, the fourth template is
+necessary to recover the $u$-band flux to better than about 15\%. In
+addition, since one of the applications of these SED determinations is
+the distribution of galaxy colors in fixed-frame magnitudes, we don't
+want to artificially reduce the dimensionality of the color space to
+only two.
 
@@
 orthogonalize the templates and over which to calculate the flux. We
-choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA and
+choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA\ and
 $\lambda_{\mathrm{max}}=7500$\AA, since this range is included for
@@
 peak of the $z$-band response is at 8700\AA). Therefore, the odd
-behavior near the edges (particularly at the red edge) should not be
-taken too seriously.
+behavior near the edges should not be taken too seriously.
 
 These reconstructed spectra do an excellent job of reproducing the
-observed galaxy fluxes. Figure \ref{model} shows the differences
-between the observed and reconstructed fluxes as a function of
-redshift and as a function of absolute magnitude (in $\band{0.1}{r}$).
-The solid lines show the median differences. The estimates of the
-fluxes in $g$, $r$, $i$, and $z$ do not correlate strongly with any of
-these parameters, while for $u$ there is a 5--10\% discrepancy for the
-most distant galaxies, in the sense that the model $u$-band flux is
-too bright. The dotted lines show the boundaries containing 80\% of
-the objects surrounding the median (roughly the $1.3\sigma$ limits if
-the distribution were Gaussian). For $g$, $r$, and $i$, the scatter is
-very small --- certainly comparable to the errors in calibration ---
-and roughly independent of redshift and absolute magnitude. For $z$,
-the scatter is somewhat larger, perhaps reflecting the poor
-signal-to-noise in $z$, but remains constant with redshift. For $u$,
-the scatter is larger, perhaps reflecting the fact that four templates
-simply is not enough to characterize galaxy SEDs simultaneously in the
-optical and near ultra-violet. After all, galaxies are known to have
-large variations in their near-ultraviolet properties (because it
-depends strongly on dust content and recent star-formation). This
-effect may become more important with redshift as $u$ probes bluer and
-bluer regions of the SED. All told, these spectra seem to be fair
-estimates of the true underlying spectra.
+observed galaxy fluxes. Figure \ref{k_model_plot} shows the
+differences between the observed and reconstructed fluxes as a
+function of redshift and as a function of absolute magnitude (in
+$\band{0.1}{r}$).  The solid lines show the median differences. The
+estimates of the fluxes in $g$, $r$, $i$, and $z$ do not correlate
+strongly with any of these parameters, while for $u$ there is a
+5--10\% discrepancy for the most distant galaxies, in the sense that
+the model $u$-band flux is too bright. The dotted lines show the
+boundaries containing 80\% of the objects surrounding the median
+(roughly the $1.3\sigma$ limits if the distribution were
+Gaussian). For $g$, $r$, and $i$, the scatter is very small ---
+certainly comparable to the errors in calibration --- and roughly
+independent of redshift and absolute magnitude. For $z$, the scatter
+is somewhat larger, perhaps reflecting the poor signal-to-noise in
+$z$, but remains constant with redshift. For $u$, the scatter is
+larger, perhaps reflecting the fact that four templates simply is not
+enough to characterize galaxy SEDs simultaneously in the optical and
+near ultra-violet. After all, galaxies are known to have large
+variations in their near-ultraviolet properties (because it depends
+strongly on dust content and recent star-formation). This effect may
+become more important with redshift as $u$ probes bluer and bluer
+regions of the SED. All told, these spectra seem to be fair estimates
+of the true underlying spectra.
 
 An important test of the reasonableness of the fits is to check that
-for a fixed absolute magnitude, the distribution of the SED fits
-depends little on redshift. Figures \ref{coeffs1}--\ref{coeffs4}
+for a fixed type of galaxy, the distribution of the SED fits
+depends little on redshift. We must be careful in performing this
+check that we are comparing Figures \ref{coeffs1}--\ref{coeffs4}
 demonstrates this to be the case for galaxies in several luminosity
@@
 \begin{equation}
-m_{AB} = -48.60 - 2.5 \log_{10}\left[
+m_{AB} = -2.41 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}

### 2002-02-13 docs/paper/paper.tex
add plots
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@35 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \begin{abstract}
-Certain astronomical questions, such as the evolution of the
-luminosity function of galaxies and the relationship between galaxy
-luminosities and galaxy colors, require transforming inherently
-redshift-dependent measurements of galaxy magnitudes into measures of
-galaxy properties which are consistent over a range of redshifts.  To
+Broad-band measurements of galaxy flux suffer from the difficulty that
+they probe different regions of the rest-frame galaxy spectrum for
+galaxies at different redshifts.  Certain astronomical questions, such
+as the evolution of the luminosity function of galaxies, require
+transforming these inherently redshift-dependent measurements of
+galaxy magnitudes into redshift-independent quantities. To prepare to
 address these astronomical questions, investigated in detail in
@@
 broad-band filters. It is further complicated by the fact that SEDs of
-distant galaxies are redshifts due to the expansion of the
+distant galaxies are redshifted due to the expansion of the
 universe. Thus, the more distant the galaxy, the further the observed
@@
 
-In the past, accounting for these effects has in general come in the
-form of ``$K$-corrections'' applied to the observed magnitudes. That
-is, a function $K(z)$ is added to the standard cosmological bolometric
+In the past, people have in general accounted for these effects using
+``$K$-corrections'' applied to the observed magnitudes. That is, a
+function $K(z)$ is added to the standard cosmological bolometric
 distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
@@
 $\vv{v}_i(\lambda)$, one can find an orthonormal set of spectra
-$\vv{b}_i(\lambda)$ spanning the space.
-
-To be more precise, we define the dot-product in the space of galaxy
-spectra to be:
+$\vv{b}_i(\lambda)$ spanning the space.  To be more precise, we define
+the dot-product in the space of galaxy spectra to be:
 \begin{equation}
@@
 defined not over the full range covered by the $\vv{v}_i(\lambda)$
-spectra, but instead over a smaller range corresponding to the range
-which is within the observed wavelength range over most of the
-redshift range of the sample. Otherwise, orthogonality is defined in
-terms of regions of the spectra which are poorly constrained, which
-make interpretation of the fits inconvenient.  This choice does not
-affect the results for the reconstructed SED fit quantitatively
-whatsoever.\footnote{If the template spectra were identical in the
-subrange chosen and only differed outside of it, this choice would
-cause computational problems, but in fact the spectra are all
-independent in the restricted, as well as the full, wavelength range.}
-We then define the basis of the SED space such that the $N_b$ template
-SEDs, $\vv{v}_i$, are linear combinations of $N_b$ basis spectra
-$\vv{b}_i$ and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These
-conditions do not fully specify the $\vv{b}_i$, which naturally may be
-rotated arbitrarily within our $N_b$-dimensional subspace of spectral
-space.
+spectra, but instead over a smaller range which is within observable
+wavelengths over most of the redshift range of the sample. If we
+defined orthogonality over the full range of $\vv{v}(\lambda)$, much
+of which is unobservable, it would complicate the interpretation of
+our fits.  This choice does not affect the results for the
+reconstructed SED fit quantitatively whatsoever.\footnote{If the
+template spectra were identical in the subrange chosen and only
+differed outside of it, this choice would cause computational
+problems, but in fact the spectra are all independent in the
+restricted, as well as the full, wavelength range.}  We then define
+the basis of the SED space such that the $N_b$ template SEDs,
+$\vv{v}_i$, are linear combinations of $N_b$ basis spectra $\vv{b}_i$
+and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These conditions do
+not fully specify the $\vv{b}_i$, which naturally may be rotated
+arbitrarily within our $N_b$-dimensional subspace of spectral space.
 
@@
 \begin{equation}
-F_{\mathrm{rec},il} = \int_0^\infty d\lambda f(\lambda) S(\lambda),
+F_{\mathrm{rec},il} = \int_0^\infty d\lambda
+f_{\mathrm{rec},i}(\lambda) S_l(\lambda),
 \end{equation}
-where $S(\lambda)$ is the response of the instrument.  We define
+where $S_l(\lambda)$ is the response of the instrument in band $l$.  We define
 \begin{equation}
@@
 are aligned with the principal axes of the ellipsoid. Furthermore, we
-normalize the vectors $\vv{e}_i$ such that the $a_0=F_t$ in units of
-ergs cm$^{-2}$ s$^{-1}$. This choice of coordinate system is
-arbitrary and does not affect our quantitative results, but it is
-convenient.
+normalize the vectors $\vv{e}_i$ such that the zeroth component
+$a_0=F_t$ is in units of ergs cm$^{-2}$ s$^{-1}$. This choice of
+coordinate system is arbitrary and does not affect our quantitative
+results, but it is convenient.
 
@@
 
-The version of the $K$-correction code (v1\_2) implementing the
+The version of the $K$-correction code ({\tt v1\_2}) implementing the
 calculations described here, along with eigentemplates and filter
@@
 for the template-fitting also exists in stand-alone C programs (which
-calls the same routines, guaranteeing consistency).
+call the same routines, guaranteeing consistency).
 
@@
 {\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
-\cr
 \end{eqnarray}

### 2002-02-14 docs/paper/paper.tex
lots
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@38 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 differences between the observed and reconstructed fluxes as a
-function of redshift and as a function of absolute magnitude (in
-$\band{0.1}{r}$).  The solid lines show the median differences. The
-estimates of the fluxes in $g$, $r$, $i$, and $z$ do not correlate
-strongly with any of these parameters, while for $u$ there is a
-5--10\% discrepancy for the most distant galaxies, in the sense that
-the model $u$-band flux is too bright. The dotted lines show the
-boundaries containing 80\% of the objects surrounding the median
-(roughly the $1.3\sigma$ limits if the distribution were
-Gaussian). For $g$, $r$, and $i$, the scatter is very small ---
-certainly comparable to the errors in calibration --- and roughly
-independent of redshift and absolute magnitude. For $z$, the scatter
-is somewhat larger, perhaps reflecting the poor signal-to-noise in
-$z$, but remains constant with redshift. For $u$, the scatter is
-larger, perhaps reflecting the fact that four templates simply is not
-enough to characterize galaxy SEDs simultaneously in the optical and
-near ultra-violet. After all, galaxies are known to have large
-variations in their near-ultraviolet properties (because it depends
-strongly on dust content and recent star-formation). This effect may
-become more important with redshift as $u$ probes bluer and bluer
-regions of the SED. All told, these spectra seem to be fair estimates
-of the true underlying spectra.
+function of redshift. There are no systematic trends with redshift,
+and the standard deviations of the differences between the observed
+and the reconstructed fluxes (shown on the right bottom corner of each
+panel) are of order the photometric errors in the sample. This
+agreement should not be overly surprising given that we have five
+galaxy fluxes and four templates to fit, and when one considers the
+fact that galaxy colors are known to be correlated with one another.
 
 An important test of the reasonableness of the fits is to check that
-for a fixed type of galaxy, the distribution of the SED fits
-depends little on redshift. We must be careful in performing this
-check that we are comparing Figures \ref{coeffs1}--\ref{coeffs4}
-demonstrates this to be the case for galaxies in several luminosity
-ranges. The left column of Figure \ref{coeffs1} shows the three
-coefficient ratios $a_i/a_0$ for each galaxy as a function of
-redshift. The right column shows the distribution of these coefficient
-ratios for two redshift ranges, showing that they are nearly
-consistent. The resulting change in the color distribution inferred
-from these coefficients is similarly small, at the level of what one
-would expect intrinsic evolution to cause.
+for a fixed type of galaxy, the distribution of the fixed frame colors
+depends little on redshift. Because the galaxies shown in Figures
+\ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
+selected --- some are Main Sample
+
+We must be careful in performing this check that
+we are comparing Figures \ref{coeffs1}--\ref{coeffs4} demonstrates
+this to be the case for galaxies in several luminosity ranges. The
+left column of Figure \ref{coeffs1} shows the three coefficient ratios
+$a_i/a_0$ for each galaxy as a function of redshift. The right column
+shows the distribution of these coefficient ratios for two redshift
+ranges, showing that they are nearly consistent. The resulting change
+in the color distribution inferred from these coefficients is
+similarly small, at the level of what one would expect intrinsic
+evolution to cause.
 

### 2002-02-14 src/color_interp.c

@@ -1,8 +1,7 @@
 #include <stdio.h>
 #include <string.h>
 #include <math.h>
-#include "utils.h"
-#include "color_interp.h"
+#include <kcorrect.h>
 
 /*
  * color_interp.c
@@ -14,9 +13,13 @@
  * 6/2001
  */
 
-#ifndef ZZERO
-#define ZZERO 0.
-#endif
+#define ZZERO 0.1
+#define MAGGIEFILE "testsp.K.dat"
+#define RFFILE "testsp.K.cibreak.dat"
+#define INCLUDEBREAK
+#define NGALS 6534
+#define FILTERPATH "../data/filters"
+#define FILTERBASE "sdss_"
 
 static int band,currband,igal;
 static int band0,band1;
@@ -27,18 +30,17 @@ static double lfilt[5][300];
 static int nfilt[5];
 
 /* Galaxy properties */
-static double galaxy_mag[5][NALL];
-static double galaxy_rfmag[5][NALL];
-static double galaxy_magerr[5][NALL];
-static double galaxy_z[NALL];
-static int galaxy_use[NALL];
+static double galaxy_maggies[5][NGALS];
+static double galaxy_rfmag[5][NGALS];
+static double galaxy_invvar[5][NGALS];
+static double galaxy_z[NGALS];
 
 static double filtfunc(double lambda)
 {
 	unsigned long i,ip1;
 	double s,val;
 
-	locate(lfilt[currband], nfilt[currband], lambda, &i);
+	k_locate(lfilt[currband], nfilt[currband], lambda, &i);
 	if(i>=nfilt[currband]-1 || i<0)
 		return(0.);
 		
@@ -74,7 +76,7 @@ static double bfunc(double lambda)
 {
 	unsigned long i,ip1;
 
-	if(lambda/(1.+ci_z)<4100.)
+	if(lambda/(1.+ci_z)<4000.)
 		return(ci_break*pow(lambda/(1.+ci_z),2.)/(1.+ci_z));
 	else 
 		return(ci_A*pow(lambda/(1.+ci_z),ci_alpha)/(1.+ci_z));
@@ -98,17 +100,17 @@ double alpha_fit(double alpha)
 	ci_alpha=alpha;
 
 	currband=band0;
-	scaleval0=qromo(scalefiltfunc,lfilt[currband][0],
[…]

### 2002-02-14 docs/paper/paper.tex
added speck
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@39 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 a pipeline created specifically for the SDSS instrumentation
-(\citealt{schlegel02a}). We also use redshifts determined by the same
-pipeline; the official SDSS redshifts are obtained from a different
+(\citealt{schlegel02a}), which also fits for the redshift of each
+spectrum. The official SDSS redshifts are obtained from a different
 pipeline (\citealt{subbarao02a}). The two independent versions provide
@@
 template SEDs, it is perfectly acceptable to use {\it any}
-well-defined part of any galaxy.  We reddening-correct both types of
-magnitudes using the dust maps of \citet{schlegel98a}.
+well-defined part of any galaxy.  For our purposes, it is more
+important to have high signal-to-noise measurement for any
+well-defined part of our galaxies, even if the region of the galaxies
+probed and their total estimated fluxes are inconsistent with
+redshift. We extinction-correct both types of magnitudes using the
+dust maps of \citet{schlegel98a}.
 
@@
 0.3$--$0.4$. These objects are invaluable for tying down the blue end
-of the templates.
+of the templates. Finally, we exclude galaxies in the redshift range
+$0.28 <z<0.32$ from the fit for the templates, for reasons which we
+will explain more fully below (nevertheless, we still can and do use
+the resulting templates to analyze galaxies in this range).
 
@@
 We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
-and still fit allow freedom to fit for the templates themselves. Since
+and still allow freedom to fit for the templates themselves. Since
 galaxies live in a low-dimensional space, we find it is worth being
@@
 \ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
-selected --- some are Main Sample
-
-We must be careful in performing this check that
-we are comparing Figures \ref{coeffs1}--\ref{coeffs4} demonstrates
-this to be the case for galaxies in several luminosity ranges. The
-left column of Figure \ref{coeffs1} shows the three coefficient ratios
-$a_i/a_0$ for each galaxy as a function of redshift. The right column
-shows the distribution of these coefficient ratios for two redshift
-ranges, showing that they are nearly consistent. The resulting change
-in the color distribution inferred from these coefficients is
-similarly small, at the level of what one would expect intrinsic
-evolution to cause.
+selected --- some are Main Sample galaxies, some are Luminous Red
+Galaxies, and some are selected from the photometric redshift plates
+--- we will split our sample into two well-defined sets of
+objects. First, we choose a set of Main Sample galaxies in the
+luminosity range $-21.5 < M_{\band{0.1}{r}} < -21$. We reconstruct the
+four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
+$^{0.1}(i-z)$ for these galaxies and plot the reconstructions as a
+function of redshift in left-hand panels of Figure
+\ref{main_colors_plot}. The plots show a general consistency in the
+fixed-frame color distribution of these objects with redshift. The
+right-hand panels show the distribution for redshifts $z<0.13$ (solid
+histogram) and $z>0.13$ (dotted histogram) . Small changes are
+discernible in the distributions, mostly attributable to the increased
+errors at higher redshift. Note that for $z<0.1$ the $^{0.1}(u-g)$
+color depends on an extrapolation of the SED in the blue, while for
+$z>0.1$ the $^{0.1}(i-z)$ color depends on an extrapolation of the SED
+in the red.
+
+Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.1}{r}}
+< -22.5$. Now we reconstruct the colors to a fixed-frame corresponding
+to the rest-frame bands used to observe a galaxy at $z=0.3$.  Near
+$z=0.3$ there is some singular behavior of the colors, which we will
+discuss in more detail later. Again, for $z<0.3$ the $^{0.3}(u-g)$
+color relies on an extrapolation, which is probably responsible for
+the scatter in that color at low redshifts. Otherwise, the
+distribution of colors is remarkably constant. $^{0.3}(g-r)$ (which is
+very similar to $^{0.0}(u-g)$) experiences a blueward shift of about
+0.1 magnitudes between $z=0.1$ and $z=0.45$, which is attributable to
+passive evolution.
 
@@
 photometry itself, seem physically reasonable, and are consistent over
-the range of redshifts we consider ($0.0. < z < 0.5$).
+the range of redshifts we consider ($0.0 < z < 0.5$). The only
+remaining problem is the aforementioned difficulties at $z=0.3$,
+discussed in more detail in the next section.
+
+\section{The $g$/$r$ Gap}
+\label{grgap}
+
+It is clear from Figure \ref{k_kcorrect_plot} that $z=0.3$ is a
+special redshift regime where our fits become unphysical. In fact, it
+is easy to demonstrate that the diagonal spur (going from the upper
+left to the lower right) around the red dot in the upper right panel
+of Figure \ref{k_coeffdist_plot} is due to galaxies at $z\sim
+0.3$. The difficulty arises because the 4000 \AA\ break lies in the
+gap between $g$ and $r$ at $z=0.3$. The algorithm requires strong
+gradients in the spectrum in this area to explain the 4000\AA break
+--- however, the gap in the bands causes a degeneracy arise.
+
+To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
+corresponding to the direction of the spur in the $a_1/a_0$-$a_3/a_0$
+plane of Figure \ref{k_coeffdist_plot}. We overplot the \band{0.3}{g}
+and \band{0.3}{r} bands. The spur has a strong peak in the gap between
+the two bands at $z=0.3$. Because of this gap, at $z\sim 0.3$ the
+algorithm tends to use this direction to help fit the $u$, $i$, and
+$z$ fluxes, with relatively little harm done to the $g$ and $r$
+fits. This explains why the reconstructed fluxes in Figure
+\ref{k_model_plot} fit well even at $z=0.3$, even though the spur at
+$z=0.3$ in the distribution of coefficients is so evident.  This
+difficulty is why, in Section \ref{data}, we exclude the regime around
+$z\sim 0.3$; otherwise the templates became distracted by the
+degeneracies in this regime.
+
+The question arises of what to do about this problem. One solution is
+to simply work around it. It is clear we cannot use the fits presented
+here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
+reliably. However, nothing prevents one from $K$-correcting galaxies
+at other redshifts {\it to} $z=0.3$. If you are considering the LRG
+sample, the median redshift is close to $z=0.3$ anyway, so this is a
+viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).
+
+A second solution, not adopted here, is to place stronger constraints
+in the SED fit on the form of the 4000 \AA\ break. One obvious way of
+doing this is to use the spectra to constrain the size of the break
+(making it a weak constraint to prevent biases). This constraint could
+be applied in a linear fashion to retain the speed of our method ---
+except for the fact that you would have to read in the spectra to
+apply the $K$-correction. For simplicity, and because the aforemention
+work-around is so easy to obey, we have not implemented this solution
+here.
 
@@
 
-With a model of the rest-frame galaxy SED in hand, it is 
-straightforward to project that SED onto a filter corresponding to 
-an observation at any redshift, and predict the so-called
-$K$-correction to the galaxy magnitude in order to compare galaxies 
-at different redshifts. One simply calculates the AB magnitude at 
-any redshift as:
+With a reconstruction of the rest-frame galaxy SED in hand, it is
+straightforward to project that SED onto a filter corresponding to an
+observation at any redshift, and predict the so-called $K$-correction
+to the galaxy magnitude in order to compare galaxies at different
+redshifts. One simply calculates the AB magnitude at any redshift as:
 \begin{equation}
-m_{AB} = -2.41 - 2.5 \log_{10}\left[
+m_{AB,\mathrm{rec}} = -2.41 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
-f(\frac{\lambda}{1+z}) R(\lambda)}
+f_{\mathrm{rec}}(\frac{\lambda}{1+z}) R(\lambda)}
 {\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].

### 2002-02-15 docs/paper/paper.tex
writing
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@43 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-\title{Modeling Galaxy Spectral Energy Distributions in the SDSS
+\title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS
 Spectroscopic Survey}
@@
 Mamoru Doi\altaffilmark{\ref{Tokyo}},
-%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
+Daniel Eisenstein\altaffilmark{\ref{Arizona}},
 James E. Gunn\altaffilmark{\ref{Princeton}}, and
@@
 \begin{abstract}
-Broad-band measurements of galaxy flux suffer from the difficulty that
-they probe different regions of the rest-frame galaxy spectrum for
-galaxies at different redshifts.  Certain astronomical questions, such
-as the evolution of the luminosity function of galaxies, require
-transforming these inherently redshift-dependent measurements of
-galaxy magnitudes into redshift-independent quantities. To prepare to
-address these astronomical questions, investigated in detail in
-subsequent papers, we implement here the method of \citet{csabai00a}
-for fitting spectral energy distributions (SEDs) to broad-band
-photometric observations, in the context of the optical observations
-of the Sloan Digital Sky Survey (SDSS). Linear combinations of four
-spectral templates can reproduce the five SDSS magnitudes of galaxies
-to the precision of the photometry. Expressed in the correct
-coordinate system, the distribution of the coefficients of the
-templates is planar, and in fact nearly linear. The resulting
-reconstructed SEDs (the sum of the templates) can be used to recover
-fixed-frame magnitudes over a range of redshifts. We show that the
-results of doing so yield consistent results within the SDSS Main
-Galaxy Sample and the SDSS Luminous Red Galaxy Sample, in the sense
-that within each sample the trend with redshift of intrinsic colors
-for similar type galaxies are small. A potential problem in the
-measurement of the SDSS filter response curves is noted. We also
-validate the results using galaxy spectrophotometry.  All code and
-templates from this paper are publicly available and well-documented;
-they are designed such that they can easily be incorporated into other
-applications. The code is easily adapted to handle galaxy observations
-in any AB magnitude system in the rest-frame optical regime.
+Broad-band measurements of galaxy fluxes suffer from the difficulty
+that for galaxies at different redshifts the observations probe
+different regions of the rest-frame galaxy spectrum. Certain
+astronomical questions, such as the evolution of the luminosity
+function of galaxies, require transforming these inherently
+redshift-dependent magnitudes into redshift-independent quantities. To
+prepare to address these astronomical questions, investigated in
+detail in subsequent papers, we implement here the method of
+\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
+broad-band photometric observations, in the context of the optical
+observations of the Sloan Digital Sky Survey (SDSS). Linear
+combinations of four spectral templates can reproduce the five SDSS
+magnitudes of all galaxies to the precision of the
+photometry. Expressed in the appropriate coordinate system, the
+distribution of the coefficients multiplying the templates is planar,
+and in fact nearly linear. The resulting reconstructed SEDs can be
+used to recover fixed-frame magnitudes over a range of redshifts. We
+show that the results of doing so yield consistent results, in the
+sense that within each sample the intrinsic colors of similar type
+galaxies are nearly constant with redshift. We note one remaining
+problem associated with the gap between the $g$ and $r$ bands. We
+compare our results to simpler interpolation methods and galaxy
+spectrophotometry.  The software which generates these results is
+publicly available and easily adapted to handle galaxy observations in
+any AB magnitude system in the rest-frame optical regime.
 \end{abstract}
@@
 care has to be taken to recover knowledge of galaxy spectral energy
-distributions (SEDs) from these observations. This paper focuses on a
-method of facing this problem which (in the optical wavelength regime)
-is fast, robust, and consistent over a large range of redshifts.
-
-This task is nontrivial because SEDs of galaxies contain important
+distributions (SEDs) from these observations.  Reconstructing galaxy
+SEDs is nontrivial because SEDs of galaxies contain important
 information on scales considerably smaller than the width of typical
-broad-band filters. It is further complicated by the fact that SEDs of
-distant galaxies are redshifted due to the expansion of the
-universe. Thus, the more distant the galaxy, the further the observed
-bandpass is blueshifted relative to the rest-frame spectral energy
-distribution of the object observed.
+broad-band filters. Furthermore, SEDs of distant galaxies are
+redshifted such that the more distant the galaxy, the further the
+observed bandpass is blueshifted relative to the rest-frame spectral
+energy distribution of the object observed.  This paper focuses on a
+method of facing these problems which (in the optical wavelength
+regime) is fast, robust, and consistent over a large range of
+redshifts.
 
 In the past, people have in general accounted for these effects using
-``$K$-corrections'' applied to the observed magnitudes. That is, a
-function $K(z)$ is added to the standard cosmological bolometric
-distance modulus $\mathrm{DM}(z)$. Sometimes a single function has
-been applied regardless of the galaxy type, though more recently it
-has become standard to use a discrete set of $K(z)$ functions
-depending on galaxy type (based either on morphology or spectral
-features). As described in \cite{oke68a} and later papers, the
-function is based on the projection of an assumed galaxy SED
+``$K$-corrections'' applied to the observed magnitudes. In these
+analyses, a function $K(z)$ is added to the standard cosmological
+bolometric distance modulus $\mathrm{DM}(z)$. Sometimes a single
+function has been applied regardless of the galaxy type, though more
+recently it has become standard to use a discrete set of $K(z)$
+functions depending on galaxy type (based either on morphology or
+spectral features). As described in \cite{oke68a} and later papers,
+the function is based on the projection of an assumed galaxy SED
 redshifted to $z$ onto the measured instrument response as a function
@@
 reflectivity of the mirrors, the filter transmission, and the response
-of the CCD device or photographic emulsion).
-
-However, galaxies do not all have the same SED, nor are they selected
-from some discrete set of SEDs. For this reason, schemes for applying
-the $K$-corrections can fail to be self-consistent: the SED assumed
-for the $K$-corrections can be significantly inconsistent with the
-observed galaxy colors! As we demand more precision from our
-astronomical data analysis in new, large multi-band surveys, and in
-particular as we try to quantify the evolution of galaxies, we must
-take a more sophisticated approach to approximating fixed-frame
-observations of galaxies.
+of the CCD device or photographic emulsion).  However, galaxies do not
+all have the same SED, nor are they selected from some discrete set of
+SEDs. For this reason, these schemes for applying the $K$-corrections
+can fail to be self-consistent: the SED assumed for the
+$K$-corrections can be significantly inconsistent with the observed
+galaxy colors! As we demand more precision from our astronomical data
+analysis in new, large multi-band surveys, and in particular as we try
+to quantify the evolution of galaxies, we must take a more
+sophisticated approach to approximating fixed-frame observations of
+galaxies.
 
@@
 recover a model SED, which can be used to synthesize the galaxy's
-magnitude in any bandpass. In addition, this method provides a natural
-way to infer an estimate of the flux within the whole optical range.
-The approach is equivalent (nearly identical) to the photometric
-metric redshift estimation methods of \citet{csabai99a} and
-\citet{budavari00a}, except we use slightly different coordinate
-systems.
+magnitude in any bandpass. Although the operation we are performing on
+the magnitudes is not strictly speaking a ``$K$-correction,'' we will
+refer to it as such in this and subseqent papers. Our approach is
+equivalent (nearly identical) to the photometric metric redshift
+estimation methods of \citet{csabai00a} and \citet{budavari00a},
+except we use slightly different coordinate systems. In fact, the
+software includes a fast and relatively accurate photometric redshift
+estimator based on their method.
 
@@
 the reader for upcoming papers which will rely heavily on the
-reliability of the SED fits described here.  However, having
-implemented the fits, there is no reason not to share them with the
-community; thus, this paper serves the secondary purpose of describing
-a piece of software being released to the astronomical community. The
-computer software created is a C shared object library, around which
-we have written both stand-alone C programs and IDL routines. We have
-made the source code available publicly, through a URL on the World
-Wide Web and through a public CVS repository. Improvements or ports to
-other languages implemented by users may be incorporated into the code
-upon request. The conditions of use for the code are that this paper
-is cited in any resulting refereed journal article and that the
-version of the code used is specified in any such paper. The version
-of the code used to make the figures for this paper is {\tt kcorrect
-v1\_2}.
+reliability of the fixed-frame magnitudes determined here.  In
+addition, having implemented the fits, there is no reason not to share
+them with the astronomical community; thus, this paper serves the
+secondary purpose of describing the release of a piece of
+software. The computer software we distribute builds itself into a C
+shared object library, around which we have written both stand-alone C
+programs and IDL routines. We have made the source code available
+publicly, through a URL on the World Wide Web and through a public CVS
+repository. Improvements or ports to other languages implemented by
+users may be incorporated into the code upon request. The conditions
+of use for the code are that this paper is cited in any resulting
+refereed journal article and that the version of the code used is
+specified in any such paper. The version of the code used to make the
+figures for this paper is {\tt kcorrect v1\_2}.
 
@@
 results, and discusses how best to use the results of such
-fits. Section \ref{conclusion} concludes and discusses future
+fits. Section \ref{conclusions} concludes and discusses future
 development of the method described here.
 
-\section{Fitting SEDs to Galaxy Broad-band Colors}
+\section{Calculating Fixed-frame Galaxy Magnitudes}
 \label{sedfit}
+
+First, we describe how we reconstruct galaxy SEDs from broad-band
+magnitude measurements. Second, we describe how we convert these SEDs
+into estimates of fixed-frame galaxy magnitudes.
+
+\subsection{Fitting SEDs to Galaxy Broad-band Magnitudes}
 
@@
 large. In principle, this should be a complete set of galaxy SEDs if
-we want to express {\it any} galaxy SED in this basis set; in
-practice, this property is not necessary.  Given a set of
-$\vv{v}_i(\lambda)$, one can find an orthonormal set of spectra
-$\vv{b}_i(\lambda)$ spanning the space.  To be more precise, we define
-the dot-product in the space of galaxy spectra to be:
+we want {\it all} galaxy SEDs to be in the space spanned by the
+$\vv{v}_i(\lambda)$; in practice, this property is not necessary.
+Given a set of $\vv{v}_i(\lambda)$, one can find an orthonormal set of
+spectra $\vv{b}_i(\lambda)$ spanning the same space.  To be more
+precise, we define the dot-product in the space of galaxy spectra to
+be:
 \begin{equation}
@@
 spectra, but instead over a smaller range which is within observable
-wavelengths over most of the redshift range of the sample. If we
-defined orthogonality over the full range of $\vv{v}(\lambda)$, much
-of which is unobservable, it would complicate the interpretation of
-our fits.  This choice does not affect the results for the
-reconstructed SED fit quantitatively whatsoever.\footnote{If the
-template spectra were identical in the subrange chosen and only
-differed outside of it, this choice would cause computational
+wavelengths over most of the redshift range of the
+sample. \footnote{Were we to define orthogonality over the full range
+of $\vv{v}(\lambda)$, much of which is unobservable, the geometry of
+the distribution of galaxies in this space would be more complicated.
+If the template spectra were identical in the subrange chosen and only
+differed outside of it, our choice would cause computational
 problems, but in fact the spectra are all independent in the
@@
 range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
-estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$. The
-locus of points of a fixed flux in the component space $a_k$ lie on a
-plane. Therefore, a natural choice for the orientation of the axes
-$\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
+estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$.
+Galaxies of some fixed $F_t$ clearly lie on a plane in the component
+space $a_k$. Therefore, a natural choice for the orientation of the
+axes $\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
 constant flux; in this manner, the coefficient $a_0$ in Equation
-(\ref{lincomb}) is directly proportional to $F$ and is thus purely a
-measure of the object's flux (in the wavelength range considered),
-while the parameters $a_i/a_0$ (where $i>0$) primarily measure the
-shape of the galaxy SED in the wavelength range considered. $F$ can be
-trivially related to $L$, the luminosity in this range, by the inverse
-square of the luminosity distance.  As it turns out, the parameters
-$a_i/a_0$ tend to be distributed very approximately in an ellipsoid,
-with not much curvature within the space defined by the template
-SEDs. Thus, we rotate the axes $\vv{e}_i$ (for $i>0$) such that they
-are aligned with the principal axes of the ellipsoid. Furthermore, we
-normalize the vectors $\vv{e}_i$ such that the zeroth component
-$a_0=F_t$ is in units of ergs cm$^{-2}$ s$^{-1}$. This choice of
-coordinate system is arbitrary and does not affect our quantitative
-results, but it is convenient.
+(\ref{lincomb}) is directly proportional to $F_t$ and is thus purely a
+measure of the object's flux (in the wavelength range
+$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$), while the
+parameters $a_i/a_0$ (where $i>0$) primarily measure the shape of the
+galaxy SED in the wavelength range $\lambda_{\mathrm{min}} <\lambda <
+\lambda_{\mathrm{max}}$. $F_t$ can be trivially related to $L_t$, the
+total luminosity in this range, by the inverse square of the
+luminosity distance (from, for example, \citealt{hogg99a}).  As it
+turns out, the parameters $a_i/a_0$ tend to be distributed
+approximately in an ellipsoid, with not much curvature within the
+space defined by the template SEDs. Thus, we rotate the axes
+$\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
+axes of the ellipsoid. Furthermore, we normalize the vectors
+$\vv{e}_i$ such that the zeroth component $a_0=F_t$ is output by the
+code in units of ergs cm$^{-2}$ s$^{-1}$. This choice of coordinate
+system is arbitrary and does not affect our quantitative results, but
+it is convenient.
 
@@
 magnitudes of galaxies. From these reconstructed SEDs one can estimate
-$K$-corrections, develop a measure of galaxy spectral type, or to
-synthesize a very broad-band measure of the galaxy flux.  The
-templates determined during the procedure can, of course, be used to
-estimate photometric redshifts, as \citet{csabai00a} and
-\citet{budavari00a} describe.
+$K$-corrections, develop a measure of galaxy spectral type, or
+synthesize other measurements of galaxy flux.  The templates
+determined during the procedure can, of course, be used to estimate
+photometric redshifts, as \citet{csabai00a} and \citet{budavari00a}
+describe.
+
+\subsection{Calculating $K$-corrections}
+
+We here define a notation which we will use in subsequent papers for
+the effective rest-frame bandpass at any redshift in band $b$, namely:
+\begin{equation}
+\band{z}{b}
+\end{equation}
+read, ``$b$-band at redshift $z$.'' The goal of ``$K$-corrections'' is
+to convert the $ugriz$ bands at redshift $z$ for each galaxy to
+$ugriz$ bands at some fixed redshift $z_0$.
+
+Given AB galaxy magnitudes $^{z}m_{AB}$ and errors for a galaxy at
+redshift $z$, one can use the method of the previous subsection to
+reconstruct the galaxy rest-frame SED.  With this reconstruction in
+hand, it is straightforward to project the SED onto bands at any
+redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
+\begin{eqnarray}
+^{z_0}m_{\mathrm{rec}} &=& -2.41 - 2.5 \log_{10}\left[
+\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
+f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)}
+{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right] \cr
+&=& -48.60 - 2.5 \log_{10}\left[
+\frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
+f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)}
+{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right] 
+\end{eqnarray}
+The simple way to think about this equation is that simply moving away
+from the Earth doing does not change the bolometric flux of the
+galaxy. So when the spectrum is stretched by the $(1+z_0)$ redshift
+factor, as in $f(\lambda/(1+z_0))$, the bolometric flux should be
+constant; that is, you must apply the extra multiplicative factor
+$1/(1+z_0)$ to counteract the stretching of the spectrum.
+
+Given the reconstructed magnitudes, one can calculate the fixed-frame
+magnitudes in one of two ways:
+\begin{enumerate}
+\item Simply use $\band{z_0}{m_{\mathrm{rec}}}$, the reconstructed
+magnitudes at redshift $z_0$, as the fixed-frame magnitudes.
+\item Use the difference
+$\band{z_0}{m_{\mathrm{rec}}}-\band{z}{m_{\mathrm{rec}}}$ as an
+estimate of the correction to apply to the observed $\band{z}{m}$.
+\end{enumerate}
+If the reconstructions $\band{z}{m_{\mathrm{rec}}}$ are very similar
+to the actual observations $\band{z}{m}$ (as we will show is the case
+for our fits), there is very little difference between these two
+methods. Having estimated $\band{z_0}{m}$, you can then calculate the
+luminosity by simply applying to the calculated magnitude the
+cosmological distance modulus (that is, the luminosity
+distance-squared law, as tabulated by, {\it e.g.}  \citealt{hogg97a}).
+
+We want to emphasize here that, while $K$-correcting to a fixed frame
+bandpass is {\it sometimes} necessary in order to achieve a scientific
+objective, it is not {\it always} necessary or appropriate. Because
+$K$-corrections are an inherently uncertain business (the broad-band
+magnitudes just do not uniquely determine the SED) they should be
+avoid or minimized when possible. For example, if one was calculating
+the evolution of clustering of red and blue galaxies separately, it
+would perhaps be wise to perform the separation not on the
+$K$-corrected colors of the galaxies but on the median color of, say,
+$M_\ast$ galaxies as a function of redshift. Similarly, in situation
+where $K$-corrections cannot be ignored, such as the calculation of
+the evolution of the luminosity function, their effect should be
+minimized by, for example, correcting to the median redshift of
+galaxies in the sample.
+
+Now, typically one's measurements in will be difficult to connect to,
+say, solar luminosities in $\band{z}{b}$ (unless $z=0$), simply
+because people have not projected the appropriate stellar
+spectrophotometry onto these bandpasses. However, it is our position
+that stars are better understood than galaxies, and that it therefore
+is in the end simpler to stay as close as possible to the system in
+which the galaxies are observed. In any case, astronomy is quickly
+reaching a level of precision for which the exact nature of the
+bandpasses used has to be known and considered in most analyses of
+observational data.
+
+The reader may ask why, if we are basing our $K$-corrections on a full
+model of the galaxy SED, we do not correct to bandpasses with simpler
+shapes (say, Gaussians). The reason is that we want the effective
+bandpasses to actually have been observed for some galaxies in the
+sample, so that the $K$-correction for those galaxies is formally and
+actually zero. This property is desirable because, as noted above,
+there are uncertainties in the $K$-corrections. Nevertheless, we note
+that synthesizing such magnitudes may be appropriate in certain
+situations, having the virtue (similarly to the $AB$ magnitude system)
+that such magnitudes are very easy to comprehend and synthesize from
+theory. For example, if we wanted to convert our results into a form
+easily comprehended by the future astronomers discussed in the first
+paragraph of Section \ref{motivation}, it would be best to calculate
+fluxes within wide top-hat filters so that the difficulties of
+understanding the magnitude system were minimized.
+
+\subsection{Public Access to the Code}
 
@@
 the appropriate version of the code from this repository are provided
-on the web page noted above. The whole of the code can be used through
-the Research Systems, Incorporated, IDL language; everything except
-for the template-fitting also exists in stand-alone C programs (which
-call the same routines, guaranteeing consistency).
+on the web page noted above). The whole of the code can be used
+through the Research Systems, Incorporated, IDL language; everything
+except for the template-fitting also exists in stand-alone C programs
+(which call the same routines, guaranteeing consistency).
 
@@
 
-In this section we describe the SDSS data set, we describe how we fit
-the templates and the coefficients in detail for this data set, and we
-describe our distribution of code to do this.
+In this section we describe how we applied the above method to the
+SDSS data set.
 
@@
 22.5$. The ultimate goal is to spectroscopically observe 900,000
-galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Bright
-Red Galaxies (BRGs; Eisenstein {\it et al.}, in preparation), and
-100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation)
-will be targeted for spectroscopic follow up using two digital
-spectrographs on the same telescope. Many of the details of the galaxy
-survey are described in a description of the galaxy target selection
-paper (\citealt{strauss02a}). Other aspects of the survey are
-described in the Early Data Release (EDR; \citealt{stoughton01a}). The
-survey has begun in earnest, and has obtained about 20\% of its
-intended data.
-
-As of October 2001, the SDSS had imaged and targeted around 1,739
-square degrees of sky and taken spectra of approximately 170,000
-objects over 1,169 square degrees of that area. As described below,
-only a subset of these galaxies are used here to develop the
+galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Luminous
+Red Galaxies (LRGs; Eisenstein {\it et al.}, in preparation), and
+100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation).
+This spectroscopic follow up uses two digital spectrographs on the
+same telescope as the imaging camera. Many of the details of the
+galaxy survey are described in a description of the galaxy target
+selection paper (\citealt{strauss02a}). Other aspects of the survey
+are described in the Early Data Release (EDR;
+\citealt{stoughton01a}). The survey has begun in earnest, and has
+obtained about 20\% of its intended data.
+
+As of February 2002, the SDSS had targeted and taken spectra of
+approximately XXX galaxies over XXX square degrees. As described
+below, only a subset of these galaxies are used here to develop the
 templates, though we calculate the SED fits for all of them. The
@@
 that used when the objects were targeted. However, for those objects
-which were in the EDR photometry, we used the better calibrations and
-photometry from the EDR.
-
-The magnitudes are calibrated $AB$, meaning that the system is
-designed such that:
+which were in the EDR photometric catalog, we used the better
+calibrations and photometry from the EDR. The versions of the SDSS
+pipeline PHOTO used for the reductions of these data ranged from {\tt
+v5\_0} to {\tt v5\_2}. The changes in the treatment of relatively
+small galaxies, which accounts for most of our sample, did not change
+substantially throughout these versions.
+
+The magnitudes are calibrated in the $AB$ system, meaning that the
+magnitudes are designed such that:
 \begin{eqnarray}
+\label{ABdef}
 m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
@@
 s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes equal to zero.  The
-difference in the zeropoint in the two equations is simply the factor
-of speed of light $c$ (expressed in \AA\ s$^{-1}$) between the
-frequency and wavelength scales.
+difference in the zeropoint in the two equations simply corresponds to
+the factor of the speed of light $c$ (expressed in \AA\ s$^{-1}$)
+between the frequency and wavelength scales.
 
@@
 mirrors, one can then model the response of the entire system. For
-each filter in the SDSS, there are six different CCDs; it has been
+each bandpass in the SDSS, there are six different CCDs; it has been
 shown that the differences between these six camera columns are
@@
 
-\subsection{Calculating $K$-corrections}
-
-With a reconstruction of the rest-frame galaxy SED in hand, it is
-straightforward to project that SED onto a filter corresponding to an
-observation at any redshift, and predict the so-called $K$-correction
-to the galaxy magnitude in order to compare galaxies at different
-redshifts. One simply calculates the AB magnitude at any redshift as:
-\begin{equation}
-m_{AB,\mathrm{rec}} = -2.41 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
-f_{\mathrm{rec}}(\frac{\lambda}{1+z}) R(\lambda)}
-{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].
-\end{equation}
-The simple way to think about this equation is that simply moving away
-from the Earth doing does not change the bolometric flux of the
-galaxy. So when the spectrum is stretched by the $(1+z)$ redshift
-factor, as in $f(\lambda/(1+z))$, the bolometric flux should be
-constant; that is, you must apply the extra multiplicative factor
-$1/(1+z)$ to counteract the stretching of the spectrum. Once you have
-have accounted for the fact that you are observing a particular
-bandpass given the redshift of the object, you can then go ahead and
-apply to the calculated magnitude the cosmological distance modulus
-(that is, the luminosity distance-squared law, as tabulated by, {\it
-e.g.} \citealt{hogg97a}).
-
-We want to emphasize here that, while $K$-correcting to a fixed frame
-bandpass is sometimes necessary in order to achieve a scientific
-objective, it is not always necessary or appropriate. Because
-$K$-corrections are an inherently uncertain business (the broad-band
-magnitudes just do not uniquely determine the SED) they should be
-avoid or minimized when possible. For example, if one was calculating
-the evolution of clustering of red and blue galaxies separately, it
-would perhaps be wise to perform the separation not on the
-$K$-corrected colors of the galaxies but on the median color of, say,
-$M_\ast$ galaxies as a function of redshift. Similarly, in situation
-where $K$-corrections cannot be ignored, such as the calculation of
-the evolution of the luminosity function, their effect should be
-minimized by, for example, correcting to the median redshift of
-galaxies in the sample.
-
-For this reason, we here define a notation which we will use in
-subsequent papers for the effective rest-frame bandpass at any
-redshift in band $b$, namely:
-\begin{equation}
-\band{z}{b}
-\end{equation}
-read, ``$b$-band at redshift $z$.'' Now, typically one's measurements
-in this band will be difficult to connect to, say, solar luminosities
-in $\band{z}{b}$, simply because people have not projected the appropriate
-stellar spectrophotometry onto these bandpasses. However, it is our
-position that stars are {\it better} understood than galaxies, and
-that it therefore is in the end simpler to stay as close as possible
-to the system in which the galaxies are observed. In any case,
-astronomy is quickly reaching a level of precision for which the exact
-nature of the bandpasses used has to be known and considered in most
-analyses of observational data.
-
-The reader may ask why, if we are basing our $K$-corrections on a full
-model of the galaxy SED, we do not correct to bandpasses with simpler
-shapes (say, Gaussians). The reason is that we want the effective
-bandpasses to actually have been observed for some galaxies in the
-sample, so that the $K$-correction for those galaxies is formally and
-actually zero. This property is desirable because, as noted above,
-there are uncertainties in the $K$-corrections.
-
 \subsection{Testing the $K$-corrections}

### 2002-02-15 docs/paper/figures.tex
add plots
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@35 75dff457-8751-0410-ab1a-cd701654ce6b
lots
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@38 75dff457-8751-0410-ab1a-cd701654ce6b
added speck
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@39 75dff457-8751-0410-ab1a-cd701654ce6b
plot tweak
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@44 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 with redshift in any band. The 5-$\sigma$ clipped estimate of the
-scatter around the observed fluxes is listed for each band. In $g$,
-$r$, and $i$ the scatter is consistent with the expected photometric
-errors in the survey. In $z$ and $u$ there is more scatter. In the
-case of $z$ this is attributable to Poisson noise in the
-observations. In the case of $u$, it is not, but it may simply be that
-there are more intrinic differences between galaxies in the $u$-band
-than elsewhere.}
+scatter around the observed fluxes is listed for each band. In $u$,
+$g$, $r$, and $i$ the scatter is consistent with the expected
+photometric errors in the survey at all redshifts. At high redshift
+the scatter in $z$ becomes large, most likely due to increasing
+photometric errors. }
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{k_coeff_plot2.ps}
-\caption{\label{k_coeff_plot1} }
+\plotone{main_colors_plot.ps}
+\caption{\label{main_colors_plot} Color distributions $K$-corrected to
+$z=0.1$ for Main Sample galaxies in the luminosity range
+$-21.5<M_{\band{0.1}{r}}<-21$. This sample is complete for
+$0.05<z<0.15$. Left panels show the colors as a function of
+redshift. Right panel shows the distributions of each color at high
+and low redshift.}
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{k_coeff_plot1.ps}
-\caption{\label{k_coeff_plot1} }
+\plotone{lrg_colors_plot.ps}
+\caption{\label{lrg_colors_plot} Same as Figure
+\ref{main_colors_plot}, now $K$-corrected to $z=0.3$ for LRG galaxies
+in the luminosity range $-22.8<M_{\band{0.3}{r}}<-22.5$. }
 \end{figure}
@@
 \figurenum{\fignum}
-%\plotone{kcorrect.sample8b15.ps}
+\plotone{spur.ps}
+\caption{\label{spur} }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{k_kcorrect_plot.ps}
 \caption{\label{kcorrect.sample8b15} $K$-corrections as a function of
@@
 \figurenum{\fignum}
-%\plotone{ciCompare.sample8b15.ps}
-\caption{\label{ciCompare.sample8b15} Difference in the
+\plotone{compareci.ps}
+\caption{\label{compareci} Difference in the
 $K$-corrections in each band between the method used in Figure
-\ref{kcorrect.sample8b15} and the method of simply interpolating
+\ref{k_kcorrect_plot} and the method of simply interpolating
 between adjacent bands fitting a power-law SED. The differences are
@@
 \figurenum{\fignum}
-%\plotone{cibreakCompare.sample8b15.ps}
-\caption{\label{cibreakCompare.sample8b15} Same as Figure
-\ref{ciCompare.sample8b15}, now comparing the $K$-corrections of Figure
-\ref{kcorrect.sample8b15} with the ``interpolation with a break''
-method. This method fits a power law between adjacent bands, except at
-4000 \AA, where we fit for the size of the 4000\AA\ break (assuming
-that $f(\lambda)\propto \lambda^2$ for $\lambda<4000$ \AA). This
-greatly improves the agreement in \band{0.1}{g} while making the
-disagreement in \band{0.0}{u} only slightly worse. These results
-indicate that it important to account for the structure in the blue
-region of the spectrum when performing $K$-corrections.}
+\plotone{k_speck_plot.fiber.0.1.ps}
+\caption{\label{k_speck_plot.fiber.0.1} }
 \end{figure}
@@
 \figurenum{\fignum}
-%\plotone{specK.ps}
-\caption{\label{specK} Similar to Figure \ref{ciCompare.sample8b15},
-now comparing the $K$-corrections of Figure \ref{kcorrect.sample8b15}
-with $K$-corrections determined from the spectra (which can only be
-calculated for the \band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}
-bands). In all bands, the $K$-corrections are very similar, giving us
-confidence in our results. This result is actually remarkable more for
-what it says about the high quality of the spectrophotometry in the
-SDSS survey.}
+\plotone{k_speck_plot.0.3.ps}
+\caption{\label{k_speck_plot.0.3} }
 \end{figure}
-
-\clearpage
-\stepcounter{thefigs}
-\begin{figure}
-\figurenum{\fignum}
-%\plotone{aperturevsz.M.sample8b15.ps}
-\caption{\label{aperturevsz.M.sample8b15} }
-\end{figure}

### 2002-02-15 docs/paper/paper.tex
plot tweak
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@44 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 David W. Hogg\altaffilmark{\ref{NYU}}
-%David J. Schlegel\altaffilmark{\ref{Princeton}}
+David J. Schlegel\altaffilmark{\ref{Princeton}}
 %Julianne Dalcanton\altaffilmark{\ref{UW}},
@@
 that synthesizing such magnitudes may be appropriate in certain
-situations, having the virtue (similarly to the $AB$ magnitude system)
-that such magnitudes are very easy to comprehend and synthesize from
-theory. For example, if we wanted to convert our results into a form
-easily comprehended by the future astronomers discussed in the first
-paragraph of Section \ref{motivation}, it would be best to calculate
-fluxes within wide top-hat filters so that the difficulties of
-understanding the magnitude system were minimized.
+situations, as they have the virtue (similarly to the $AB$ magnitude
+system) that such magnitudes are very easy to comprehend and
+synthesize from theory. For example, if we wanted to convert our
+results into a form easily comprehended by the hypothetical future
+astronomers discussed in the first paragraph of Section
+\ref{motivation}, it would be best to calculate fluxes within wide
+top-hat filters so that the difficulties of understanding the
+magnitude system were minimized.
 
@@
 \citealt{stoughton01a}). The survey has begun in earnest, and has
-obtained about 20\% of its intended data.
-
-As of February 2002, the SDSS had targeted and taken spectra of
-approximately XXX galaxies over XXX square degrees. As described
-below, only a subset of these galaxies are used here to develop the
-templates, though we calculate the SED fits for all of them. The
-results of the photometric pipeline for all of these objects were
-extracted from the SDSS Operational Database {\bf ref munn?}. The
+obtained about 20\% of its intended data.  As of February 2002, the
+SDSS had targeted and taken spectra of approximately XXX galaxies over
+XXX square degrees.
+
+As described below, only a subset of these galaxies are used here to
+develop the templates, though we calculate the SED fits for all of
+them. The results of the photometric pipeline for all of these objects
+were extracted from the SDSS Operational Database {\bf ref munn?}. The
 photometry used here for the bulk of these objects was the same as
@@
 
-The response functions $R(\lambda)$ have been measured using a
-monochrometer by {\bf how ref Doi}.  Using a model for the atmospheric
-transmission and the reflectivity of the primary and secondary
-mirrors, one can then model the response of the entire system. For
-each bandpass in the SDSS, there are six different CCDs; it has been
-shown that the differences between these six camera columns are
-small. The resulting set of filter curves is shown in Figure
-\ref{response_sdss}, in comparison to a model of a galaxy spectral
-energy distribution observed at $z=0$ (a 4 Gyr old instantaneous burst
-from the models of \citealt{bruzual93a}).
-
+The response functions of the filter-CCD combinations have been
+measured using a monochrometer by {\bf how ref Doi}.  Using a model
+for the atmospheric transmission and the reflectivity of the primary
+and secondary mirrors, one can then model the response of the entire
+system. For each bandpass in the SDSS, there are six different CCDs;
+it has been shown that the differences between these six camera
+columns are small; what differences exist are largely fit out during
+the survey calibration. The resulting set of filter curves is shown in
+Figure \ref{response_sdss}, in comparison to a model of a galaxy
+spectral energy distribution observed at $z=0$ (a 4 Gyr old
+instantaneous burst from the models of \citealt{bruzual93a}).
+
+The results of the spectroscopic observations are treated as follows.
 We extract one-dimensional spectra from the two-dimension images using
-a pipeline created specifically for the SDSS instrumentation
-(\citealt{schlegel02a}), which also fits for the redshift of each
-spectrum. The official SDSS redshifts are obtained from a different
-pipeline (\citealt{subbarao02a}). The two independent versions provide
-a consistency check on the redshift determination. They are consistent
-(for galaxies) at over the 99\% level.\footnote{The official pipeline
-tends to work better for objects with unusual spectra, such as certain
-types of stars and QSOs.}.
+a pipeline ({\tt specBS v4\_8}) created specifically for the SDSS
+instrumentation (\citealt{schlegel02a}), which also fits for the
+redshift of each spectrum. The official SDSS redshifts are obtained
+from a different pipeline (\citealt{subbarao02a}). The two independent
+versions provide a consistency check on the redshift
+determination. They are consistent (for galaxies) at over the 99\%
+level.\footnote{The official pipeline tends to work better for objects
+with unusual spectra, such as certain types of stars and QSOs.}.
 
@@
 higher signal-to-noise ``model magnitudes.'' Model magnitudes are
-calculated in all bands using a single weighted aperture. The weighted
-aperture is the better fitting model (pure exponential or pure de
-Vaucouleurs) to the galaxy image in the $r$-band. In
-\citet{stoughton01a} we are explicitly warned (in paragraphs written
-by the lead author of this paper!)  not to mix Petrosian and model
-magnitudes in a single analysis, since they measure galaxies in very
-different ways. Nevertheless, for the purposes of contraining galaxy
-template SEDs, it is perfectly acceptable to use {\it any}
-well-defined part of any galaxy.  For our purposes, it is more
-important to have high signal-to-noise measurement for any
-well-defined part of our galaxies, even if the region of the galaxies
-probed and their total estimated fluxes are inconsistent with
-redshift. We extinction-correct both types of magnitudes using the
-dust maps of \citet{schlegel98a}.
+calculated in all bands using a single weighted aperture; thus, the
+measured colors again correspond to an identifiable region of the
+galaxy, though a different one than would be measured by the Petrosian
+magnitude. The weighted aperture is the better fitting model (pure
+exponential or pure de Vaucouleurs) to the galaxy image in the
+$r$-band. In \citet{stoughton01a} we are explicitly warned (in a
+paragraph written by the lead author of this paper!)  not to mix
+Petrosian and model magnitudes in a single analysis, since they
+measure galaxies in very different ways. Nevertheless, for the
+purposes of contraining galaxy template SEDs, it is perfectly
+acceptable to use {\it any} well-defined part of any galaxy.  For our
+purposes, it is more important to have high signal-to-noise
+measurements of galaxy colors than to probe exactly the same regions
+of galaxies at low and high redshift.  We extinction-correct both
+types of magnitudes using the dust maps of \citet{schlegel98a}. {\tt
+describe fiber mags}
 
@@
 10^{10}$ years, five with metallicity $Z=0.02$ and five with
-$Z=0.004$, all assuming a Salpeter Initial Mass Function. We find that
-we cannot recover information about the SED below a certain scale,
-except perhaps around the 4000\AA\ break. Thus, we smooth most of the
-wavelength regime of the templates using a Gaussian with a standard
-deviation of $\sigma = 300$\AA; around 4000\AA, we smooth only with
-$\sigma = 150$\AA.
+$Z=0.004$, all assuming a Salpeter Initial Mass Function. Since we
+cannot recover information about the SED below a certain scale, we
+smooth most of the wavelength regime of the templates using a Gaussian
+with a standard deviation of $\sigma = 300$\AA; for the region of the
+spectrum which contains the sharpest gradients, around 4000\AA, we
+smooth only with $\sigma = 150$\AA. {\bf not clear anymore that this
+is the best thing to do; see the g/r gap discusssion}
 
@@
 The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
-joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$. In the
-bottom panel, three spectra taken from a one-dimensional sequence
-along the galay locus are shown, showing that the spectra become
-progressively bluer along that sequence. Note that constraints on the
-SED become poor at the bluest and reddest edges of this diagram (the
-peak of the $z$-band response is at 8700\AA). Therefore, the odd
-behavior near the edges should not be taken too seriously.
+joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$ for a
+random subset of about 10,000 of the galaxies in the SDSS sample (not
+only the ones which we used to fit the templates). In the bottom
+panel, three spectra taken from a one-dimensional sequence along the
+galay locus are shown, showing that the spectra become progressively
+bluer along that sequence. Note that constraints on the SED become
+poor at the bluest and reddest edges of this diagram (the peak of the
+$z$-band response is at 8700\AA). Therefore, the odd behavior near the
+edges should not be taken too seriously. In addition, note the spur
+extending from upper left to lower right through the left edge of the
+($a3/a0$)-($a1/a0$) plane. Most of the objects in this extendend spur
+are at around $z=0.3$. We discuss in Section \ref{grgap} why this spur
+arises and what to do about it.
 
@@
 for a fixed type of galaxy, the distribution of the fixed frame colors
-depends little on redshift. Because the galaxies shown in Figures
+depends only weakly on redshift. Because the galaxies shown in Figures
 \ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
@@
 Galaxies, and some are selected from the photometric redshift plates
---- we will split our sample into two well-defined sets of
-objects. First, we choose a set of Main Sample galaxies in the
-luminosity range $-21.5 < M_{\band{0.1}{r}} < -21$. We reconstruct the
-four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
+--- we will split our sample into two well-defined sets of objects to
+perform this test. First, we choose a set of Main Sample galaxies in
+the luminosity range $-21.5 < M_{\band{0.1}{r}} < -21$. We reconstruct
+the four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
 $^{0.1}(i-z)$ for these galaxies and plot the reconstructions as a
@@
 
-Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.1}{r}}
-< -22.5$. Now we reconstruct the colors to a fixed-frame corresponding
-to the rest-frame bands used to observe a galaxy at $z=0.3$.  Near
+Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.3}{r}}
+< -22.5$. For this sample we reconstruct the colors at $z=0.3$.  Near
 $z=0.3$ there is some singular behavior of the colors, which we will
-discuss in more detail later. Again, for $z<0.3$ the $^{0.3}(u-g)$
-color relies on an extrapolation, which is probably responsible for
-the scatter in that color at low redshifts. Otherwise, the
-distribution of colors is remarkably constant. $^{0.3}(g-r)$ (which is
-very similar to $^{0.0}(u-g)$) experiences a blueward shift of about
-0.1 magnitudes between $z=0.1$ and $z=0.45$, which is attributable to
-passive evolution.
+discuss in more detail in Section \ref{grgap}. Again, for $z<0.3$ the
+$^{0.3}(u-g)$ color relies on an extrapolation, which is probably
+responsible for the scatter in that color at low redshifts. Otherwise,
+the distribution of colors is remarkably constant. $^{0.3}(g-r)$
+(which is very similar to $^{0.0}(u-g)$) experiences a blueward shift
+of about 0.1 magnitudes between $z=0.1$ and $z=0.45$, which is
+attributable to passive evolution.
 
@@
 remaining problem is the aforementioned difficulties at $z=0.3$,
-discussed in more detail in the next section.
+discussed in more detail in the next subsection.
 
@@
 left to the lower right) around the red dot in the upper right panel
-of Figure \ref{k_coeffdist_plot} is due to galaxies at $z\sim
-0.3$. The difficulty arises because the 4000 \AA\ break lies in the
-gap between $g$ and $r$ at $z=0.3$. The algorithm requires strong
-gradients in the spectrum in this area to explain the 4000\AA break
---- however, the gap in the bands causes a degeneracy arise.
+of Figure \ref{k_coeffdist_plot} is due almost completely to galaxies
+at $z\sim 0.3$. The spur arises because the 4000 \AA\ break lies in
+the gap between $g$ and $r$ at $z=0.3$, visible in Figure
+\ref{response_sdss}. Our algorithm has strong gradients in the
+spectrum around the 4000\AA\ break, which is necessary to properly
+model galaxy spectra. In the presence of these sharp gradients, the
+gap in the bands causes certain combinations of the template spectra
+to be degenerate at around $z=0.3$.
 
 To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
-corresponding to the direction of the spur in the $a_1/a_0$-$a_3/a_0$
-plane of Figure \ref{k_coeffdist_plot}. We overplot the \band{0.3}{g}
-and \band{0.3}{r} bands. The spur has a strong peak in the gap between
-the two bands at $z=0.3$. Because of this gap, at $z\sim 0.3$ the
-algorithm tends to use this direction to help fit the $u$, $i$, and
-$z$ fluxes, with relatively little harm done to the $g$ and $r$
-fits. This explains why the reconstructed fluxes in Figure
+corresponding to the direction of the spur in the
+$(a_1/a_0)$-$(a_3/a_0)$ plane of Figure \ref{k_coeffdist_plot}; this
+direction is represented by the vector $-0.6 \vv{e}_1(\lambda) +
+\vv{e}_3$. We overplot the \band{0.3}{g} and \band{0.3}{r} bands. The
+spur has a strong peak in the gap between the two bands at
+$z=0.3$. Because of this gap, at $z\sim 0.3$ the algorithm tends to a
+component in this direction to help fit the $u$, $i$, and $z$ fluxes,
+with relatively little extra $\chi^2$ contribution from the $g$ and
+$r$ observations. Thus the reconstructed fluxes in Figure
 \ref{k_model_plot} fit well even at $z=0.3$, even though the spur at
@@
 here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
-reliably. However, nothing prevents one from $K$-correcting galaxies
+reliably --- their coefficients $a_k$ are unusual and probably
+unphysical. However, nothing prevents one from $K$-correcting galaxies
 at other redshifts {\it to} $z=0.3$. If you are considering the LRG
@@
 
-A second solution, not adopted here, is to place stronger constraints
+A second solution is to place extra constraints on the fit which
+prevent the spur in the coefficients. One can imagine fitting the
+distribution in the $(a_1/a_0)$-$(a_3/a_0)$ plane with a simple model,
+such as a multigaussian fit, and using the resulting fit as a prior in
+one's fit (with gaussians, the extra constraints would be linear and
+thus easily accounted for). One could design these constraints to be
+strong enough (at least near the redshifts in question) to prevent the
+spur. However, the disadvantage of this solution is that for some
+scientific projects the change in the galaxy SED is exactly what one
+is trying to measure. In these cases, it may be inappropriate to be
+putting explicit {\it a priori} constraints on exactly what one is
+trying to measure.
+
+A third solution, not adopted here, is to place stronger constraints
 in the SED fit on the form of the 4000 \AA\ break. One obvious way of
@@
 
-\section{$K$-corrections}
-\label{kcorrection}
-
-\subsection{Testing the $K$-corrections}
-
-We show, in Figure \ref{kcorrect}, the $K$-corrections to $z=0.1$
-inferred from this method, for all five SDSS bands. Note that the
-$K$-corrections are largest (and thus most uncertain) in
-$\band{0.1}{u}$ and $\band{0.1}{g}$.
+\subsection{$K$-corrections in the SDSS}
+
+We show, in Figure \ref{k_kcorrect_plot}, the resulting $K$-corrections to
+$z=0.3$ inferred from this method, for all five SDSS bands. Note that
+the $K$-corrections are largest (and thus most uncertain) in
+$\band{0.3}{u}$ and $\band{0.3}{g}$. Remember that the $K$-corrections
+to $\band{0.3}{u}$ are extrapolations for $z<0.3$ and that the
+$K$-corrections to $\band{0.3}{z}$ are extrapolations for $z>0.3$, so
+those results should not be taken too seriously. (Although the
+$K$-corrections given are not {\it too} unreasonable).
 
@@
 slope and amplitude to the fluxes in the two adjacent bandpasses
-(extrapolating when necessary). Figure \ref{ciCompare.sample8b15}
+(extrapolating when necessary). Figure \ref{comparecibreak}
 shows the differences in the $K$-corrections inferred from this method
 and those inferred from the method of Figure
-\ref{kcorrect.sample8b15}, as a function of redshift. $r$, $i$, and
+\ref{k_kcorrect_plot}, as a function of redshift. $r$, $i$, and
 $z$ are all reasonably similar in either method; $u$ and $g$, however,
-have distinct trends with redshift, mostly due to the sharper
-gradients in the spectra in this regime. To show this fact, we perform
-a similar power-law fit, only this time including a break in the
-spectrum at 4000\AA. We use the $u$-$g$ color to fit the break,
-assuming that the slope blueward of 4000\AA\ is $f(\lambda)\propto
-\lambda^{2}$.  Figure \ref{cibreakCompare.sample8b15} shows the
-results of this fit; the redshift trend in $g$ is greatly reduced, as
-is the trend in $u$, but a large amount of scatter remains in the $u$
-band. This results from the fact that you can {\it either} fit the
-slope of the SED below the 4000 \AA\ break {\it or} the size of the
-4000 \AA\ break itself; it is not possible with this data to constrain
-both in an individual spectrum, which is a fundamental limitation of
-our analysis of a fixed frame $u$-band luminosity function.
+have distinct trends with redshift, mostly due to the non-power law
+nature of the template galaxy spectra (and probably actual galaxy
+spectra) in this regime. To show this fact, we perform a similar
+power-law fit, only this time including a break in the spectrum at
+4000\AA. We use the $u$-$g$ color to fit the break, assuming that the
+slope blueward of 4000\AA\ is $f(\lambda)\propto \lambda^{2}$.  Figure
+\ref{comparecibreak} shows the results of this fit; the
+redshift trend in $g$ is greatly reduced, as is the trend in $u$, but
+a large amount of scatter remains in the $u$ band. This results from
+the fact that you can {\it either} fit the slope of the SED below the
+4000 \AA\ break {\it or} the size of the 4000 \AA\ break itself; it is
+not possible with this data to constrain both in an individual
+spectrum, which is a fundamental limitation of our determination of 
+fixed frame $u$-band magnitudes.
 
 Finally, it is possible to use the galaxy spectra obtained with the
-spectrograph to estimate the $K$-correction for each object (in
-\band{0.1}{g}, \band{0.1}{r}, and \band{0.1}{i}). However, this
-procedure requires trusting the spectrophotometry over a wide
+spectrograph to estimate the $K$-correction for each object.  However,
+this procedure requires trusting the spectrophotometry over a wide
 wavelength range. In addition, the spectra only probe the inner parts
@@
 a different region of the galaxy is probed as a function of
-redshift. Nevertheless, in Figure \ref{specK}, we compare the
+redshift. However, we can test the $K$-correction method by applying
+it to the SDSS fiber magnitudes. We again group the galaxies into Main
+Sample galaxies and LRG galaxies, and do this problem separately for
+objects $K$-corrected to $z=0.1$ and objects $K$-corrections to
+$z=0.3$, respectively.
+
+Nevertheless, in Figure \ref{k_speck_plot.fiber.0.1}, we compare the
 $K$-corrections calculated based on the spectra to those of Figure

### 2002-02-16 docs/paper/figures.tex
finishing paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@45 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 variable parameter and thus is the better separator of galaxy
-type. {\it Bottom panel}: At fixed $a_2/a_0$ and $a_3/a_0$, the
+type. The spur extending from the upper left to the lower right from
+the red dot in the $(a_3/a_0)$-$(a_1/a_0)$ plane is due to a
+degeneracy for galaxies at $z\sim 0.3$, described in detail in Section
+\ref{grgap}. {\it Bottom panel}: At fixed $a_2/a_0$ and $a_3/a_0$, the
 inferred spectra corresponding to various values of $a_3/a_0$. Near
@@
 \caption{\label{main_colors_plot} Color distributions $K$-corrected to
-$z=0.1$ for Main Sample galaxies in the luminosity range
-$-21.5<M_{\band{0.1}{r}}<-21$. This sample is complete for
-$0.05<z<0.15$. Left panels show the colors as a function of
+$z=0.1$ for SDSS Main Sample galaxies in the luminosity range
+$-21.5<M_{\band{0.1}{r}}<-21.2$. This sample is complete for
+$0.05<z<0.17$. Left panels show the colors as a function of
 redshift. Right panel shows the distributions of each color at high
-and low redshift.}
+and low redshift within the volume-limited subsample. There is very
+little dependence of the colors on redshift, even for the
+$\band{0.1}{u}$ band, where the low-redshift end is an extrapolation
+of the data.}
 \end{figure}
@@
 \ref{main_colors_plot}, now $K$-corrected to $z=0.3$ for LRG galaxies
-in the luminosity range $-22.8<M_{\band{0.3}{r}}<-22.5$. }
+in the luminosity range $-22.8<M_{\band{0.3}{r}}<-22.5$. Now
+considerably more dependence on redshift is apparent. First of all, a
+degeneracy at $z\sim 0.3$ appears (described in detail in Section
+\ref{grgap}). Second of all, in $\band{0.3}{(g-r)}$ an overall trend
+is apparent; LRGs at $z=0.4$ are about 0.1 magnitudes bluer than LRGs
+at $z=0.2$. This change is attributable to galaxy evolution.}
 \end{figure}
@@
 \plotone{spur.ps}
-\caption{\label{spur} }
+\caption{\label{spur} The spectrum corresponding to the direction of
+the spur in the upper right panel of Figure
+\ref{k_coeffdist_plot}. Note the strong feature near 4000
+\AA. Overplotted are the bandpasses for $\band{0.3}{g}$ and
+$\band{0.3}{r}$. The strong features fall in the gap between the
+bandpasses. Thus, in a linear fit to a galaxy at $z=0.3$, this
+component can be used to fit the observed magnitudes without being
+constrained to have reasonable behavior around 4000 \AA. This fact
+explains the singular behavior at $z=0.3$ in Figure
+\ref{lrg_colors_plot}.}
 \end{figure}
@@
 \plotone{k_kcorrect_plot.ps}
-\caption{\label{kcorrect.sample8b15} $K$-corrections as a function of
-redshift in all five bands for a random subsample consisting of 8,000
-of the SDSS galaxies. The $K$-corrections are largest, and therefore
-the most uncertain, for the \band{0.0}{u} and \band{0.1}{g} bands. The
-range of $K$-corrections at each redshift reflects the range of galaxy
-types at each redshift.}
+\caption{\label{k_kcorrect_plot} $K$-corrections to $z=0.3$ as a
+function of redshift in all five bands for a random subsample
+consisting around 10,000 of the SDSS galaxies.  The range of
+$K$-corrections at each redshift reflects the range of galaxy types at
+each redshift.  The $K$-corrections are largest, and therefore the
+most uncertain, for the \band{0.3}{u} and \band{0.3}{g} bands. While
+we show the $K$-corrections for \band{0.3}{u} at $z<0.3$ and for
+\band{0.3}{z} at $z>0.3$, and indeed these $K$-corrections are fairly
+well-behaved, we do not recommend using these extrapolated results for
+scientific purposes. Note the degeneracy at $z=0.3$.}
 \end{figure}
@@
 \plotone{compareci.ps}
-\caption{\label{compareci} Difference in the
-$K$-corrections in each band between the method used in Figure
-\ref{k_kcorrect_plot} and the method of simply interpolating
-between adjacent bands fitting a power-law SED. The differences are
-small in \band{0.1}{r}, \band{0.1}{i}, and \band{0.2}{z}, where galaxy
-SEDs have simple shapes. There are large differences in \band{0.0}{u}
-and \band{0.1}{g}, for which the 4000 \AA\ break is important. Note
-particularly the systematic differences in \band{0.1}{g} with
-redshift.}
+\caption{\label{compareci} Difference in the $K$-corrections to
+$z=0.1$ in each band between the method used in Figure
+\ref{k_kcorrect_plot} and the method of simply interpolating between
+adjacent bands fitting a power-law SED. The differences are small in
+\band{0.1}{r}, \band{0.1}{i}, and \band{0.1}{z}, where galaxy SEDs
+have simple shapes. There are large systematic differences in
+\band{0.1}{u} and \band{0.1}{g}, for which the 4000 \AA\ break is
+important in the spectral templates used. }
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{k_speck_plot.fiber.0.1.ps}
-\caption{\label{k_speck_plot.fiber.0.1} }
+\plotone{comparecibreak.ps}
+\caption{\label{comparecibreak} Same as Figure \ref{comparecibreak},
+now comparing to a method of interpolating the bandpasses using
+power-laws, and fitting for the 4000 \AA break. The systematic trends
+in \band{0.1}{u} and \band{0.1}{g} are gone (though there is
+considerable scatter in \band{0.1}{u}). }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{k_speck_plot.0.1.ps}
+\caption{\label{k_speck_plot.0.1} Difference between the
+$K$-corrections to $z=0.1$ determined from the spectroscopy and those
+determined from the analysis of the broad-band Petrosian magnitudes, for
+Main Sample galaxies.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{k_speck_plot.fitfib.0.1.ps}
+\caption{\label{k_speck_plot.fitfib.0.1} Difference between the
+$K$-corrections to $z=0.1$ determined from the spectroscopy and those
+determined from the analysis of broad-band magnitudes {\it
+synthesized} from the same spectra.}
 \end{figure}
@@
 \plotone{k_speck_plot.0.3.ps}
-\caption{\label{k_speck_plot.0.3} }
+\caption{\label{k_speck_plot.0.3} Difference between the
+$K$-corrections to $z=0.3$ determined from the spectroscopy and those
+determined from the analysis of the broad-band model magnitudes, for
+LRGs.}
 \end{figure}

### 2002-02-16 docs/paper/paper.tex
finishing paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@45 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS
-Spectroscopic Survey}
+Spectroscopic Survey$^1$}
 
@@
 Daniel Eisenstein\altaffilmark{\ref{Arizona}},
-James E. Gunn\altaffilmark{\ref{Princeton}}, and
-David W. Hogg\altaffilmark{\ref{NYU}}
+James E. Gunn\altaffilmark{\ref{Princeton}}
+David W. Hogg\altaffilmark{\ref{NYU}}, and
 David J. Schlegel\altaffilmark{\ref{Princeton}}
@@
 %\label{Pitt}}
-%\altaffiltext{\theaddress}{
-%\stepcounter{address}
-%rinceton University Observatory, Princeton,
-%NJ 08544
-%\label{Princeton}}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+Department of Astronomy and Research Center for 
+the Early Universe,
+School of Science, University of Tokyo,
+Tokyo 113-0033, Japan
+\label{Tokyo}}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+Steward Observatory, 
+933 N. Cherry Ave., Tucson, AZ
+85721
+\label{Arizona}}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+Princeton University Observatory, Princeton,
+NJ 08544
+\label{Princeton}}
 %\addtocounter{address}{1}
@@
 %\label{Pyrenees}}
-%\altaffiltext{\theaddress}{
-%\stepcounter{address}
-%Department of Astronomy and Research Center for 
-%the Early Universe,
-%School of Science, University of Tokyo,
-%Tokyo 113-0033, Japan
-%\label{Tokyo}}
 %\altaffiltext{\theaddress}{
@@
 Section \ref{sedfit} describes our method of fitting galaxy SEDs to
-broad-band photometry. Section \ref{data} applies the method to
-galaxies in the SDSS, showing that the fits are robust. Section
-\ref{kcorrection} shows how one derives $K$-corrections from the
-results, and discusses how best to use the results of such
-fits. Section \ref{conclusions} concludes and discusses future
-development of the method described here.
+broad-band photometry and how to calculate $K$-corrections. Section
+\ref{data} applies the method to galaxies in the SDSS, showing that
+the fits are robust.  Section \ref{conclusions} concludes and
+discusses future development of the method described here.
 
@@
 cosmological distance modulus (that is, the luminosity
-distance-squared law, as tabulated by, {\it e.g.}  \citealt{hogg97a}).
+distance-squared law, as tabulated by, {\it e.g.}  \citealt{hogg99a}).
 
@@
 magnitudes just do not uniquely determine the SED) they should be
-avoid or minimized when possible. For example, if one was calculating
+avoided or minimized when possible. For example, if one was calculating
 the evolution of clustering of red and blue galaxies separately, it
@@
 of galaxies at low and high redshift.  We extinction-correct both
-types of magnitudes using the dust maps of \citet{schlegel98a}. {\tt
-describe fiber mags}
+types of magnitudes using the dust maps of \citet{schlegel98a}.
 
@@
 
+In addition, the photometric errors for most objects in the
+spectroscopic galaxy sample of the SDSS are not dominated by the
+estimated errors on the photometry listed in the catalog. Instead, the
+errors are dominated by local calibration errors and other systematic
+effects, which are poorly known. To account for these errors, we add
+extra error terms in quadrature with the errors listed in the catalog
+(0.05 mags, 0.02 mags 0.02 mags, 0.02 mags, and 0.03 mags for $ugriz$,
+respectively). Not accounting for this extra source of errors can
+cause the fits to be ill-behaved.
+
 For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
@@
 unphysical trends of galaxy SED versus redshift ({\it cf.}  Figure
-\ref{constantz} below). Using three templates does nearly as well as
+\ref{main_colors_plot} below). Using three templates does nearly as well as
 four templates in the sense that the resulting templates reproduce the
@@
 
-\section{The $g$/$r$ Gap}
+\subsection{The $g$/$r$ Gap}
 \label{grgap}
 
-It is clear from Figure \ref{k_kcorrect_plot} that $z=0.3$ is a
+It is clear from Figure \ref{lrg_colors_plot} that $z=0.3$ is a
 special redshift regime where our fits become unphysical. In fact, it
@@
 this procedure requires trusting the spectrophotometry over a wide
-wavelength range. In addition, the spectra only probe the inner parts
-of galaxies, and because the fiber aperture is fixed in angular space,
-a different region of the galaxy is probed as a function of
-redshift. However, we can test the $K$-correction method by applying
-it to the SDSS fiber magnitudes. We again group the galaxies into Main
-Sample galaxies and LRG galaxies, and do this problem separately for
-objects $K$-corrected to $z=0.1$ and objects $K$-corrections to
-$z=0.3$, respectively.
-
-Nevertheless, in Figure \ref{k_speck_plot.fiber.0.1}, we compare the
-$K$-corrections calculated based on the spectra to those of Figure
-\ref{kcorrect}, finding that they are extremely similar. The fact that
-they are so similar might prove useful, in the sense that one could
-combine the photometric and spectroscopic results, using the spectra
-to constrain small-scale features such as the 4000 \AA\ break and the
-photometry to cover a large wavelength range and to probe the full
-extent of the galaxy. Such a procedure might resolve the difficulties
-in handling the $u$-band discussed in the previous paragraph. We will
-not pursue this issue further in this paper and leave it for future
-investigation.
+wavelength range. In addition, the region of the galaxy which the
+spectra probe is not completely well defined. The small scale
+wavelength features are set by the fiber aperture ($3''$ in diameter)
+which large scale wavelength features are constrained by a short
+``smear exposure'' which probes a larger area. The smear exposure
+corrections are clearly the correct approach for point sources, but
+for galaxies it does cause the small scale and large scale features of
+the spectrum to be determined by somewhat different regions of the
+galaxy.
+
+Nevertheless, in Figure \ref{k_speck_plot.0.1}, we compare the
+$K$-corrections to $z=0.1$ of Main Sample galaxies calculated based on
+the spectra to those of Figure \ref{k_kcorrect_plot}, finding that
+they are extremely similar. There is a trend within redshift in the
+$g$-band, which is due to the fact that on average the $g-r$ color is
+about 0.1 magnitudes redder in the fiber aperture than in the
+Petrosian aperture. The implication is that the $K$-corrections {\it
+are} stronger for the fiber aperture than for the Petrosian aperture.
+Correspondingly we repeat the exercise using the SDSS fiber magnitudes
+to determine the $K$-corrections, we find consistent results for the
+$\band{0.1}{g}$-band.
+
+Skeptics may worry that even if we {\it did} have photometry that
+corresponded precisely to the same region of the galaxy that the
+spectroscopy did, we still might be calculating the $g$-band
+$K$-corrections in some biased way. However, we can test whether this
+is the case by synthesizing the observed broad-band photometry from
+the spectra, and trying to recover the $K$-corrections in a case where
+we know the spectrum completely. Since we cannot synthesize the
+observed $u$ and $z$ bands, we simply base them on the synthesized $g$
+and $i$ band magnitudes plus the $u-g$ and $z-i$ colors. (This
+procedure is unlikely to make the photometric estimations of the
+$K$-corrections {\it better}). Finally, we added 2\% random errors to
+all of the measurements. Figure \ref{k_speck_plot.fitfib.0.1} compares
+the $K$-corrections from the synthesized photometry to that calculated
+from the spectroscopy itself. The agreement is now very good,
+suggesting that our method can indeed recover the correct
+$K$-corrections based only on broad-band magnitudes (and furthermore
+that for the $g$-band one should not rely on $K$-corrections based on
+the spectroscopy alone).
+
+For completeness, Figure \ref{k_speck_plot.0.3} compares the
+spectroscopic and photometric $K$-corrections to $z=0.3$ of LRGs. The
+\band{0.3}{r} and \band{0.3}{i} bands agree very well. However, there
+is considerable scatter in the \band{0.3}{g} band (nearly equivalent
+to the \band{0.0}{u} band), at around 20\%. Because the LRGs tend to
+have small color gradients, the issues of fiber sizes and smear
+exposures are not important in this context.
 
@@
 
-Some of the scatter in the three-dimension space describing the shape
-of the galaxy SED is probably due to dust. It may be possible to
-evaluate the effects of reddening in this space and, by assuming that
-galaxies corrected for internal reddening live in an even lower
-dimensional space, perform a reddening correction. We will be
-investigating this question in the near future.
-
-Furthermore, one may want to include the effects of evolution in the
-$K$-corrections. One approach to this calculation is to define a
-one-dimensional locus, and record which position on this locus is
-closest to each galaxy. Then, the one-dimensional locus provides a set
-of spectra, for which estimates of evolution can be calculated. This
-procedure avoids fitting star-formation histories to a huge number of
-different positions in galaxy SED space. Since their will be
-degeneracies between age and metallicity in these fits in any case, it
-may be best to simplify the problem to one dimension.
+We have not discussed in detail the question of error analysis. There
+is no answer to this question which is simple, general, {\it and}
+realistic. The photometric errors can of course be propagated to the
+errors on the reconstructed magnitudes, but these estimates do not
+describe the errors associated with the restriction to a
+three-dimensional space of SED shapes. {\bf maybe we should implement
+this anyway, though}. An idea of the level of errors can be gleaned
+from the comparison of our method to other methods of
+$K$-correction. Clearly, $r$, $i$ and $z$ can be reconstructed to the
+level of the photometry. Judging from Figure
+\ref{k_speck_plot.fitfib.0.1}, the $g$ band can be reconstructed to
+about 5\% accuracy. The $u$ band is more difficult to evaluate.
+Judging from the poor reconstructions of the $\band{0.3}{g}$ band
+(nearly equal to the $\band{0.0}{u}$ band, the reconstructions may be
+as bad as 20\% accuracy.
+
+Topics we have not discussed here are dust- and evolution-correction
+of the magnitudes. Some of the scatter in the three-dimensional space
+describing the shape of the galaxy SED is probably due to dust. It may
+be possible to evaluate the effects of reddening in this space and, by
+assuming that galaxies corrected for internal reddening live in an
+even lower dimensional space, perform a reddening correction. We will
+be investigating this question in the near future. One can perform a
+similar exercise with evolution estimates, relating the position of a
+galaxy along the $a_1$ axis in Figure \ref{k_coeffdist_plot} to a
+particular star-formation history, and estimating the evolution of the
+object from that history.
 
 The distribution of galaxies in the three-dimensional space of galaxy
-SEDs described here may be useful for other purposes, as well. As one
-example, which has already been explored in a separate project, one
-can calculate photometric redshifts (if you are willing to restrict
-your galaxy models to a one-dimensional, though possibly curved, locus
-within the three-dimensional space). However, there is a fairly severe
-redshift-type degeneracy, in the sense that blue galaxies at high
-redshift can be confused with red galaxies at lower redshift. These
-problems can be hazardous to someone trying to estimate the evolution
-of the luminosity function from a photometric redshift sample (such
-studies rarely include these correlated errors in their estimate of
-their uncertainties). However, one can make an alternative set of
-assumptions by modeling the distribution of the galaxies in the
-three-dimensional spectral space as a pair of gaussians. For such a
-model, and a model for the position and luminosity evolution of each
-gaussian, one can calculate the probability of observing any given
-galaxy. Thus, one could use the galaxy photometry to constrain the
-luminosity evolution of galaxies in each gaussian, as well as the
-evolution of the position of the gaussians. This method has its own
-systematic errors and its own restrictive set of assumptions, but
-provides an alternative path to calculating the luminosity function
-directly from photometric redshifts.
+SEDs described here may be useful for other purposes, as well. For
+example, you can calculate photometric redshifts using the method
+described here. {\tt kcorrect v1\_2} has a simple (and very fast)
+photometric redshift estimator with very tight core of residuals
+($\sigma \sim 0.05$) in a comparison to SDSS galaxies with redshifts.
+Before this code can be used as a reliable photometric redshift
+indicator, more work has to be done (which the reader is invited to do
+using the released code) to identify galaxies which are likely to be
+redshift outliers.
 
@@
 
-\include{tables}
+%\include{tables}
 

### 2002-03-08 docs/paper/paper.tex
altering docs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@87 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 them. The results of the photometric pipeline for all of these objects
-were extracted from the SDSS Operational Database {\bf ref munn?}. The
+were extracted from the SDSS Operational Database. The
 photometry used here for the bulk of these objects was the same as
@@
 The response functions of the filter-CCD combinations have been
-measured using a monochrometer by {\bf how ref Doi}.  Using a model
+measured using a monochrometer by Mamoru Doi.  Using a model
 for the atmospheric transmission and the reflectivity of the primary
@@
 (totalling about 1,000 objects) which were selected by a photometric
-redshift algorithm ({\bf ref who?}) to be at around $z\sim
+redshift algorithm ({\bf ref what?}) to be at around $z\sim
 0.3$--$0.4$. These objects are invaluable for tying down the blue end
@@
 extra error terms in quadrature with the errors listed in the catalog
-(0.05 mags, 0.02 mags 0.02 mags, 0.02 mags, and 0.03 mags for $ugriz$,
-respectively). Not accounting for this extra source of errors can
-cause the fits to be ill-behaved.
+(0.05 mags, 0.02 mags, 0.02 mags, 0.02 mags, and 0.03 mags for
+$ugriz$, respectively). Not accounting for this extra source of errors
+can cause the fits to be ill-behaved.
 
@@
 spectrum which contains the sharpest gradients, around 4000\AA, we
-smooth only with $\sigma = 150$\AA. {\bf not clear anymore that this
-is the best thing to do; see the g/r gap discusssion}
+smooth only with $\sigma = 150$\AA.
 
@@
 describe the errors associated with the restriction to a
-three-dimensional space of SED shapes. {\bf maybe we should implement
-this anyway, though}. An idea of the level of errors can be gleaned
-from the comparison of our method to other methods of
+three-dimensional space of SED shapes. An idea of the level of errors
+can be gleaned from the comparison of our method to other methods of
 $K$-correction. Clearly, $r$, $i$ and $z$ can be reconstructed to the

### 2002-04-14 docs/paper/paper.tex
changed
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@92 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
-\newcommand{\band}[2]{\ensuremath{^{#1}{#2}}}
+\newcommand{\band}[2]{\ensuremath{^{#1}\!{#2}}}
 
@@
 
+\subsection{The $g$/$r$ Gap}
+\label{grgap}
+
+In Figure \ref{response_sdss}, there is clearly a gap between the $g$
+and $r$ bands. Linear fits to the data allow considerable freedom in
+the resulting reconstructed SEDs. As it happens, one of the directions
+in our best fit coefficient space $a_i/a_0$ corresponds to a large
+spike at around 4000 \AA; that this direction is important is not
+surprising, since one of the most variable qualities of galaxy SEDs is
+the size of the 4000 \AA\ break. However, the gap between the $g$ and
+$r$ bands corresponds to the 4000 \AA\ break at around $z=0.3$. The
+existence of this spike in one direction in our four-dimensional space
+means that galaxies at $z=0.3$ can vary along this direction in order
+to better fit $u$, $i$, and $z$, with little effect on the $g-r$ color
+of the fit. This results in unphysical SED fits to galaxies near
+$z=0.3$.
+
+To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
+corresponding to the direction represented by the vector $-0.6
+\vv{e}_1(\lambda) + \vv{e}_3$. We overplot the \band{0.3}{g} and
+\band{0.3}{r} bands. This direction has a strong peak in the gap
+between the two bands at $z=0.3$.  This difficulty is why, in Section
+\ref{data}, we exclude the regime around $z\sim 0.3$; otherwise the
+templates became distracted by the degeneracies in this range.
+
+We have explored using a larger set of input spectrum, by including
+reddenned versions of all our templates, to test whether a more
+complete space of SEDs would cause our fit to choose a more realistic
+subspace. In addition, we have tried removing the increased resolution
+of the template SEDs around the 4000 \AA break. Neither of these tests
+yielded better results.
+
+Our view is that the ideal solution to this problem would be to
+minimize $\chi^2$ under the constraint that none of the original
+Bruzual-Charlot spectra defining our SED space contribute negatively
+to the SED fit. Such a solution would almost certainly have better
+properties than our linear approach here: it would almost certainly
+avoid the degeneracies, it would yield reasonable fit SED over a
+larger wavelength range, it would associate a reasonable
+star-formation history with each object (which could be used to
+evolution-correct the magnitudes).
+
+However, we have not tackled this approach yet. We have implemented
+(and used for the figures shown here) a temporary fix for those
+interested in galaxies at $z=0.3$, which is to artificially constrains
+the fit as follows. We pretend that there is an extra band between $g$
+and $r$, which has a FWHM of 2000 \AA\ and a center of 5500 \AA. The
+``measured'' value we use in this band is the mean of the $g$ and $r$
+fluxes, and is given $4\sigma$ significance. This extra constraint
+changes the K-corrections for galaxies outside the range $0.27 < z <
+0.33$ by far less than 1 percent. The constraint helps to clean up the
+fits but it is not sufficient; the coefficients are still clearly
+wrong around $z=0.3$. It helps considerably to $K$-correct to $z=0.3$,
+which we recommend for the moment for studies of LRGs; this minimizes
+the corrections which are performed using our most unphysical fits.
+
+Remember, you really only need to worry about this effect in the SDSS
+when you are dealing with the LRGs, for galaxies in the range about
+0.27 < z < 0.33. The Main Sample results are not significantly
+affected at all.  Furthermore, one can still observe a galaxy at
+$z=0.1$ and reliably infer what it would look like at $z=0.3$; it is
+only the reverse process which is difficult.
+
+%It is clear from Figure \ref{lrg_colors_plot} that $z=0.3$ is a
+%special redshift regime where our fits become unphysical. In fact, it
+%is easy to demonstrate that the diagonal spur (going from the upper
+%left to the lower right) around the red dot in the upper right panel
+%of Figure \ref{k_coeffdist_plot} is due almost completely to galaxies
+%at $z\sim 0.3$. The spur arises because the 4000 \AA\ break lies in
+%the gap between $g$ and $r$ at $z=0.3$, visible in Figure
+%\ref{response_sdss}. Our algorithm has strong gradients in the
+%spectrum around the 4000\AA\ break, which is necessary to properly
+%model galaxy spectra. In the presence of these sharp gradients, the
+%gap in the bands causes certain combinations of the template spectra
+%to be degenerate at around $z=0.3$.
+
+%The question arises of what to do about this problem. One solution is
+%to simply work around it. It is clear we cannot use the fits presented
+%here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
+%reliably --- their coefficients $a_k$ are unusual and probably
+%unphysical. However, nothing prevents one from $K$-correcting galaxies
+%at other redshifts {\it to} $z=0.3$. If you are considering the LRG
+%sample, the median redshift is close to $z=0.3$ anyway, so this is a
+%viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).
+
+%A second solution is to place extra constraints on the fit which
+%prevent the spur in the coefficients. One can imagine fitting the
+%distribution in the $(a_1/a_0)$-$(a_3/a_0)$ plane with a simple model,
+%such as a multigaussian fit, and using the resulting fit as a prior in
+%one's fit (with gaussians, the extra constraints would be linear and
+%thus easily accounted for). One could design these constraints to be
+%strong enough (at least near the redshifts in question) to prevent the
+%spur. However, the disadvantage of this solution is that for some
+%scientific projects the change in the galaxy SED is exactly what one
+%is trying to measure. In these cases, it may be inappropriate to be
+%putting explicit {\it a priori} constraints on exactly what one is
+%trying to measure.
+
+%A third solution, not adopted here, is to place stronger constraints
+%in the SED fit on the form of the 4000 \AA\ break. One obvious way of
+%doing this is to use the spectra to constrain the size of the break
+%(making it a weak constraint to prevent biases). This constraint could
+%be applied in a linear fashion to retain the speed of our method ---
+%except for the fact that you would have to read in the spectra to
+%apply the $K$-correction. For simplicity, and because the aforemention
+%work-around is so easy to obey, we have not implemented this solution
+%here.
+
 \subsection{Results}

### 2002-04-14 include/kcorrect.h

@@ -24,6 +24,10 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 											double *galaxy_maggies, /* galaxy maggies */
 											double *galaxy_invvar,
 											double *galaxy_z,
+											double *constraints_amp,
+											double *constraints_mean,
+											double *constraints_invvar,
+											int nconstraints,
 											IDL_LONG ngalaxy);
 
 /* finds best fit eigentemplates, given the galaxy properties and 

### 2002-04-14 src/fit_coeffs.c

@@ -120,8 +120,9 @@ int main(int argc,
 			fscanf(stdin,"%lf",&(galaxy_z[i+1]));
 		} /* end for i */
 		ncurrchunk=i;
+		/* no direct constraints on the coeffs are included in this fit */
 		k_fit_coeffs(ematrix,nt,zvals,nz,rmatrix,nk,nb,coeffs,galaxy_maggies,
-								 galaxy_invvar,galaxy_z,ncurrchunk);
+								 galaxy_invvar,galaxy_z,NULL,NULL,NULL,0,ncurrchunk);
 		for(i=0;i<ncurrchunk;i++) {
 			for(j=0;j<nt;j++)
 				fprintf(stdout,"%e ",coeffs[i*nt+j]);

### 2002-04-14 src/idl_k_fit_coeffs.c

@@ -13,9 +13,10 @@ IDL_LONG idl_k_fit_coeffs
   (int      argc,
    void *   argv[])
 {
-   IDL_LONG nt,nz,nk,nb,ngalaxy;
+   IDL_LONG nt,nz,nk,nb,ngalaxy,nconstraints;
    double *ematrix,*zvals,*rmatrix,*amatrix;
 	 double *galaxy_maggies,*galaxy_invvar,*galaxy_z;
+	 double *constraints_amp,*constraints_mean,*constraints_invvar;
 
 	 IDL_LONG i;
 	 IDL_LONG retval=1;
@@ -33,13 +34,18 @@ IDL_LONG idl_k_fit_coeffs
 	 galaxy_maggies=(double *)argv[i]; i++;
 	 galaxy_invvar=(double *)argv[i]; i++;
 	 galaxy_z=(double *)argv[i]; i++;
+	 constraints_amp=(double *)argv[i]; i++;
+	 constraints_mean=(double *)argv[i]; i++;
+	 constraints_invvar=(double *)argv[i]; i++;
+	 nconstraints=*((IDL_LONG *)argv[i]); i++;
    ngalaxy=*((IDL_LONG *)argv[i]); i++;
 
 	 /* 1. run the fitting routine */
 	 retval=(IDL_LONG) k_fit_coeffs(ematrix,nt,zvals,nz,rmatrix,nk,nb,
 																	amatrix,galaxy_maggies,galaxy_invvar,
-																	galaxy_z,ngalaxy);
-
+																	galaxy_z,constraints_amp,constraints_mean,
+																	constraints_invvar,nconstraints,ngalaxy);
+	 
 	 /* 2. free memory and leave */
 	 free_memory();
    return retval;

### 2002-04-14 src/k_fit_coeffs.c

@@ -34,9 +34,13 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 																							redshifts */
 											double *galaxy_invvar,
 											double *galaxy_z,
+											double *constraints_amp,
+											double *constraints_mean,
+											double *constraints_invvar,
+											int nconstraints,
 											IDL_LONG ngalaxy)
 {
-	double currz;
+	double currz,a0scale;
 	char uplo;
 	IDL_LONG i,j,jp,k,b,indx,unity,info;
 	
@@ -76,7 +80,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 	rhs=(double *) malloc(nt*sizeof(double));
 	for(i=0;i<ngalaxy;i++) {
 
-		/* 2b. Create A */
+		/* 2b. Create A based on data */
 		for(j=0;j<nt;j++)
 			for(jp=0;jp<nt;jp++) {
 				covar[j*nt+jp]=0.;
@@ -84,7 +88,7 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 					covar[j*nt+jp]+=cmatrix[i*nk*nt+k*nt+j]
 						*cmatrix[i*nk*nt+k*nt+jp]*galaxy_invvar[k+i*nk];
 			} /* end for j jp */
-																		
+
 		/* 2c. Create b */
 		for(j=0;j<nt;j++) {
 			rhs[j]=0.;
@@ -92,6 +96,28 @@ IDL_LONG k_fit_coeffs(double *ematrix,    /* eigentemplates */
 				rhs[j]+=galaxy_maggies[k+i*nk]*cmatrix[i*nk*nt+k*nt+j]
 						*galaxy_invvar[k+i*nk];
 		} /* end for j */
+		
+		/* 2d. Add constraints to A and b */
+		if(nconstraints>0) {
+			/* first find appropriate a0 */
+			a0scale=rhs[0]/covar[0];
+			
+			/* then add each constraint */
+			for(k=0;k<nconstraints;k++) {
+				for(j=1;j<nt;j++)
+					for(jp=1;jp<nt;jp++) 
+						covar[j*nt+jp]+=constraints_amp[k]* 
+							constraints_invvar[k*(nt-1)*(nt-1)+(j-1)*(nt-1)+(jp-1)]/
+							(a0scale*a0scale);
[…]

### 2002-04-14 src/k_fit_photoz.c

@@ -49,10 +49,10 @@ double pz_fit_coeffs(double z)
 	double chi2,az,bz,cz;
 	IDL_LONG i,k;
 	
-	/* fit coeffs */
+	/* fit coeffs; no direct constraints are put on the coeffs in this case */
 	k_fit_coeffs(pz_ematrix,pz_nt,pz_zvals,pz_nz,pz_rmatrix,
 							 pz_nk,pz_nb,pz_coeffs,pz_galaxy_maggies,pz_galaxy_invvar,
-							 &z,1);
+							 &z,NULL,NULL,NULL,0,1);
 	
 	chi2=pz_calc_chi2(z);
 	return(chi2);

### 2002-04-15 docs/paper/paper.tex
final figs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@95 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-\slugcomment{Submitted to \aj}
+\slugcomment{To be submitted to \aj}
 
@@
 
-\title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS
-Spectroscopic Survey$^1$}
+\title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS$^1$}
 
@@
 \begin{abstract}
-Broad-band measurements of galaxy fluxes suffer from the difficulty
-that for galaxies at different redshifts the observations probe
-different regions of the rest-frame galaxy spectrum. Certain
+Broad-band measurements of flux for galaxies at different redshifts
+measure different regions of the rest-frame galaxy spectrum. Certain
 astronomical questions, such as the evolution of the luminosity
@@
 and in fact nearly linear. The resulting reconstructed SEDs can be
-used to recover fixed-frame magnitudes over a range of redshifts. We
-show that the results of doing so yield consistent results, in the
-sense that within each sample the intrinsic colors of similar type
-galaxies are nearly constant with redshift. We note one remaining
-problem associated with the gap between the $g$ and $r$ bands. We
-compare our results to simpler interpolation methods and galaxy
-spectrophotometry.  The software which generates these results is
-publicly available and easily adapted to handle galaxy observations in
-any AB magnitude system in the rest-frame optical regime.
+used to recover fixed-frame magnitudes over a range of redshifts. This
+process yields consistent results, in the sense that within each
+sample the intrinsic colors of similar type galaxies are nearly
+constant with redshift. We compare our results to simpler
+interpolation methods and galaxy spectrophotometry.  The software that
+generates these results is publicly available and easily adapted to
+handle a wide range of galaxy observations.
 \end{abstract}
@@
 
-In the future, we expect that observations of galaxies (and indeed of
-any astronomical sources) will be performed using integral field
+In the future, observations of galaxies (and indeed of any
+astronomical sources) will be performed using integral field
 spectrographs or equivalent devices which combine high spatial
@@
 energy distribution of the object observed.  This paper focuses on a
-method of facing these problems which (in the optical wavelength
-regime) is fast, robust, and consistent over a large range of
-redshifts.
-
-In the past, people have in general accounted for these effects using
+solution to these problems which (in the optical wavelength regime) is
+fast, robust, and consistent over a large range of redshifts.
+
+In the past, people have accounted for these effects using
 ``$K$-corrections'' applied to the observed magnitudes. In these
@@
 reflectivity of the mirrors, the filter transmission, and the response
-of the CCD device or photographic emulsion).  However, galaxies do not
-all have the same SED, nor are they selected from some discrete set of
-SEDs. For this reason, these schemes for applying the $K$-corrections
-can fail to be self-consistent: the SED assumed for the
-$K$-corrections can be significantly inconsistent with the observed
-galaxy colors! As we demand more precision from our astronomical data
-analysis in new, large multi-band surveys, and in particular as we try
-to quantify the evolution of galaxies, we must take a more
-sophisticated approach to approximating fixed-frame observations of
-galaxies.
+of the CCD device or photographic emulsion). For example, for the
+Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
+\citet{fukugita95a} have both presented results for the
+$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
+(called $k(z)$ by \citealt{frei94a}) for galaxies of various
+morphological types, based on galaxy spectrophotometry.\footnote{Note
+that the description on p. 57 of \citet{binney98a} of the meaning of
+$k(z)$ in \citet{frie94a} is incorrect; in fact, $k(z)\equiv K(z)$.}
+
+However, galaxies do not all have the same SED, nor are they selected
+from some discrete set of SEDs. For this reason, these schemes for
+applying the $K$-corrections can fail to be self-consistent: the SED
+assumed for the $K$-corrections can be significantly inconsistent with
+the observed galaxy colors! As we demand more precision from our
+astronomical data analysis in new, large multi-band surveys, and in
+particular as we try to quantify the evolution of galaxies, we must
+take a more sophisticated approach to approximating fixed-frame
+observations of galaxies.
 
@@
 magnitude in any bandpass. Although the operation we are performing on
-the magnitudes is not strictly speaking a ``$K$-correction,'' we will
-refer to it as such in this and subseqent papers. Our approach is
-equivalent (nearly identical) to the photometric metric redshift
-estimation methods of \citet{csabai00a} and \citet{budavari00a},
-except we use slightly different coordinate systems. In fact, the
-software includes a fast and relatively accurate photometric redshift
-estimator based on their method.
-
-We implemented this system and are publishing it in order to prepare
-the reader for upcoming papers which will rely heavily on the
-reliability of the fixed-frame magnitudes determined here.  In
-addition, having implemented the fits, there is no reason not to share
-them with the astronomical community; thus, this paper serves the
-secondary purpose of describing the release of a piece of
+the magnitudes is not a ``$K$-correction'' in the historical sense of
+the term, we will refer to it as such in this and subsequent
+papers. Our approach is equivalent (nearly identical) to the
+photometric metric redshift estimation methods of \citet{csabai00a},
+except we use slightly different coordinate systems for our SED
+template space. In fact, the software includes a fast and relatively
+accurate photometric redshift estimator based on their method.
+
+We implemented this system and are publishing it in order to lay the
+groundwork for upcoming papers which will rely heavily on the
+reliability of the fixed-frame magnitudes determined here.  Because
+the SDSS is one of the largest astronomical surveys to date, our
+method will be useful to a large number of other investigators, both
+inside and outside the SDSS collaboration.  Thus, this paper also
+serves the purpose of describing the release of a piece of
 software. The computer software we distribute builds itself into a C
@@
 programs and IDL routines. We have made the source code available
-publicly, through a URL on the World Wide Web and through a public CVS
+publicly, through a web page and through a public CVS
 repository. Improvements or ports to other languages implemented by
@@
 specified in any such paper. The version of the code used to make the
-figures for this paper is {\tt kcorrect v1\_2}.
+figures for this paper is {\tt kcorrect v1\_9}.
+
+We refer throughout this paper to $AB$ magnitudes, meaning that the
+magnitudes are designed such that ():
+\begin{eqnarray}
+\label{ABdef}
+m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
+\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
+{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
+\cr
+&=& -48.60 - 2.5 \log_{10}\left[
+\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
+{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
+\end{eqnarray}
+where $R(\lambda)$ is the fraction of photons entering the Earth's
+atmosphere which are detected as a function of wavelength (a unitless
+quantity). In this equation, $f(\lambda)$ is in units of ergs
+cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
+cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
+that an object with $f(\nu) = 3631\mathrm{~Jy} = 3.631 \times
+10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes
+equal to zero.  The difference in the zeropoint in the two equations
+simply corresponds to the factor of the speed of light $c$ (expressed
+in \AA\ s$^{-1}$) between the frequency and wavelength scales.
 
@@
 are known to have significant structure over wavelength ranges small
-compared to our bandpass, this task is an ill-posed, inverse
-problem. However, we are not completely ignorant about the forms which
-galaxy SEDs take, so we can attempt to use what we know about galaxy
-SEDs to simplify our task.  The method described here for doing so
-follows closely the methods of estimating photometric redshifts used
-by \citet{csabai00a} and \citet{budavari00a}. It is designed to take
-advantage of what we already understand about galaxy spectral energy
-distributions.
+compared to our bandpass, this task is an ill-posed, inverse problem:
+given any set of observed galaxy colors, there is not a unique SED to
+which those colors correspond, even in the absence of errors. However,
+we are not completely ignorant about the forms which galaxy SEDs take,
+so we can attempt to use what we know about galaxy SEDs to simplify
+our task.  The method described here for doing so follows closely the
+methods of estimating photometric redshifts used by \citet{csabai00a}
+and \citet{budavari00a}. It is designed to take advantage of what we
+already understand about galaxy spectral energy distributions.
 %In addition, as implemented here, the method is
@@
 
-Let us begin with an SED space defined by $N_b$ template galaxy SEDs
+Begin with an SED space defined by $N_b$ template galaxy SEDs
 $\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
@@
 axes $\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
-constant flux; in this manner, the coefficient $a_0$ in Equation
+constant $F_t$; in this manner, the coefficient $a_0$ in Equation
 (\ref{lincomb}) is directly proportional to $F_t$ and is thus purely a
@@
 approximately in an ellipsoid, with not much curvature within the
-space defined by the template SEDs. Thus, we rotate the axes
+space defined by the template SEDs. Thus, we transform the axes such
+that $a_i/a0 = 0$ is the mean of the distribution and rotate the axes
 $\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
@@
 $\vv{e}_i$ such that the zeroth component $a_0=F_t$ is output by the
-code in units of ergs cm$^{-2}$ s$^{-1}$. This choice of coordinate
-system is arbitrary and does not affect our quantitative results, but
-it is convenient.
+code in units of ergs cm$^{-2}$ s$^{-1}$ (within the range
+$\ambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$).
 
@@
 determined during the procedure can, of course, be used to estimate
-photometric redshifts, as \citet{csabai00a} and \citet{budavari00a}
-describe.
+photometric redshifts, as \citet{csabai00a} describes.
+
+\subsection{Applying Direct Constraints on $a_i/a0$}
+\label{direct}
+
+Sometimes we will want to apply direct constraints on the coefficients
+$a_i/a_0$ of the galaxy we are fitting. For example, in this paper we
+fit our templates to a set of galaxies with high quality photometry in
+all bands. However, we would like to use these templates to fit for
+the SEDs of other galaxies, which might have much worse photometry in
+one or more bands. Since the errors in the photometry can be
+catastrophic (that is, non-gaussian), when we fit the SEDs of the
+lower signal-to-noise galaxies it makes sense to at least weakly
+constrain these galaxies to have reasonable SEDs, in the sense that
+the $a_i/a_0$ ought not to be too different than for well-measured
+galaxies. Here we describe how to implement direct constraints on the
+coefficients $a_i/a_0$ which describe the galaxy SED.
+
+We will assume that any constraints can be expressed as a combination
+of gaussian constraints with means $a_{c,j}/a_{c,0}$, variance matrices
+$V_{c,jl}$, weighted by amplitudes $A_c$.  We calculate a
+preliminary value of $a_{c,0}$, assuming $a_i = 0$. Then we constrain
+our fit by adding a term to $\chi^2$ for each constraint:
+\begin{equation}
+\sum_j \sum_l V^{-1}_{jl} (a_l/a_{c,0}-a_{c,l}/a_{c,0})
+(a_j/a_{c,0}-a_{c,j}/a_{c,0})  
+\end{equation}
+The resulting $\chi^2$ can still be minimized linearly, so we have
+sacrificed very little in speed.
+
+Note that by using this constraint, one can calculate $K$-corrections
+even for galaxies with only one or two bands measured, and still
+obtain sensible results.
 
@@
 \end{equation}
-read, ``$b$-band at redshift $z$.'' The goal of ``$K$-corrections'' is
-to convert the $ugriz$ bands at redshift $z$ for each galaxy to
-$ugriz$ bands at some fixed redshift $z_0$.
+read, ``$b$-band at redshift $z$.'' For example, if we measure the
+flux of a galaxy at $z=0.22$ using a $g$-band filter, we have obtained
+the $\band{0.22}{g}$-band magnitude of the galaxy. The goal of
+``$K$-corrections'' is to convert the $ugriz$ bands at redshift $z$
+for each galaxy to $ugriz$ bands at some fixed redshift $z_0$.
 
@@
 reconstruct the galaxy rest-frame SED.  With this reconstruction in
-hand, it is straightforward to project the SED onto bands at any
+hand, it is straightforward to project the SED onto any bands at any
 redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
 \begin{eqnarray}
-^{z_0}m_{\mathrm{rec}} &=& -2.41 - 2.5 \log_{10}\left[
+\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
 \frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
-f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)}
-{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right] \cr
-&=& -48.60 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
-f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)}
-{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right] 
+f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
+d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
+\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
+f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
+R(\nu)}\right]
 \end{eqnarray}
-The simple way to think about this equation is that simply moving away
-from the Earth doing does not change the bolometric flux of the
-galaxy. So when the spectrum is stretched by the $(1+z_0)$ redshift
-factor, as in $f(\lambda/(1+z_0))$, the bolometric flux should be
-constant; that is, you must apply the extra multiplicative factor
-$1/(1+z_0)$ to counteract the stretching of the spectrum.
+The simple way to remember where the factors of $(1+z)$ appear in
+these equations is to remember that $K$-corrections are defined to be
+the change in magnitude with redshift for {\it fixed bolometric flux}.
+So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
+in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
+is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
+counteract the stretching of the spectrum.
 
@@
 for our fits), there is very little difference between these two
-methods. Having estimated $\band{z_0}{m}$, you can then calculate the
-luminosity by simply applying to the calculated magnitude the
-cosmological distance modulus (that is, the luminosity
+methods. However, if the reconstructions are poor, there will be large
+differences between the methods; in particular, method (1) will yield
+a color distribution of galaxies with artificially low
+dimensionality. Having estimated $\band{z_0}{m}$, you can then
+calculate the luminosity by simply applying to the calculated
+magnitude the cosmological distance modulus (that is, the luminosity
 distance-squared law, as tabulated by, {\it e.g.}  \citealt{hogg99a}).
@@
 objective, it is not {\it always} necessary or appropriate. Because
-$K$-corrections are an inherently uncertain business (the broad-band
-magnitudes just do not uniquely determine the SED) they should be
-avoided or minimized when possible. For example, if one was calculating
-the evolution of clustering of red and blue galaxies separately, it
-would perhaps be wise to perform the separation not on the
-$K$-corrected colors of the galaxies but on the median color of, say,
-$M_\ast$ galaxies as a function of redshift. Similarly, in situation
-where $K$-corrections cannot be ignored, such as the calculation of
-the evolution of the luminosity function, their effect should be
-minimized by, for example, correcting to the median redshift of
-galaxies in the sample.
-
-Now, typically one's measurements in will be difficult to connect to,
-say, solar luminosities in $\band{z}{b}$ (unless $z=0$), simply
-because people have not projected the appropriate stellar
-spectrophotometry onto these bandpasses. However, it is our position
-that stars are better understood than galaxies, and that it therefore
-is in the end simpler to stay as close as possible to the system in
-which the galaxies are observed. In any case, astronomy is quickly
-reaching a level of precision for which the exact nature of the
-bandpasses used has to be known and considered in most analyses of
-observational data.
+$K$-corrections are inherently uncertain (the broad-band magnitudes
+do not uniquely determine the SED) they should be avoided or
+minimized when possible. For example, if one were calculating the
+evolution of clustering of red and blue galaxies separately, it would
+perhaps be wise to perform the separation not on the $K$-corrected
+colors of the galaxies but on the median color of, say, $M_\ast$
+galaxies as a function of redshift. Similarly, in situation where
+$K$-corrections cannot be ignored, such as the calculation of the
+evolution of the luminosity function, their effect should be minimized
+by, for example, correcting to the median redshift of galaxies in the
+sample.
+
+Typically one's measurements will be difficult to connect to solar
+luminosities in $\band{z}{b}$ (unless $z=0$), simply because nobody
+has projected the appropriate stellar spectrophotometry onto these
+bandpasses. However, it is our position that stars are better
+understood than galaxies, and that it is therefore simpler in the end
+to stay as close as possible to the system in which the galaxies are
+observed. In any case, astronomy is quickly reaching a level of
+precision for which the exact nature of the bandpasses used has to be
+known and considered in most analyses of observational data.
 
@@
 model of the galaxy SED, we do not correct to bandpasses with simpler
-shapes (say, Gaussians). The reason is that we want the effective
+shapes (say, top hats). The reason is that we want the effective
 bandpasses to actually have been observed for some galaxies in the
-sample, so that the $K$-correction for those galaxies is formally and
+sample, so that the $K$-correction for those galaxies are formally and
 actually zero. This property is desirable because, as noted above,
-there are uncertainties in the $K$-corrections. Nevertheless, we note
-that synthesizing such magnitudes may be appropriate in certain
-situations, as they have the virtue (similarly to the $AB$ magnitude
-system) that such magnitudes are very easy to comprehend and
-synthesize from theory. For example, if we wanted to convert our
-results into a form easily comprehended by the hypothetical future
-astronomers discussed in the first paragraph of Section
-\ref{motivation}, it would be best to calculate fluxes within wide
-top-hat filters so that the difficulties of understanding the
-magnitude system were minimized.
+there are inherent uncertainties in the $K$-corrections. Nevertheless,
+we note that synthesizing such magnitudes may be appropriate in
+certain situations, as they have the virtue (similarly to the $AB$
+magnitude system) that they are very easy to comprehend and synthesize
+from theory.
+%For example, if we wanted to convert our results into a
+%form easily comprehended by the hypothetical future astronomers
+%discussed in the first paragraph of the introduction, it would be best
+%to calculate fluxes within wide top-hat filters so that the
+%difficulties of understanding the magnitude system were minimized.
 
@@
 
-The version of the $K$-correction code ({\tt v1\_2}) implementing the
+The version of the $K$-correction code ({\tt v1\_9}) implementing the
 calculations described here, along with eigentemplates and filter
 curves is publicly available on the World Wide Web at {\tt
-http://physics.nyu.edu/\~\ mb144/kcorrect}. In addition, a public CVS
-repository can be accessed at {\tt xxx} (instructions for exporting
-the appropriate version of the code from this repository are provided
-on the web page noted above). The whole of the code can be used
-through the Research Systems, Incorporated, IDL language; everything
-except for the template-fitting also exists in stand-alone C programs
-(which call the same routines, guaranteeing consistency).
+http://physics.nyu.edu/\~\ mb144/kcorrect}.  The whole of the code can
+be used through the Research Systems, Incorporated, IDL language;
+everything except for the template-fitting also exists in stand-alone
+C programs (which call the same routines, guaranteeing consistency).
 
@@
 galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Luminous
-Red Galaxies (LRGs; Eisenstein {\it et al.}, in preparation), and
+Red Galaxies (LRGs; \citealt{eisenstein01a}), and
 100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation).
@@
 are described in the Early Data Release (EDR;
-\citealt{stoughton01a}). The survey has begun in earnest, and has
-obtained about 20\% of its intended data.  As of February 2002, the
-SDSS had targeted and taken spectra of approximately XXX galaxies over
-XXX square degrees.
-
-As described below, only a subset of these galaxies are used here to
-develop the templates, though we calculate the SED fits for all of
-them. The results of the photometric pipeline for all of these objects
-were extracted from the SDSS Operational Database. The
-photometry used here for the bulk of these objects was the same as
-that used when the objects were targeted. However, for those objects
-which were in the EDR photometric catalog, we used the better
-calibrations and photometry from the EDR. The versions of the SDSS
-pipeline PHOTO used for the reductions of these data ranged from {\tt
-v5\_0} to {\tt v5\_2}. The changes in the treatment of relatively
-small galaxies, which accounts for most of our sample, did not change
-substantially throughout these versions.
-
-The magnitudes are calibrated in the $AB$ system, meaning that the
-magnitudes are designed such that:
-\begin{eqnarray}
-\label{ABdef}
-m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
-{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
-\cr
-&=& -48.60 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
-{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
-\end{eqnarray}
-where $R(\lambda)$ is the fraction of photons entering the Earth's
-atmosphere which are detected as a function of wavelength (a unitless
-quantity). In this equation, $f(\lambda)$ is in units of ergs
-cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
-cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
-that an object with $f(\nu) = 3.631 \times 10^{-20}$ ergs cm$^{-2}$
-s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes equal to zero.  The
-difference in the zeropoint in the two equations simply corresponds to
-the factor of the speed of light $c$ (expressed in \AA\ s$^{-1}$)
-between the frequency and wavelength scales.
-
-The response functions of the filter-CCD combinations have been
-measured using a monochrometer by Mamoru Doi.  Using a model
+\citealt{stoughton01a}).
+
+The results of the photometric pipeline for all of the galaxies were
+extracted from the SDSS Operational Database. The photometry used here
+for the bulk of these objects was the same as that used when the
+objects were targeted. However, for those objects which were in the
+EDR photometric catalog, we used the better calibrations and
+photometry from the EDR. The versions of the SDSS pipeline PHOTO used
+for the reductions of these data ranged from {\tt v5\_0} to {\tt
+v5\_2}. The treatment of relatively small galaxies, which account for
+most of our sample, did not change substantially throughout these
+versions. The magnitudes are calibrated approximately in the $AB$
+system, described in Section \ref{motivation}.
+
+The response functions $R(\lambda)$ of the filter-CCD combinations
+have been measured using a monochrometer by Mamoru Doi.  Using a model
 for the atmospheric transmission and the reflectivity of the primary
@@
 determination. They are consistent (for galaxies) at over the 99\%
-level.\footnote{The official pipeline tends to work better for objects
-with unusual spectra, such as certain types of stars and QSOs.}.
-
-We use two types of magnitudes determined by the SDSS. First, the
-modified form of the magnitude described by \citet{petrosian76a}, as
-described in \citet{strauss02a}.  These magnitudes have a nearly
-constant metric aperture as a function redshift; in addition, all the
-bands use the same aperture, so the measured SED corresponds (to
-within the effects of seeing) to the SED of an identifiable region of
-the galaxy. However, for faint objects, the Petrosian magnitudes tend
-to become noisy. Thus, for galaxies with $z>0.25$ we instead use the
-higher signal-to-noise ``model magnitudes.'' Model magnitudes are
-calculated in all bands using a single weighted aperture; thus, the
-measured colors again correspond to an identifiable region of the
-galaxy, though a different one than would be measured by the Petrosian
-magnitude. The weighted aperture is the better fitting model (pure
-exponential or pure de Vaucouleurs) to the galaxy image in the
-$r$-band. In \citet{stoughton01a} we are explicitly warned (in a
-paragraph written by the lead author of this paper!)  not to mix
-Petrosian and model magnitudes in a single analysis, since they
-measure galaxies in very different ways. Nevertheless, for the
-purposes of contraining galaxy template SEDs, it is perfectly
-acceptable to use {\it any} well-defined part of any galaxy.  For our
-purposes, it is more important to have high signal-to-noise
-measurements of galaxy colors than to probe exactly the same regions
-of galaxies at low and high redshift.  We extinction-correct both
-types of magnitudes using the dust maps of \citet{schlegel98a}.
+level.\footnote{The official pipeline currently tends to work better
+for objects with unusual spectra, such as certain types of stars and
+QSOs.}.
+
+We use two types of magnitudes determined by the SDSS. First, the SDSS
+Petrosian magnitudes, a modified form of the magnitude described by
+\citet{petrosian76a}, as described in \citet{strauss02a}.  These
+magnitudes have a nearly constant metric aperture as a function
+redshift; in addition, all the bands use the same aperture, so the
+measured SED corresponds (to within the effects of seeing) to the SED
+of an identifiable region of the galaxy. However, for faint objects,
+the Petrosian magnitudes tend to become noisy. Thus, for galaxies with
+$z>0.25$ we instead use the higher signal-to-noise ``model
+magnitudes.'' Model magnitudes are calculated in all bands using a
+single weighted aperture; thus, the measured colors again correspond
+to an identifiable region of the galaxy, though a different one than
+would be measured by the Petrosian magnitude. The weighted aperture is
+the better fitting model (pure exponential or pure de Vaucouleurs) to
+the galaxy image in the $r$-band. In \citet{stoughton01a} we are
+explicitly warned not to mix Petrosian and model magnitudes in a
+single analysis, since they measure galaxies in very different
+ways. Nevertheless, for the purposes of contraining galaxy template
+SEDs, it is perfectly acceptable to use {\it any} well-defined part of
+any galaxy.  For our purposes, it is more important to have high
+signal-to-noise measurements of galaxy colors than to probe exactly
+the same regions of galaxies at low and high redshift.  We
+extinction-correct both types of magnitudes using the dust maps of
+\citet{schlegel98a}.
 
@@
 (totalling about 1,000 objects) which were selected by a photometric
-redshift algorithm ({\bf ref what?}) to be at around $z\sim
-0.3$--$0.4$. These objects are invaluable for tying down the blue end
-of the templates. Finally, we exclude galaxies in the redshift range
-$0.28 <z<0.32$ from the fit for the templates, for reasons which we
-will explain more fully below (nevertheless, we still can and do use
-the resulting templates to analyze galaxies in this range).
+redshift algorithm (Csabai {\it et al.}, in preparation) to be at
+around $z\sim 0.3$--$0.4$. These objects are invaluable for tying down
+the blue end of the templates. Finally, we exclude galaxies in the
+redshift range $0.28 <z<0.32$ from the fit for the templates, for
+reasons which we will explain more fully below (nevertheless, we still
+can and do use the resulting templates to analyze galaxies in this
+range).
 
@@
 the photon noise of the sky. We identify such cases as magnitude
-errors greater than 0.8 or magnitudes fainter than 22.5 in any
+errors greater than 2.0 or magnitudes fainter than 24.0 in any
 band. We ignore these objects entirely when fitting for the
 templates. However, we still want to determine a best-fit SED for each
-object. For this purpose, we replace a ``bad'' magnitude based on an
-adjacent (usually the redder) band and the average color in those
-bands of objects at that redshift. We give these measurements a large
-error bar, of 0.8 magnitudes. Thus, we account for the missing
-information by simply requiring that the object SED have
-``reasonable'' properties.
+object, once the templates are determined. Thus, we constrain all the
+SED fits to galaxies using the method described in Section
+\ref{direct}. As a constraint, we use the covariance matrix and mean
+value (essentially zero, by design) of the coefficients of the
+well-measured set of galaxies; we multiply all the elements of the
+covariance matrix by a factor of our to weaken our constraint ((which
+is appropriate, since the actual distribution of coefficients is not
+well described by a single gaussian). For this purpose, we replace a
+``bad'' magnitude based on an adjacent (usually the redder) band and
+the average color in those bands of objects at that redshift. We give
+these measurements a large error bar, of 0.8 magnitudes. Thus, we
+account for the missing information by simply requiring that the
+object SED have ``reasonable'' properties.
 
@@
 spectrum which contains the sharpest gradients, around 4000\AA, we
-smooth only with $\sigma = 150$\AA.
+smooth only with $\sigma = 150$\AA. None of our results change
+dramatically if we vary our smoothing procedure or if we add reddened
+templates to our allowed SED space.
 
 We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
-and still allow freedom to fit for the templates themselves. Since
-galaxies live in a low-dimensional space, we find it is worth being
-able to fit for the templates. Simply using five templates (which
-obviously reproduces all the magnitudes exactly) tends to yield
-unphysical trends of galaxy SED versus redshift ({\it cf.}  Figure
-\ref{main_colors_plot} below). Using three templates does nearly as well as
-four templates in the sense that the resulting templates reproduce the
-$griz$ magnitudes nearly as well. However, the fourth template is
-necessary to recover the $u$-band flux to better than about 15\%. In
-addition, since one of the applications of these SED determinations is
-the distribution of galaxy colors in fixed-frame magnitudes, we don't
-want to artificially reduce the dimensionality of the color space to
-only two.
+and still allow freedom to fit for the templates themselves.  Simply
+using five templates (which obviously reproduces all the magnitudes
+exactly) tends to yield unphysical trends of galaxy SED versus
+redshift ({\it cf.}  Figure \ref{main_colors_plot} below). Using three
+templates does nearly as well as four templates in the sense that the
+resulting templates reproduce the $griz$ magnitudes nearly as
+well. However, the fourth template is necessary to recover the
+$u$-band flux to any better than about 15\%. In addition, since one of
+the applications of these SED determinations is the distribution of
+galaxy colors in fixed-frame magnitudes, we don't want to artificially
+reduce the dimensionality of the color space to only two.
 
@@
 spike at around 4000 \AA; that this direction is important is not
-surprising, since one of the most variable qualities of galaxy SEDs is
+surprising, since one of the most variable quantities of galaxy SEDs is
 the size of the 4000 \AA\ break. However, the gap between the $g$ and
@@
 To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
-corresponding to the direction represented by the vector $-0.6
-\vv{e}_1(\lambda) + \vv{e}_3$. We overplot the \band{0.3}{g} and
+corresponding to the direction represented by the vector $
+\vv{e}_2(\lambda) + 0.7 \vv{e}_3$. We overplot the \band{0.3}{g} and
 \band{0.3}{r} bands. This direction has a strong peak in the gap
@@
 
-However, we have not tackled this approach yet. We have implemented
-(and used for the figures shown here) a temporary fix for those
-interested in galaxies at $z=0.3$, which is to artificially constrains
-the fit as follows. We pretend that there is an extra band between $g$
-and $r$, which has a FWHM of 2000 \AA\ and a center of 5500 \AA. The
-``measured'' value we use in this band is the mean of the $g$ and $r$
-fluxes, and is given $4\sigma$ significance. This extra constraint
-changes the K-corrections for galaxies outside the range $0.27 < z <
-0.33$ by far less than 1 percent. The constraint helps to clean up the
-fits but it is not sufficient; the coefficients are still clearly
-wrong around $z=0.3$. It helps considerably to $K$-correct to $z=0.3$,
-which we recommend for the moment for studies of LRGs; this minimizes
-the corrections which are performed using our most unphysical fits.
+However, we have not tackled this approach yet. As it happens, the the
+direct constraint on the coefficients described in the previous
+section is sufficient to suppress the $g$/$r$ gap problem sufficiently
+for many purposes, as long as one $K$-corrections the magnitudes to
+$z=0.3$. In this way, one minimizes the $K$-corrections for the
+galaxies which one is least certain of.
 
@@
 when you are dealing with the LRGs, for galaxies in the range about
-0.27 < z < 0.33. The Main Sample results are not significantly
+$0.27 < z < 0.33$. The Main Sample results are not significantly
 affected at all.  Furthermore, one can still observe a galaxy at
@@
 $z$-band response is at 8700\AA). Therefore, the odd behavior near the
-edges should not be taken too seriously. In addition, note the spur
-extending from upper left to lower right through the left edge of the
-($a3/a0$)-($a1/a0$) plane. Most of the objects in this extendend spur
-are at around $z=0.3$. We discuss in Section \ref{grgap} why this spur
-arises and what to do about it.
+edges should not be taken too seriously. In addition, note the small
+spur extending from upper left to lower right through the right edge
+of the ($a_3/a_0$)-($a_2/a_0$) plane. Many of the objects in this spur are
+at around $z=0.3$, and this spur is the result of the degeracy due to
+the $g$/$r$ gap.
 
@@
 and the reconstructed fluxes (shown on the right bottom corner of each
-panel) are of order the photometric errors in the sample. This
-agreement should not be overly surprising given that we have five
-galaxy fluxes and four templates to fit, and when one considers the
-fact that galaxy colors are known to be correlated with one another.
+panel) are of order the photometric errors in the sample. In
+particular, for our 30,000 galaxies (and thus approximately 30,000
+degrees of freedom) we find $\chi^2 \sim 55,000$, indicating that our
+fit is reasonable.
 
@@
 perform this test. First, we choose a set of Main Sample galaxies in
-the luminosity range $-21.5 < M_{\band{0.1}{r}} < -21$. We reconstruct
+the luminosity range $-21.5 < M_{\band{0.1}{r}} < -21.2$. We reconstruct
 the four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
@@
 Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.3}{r}}
-< -22.5$. For this sample we reconstruct the colors at $z=0.3$.  Near
-$z=0.3$ there is some singular behavior of the colors, which we will
-discuss in more detail in Section \ref{grgap}. Again, for $z<0.3$ the
-$^{0.3}(u-g)$ color relies on an extrapolation, which is probably
-responsible for the scatter in that color at low redshifts. Otherwise,
-the distribution of colors is remarkably constant. $^{0.3}(g-r)$
-(which is very similar to $^{0.0}(u-g)$) experiences a blueward shift
-of about 0.1 magnitudes between $z=0.1$ and $z=0.45$, which is
-attributable to passive evolution.
+< -22.5$. For this sample we reconstruct the colors at $z=0.3$.
+Again, for $z<0.3$ the $^{0.3}(u-g)$ color relies on an extrapolation,
+which is probably responsible for the scatter in that color at low
+redshifts. Otherwise, the distribution of colors is remarkably
+constant. $^{0.3}(g-r)$ (which is very similar to $^{0.0}(u-g)$)
+experiences a blueward shift of about 0.1 magnitudes between $z=0.1$
+and $z=0.45$, which is attributable to passive evolution.
 
@@
 discussed in more detail in the next subsection.
-
-\subsection{The $g$/$r$ Gap}
-\label{grgap}
-
-It is clear from Figure \ref{lrg_colors_plot} that $z=0.3$ is a
-special redshift regime where our fits become unphysical. In fact, it
-is easy to demonstrate that the diagonal spur (going from the upper
-left to the lower right) around the red dot in the upper right panel
-of Figure \ref{k_coeffdist_plot} is due almost completely to galaxies
-at $z\sim 0.3$. The spur arises because the 4000 \AA\ break lies in
-the gap between $g$ and $r$ at $z=0.3$, visible in Figure
-\ref{response_sdss}. Our algorithm has strong gradients in the
-spectrum around the 4000\AA\ break, which is necessary to properly
-model galaxy spectra. In the presence of these sharp gradients, the
-gap in the bands causes certain combinations of the template spectra
-to be degenerate at around $z=0.3$.
-
-To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
-corresponding to the direction of the spur in the
-$(a_1/a_0)$-$(a_3/a_0)$ plane of Figure \ref{k_coeffdist_plot}; this
-direction is represented by the vector $-0.6 \vv{e}_1(\lambda) +
-\vv{e}_3$. We overplot the \band{0.3}{g} and \band{0.3}{r} bands. The
-spur has a strong peak in the gap between the two bands at
-$z=0.3$. Because of this gap, at $z\sim 0.3$ the algorithm tends to a
-component in this direction to help fit the $u$, $i$, and $z$ fluxes,
-with relatively little extra $\chi^2$ contribution from the $g$ and
-$r$ observations. Thus the reconstructed fluxes in Figure
-\ref{k_model_plot} fit well even at $z=0.3$, even though the spur at
-$z=0.3$ in the distribution of coefficients is so evident.  This
-difficulty is why, in Section \ref{data}, we exclude the regime around
-$z\sim 0.3$; otherwise the templates became distracted by the
-degeneracies in this regime.
-
-The question arises of what to do about this problem. One solution is
-to simply work around it. It is clear we cannot use the fits presented
-here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
-reliably --- their coefficients $a_k$ are unusual and probably
-unphysical. However, nothing prevents one from $K$-correcting galaxies
-at other redshifts {\it to} $z=0.3$. If you are considering the LRG
-sample, the median redshift is close to $z=0.3$ anyway, so this is a
-viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).
-
-A second solution is to place extra constraints on the fit which
-prevent the spur in the coefficients. One can imagine fitting the
-distribution in the $(a_1/a_0)$-$(a_3/a_0)$ plane with a simple model,
-such as a multigaussian fit, and using the resulting fit as a prior in
-one's fit (with gaussians, the extra constraints would be linear and
-thus easily accounted for). One could design these constraints to be
-strong enough (at least near the redshifts in question) to prevent the
-spur. However, the disadvantage of this solution is that for some
-scientific projects the change in the galaxy SED is exactly what one
-is trying to measure. In these cases, it may be inappropriate to be
-putting explicit {\it a priori} constraints on exactly what one is
-trying to measure.
-
-A third solution, not adopted here, is to place stronger constraints
-in the SED fit on the form of the 4000 \AA\ break. One obvious way of
-doing this is to use the spectra to constrain the size of the break
-(making it a weak constraint to prevent biases). This constraint could
-be applied in a linear fashion to retain the speed of our method ---
-except for the fact that you would have to read in the spectra to
-apply the $K$-correction. For simplicity, and because the aforemention
-work-around is so easy to obey, we have not implemented this solution
-here.
 
@@
 $K$-corrections given are not {\it too} unreasonable).
+
+An important test of our method is to synthesize broad-band photometry
+from spectra, and then try to recover the $K$-corrections in a case
+where we know the spectrum completely. For this purpose, we use as
+example spectra the spectra of galaxies in the SDSS.  Since we cannot
+synthesize the observed $u$ and $z$ bands, we simply base them on the
+synthesized $g$ and $i$ band magnitudes plus the $u-g$ and $z-i$
+colors. (This procedure is unlikely to make the photometric
+estimations of the $K$-corrections {\it better}). Finally, we added
+2\% random errors to all of the measurements. Figure
+\ref{k_speck_plot.fitfib.0.1} compares the $K$-corrections from the
+synthesized photometry to that calculated from the spectrum
+itself: 
+\begin{equation}
+\Delta K = K_{\mathrm{spec}} - K_{\mathrm{photo}}
+\end{equation}
+The agreement is now very good, suggesting that our method can
+indeed recover the correct $K$-corrections based only on broad-band
+magnitudes.
 
@@
 Finally, it is possible to use the galaxy spectra obtained with the
-spectrograph to estimate the $K$-correction for each object.  However,
-this procedure requires trusting the spectrophotometry over a wide
-wavelength range. In addition, the region of the galaxy which the
-spectra probe is not completely well defined. The small scale
-wavelength features are set by the fiber aperture ($3''$ in diameter)
-which large scale wavelength features are constrained by a short
-``smear exposure'' which probes a larger area. The smear exposure
-corrections are clearly the correct approach for point sources, but
-for galaxies it does cause the small scale and large scale features of
-the spectrum to be determined by somewhat different regions of the
-galaxy.
+spectrograph to estimate the $K$-correction for each object. 
+However, this procedure requires trusting the
+spectrophotometry over a wide wavelength range. In addition, the
+region of the galaxy which the spectra probe is not completely well
+defined. The small scale wavelength features are set by the fiber
+aperture ($3''$ in diameter) while large scale wavelength features are
+constrained by a short ``smear exposure'' which probes a larger
+area. The smear exposure corrections are clearly the correct approach
+for point sources, but for galaxies it does cause the small scale and
+large scale features of the spectrum to be determined by somewhat
+different regions of the galaxy.
 
@@
 Petrosian aperture. The implication is that the $K$-corrections {\it
-are} stronger for the fiber aperture than for the Petrosian aperture.
-Correspondingly we repeat the exercise using the SDSS fiber magnitudes
-to determine the $K$-corrections, we find consistent results for the
-$\band{0.1}{g}$-band.
-
-Skeptics may worry that even if we {\it did} have photometry that
-corresponded precisely to the same region of the galaxy that the
-spectroscopy did, we still might be calculating the $g$-band
-$K$-corrections in some biased way. However, we can test whether this
-is the case by synthesizing the observed broad-band photometry from
-the spectra, and trying to recover the $K$-corrections in a case where
-we know the spectrum completely. Since we cannot synthesize the
-observed $u$ and $z$ bands, we simply base them on the synthesized $g$
-and $i$ band magnitudes plus the $u-g$ and $z-i$ colors. (This
-procedure is unlikely to make the photometric estimations of the
-$K$-corrections {\it better}). Finally, we added 2\% random errors to
-all of the measurements. Figure \ref{k_speck_plot.fitfib.0.1} compares
-the $K$-corrections from the synthesized photometry to that calculated
-from the spectroscopy itself. The agreement is now very good,
-suggesting that our method can indeed recover the correct
-$K$-corrections based only on broad-band magnitudes (and furthermore
-that for the $g$-band one should not rely on $K$-corrections based on
-the spectroscopy alone).
+are} stronger for the aperture which corresponds to the spectrum than
+for the Petrosian aperture. This indicates that for work which
+requires precision, it is best to avoid using the $K$-corrections
+determined from the spectra. (Note the in the second paragraph of this
+section we used the SDSS spectra as {\it example} spectra to validate
+the method; in this paragraph we have evaluate whether to rely on the
+spectra to calculate $K$-corrections).
 
@@
 using the released code) to identify galaxies which are likely to be
-redshift outliers.
+redshift outliers. Another use of the three-dimensional space of
+galaxy SEDs is to create mock samples of galaxies observed in any
+bandpass at any redshift. We are working on quantifying the
+correlations between these coefficients and luminosity,
+surface-brighntess and radial profile in order to create mock catalogs
+which truly reflect the correlations between galaxy properties found
+in the data.
 

### 2002-04-15 src/color_interp.c

@@ -15,11 +15,13 @@
 
 #define ZZERO 0.1
 #define MAGGIEFILE "testsp.K.dat"
-#define RFFILE "testsp.K.cibreak.dat"
+#define RFFILE "testsp.K.ci.dat"
 #define INCLUDEBREAK
+#undef INCLUDEBREAK
 #define NGALS 6534
 #define FILTERPATH "../data/filters"
 #define FILTERBASE "sdss_"
+
 
 static int band,currband,igal;
 static int band0,band1;

### 2002-04-15 docs/paper/paper.tex
changes
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@100 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \newcommand{\band}[2]{\ensuremath{^{#1}\!{#2}}}
+
+\newcommand{\kversion}{{\tt v1\_10}}
 
@@
 that the description on p. 57 of \citet{binney98a} of the meaning of
-$k(z)$ in \citet{frie94a} is incorrect; in fact, $k(z)\equiv K(z)$.}
+$k(z)$ in \citet{frei94a} is incorrect; in fact, $k(z)\equiv K(z)$.}
 
@@
 specified in any such paper. The version of the code used to make the
-figures for this paper is {\tt kcorrect v1\_9}.
+figures for this paper is {\tt kcorrect \kversion}.
 
@@
 code in units of ergs cm$^{-2}$ s$^{-1}$ (within the range
-$\ambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$).
+$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$).
 
@@
 
-The version of the $K$-correction code ({\tt v1\_9}) implementing the
+The version of the $K$-correction code (\kversion) implementing the
 calculations described here, along with eigentemplates and filter

### 2002-04-18 docs/paper/figures.tex
chjamnges
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@104 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \caption{\label{response_sdss} Estimated filter response for all five
-bands in the SDSS, as a function of observed wavelength. A 4 gigayear
-old instantaneous burst using the models of \citet{bruzual93a} (and
-observed at $z=0$) is shown for reference.}
+bands in the SDSS, as a function of observed wavelength, as measured
+by Mamoru Doi. A 4 gigayear old instantaneous burst using the models
+of \citet{bruzual93a} (and observed at $z=0$) is shown for reference.}
 %at different redshifts. A 4 gigayear old instantaneous burst using the
@@
 %determine \band{0.0}{u} and \band{0.1}{z}. 
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{spur.ps}
+\caption{\label{spur} The spectrum corresponding to the direction of
+the spur in the upper right panel of Figure
+\ref{k_coeffdist_plot}. Note the strong feature near 4000
+\AA. Overplotted are the bandpasses for $\band{0.3}{g}$ and
+$\band{0.3}{r}$. The strong features fall in the gap between the
+bandpasses. Thus, in a linear fit to a galaxy at $z=0.3$, this
+component can be used to fit the observed magnitudes without being
+constrained to have reasonable behavior around 4000 \AA. }
 \end{figure}
@@
 \figurenum{\fignum}
+\plotone{main_colors_plot.z.ps}
+\caption{\label{main_colors_plot.z} Color distributions in the
+observed frame for SDSS Main Sample galaxies in the luminosity range
+$-21.5<M_{\band{0.1}{r}}<-21.2$. This sample is complete (that is,
+volume limited) for $0.05<z<0.17$. Left panels show the colors as a
+function of redshift. Right panel shows the distributions of each
+color at high and low redshift within the volume-limited
+subsample. The observed colors clearly depend strongly on redshift,
+even for the $\band{0.1}{u}$ band, where the low-redshift end is an
+extrapolation of the data.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
 \plotone{main_colors_plot.ps}
-\caption{\label{main_colors_plot} Color distributions $K$-corrected to
-$z=0.1$ for SDSS Main Sample galaxies in the luminosity range
-$-21.5<M_{\band{0.1}{r}}<-21.2$. This sample is complete for
-$0.05<z<0.17$. Left panels show the colors as a function of
-redshift. Right panel shows the distributions of each color at high
-and low redshift within the volume-limited subsample. There is very
-little dependence of the colors on redshift, even for the
-$\band{0.1}{u}$ band, where the low-redshift end is an extrapolation
-of the data.}
+\caption{\label{main_colors_plot} Similar to Figure
+\ref{main_colors_plot.z}, but now the colors are $K$-corrected to
+$z=0.1$.  There is very little dependence of the colors on redshift,
+even for the $\band{0.1}{u}$ band, where the low-redshift end is an
+extrapolation of the data.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{lrg_colors_plot.z.ps}
+\caption{\label{lrg_colors_plot.z} Similar to Figure
+\ref{main_colors_plot.z}, now showing LRG galaxies (Cut I) in the
+luminosity range $-22.8<M_{\band{0.3}{r}}<-22.5$. Again, there is a
+strong dependence on redshift.}
 \end{figure}
@@
 \caption{\label{lrg_colors_plot} Same as Figure
-\ref{main_colors_plot}, now $K$-corrected to $z=0.3$ for LRG galaxies
-in the luminosity range $-22.8<M_{\band{0.3}{r}}<-22.5$. Now
-considerably more dependence on redshift is apparent. First of all, a
-degeneracy at $z\sim 0.3$ appears (described in detail in Section
-\ref{grgap}). Second of all, in $\band{0.3}{(g-r)}$ an overall trend
-is apparent; LRGs at $z=0.4$ are about 0.1 magnitudes bluer than LRGs
-at $z=0.2$. This change is attributable to galaxy evolution.}
-\end{figure}
-
-\clearpage
-\stepcounter{thefigs}
-\begin{figure}
-\figurenum{\fignum}
-\plotone{spur.ps}
-\caption{\label{spur} The spectrum corresponding to the direction of
-the spur in the upper right panel of Figure
-\ref{k_coeffdist_plot}. Note the strong feature near 4000
-\AA. Overplotted are the bandpasses for $\band{0.3}{g}$ and
-$\band{0.3}{r}$. The strong features fall in the gap between the
-bandpasses. Thus, in a linear fit to a galaxy at $z=0.3$, this
-component can be used to fit the observed magnitudes without being
-constrained to have reasonable behavior around 4000 \AA. This fact
-explains the singular behavior at $z=0.3$ in Figure
-\ref{lrg_colors_plot}.}
+\ref{lrg_colors_plot.z}, now $K$-correcting the LRG galaxies to
+$z=0.3$.  The redshift dependence is greatly reduced for the LRGs in
+comparison to Figure \ref{lrg_colors_plot.z}; on the other hand, there
+are distinct trends of restframe color with redshift.  In
+$\band{0.3}{(g-r)}$ an overall trend is apparent; LRGs at $z=0.4$ are
+about 0.1 magnitudes bluer than LRGs at $z=0.2$. A change of this
+magnitude is attributable to passive galaxy evolution, though
+considerably more work needs to be done to show that this is
+occurring. In $\band{0.3}{(r-i)}$, a blueward shift also occurs,
+though at a much smaller level.}
 \end{figure}
@@
 scientific purposes. Note the degeneracy at $z=0.3$.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{k_speck_plot.fitfib.0.1.ps}
+\caption{\label{k_speck_plot.fitfib.0.1} Difference between the
+$K$-corrections to $z=0.1$ determined from the spectroscopy and those
+determined from the analysis of broad-band magnitudes {\it
+synthesized} from the same spectra.}
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{k_speck_plot.fitfib.0.1.ps}
-\caption{\label{k_speck_plot.fitfib.0.1} Difference between the
-$K$-corrections to $z=0.1$ determined from the spectroscopy and those
-determined from the analysis of broad-band magnitudes {\it
-synthesized} from the same spectra.}
-\end{figure}
-
-\clearpage
-\stepcounter{thefigs}
-\begin{figure}
-\figurenum{\fignum}
 \plotone{k_speck_plot.0.3.ps}

### 2002-04-18 docs/paper/paper.tex
chjamnges
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@104 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-\title{Estimating Fixed-frame Galaxy Magnitudes in the SDSS$^1$}
+\title{Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$}
 
@@
 prepare to address these astronomical questions, investigated in
-detail in subsequent papers, we implement here the method of
-\citet{csabai00a} for fitting spectral energy distributions (SEDs) to
-broad-band photometric observations, in the context of the optical
-observations of the Sloan Digital Sky Survey (SDSS). Linear
+detail in subsequent papers, we fit spectral energy distributions
+(SEDs) to broad band photometric observations, in the context of the
+optical observations of the Sloan Digital Sky Survey (SDSS). Linear
 combinations of four spectral templates can reproduce the five SDSS
@@
 and in fact nearly linear. The resulting reconstructed SEDs can be
-used to recover fixed-frame magnitudes over a range of redshifts. This
+used to recover fixed frame magnitudes over a range of redshifts. This
 process yields consistent results, in the sense that within each
@@
 constant with redshift. We compare our results to simpler
-interpolation methods and galaxy spectrophotometry.  The software that
-generates these results is publicly available and easily adapted to
-handle a wide range of galaxy observations.
+interpolation methods and galaxy spectrophotometry from the SDSS.  The
+software that generates these results is publicly available and easily
+adapted to handle a wide range of galaxy observations.
 \end{abstract}
@@
 resolution and high spectral resolution. By that time, the phenomena
-of broad-band filters and the quaint terminology surrounding their
+of broad band filters and the quaint terminology surrounding their
 usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
@@
 information on scales considerably smaller than the width of typical
-broad-band filters. Furthermore, SEDs of distant galaxies are
+broad band filters. Furthermore, SEDs of distant galaxies are
 redshifted such that the more distant the galaxy, the further the
@@
 analyses, a function $K(z)$ is added to the standard cosmological
-bolometric distance modulus $\mathrm{DM}(z)$. Sometimes a single
-function has been applied regardless of the galaxy type, though more
-recently it has become standard to use a discrete set of $K(z)$
-functions depending on galaxy type (based either on morphology or
-spectral features). As described in \cite{oke68a} and later papers,
-the function is based on the projection of an assumed galaxy SED
-redshifted to $z$ onto the measured instrument response as a function
-of wavelength (including the effects of the atmosphere, the
+bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
+relationship between the apparent magnitude $m$ and the absolute
+magnitude $M$: 
+\begin{equation}
+m = M + \mathrm{DM}(z) + K(z)
+\end{equation}
+Sometimes a single function has been applied regardless of the galaxy
+type, though more recently it has become standard to use a discrete
+set of $K(z)$ functions depending on galaxy type (based either on
+morphology or spectral features). As described in \cite{oke68a} and
+later papers, the function is based on the projection of an assumed
+galaxy SED redshifted to $z$ onto the measured instrument response as
+a function of wavelength (including the effects of the atmosphere, the
 reflectivity of the mirrors, the filter transmission, and the response
@@
 that the description on p. 57 of \citet{binney98a} of the meaning of
-$k(z)$ in \citet{frei94a} is incorrect; in fact, $k(z)\equiv K(z)$.}
+$k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
+exactly equivalent to our $K(z)$.}.
 
@@
 particular as we try to quantify the evolution of galaxies, we must
-take a more sophisticated approach to approximating fixed-frame
+take a more sophisticated approach to approximating fixed frame
 observations of galaxies.
@@
 papers. Our approach is equivalent (nearly identical) to the
-photometric metric redshift estimation methods of \citet{csabai00a},
+photometric redshift estimation methods of \citet{csabai00a},
 except we use slightly different coordinate systems for our SED
@@
 groundwork for upcoming papers which will rely heavily on the
-reliability of the fixed-frame magnitudes determined here.  Because
+reliability of the fixed frame magnitudes determined here.  Because
 the SDSS is one of the largest astronomical surveys to date, our
@@
 We refer throughout this paper to $AB$ magnitudes, meaning that the
-magnitudes are designed such that ():
+magnitudes are designed such that (\citealt{fukugita96a}):
 \begin{eqnarray}
@@
 Section \ref{sedfit} describes our method of fitting galaxy SEDs to
-broad-band photometry and how to calculate $K$-corrections. Section
+broad band photometry and how to calculate $K$-corrections. Section
 \ref{data} applies the method to galaxies in the SDSS, showing that
@@
 
-\section{Calculating Fixed-frame Galaxy Magnitudes}
+\section{Calculating Fixed Frame Galaxy Magnitudes}
 \label{sedfit}
 
-First, we describe how we reconstruct galaxy SEDs from broad-band
+First, we describe how we reconstruct galaxy SEDs from broad band
 magnitude measurements. Second, we describe how we convert these SEDs
-into estimates of fixed-frame galaxy magnitudes.
+into estimates of fixed frame galaxy magnitudes.
 
@@
 
-Our task is to recover a model for the galaxy SED from broad-band
+Our task is to recover a model for the galaxy SED from broad band
 photometric measurements. Since a set of broad band galaxy fluxes does
@@
 are known to have significant structure over wavelength ranges small
-compared to our bandpass, this task is an ill-posed, inverse problem:
-given any set of observed galaxy colors, there is not a unique SED to
-which those colors correspond, even in the absence of errors. However,
-we are not completely ignorant about the forms which galaxy SEDs take,
-so we can attempt to use what we know about galaxy SEDs to simplify
-our task.  The method described here for doing so follows closely the
-methods of estimating photometric redshifts used by \citet{csabai00a}
-and \citet{budavari00a}. It is designed to take advantage of what we
-already understand about galaxy spectral energy distributions.
+compared to our bandpass, the inverse problem of reconstructing SEDs
+from broad band fluxes is ill-posed.  However, we are not completely
+ignorant about the forms which galaxy SEDs take, so we can attempt to
+use what we know about galaxy SEDs to simplify our task.  The method
+described here for doing so follows closely the method of estimating
+photometric redshifts used by \citet{csabai00a}.  It is designed to
+take advantage of what we already understand about galaxy spectral
+energy distributions.
 %In addition, as implemented here, the method is
@@
 Begin with an SED space defined by $N_b$ template galaxy SEDs
-$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), where $N_b$ is
+$\vv{v}_i(\lambda)$ (for example from \citealt{bruzual93a}), where $N_b$ is
 large. In principle, this should be a complete set of galaxy SEDs if
@@
 spectra, but instead over a smaller range which is within observable
-wavelengths over most of the redshift range of the
-sample. \footnote{Were we to define orthogonality over the full range
-of $\vv{v}(\lambda)$, much of which is unobservable, the geometry of
-the distribution of galaxies in this space would be more complicated.
+wavelengths over most of the redshift range of the sample.\footnote{
 If the template spectra were identical in the subrange chosen and only
-differed outside of it, our choice would cause computational
-problems, but in fact the spectra are all independent in the
-restricted, as well as the full, wavelength range.}  We then define
-the basis of the SED space such that the $N_b$ template SEDs,
-$\vv{v}_i$, are linear combinations of $N_b$ basis spectra $\vv{b}_i$
-and that $\vv{b}_i \cdot \vv{b}_j = \delta_{ij}$. These conditions do
-not fully specify the $\vv{b}_i$, which naturally may be rotated
-arbitrarily within our $N_b$-dimensional subspace of spectral space.
+differed outside of it, our choice would cause computational problems,
+but in fact the spectra are all independent in the restricted, as well
+as the full, wavelength range.}  We then define the basis of the SED
+space such that the $N_b$ template SEDs, $\vv{v}_i$, are linear
+combinations of $N_b$ basis spectra $\vv{b}_i$ and that $\vv{b}_i
+\cdot \vv{b}_j = \delta_{ij}$. These conditions do not fully specify
+the $\vv{b}_i$, which naturally may be rotated arbitrarily within our
+$N_b$-dimensional subspace of spectral space.
 
@@
 cannot determine an individual galaxy's position in this
-$N_b$-dimensional space from its broad-band colors alone.  However, as
+$N_b$-dimensional space from its broad band colors alone.  However, as
 outlined by \citet{csabai00a}, one can find a lower dimensional space
@@
 space defined by the template SEDs. Thus, we transform the axes such
-that $a_i/a0 = 0$ is the mean of the distribution and rotate the axes
+that $a_i/a_0 = 0$ is the mean of the distribution and rotate the axes
 $\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
@@
 
-In this way, we can reconstruct SEDs based on the broad-band
+In this way, we can reconstruct SEDs based on the broad band
 magnitudes of galaxies. From these reconstructed SEDs one can estimate
@@
 
-\subsection{Applying Direct Constraints on $a_i/a0$}
+\subsection{Applying Direct Constraints on $a_i/a_0$}
 \label{direct}
@@
 constrain these galaxies to have reasonable SEDs, in the sense that
-the $a_i/a_0$ ought not to be too different than for well-measured
-galaxies. Here we describe how to implement direct constraints on the
-coefficients $a_i/a_0$ which describe the galaxy SED.
-
-We will assume that any constraints can be expressed as a combination
-of gaussian constraints with means $a_{c,j}/a_{c,0}$, variance matrices
-$V_{c,jl}$, weighted by amplitudes $A_c$.  We calculate a
-preliminary value of $a_{c,0}$, assuming $a_i = 0$. Then we constrain
-our fit by adding a term to $\chi^2$ for each constraint:
+the $a_i/a_0$ ought not to be too different than those for
+well-measured galaxies. We can assume a gaussian prior probability
+distribution with a mean $\avg{a_{c,j}/a_{c,0}}$, a variance matrix
+$V_{c,jl}$, and a weight $A_c$.  We calculate a preliminary value of
+${\hat{a}}_{c,0}$, assuming $a_i = 0$. Then we constrain our fit by
+adding a term to $\chi^2$ as follows:
 \begin{equation}
-\sum_j \sum_l V^{-1}_{jl} (a_l/a_{c,0}-a_{c,l}/a_{c,0})
-(a_j/a_{c,0}-a_{c,j}/a_{c,0})  
+\sum_j \sum_l A_c V^{-1}_{jl}
+\left(\frac{a_l}{{\hat{a}}_{c,0}}-\Avg{\frac{a_{c,l}}{a_{c,0}}}\right)
+\left(\frac{a_j}{{\hat{a}}_{c,0}}-\Avg{\frac{a_{c,j}}{a_{c,0}}}\right)
 \end{equation}
 The resulting $\chi^2$ can still be minimized linearly, so we have
-sacrificed very little in speed.
-
-Note that by using this constraint, one can calculate $K$-corrections
-even for galaxies with only one or two bands measured, and still
-obtain sensible results.
+sacrificed very little in speed.  Note that with such a constraint,
+one can calculate $K$-corrections even for galaxies with only one or
+two bands measured, and still obtain sensible results.
 
@@
 
-Given AB galaxy magnitudes $^{z}m_{AB}$ and errors for a galaxy at
+Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
 redshift $z$, one can use the method of the previous subsection to
@@
 
-Given the reconstructed magnitudes, one can calculate the fixed-frame
+Given the reconstructed magnitudes, one can calculate the fixed frame
 magnitudes in one of two ways:
@@
 \item Simply use $\band{z_0}{m_{\mathrm{rec}}}$, the reconstructed
-magnitudes at redshift $z_0$, as the fixed-frame magnitudes.
+magnitudes at redshift $z_0$, as the fixed frame magnitudes.
 \item Use the difference
@@
 magnitude the cosmological distance modulus (that is, the luminosity
-distance-squared law, as tabulated by, {\it e.g.}  \citealt{hogg99a}).
+distance-squared law, as tabulated by, {\it e.g.},  \citealt{hogg99a}).
 
@@
 objective, it is not {\it always} necessary or appropriate. Because
-$K$-corrections are inherently uncertain (the broad-band magnitudes
+$K$-corrections are inherently uncertain (the broad band magnitudes
 do not uniquely determine the SED) they should be avoided or
@@
 perhaps be wise to perform the separation not on the $K$-corrected
-colors of the galaxies but on the median color of, say, $M_\ast$
+colors of the galaxies but on the median color of $M_\ast$
 galaxies as a function of redshift. Similarly, in situation where
@@
 model of the galaxy SED, we do not correct to bandpasses with simpler
-shapes (say, top hats). The reason is that we want the effective
+shapes (for example, top hats). The reason is that we want the effective
 bandpasses to actually have been observed for some galaxies in the
@@
 there are inherent uncertainties in the $K$-corrections. Nevertheless,
-we note that synthesizing such magnitudes may be appropriate in
+we note that synthesizing top hat magnitudes may be appropriate in
 certain situations, as they have the virtue (similarly to the $AB$
@@
 any galaxy.  For our purposes, it is more important to have high
-signal-to-noise measurements of galaxy colors than to probe exactly
+signal-to-noise measurements of galaxy colors than to measure exactly
 the same regions of galaxies at low and high redshift.  We
@@
 \ref{sedfit} to SDSS data, we take a subsample of the data consisting
-of around 25,000 objects in the range $0.0<z<0.5$. The sample includes
+of around 30,000 objects in the range $0.0<z<0.5$. The sample includes
 both the main sample and the LRGs; we sample these spectra such that
@@
 templates. However, we still want to determine a best-fit SED for each
-object, once the templates are determined. Thus, we constrain all the
-SED fits to galaxies using the method described in Section
-\ref{direct}. As a constraint, we use the covariance matrix and mean
-value (essentially zero, by design) of the coefficients of the
-well-measured set of galaxies; we multiply all the elements of the
-covariance matrix by a factor of our to weaken our constraint ((which
+object, once the templates are determined. Thus, after the templates
+have been determined, we constrain all the SED fits to all galaxies
+using the method described in Section \ref{direct}. As a constraint,
+we use the covariance matrix and mean value (essentially zero, by
+design) of the coefficients of the well-measured set of galaxies used
+to determine the templates; we multiply all the elements of the
+covariance matrix by a factor of four to weaken our constraint (which
 is appropriate, since the actual distribution of coefficients is not
-well described by a single gaussian). For this purpose, we replace a
-``bad'' magnitude based on an adjacent (usually the redder) band and
-the average color in those bands of objects at that redshift. We give
-these measurements a large error bar, of 0.8 magnitudes. Thus, we
-account for the missing information by simply requiring that the
-object SED have ``reasonable'' properties.
+well described by a single gaussian).  Thus, we account for the
+missing information by simply requiring that the object SED have
+``reasonable'' properties.
 
@@
 the applications of these SED determinations is the distribution of
-galaxy colors in fixed-frame magnitudes, we don't want to artificially
+galaxy colors in fixed frame magnitudes, we don't want to artificially
 reduce the dimensionality of the color space to only two.
@@
 To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
-corresponding to the direction represented by the vector $
-\vv{e}_2(\lambda) + 0.7 \vv{e}_3$. We overplot the \band{0.3}{g} and
-\band{0.3}{r} bands. This direction has a strong peak in the gap
-between the two bands at $z=0.3$.  This difficulty is why, in Section
-\ref{data}, we exclude the regime around $z\sim 0.3$; otherwise the
-templates became distracted by the degeneracies in this range.
+corresponding to the direction represented by the vector $ \vv{e}_2 +
+0.7 \vv{e}_3$, for our best fit templates. We overplot the
+\band{0.3}{g} and \band{0.3}{r} bands. This direction has a strong
+peak in the gap between the two bands at $z=0.3$.  This difficulty is
+why, in Section \ref{data}, we exclude the regime around $z\sim 0.3$;
+otherwise the templates became distracted by the degeneracies in this
+range.
 
@@
 subspace. In addition, we have tried removing the increased resolution
-of the template SEDs around the 4000 \AA break. Neither of these tests
+of the template SEDs around the 4000 \AA\ break. Neither of these tests
 yielded better results.
@@
 
-However, we have not tackled this approach yet. As it happens, the the
+However, we have not tackled this approach yet. As it happens, the
 direct constraint on the coefficients described in the previous
 section is sufficient to suppress the $g$/$r$ gap problem sufficiently
-for many purposes, as long as one $K$-corrections the magnitudes to
+for many purposes, as long as one $K$-corrected the magnitudes to
 $z=0.3$. In this way, one minimizes the $K$-corrections for the
@@
 
-Remember, you really only need to worry about this effect in the SDSS
-when you are dealing with the LRGs, for galaxies in the range about
-$0.27 < z < 0.33$. The Main Sample results are not significantly
-affected at all.  Furthermore, one can still observe a galaxy at
-$z=0.1$ and reliably infer what it would look like at $z=0.3$; it is
-only the reverse process which is difficult.
+One really only needs to worry about this effect in the SDSS when you
+are dealing with the LRGs, for galaxies in the range about $0.27 < z <
+0.33$. The Main Sample results are not significantly affected at all.
+Furthermore, one can still observe a galaxy at $z=0.1$ and reliably
+infer what it would look like at $z=0.3$; it is only the reverse
+process which is difficult.
 
@@
 edges should not be taken too seriously. In addition, note the small
-spur extending from upper left to lower right through the right edge
+spur extending from lower left to upper right through the right edge
 of the ($a_3/a_0$)-($a_2/a_0$) plane. Many of the objects in this spur are
@@
 
-An important test of the reasonableness of the fits is to check that
-for a fixed type of galaxy, the distribution of the fixed frame colors
+An important test of the consistency of the fits is to check that for
+a fixed type of galaxy, the distribution of the fixed frame colors
 depends only weakly on redshift. Because the galaxies shown in Figures
 \ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
-selected --- some are Main Sample galaxies, some are Luminous Red
-Galaxies, and some are selected from the photometric redshift plates
---- we will split our sample into two well-defined sets of objects to
-perform this test. First, we choose a set of Main Sample galaxies in
-the luminosity range $-21.5 < M_{\band{0.1}{r}} < -21.2$. We reconstruct
-the four colors $^{0.1}(u-g)$, $^{0.1}(g-r)$, $^{0.1}(r-i)$, and
-$^{0.1}(i-z)$ for these galaxies and plot the reconstructions as a
-function of redshift in left-hand panels of Figure
-\ref{main_colors_plot}. The plots show a general consistency in the
-fixed-frame color distribution of these objects with redshift. The
-right-hand panels show the distribution for redshifts $z<0.13$ (solid
-histogram) and $z>0.13$ (dotted histogram) . Small changes are
-discernible in the distributions, mostly attributable to the increased
-errors at higher redshift. Note that for $z<0.1$ the $^{0.1}(u-g)$
-color depends on an extrapolation of the SED in the blue, while for
-$z>0.1$ the $^{0.1}(i-z)$ color depends on an extrapolation of the SED
+selected --- some are Main Sample galaxies, some are LRGs, and some
+are selected from the photometric redshift plates --- we will split
+our sample into two well-defined sets of objects to perform this
+test. First, we choose a set of Main Sample galaxies in the luminosity
+range $-21.5 < M_{\band{0.1}{r}} < -21.2$. Figure
+\ref{main_colors_plot.z} shows the observed-frame colors of all the
+galaxies as a function of redshift.  The right-hand panels show the
+distribution for redshifts $0.05<z<0.13$ (solid histogram) and
+$0.1<z<0.173$ (dotted histogram). A trend exists in all colors, most
+strongly in $\band{z}{(g-r)}$. Figure \ref{main_colors_plot} shows the
+$K$-corrected colors $\band{0.1}{(u-g)}$, $\band{0.1}{(g-r)}$,
+$\band{0.1}{(r-i)}$, and $\band{0.1}{(i-z)}$ in the same manner.  The
+plots show a general consistency in the fixed frame color distribution
+of these objects with redshift.  The right-hand panels show the
+distribution for redshifts $0.05<z<0.13$ (solid histogram) and
+$0.1<z<0.173$ (dotted histogram) .  Small changes are discernible in
+the distributions, mostly attributable to the increased errors at
+higher redshift. Note that for $z<0.1$ the $\band{0.1}{(u-g)}$ color
+depends on an extrapolation of the SED in the blue, while for $z>0.1$
+the $\band{0.1}{(i-z)}$ color depends on an extrapolation of the SED
 in the red.
@@
 Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.3}{r}}
-< -22.5$. For this sample we reconstruct the colors at $z=0.3$.
-Again, for $z<0.3$ the $^{0.3}(u-g)$ color relies on an extrapolation,
-which is probably responsible for the scatter in that color at low
-redshifts. Otherwise, the distribution of colors is remarkably
-constant. $^{0.3}(g-r)$ (which is very similar to $^{0.0}(u-g)$)
+< -22.5$. Figure \ref{lrg_colors_plot.z} shows the observed colors as
+a functino of redshift for these objects. Again, there is a strong
+redshift dependence. Figure \ref{lrg_colors_plot} shows the colors
+$K$-corrected to $z=0.3$, which appear approximately constant with
+redshift.  
+%Again, for $z<0.3$ the $\band{{0.3}}{(u-g)}$ color relies
+%on an extrapolation, which is probably responsible for the scatter in
+%that color at low redshifts. Otherwise, the distribution of colors is
+%remarkably constant. 
+$\band{0.3}{(g-r)}$ (which is very similar to $\band{0.0}{(u-g)}$)
 experiences a blueward shift of about 0.1 magnitudes between $z=0.1$
@@
 photometry itself, seem physically reasonable, and are consistent over
-the range of redshifts we consider ($0.0 < z < 0.5$). The only
-remaining problem is the aforementioned difficulties at $z=0.3$,
-discussed in more detail in the next subsection.
+the range of redshifts we consider ($0.0 < z < 0.5$).
 
@@
 
-An important test of our method is to synthesize broad-band photometry
+An important test of our method is to synthesize broad band photometry
 from spectra, and then try to recover the $K$-corrections in a case
@@
 The agreement is now very good, suggesting that our method can
-indeed recover the correct $K$-corrections based only on broad-band
+indeed recover the correct $K$-corrections based only on broad band
 magnitudes.
@@
 To test how robust these $K$-correction results are to our model
-assumptions, we compare them to other methods of calculating
-fixed-frame galaxy magnitudes.  An extremely simple method is to
-calculate the flux in any desired bandpass by fitting a power-law
-slope and amplitude to the fluxes in the two adjacent bandpasses
-(extrapolating when necessary). Figure \ref{comparecibreak}
-shows the differences in the $K$-corrections inferred from this method
-and those inferred from the method of Figure
-\ref{k_kcorrect_plot}, as a function of redshift. $r$, $i$, and
-$z$ are all reasonably similar in either method; $u$ and $g$, however,
-have distinct trends with redshift, mostly due to the non-power law
-nature of the template galaxy spectra (and probably actual galaxy
-spectra) in this regime. To show this fact, we perform a similar
-power-law fit, only this time including a break in the spectrum at
-4000\AA. We use the $u$-$g$ color to fit the break, assuming that the
-slope blueward of 4000\AA\ is $f(\lambda)\propto \lambda^{2}$.  Figure
-\ref{comparecibreak} shows the results of this fit; the
-redshift trend in $g$ is greatly reduced, as is the trend in $u$, but
-a large amount of scatter remains in the $u$ band. This results from
-the fact that you can {\it either} fit the slope of the SED below the
-4000 \AA\ break {\it or} the size of the 4000 \AA\ break itself; it is
-not possible with this data to constrain both in an individual
-spectrum, which is a fundamental limitation of our determination of 
-fixed frame $u$-band magnitudes.
+assumptions, we compare them to other methods of calculating fixed
+frame galaxy magnitudes.  An extremely simple method is to calculate
+the flux in any desired bandpass by fitting a power-law slope and
+amplitude to the fluxes in the two adjacent bandpasses (extrapolating
+when necessary). Figure \ref{comparecibreak} shows the differences in
+the $K$-corrections inferred from this method and those inferred from
+the method of Figure \ref{k_kcorrect_plot}, as a function of
+redshift. $r$, $i$, and $z$ are all reasonably similar in either
+method; $u$ and $g$, however, have distinct trends with redshift,
+mostly due to the non-power law nature of the template galaxy spectra
+(and probably actual galaxy spectra) in this regime. To show this
+fact, we perform a similar power-law fit, only this time including a
+break in the spectrum at 4000\AA. We use the $u$-$g$ color to fit the
+break, assuming that the slope blueward of 4000\AA\ is
+$f(\lambda)\propto \lambda^{2}$.  Figure \ref{comparecibreak} shows
+the results of this fit; the redshift trend in $g$ is greatly reduced,
+as is the trend in $u$, but a large amount of scatter remains in the
+$u$ band. This results from the fact that you can {\it either} fit the
+slope of the SED below the 4000 \AA\ break {\it or} the size of the
+4000 \AA\ break itself; it is not possible with this data to constrain
+both in an individual spectrum, which is a limitation of our
+determination of fixed frame $u$-band magnitudes.
 
 Finally, it is possible to use the galaxy spectra obtained with the
-spectrograph to estimate the $K$-correction for each object. 
-However, this procedure requires trusting the
-spectrophotometry over a wide wavelength range. In addition, the
-region of the galaxy which the spectra probe is not completely well
-defined. The small scale wavelength features are set by the fiber
-aperture ($3''$ in diameter) while large scale wavelength features are
-constrained by a short ``smear exposure'' which probes a larger
-area. The smear exposure corrections are clearly the correct approach
-for point sources, but for galaxies it does cause the small scale and
-large scale features of the spectrum to be determined by somewhat
-different regions of the galaxy.
+spectrograph to estimate the $K$-correction for each object.  However,
+this procedure requires trusting the spectrophotometry over a wide
+wavelength range. In addition, the region of the galaxy which the
+spectra cover is not completely well defined. The small scale
+wavelength features are set by the fiber aperture ($3''$ in diameter)
+while large scale wavelength features are constrained by a short
+``smear exposure'' which covers a larger area. The smear exposure
+corrections are clearly the correct approach for point sources, but
+for galaxies it causes the small scale and large scale features of the
+spectrum to be determined by somewhat different regions of the galaxy.
 
@@
 the spectra to those of Figure \ref{k_kcorrect_plot}, finding that
-they are extremely similar. There is a trend within redshift in the
+they are extremely similar. There is a trend with redshift in the
 $g$-band, which is due to the fact that on average the $g-r$ color is
 about 0.1 magnitudes redder in the fiber aperture than in the
-Petrosian aperture. The implication is that the $K$-corrections {\it
-are} stronger for the aperture which corresponds to the spectrum than
-for the Petrosian aperture. This indicates that for work which
-requires precision, it is best to avoid using the $K$-corrections
-determined from the spectra. (Note the in the second paragraph of this
-section we used the SDSS spectra as {\it example} spectra to validate
-the method; in this paragraph we have evaluate whether to rely on the
+Petrosian aperture. The implication is that the $K$-corrections are
+stronger for the aperture which corresponds to the spectrum than for
+the Petrosian aperture. This indicates that for work which requires
+precision, it is best to avoid using the $K$-corrections determined
+from the spectra. (Note that in the second paragraph of this section
+we used the SDSS spectra as {\it example} spectra to validate the
+method; in this paragraph we have evaluated whether to rely on such
 spectra to calculate $K$-corrections).
@@
 We have presented a method and an implementation for estimating galaxy
-SEDs for the purpose of calculating fixed-frame galaxy magnitudes over
+SEDs for the purpose of calculating fixed frame galaxy magnitudes over
 a range of redshifts. We have demonstrated that it gives sensible and
@@
 example, you can calculate photometric redshifts using the method
-described here. {\tt kcorrect v1\_2} has a simple (and very fast)
+described here. {\tt kcorrect \kversion} has a simple (and very fast)
 photometric redshift estimator with very tight core of residuals

### 2002-05-15 docs/paper/figures.tex
final edit before submit
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@105 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 bands in the SDSS, as a function of observed wavelength, as measured
-by Mamoru Doi. A 4 gigayear old instantaneous burst using the models
+by Mamoru Doi. A 4 Gyr old instantaneous burst using the models
 of \citet{bruzual93a} (and observed at $z=0$) is shown for reference.}
-%at different redshifts. A 4 gigayear old instantaneous burst using the
+%at different redshifts. A 4 Gyear old instantaneous burst using the
 %models of \citet{bruzual93a} is shown for reference. {\it Top panel:}
@@
 photometry for a random subsample consisting of 10,000 of the SDSS
-galaxies. $a_0$ is linearly proportional to the flux between $3500\AA$
-and $7000\AA$, while $a_1$, $a_2$, and $a_3$ contribute no flux in
+galaxies. $a_0$ is linearly proportional to the flux between 3500 \AA
+and 7500 \AA, while $a_1$, $a_2$, and $a_3$ contribute no flux in
 this range. Thus, the ratios $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$
 describe the spectral type of the galaxy. $a_1/a_0$ is the most
-variable parameter and thus is the better separator of galaxy
+variable parameter and thus is the best separator of galaxy
 type. The spur extending from the upper left to the lower right from
@@
 color at high and low redshift within the volume-limited
-subsample. The observed colors clearly depend strongly on redshift,
-even for the $\band{0.1}{u}$ band, where the low-redshift end is an
-extrapolation of the data.}
+subsample. The observed colors clearly depend strongly on redshift.}
 \end{figure}
@@
 well-behaved, we do not recommend using these extrapolated results for
-scientific purposes. Note the degeneracy at $z=0.3$.}
+scientific purposes. }
 \end{figure}
@@
 \plotone{comparecibreak.ps}
-\caption{\label{comparecibreak} Same as Figure \ref{comparecibreak},
+\caption{\label{comparecibreak} Same as Figure \ref{compareci},
 now comparing to a method of interpolating the bandpasses using
-power-laws, and fitting for the 4000 \AA break. The systematic trends
+power-laws, and fitting for the 4000 \AA\ break. The systematic trends
 in \band{0.1}{u} and \band{0.1}{g} are gone (though there is

### 2002-05-15 docs/paper/paper.tex
final edit before submit
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@105 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 %Andrew J. Connolly\altaffilmark{\ref{Pitt}},
+J.~Brinkmann\altaffilmark{\ref{APO}},
 Istv\'an Csabai\altaffilmark{\ref{JHU}},
@@
 Daniel Eisenstein\altaffilmark{\ref{Arizona}},
+Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
 James E. Gunn\altaffilmark{\ref{Princeton}}
@@
 %Mariangela Bernardi\altaffilmark{\ref{Chicago}},
-%J. Brinkmann\altaffilmark{\ref{APO}},
 %Robert J. Brunner\altaffilmark{\ref{Caltech}},
@@
 %Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
-%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
 %G. S. Hennessy\altaffilmark{\ref{USNO}},
@@
 %\label{Ohio}}
-%\altaffiltext{\theaddress}{
-%\stepcounter{address}
-%Apache Point Observatory,
-%2001 Apache Point Road,
-%P.O. Box 59, Sunspot, NM 88349-0059
-%\label{APO}}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+Apache Point Observatory,
+2001 Apache Point Road,
+P.O. Box 59, Sunspot, NM 88349-0059
+\label{APO}}
 %\altaffiltext{\theaddress}{
@@
 %\label{Berkeley}}
-%\altaffiltext{\theaddress}{
-%\stepcounter{address}
-%Institute for Cosmic Ray Research, University of
-%Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
-%\label{CosmicRay}}
-%\altaffiltext{\theaddress}{
-%\stepcounter{address}
-%Institute for Advanced Study, Olden Lane,
-%Princeton, NJ 08540
-%\label{IAS}}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+Institute for Cosmic Ray Research, University of
+Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
+\label{CosmicRay}}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+Institute for Advanced Study, Olden Lane,
+Princeton, NJ 08540
+\label{IAS}}
 %\altaffiltext{\theaddress}{
@@
 photometry. Expressed in the appropriate coordinate system, the
-distribution of the coefficients multiplying the templates is planar,
+locus of the coefficients multiplying the templates is planar,
 and in fact nearly linear. The resulting reconstructed SEDs can be
@@
 In the future, observations of galaxies (and indeed of any
-astronomical sources) will be performed using integral field
-spectrographs or equivalent devices which combine high spatial
-resolution and high spectral resolution. By that time, the phenomena
-of broad band filters and the quaint terminology surrounding their
-usage --- magnitudes, $K$-corrections, color terms, {\it etc.} ---
-will have long since been forgotten. However, today, virtually all
-observations of galaxies are made through such filters, and special
-care has to be taken to recover knowledge of galaxy spectral energy
-distributions (SEDs) from these observations.  Reconstructing galaxy
-SEDs is nontrivial because SEDs of galaxies contain important
-information on scales considerably smaller than the width of typical
-broad band filters. Furthermore, SEDs of distant galaxies are
-redshifted such that the more distant the galaxy, the further the
-observed bandpass is blueshifted relative to the rest-frame spectral
-energy distribution of the object observed.  This paper focuses on a
-solution to these problems which (in the optical wavelength regime) is
-fast, robust, and consistent over a large range of redshifts.
+astronomical sources) will be performed using devices which combine
+high spatial resolution and high spectral resolution. By that time,
+the phenomena of broad band filters and the quaint terminology
+surrounding their usage --- magnitudes, $K$-corrections, color terms,
+{\it etc.} --- will have long since been forgotten. However, today,
+virtually all observations of galaxies are made through such filters,
+and special care has to be taken to recover knowledge of galaxy
+spectral energy distributions (SEDs) from these observations.
+Reconstructing galaxy SEDs is nontrivial because SEDs of galaxies
+contain important information on scales considerably smaller than the
+width of typical broad band filters. Furthermore, SEDs of distant
+galaxies are redshifted such that the more distant the galaxy, the
+further the observed bandpass is blueshifted relative to the
+rest-frame spectral energy distribution of the object observed.  This
+paper focuses on a solution to these problems which (in the optical
+wavelength regime) is fast, robust, and consistent over a large range
+of redshifts.
 
@@
 $k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
-exactly equivalent to our $K(z)$.}.
+exactly equivalent to our $K(z)$.}
 
@@
 
-We refer throughout this paper to $AB$ magnitudes, meaning that the
-magnitudes are designed such that (\citealt{fukugita96a}):
+We refer throughout this paper to $AB$ magnitudes, first defined 
+by \citet{oke83a} such that:
 \begin{eqnarray}
@@
 compared to our bandpass, the inverse problem of reconstructing SEDs
-from broad band fluxes is ill-posed.  However, we are not completely
-ignorant about the forms which galaxy SEDs take, so we can attempt to
-use what we know about galaxy SEDs to simplify our task.  The method
-described here for doing so follows closely the method of estimating
-photometric redshifts used by \citet{csabai00a}.  It is designed to
-take advantage of what we already understand about galaxy spectral
-energy distributions.
+from broad band fluxes is ill-posed. The task is made yet more
+difficult by the fact that the separation of the fiters are large
+compared to their widths; the SEDs are thus not well-sampled in the
+Nyquist sense of that term. However, we are not completely ignorant
+about the forms which galaxy SEDs take, so we can attempt to use what
+we know about galaxy SEDs to appropriately regularize our fits.  The
+method described here for doing so follows closely the method of
+estimating photometric redshifts used by \citet{csabai00a}.  It is
+designed to take advantage of what we already understand about galaxy
+spectral energy distributions.
 %In addition, as implemented here, the method is
@@
 the $\band{0.22}{g}$-band magnitude of the galaxy. The goal of
-``$K$-corrections'' is to convert the $ugriz$ bands at redshift $z$
+``$K$-correction'' is to convert the $ugriz$ bands at redshift $z$
 for each galaxy to $ugriz$ bands at some fixed redshift $z_0$.
@@
 calculations described here, along with eigentemplates and filter
-curves is publicly available on the World Wide Web at {\tt
+curves, is publicly available on the World Wide Web at {\tt
 http://physics.nyu.edu/\~\ mb144/kcorrect}.  The whole of the code can
@@
 are described in the Early Data Release (EDR;
-\citealt{stoughton01a}).
+\citealt{stoughton02a}).
 
@@
 most of our sample, did not change substantially throughout these
-versions. The magnitudes are calibrated approximately in the $AB$
-system, described in Section \ref{motivation}.
+versions. As described in \citet{smith02a}, magnitudes are calibrated to a
+standard star network approximately in the $AB$ system, described in
+Section \ref{motivation}.
 
@@
 for objects with unusual spectra, such as certain types of stars and
-QSOs.}.
+QSOs.}
 
@@
 the better fitting model (pure exponential or pure de Vaucouleurs) to
-the galaxy image in the $r$-band. In \citet{stoughton01a} we are
+the galaxy image in the $r$-band. In \citet{stoughton02a} we are
 explicitly warned not to mix Petrosian and model magnitudes in a
@@
 missing information by simply requiring that the object SED have
-``reasonable'' properties.
+``reasonable'' properties. The constraint does not affect the results
+for galaxies which have well-measured magnitudes in all bands.
 
@@
 $Z=0.004$, all assuming a Salpeter Initial Mass Function. Since we
-cannot recover information about the SED below a certain scale, we
-smooth most of the wavelength regime of the templates using a Gaussian
-with a standard deviation of $\sigma = 300$\AA; for the region of the
-spectrum which contains the sharpest gradients, around 4000\AA, we
-smooth only with $\sigma = 150$\AA. None of our results change
-dramatically if we vary our smoothing procedure or if we add reddened
-templates to our allowed SED space.
+cannot recover information about the SED below a certain wavelength
+scale, we smooth most of the wavelength regime of the templates using
+a Gaussian with a standard deviation of $\sigma = 300$\AA; for the
+region of the spectrum which contains the sharpest gradients, around
+4000\AA, we smooth only with $\sigma = 150$\AA. None of our results
+change dramatically if we vary our smoothing procedure or if we add
+reddened templates to our allowed SED space.
 
@@
 direct constraint on the coefficients described in the previous
-section is sufficient to suppress the $g$/$r$ gap problem sufficiently
+section is sufficient to suppress the $g$/$r$ gap problem 
 for many purposes, as long as one $K$-corrected the magnitudes to
@@
 panel, three spectra taken from a one-dimensional sequence along the
-galay locus are shown, showing that the spectra become progressively
+galaxy locus are shown, showing that the spectra become progressively
 bluer along that sequence. Note that constraints on the SED become
 poor at the bluest and reddest edges of this diagram (the peak of the
-$z$-band response is at 8700\AA). Therefore, the odd behavior near the
+$z$-band response is at 8700~\AA restframe for a galaxy observed at
+$z=0$). Therefore, the odd behavior near the
 edges should not be taken too seriously. In addition, note the small
@@
 galaxies as a function of redshift.  The right-hand panels show the
-distribution for redshifts $0.05<z<0.13$ (solid histogram) and
-$0.1<z<0.173$ (dotted histogram). A trend exists in all colors, most
+distribution for redshifts $0.05<z<0.10$ (solid histogram) and
+$0.10<z<0.17$ (dotted histogram). A trend exists in all colors, most
 strongly in $\band{z}{(g-r)}$. Figure \ref{main_colors_plot} shows the
@@
 plots show a general consistency in the fixed frame color distribution
-of these objects with redshift.  The right-hand panels show the
-distribution for redshifts $0.05<z<0.13$ (solid histogram) and
-$0.1<z<0.173$ (dotted histogram) .  Small changes are discernible in
-the distributions, mostly attributable to the increased errors at
-higher redshift. Note that for $z<0.1$ the $\band{0.1}{(u-g)}$ color
-depends on an extrapolation of the SED in the blue, while for $z>0.1$
-the $\band{0.1}{(i-z)}$ color depends on an extrapolation of the SED
-in the red.
-
-Second, we choose a set of Cut I LRGs with $-22.8 < M_{\band{0.3}{r}}
-< -22.5$. Figure \ref{lrg_colors_plot.z} shows the observed colors as
-a functino of redshift for these objects. Again, there is a strong
-redshift dependence. Figure \ref{lrg_colors_plot} shows the colors
+of these objects with redshift.  Small changes are discernible in the
+distributions, mostly attributable to the increased errors at higher
+redshift. Note that for $z<0.1$ the $\band{0.1}{(u-g)}$ color depends
+on an extrapolation of the SED in the blue, while for $z>0.1$ the
+$\band{0.1}{(i-z)}$ color depends on an extrapolation of the SED in
+the red.
+
+Second, we choose a set of Cut I LRGs (see \citealt{eisenstein01a} for
+details) with $-22.8 < M_{\band{0.3}{r}} < -22.5$. Figure
+\ref{lrg_colors_plot.z} shows the observed colors as a function of
+redshift for these objects. Again, there is a strong redshift
+dependence. Figure \ref{lrg_colors_plot} shows the colors
 $K$-corrected to $z=0.3$, which appear approximately constant with
@@
 \end{equation}
-The agreement is now very good, suggesting that our method can
+The agreement is very good, suggesting that our method can
 indeed recover the correct $K$-corrections based only on broad band
@@
 amplitude to the fluxes in the two adjacent bandpasses (extrapolating
-when necessary). Figure \ref{comparecibreak} shows the differences in
+when necessary). Figure \ref{compareci} shows the differences in
 the $K$-corrections inferred from this method and those inferred from
@@
 fact, we perform a similar power-law fit, only this time including a
-break in the spectrum at 4000\AA. We use the $u$-$g$ color to fit the
-break, assuming that the slope blueward of 4000\AA\ is
+break in the spectrum at 4000 \AA. We use the $u$-$g$ color to fit the
+break, assuming that the slope blueward of 4000 \AA\ is
 $f(\lambda)\propto \lambda^{2}$.  Figure \ref{comparecibreak} shows
@@
 
+MB is partially supported by {\bf NYU grants}.
 MB was supported at the beginning of this work by the DOE and NASA
@@
 
-The Sloan Digital Sky Survey (SDSS) is a joint project of The
+Funding for the creation and distribution of the SDSS Archive has been
+provided by the Alfred P. Sloan Foundation, the Participating
+Institutions, the National Aeronautics and Space Administration, the
+National Science Foundation, the U.S. Department of Energy, the
+Japanese Monbukagakusho, and the Max Planck Society. The SDSS Web site
+is {\tt http://www.sdss.org/}.
+
+The SDSS is managed by the Astrophysical Research Consortium (ARC) for
+the Participating Institutions. The Participating Institutions are The
 University of Chicago, Fermilab, the Institute for Advanced Study, the
-Japan Participation Group, the Johns Hopkins University, the
-Max-Planck-Institute for Astronomy, New Mexico State University,
-Princeton University, the United States Naval Observatory, and the
-University of Washington. Apache Point Observatory, site of the SDSS
-telescopes, is operated by the Astrophysical Research Consortium
-(ARC).  Funding for the project has been provided by the Alfred
-P.~Sloan Foundation, the SDSS member institutions, the National
-Aeronautics and Space Administration, the National Science Foundation,
-the U.~S.~Department of Energy, Monbusho, and the Max Planck
-Society. The SDSS Web site is http://www.sdss.org/.
+Japan Participation Group, The Johns Hopkins University, Los Alamos
+National Laboratory, the Max-Planck-Institute for Astronomy (MPIA),
+the Max-Planck-Institute for Astrophysics (MPA), New Mexico State
+University, Princeton University, the United States Naval Observatory,
+and the University of Washington.
 

### 2002-07-05 docs/paper/figures.tex
ah ha
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@107 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \plotone{response_sdss.ps}
-\caption{\label{response_sdss} Estimated filter response for all five
-bands in the SDSS, as a function of observed wavelength, as measured
-by Mamoru Doi. A 4 Gyr old instantaneous burst using the models
-of \citet{bruzual93a} (and observed at $z=0$) is shown for reference.}
+\caption{\label{response_sdss} Estimated response for all five bands
+in the SDSS (accounting for the atmosphere, the mirrors, the filters,
+and the CCDs), as a function of observed wavelength, as measured by
+Mamoru Doi. The predicted spectrum $f(\lambda)$ for a 4 Gyr old
+instantaneous burst using the models of \citet{bruzual93a} (and
+observed at $z=0$) is shown for reference.}
 %at different redshifts. A 4 Gyear old instantaneous burst using the
@@
 this range. Thus, the ratios $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$
-describe the spectral type of the galaxy. $a_1/a_0$ is the most
+describe the spectral type of the galaxy. $a_3/a_0$ is the most
 variable parameter and thus is the best separator of galaxy
-type. The spur extending from the upper left to the lower right from
-the red dot in the $(a_3/a_0)$-$(a_1/a_0)$ plane is due to a
+type. The spur extending from the lower left to the upper right from
+the red dot in the $(a_3/a_0)$-$(a_2/a_0)$ plane is due to a
 degeneracy for galaxies at $z\sim 0.3$, described in detail in Section
-\ref{grgap}. {\it Bottom panel}: At fixed $a_2/a_0$ and $a_3/a_0$, the
+\ref{grgap}. {\it Bottom panel}: At fixed $a_1/a_0$ and $a_2/a_0$, the
 inferred spectra corresponding to various values of $a_3/a_0$. Near

### 2002-07-05 docs/paper/paper.tex
final edits
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@106 75dff457-8751-0410-ab1a-cd701654ce6b
ah ha
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@107 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
-James E. Gunn\altaffilmark{\ref{Princeton}}
+James E. Gunn\altaffilmark{\ref{Princeton}},
 David W. Hogg\altaffilmark{\ref{NYU}}, and
@@
 surrounding their usage --- magnitudes, $K$-corrections, color terms,
-{\it etc.} --- will have long since been forgotten. However, today,
-virtually all observations of galaxies are made through such filters,
-and special care has to be taken to recover knowledge of galaxy
-spectral energy distributions (SEDs) from these observations.
-Reconstructing galaxy SEDs is nontrivial because SEDs of galaxies
-contain important information on scales considerably smaller than the
-width of typical broad band filters. Furthermore, SEDs of distant
-galaxies are redshifted such that the more distant the galaxy, the
-further the observed bandpass is blueshifted relative to the
-rest-frame spectral energy distribution of the object observed.  This
-paper focuses on a solution to these problems which (in the optical
-wavelength regime) is fast, robust, and consistent over a large range
-of redshifts.
+{\it etc.} --- will have long since been forgotten. However, until
+such time as wide field spatially-resolved spectroscopy is cheaply
+available, virtually all observations of galaxies will be made through
+broad band filters, and special care has to be taken to recover
+knowledge of galaxy spectral energy distributions (SEDs) from these
+observations.  Reconstructing galaxy SEDs is nontrivial because SEDs
+of galaxies contain important information on scales considerably
+smaller than the width of typical broad band filters. Furthermore,
+SEDs of distant galaxies are redshifted such that the more distant the
+galaxy, the further the observed bandpass is blueshifted relative to
+the rest-frame spectral energy distribution of the object observed.
+This paper focuses on a solution to these problems which (in the
+optical wavelength regime) is fast, robust, and consistent over a
+large range of redshifts.
 
@@
 users may be incorporated into the code upon request. The conditions
-of use for the code are that this paper is cited in any resulting
+of use for the code are that this paper be cited in any resulting
 refereed journal article and that the version of the code used is
@@
 
-We refer throughout this paper to $AB$ magnitudes, first defined 
-by \citet{oke83a} such that:
+We refer throughout this paper to $AB$ magnitudes, first defined by
+\citet{oke83a} to measure the ratio of the number of photons included
+in the signal of the detector relative to that number for a flat
+spectrum source with $g(\nu)=3.631 \times 10^{-20}$ ergs cm$^{-2}$
+s$^{-1}$ Hz$^{-1}$. For a source with a spectrum $f(\nu)$ the $AB$
+magnitude should be (for a perfectly calibrated $AB$ system)
 \begin{eqnarray}
@@
 where $R(\lambda)$ is the fraction of photons entering the Earth's
-atmosphere which are detected as a function of wavelength (a unitless
-quantity). In this equation, $f(\lambda)$ is in units of ergs
-cm$^{-2}$ s$^{-1}$ \AA$^{-1}$, and $f(\nu)$ is in units of ergs
-cm$^{-2}$ s$^{-1}$ Hz$^{-1}$. The normalizations defined here mean
-that an object with $f(\nu) = 3631\mathrm{~Jy} = 3.631 \times
-10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB magnitudes
-equal to zero.  The difference in the zeropoint in the two equations
+atmosphere which are included in the signal as a function of
+wavelength (a unitless quantity). Note that $R(\lambda)$ can be
+defined even for devices which do not count photons directly (such as
+bolometers). This equation is written such that $f(\lambda)$ is in
+units of ergs cm$^{-2}$ s$^{-1}$ \AA$^{-1}$ and $f(\nu)$ is in units
+of ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$, while $\lambda$ is expressed in
+\AA\ and $\nu$ is expressed in Hz. The normalizations defined here
+mean that an object with $f(\nu) = g(\nu) = 3631\mathrm{~Jy} = 3.631
+\times 10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB
+magnitudes equal to zero.  The $\lambda^{-1}$ appears in the integrand
+of the denominator of the first equation because
+$g(\lambda)=c/\lambda^2$ for a ``flat spectrum'' source with
+$g(\nu)=1$.  The difference in the zeropoints of the two equations
 simply corresponds to the factor of the speed of light $c$ (expressed
-in \AA\ s$^{-1}$) between the frequency and wavelength scales.
+in \AA\ s$^{-1}$) in that expression for $g(\lambda)$.
 
@@
 from broad band fluxes is ill-posed. The task is made yet more
-difficult by the fact that the separation of the fiters are large
+difficult by the fact that the separation of the filters are large
 compared to their widths; the SEDs are thus not well-sampled in the
@@
 combinations of $N_b$ basis spectra $\vv{b}_i$ and that $\vv{b}_i
-\cdot \vv{b}_j = \delta_{ij}$. These conditions do not fully specify
-the $\vv{b}_i$, which naturally may be rotated arbitrarily within our
+\cdot \vv{b}_j = \delta_{ij}$ (where $\delta_{ij}$ represents the
+Kronecker delta). These conditions do not fully specify the
+$\vv{b}_i$, which naturally may be rotated arbitrarily within our
 $N_b$-dimensional subspace of spectral space.
@@
 ``$K$-correction'' is to convert the $ugriz$ bands at redshift $z$
-for each galaxy to $ugriz$ bands at some fixed redshift $z_0$.
+for each galaxy to $ugriz$ bands at some fixed redshift $z_0$. While
+traditionally $K$-corrections are performed to $z_0=0$, this is not
+always desirable, particularly for surveys like the SDSS for which
+most of the galaxies are not observed near $z=0$.
 
@@
 colors of the galaxies but on the median color of $M_\ast$
-galaxies as a function of redshift. Similarly, in situation where
+galaxies as a function of redshift. Similarly, in situations where
 $K$-corrections cannot be ignored, such as the calculation of the
@@
 (\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
-22.5$. The ultimate goal is to spectroscopically observe 900,000
-galaxies, (down to $r_{\mathrm{lim}}'\approx 17.65$), 100,000 Luminous
-Red Galaxies (LRGs; \citealt{eisenstein01a}), and
-100,000 QSOs (\citealt{fan99a}; Newberg {\it et al.}, in preparation).
-This spectroscopic follow up uses two digital spectrographs on the
-same telescope as the imaging camera. Many of the details of the
-galaxy survey are described in a description of the galaxy target
-selection paper (\citealt{strauss02a}). Other aspects of the survey
-are described in the Early Data Release (EDR;
-\citealt{stoughton02a}).
+22.5$. The ultimate goal is to spectroscopically observe Main Sample
+galaxies 900,000 galaxies (down to $r_{\mathrm{lim}}\approx 17.77$;
+\citealt{strauss02a}), 100,000 Luminous Red Galaxies (LRGs;
+\citealt{eisenstein01a}), and 100,000 QSOs (\citealt{fan99a}; Newberg
+{\it et al.}, in preparation).  This spectroscopic follow up uses two
+digital spectrographs on the same telescope as the imaging
+camera. Many of the details of the galaxy survey are described in a
+description of the galaxy target selection paper
+(\citealt{strauss02a}). Other aspects of the survey are described in
+the Early Data Release (EDR; \citealt{stoughton02a}).
 
@@
 system. For each bandpass in the SDSS, there are six different CCDs;
-it has been shown that the differences between these six camera
-columns are small; what differences exist are largely fit out during
-the survey calibration. The resulting set of filter curves is shown in
-Figure \ref{response_sdss}, in comparison to a model of a galaxy
-spectral energy distribution observed at $z=0$ (a 4 Gyr old
-instantaneous burst from the models of \citealt{bruzual93a}).
+it has been shown that the differences are small between these CCD and
+filter responses for each bandpass.  The resulting set of filter
+curves is shown in Figure \ref{response_sdss}, in comparison to a
+model of a galaxy spectral energy distribution observed at $z=0$ (a 4
+Gyr old instantaneous burst from the models of \citealt{bruzual93a}).
 
@@
 from a different pipeline (\citealt{subbarao02a}). The two independent
-versions provide a consistency check on the redshift
-determination. They are consistent (for galaxies) at over the 99\%
-level.\footnote{The official pipeline currently tends to work better
-for objects with unusual spectra, such as certain types of stars and
-QSOs.}
+versions provide a consistency check on the redshift determination,
+and for galaxies they agree on more than 99\% of the
+objects.\footnote{The disagreement is more significant for objects
+with unusual spectra, such as certain types of stars and QSOs.}
 
@@
 Petrosian magnitudes, a modified form of the magnitude described by
-\citet{petrosian76a}, as described in \citet{strauss02a}.  These
-magnitudes have a nearly constant metric aperture as a function
-redshift; in addition, all the bands use the same aperture, so the
-measured SED corresponds (to within the effects of seeing) to the SED
-of an identifiable region of the galaxy. However, for faint objects,
-the Petrosian magnitudes tend to become noisy. Thus, for galaxies with
-$z>0.25$ we instead use the higher signal-to-noise ``model
-magnitudes.'' Model magnitudes are calculated in all bands using a
-single weighted aperture; thus, the measured colors again correspond
-to an identifiable region of the galaxy, though a different one than
-would be measured by the Petrosian magnitude. The weighted aperture is
-the better fitting model (pure exponential or pure de Vaucouleurs) to
-the galaxy image in the $r$-band. In \citet{stoughton02a} we are
-explicitly warned not to mix Petrosian and model magnitudes in a
-single analysis, since they measure galaxies in very different
-ways. Nevertheless, for the purposes of contraining galaxy template
-SEDs, it is perfectly acceptable to use {\it any} well-defined part of
-any galaxy.  For our purposes, it is more important to have high
-signal-to-noise measurements of galaxy colors than to measure exactly
-the same regions of galaxies at low and high redshift.  We
-extinction-correct both types of magnitudes using the dust maps of
-\citet{schlegel98a}.
+\citet{petrosian76a}, as described in \citet{strauss02a}.  Petrosian
+magnitudes are circular aperture magnitudes with an aperture size
+which is determined by the shape of the radial profile (not its
+amplitude). The resulting apertures are empirically nearly constant in
+metric size as a function of redshift; in addition, all the bands use
+the same aperture, so the measured SED corresponds (to within the
+effects of seeing) to the SED of an identifiable region of the
+galaxy. However, for faint objects, the Petrosian magnitudes tend to
+become noisy. Thus, for galaxies with $z>0.25$ we instead use the
+higher signal-to-noise ``model magnitudes.'' Model magnitudes are
+calculated in all bands using a single weighted aperture convolved
+with the point spread function; thus, the measured colors again
+correspond to an identifiable region of the galaxy, though a different
+one than would be measured by the Petrosian magnitude. The weighted
+aperture is the better fitting model (pure exponential or pure de
+Vaucouleurs) to the galaxy image in the $r$-band. Thus, the model
+magnitudes weight the centers of galaxies more strongly than do
+Petrosian magnitudes. In \citet{stoughton02a} we are explicitly warned
+not to mix Petrosian and model magnitudes in a single analysis, since
+they measure galaxies in very different ways. Nevertheless, for the
+purposes of contraining galaxy template SEDs, it is perfectly
+acceptable to use {\it any} well-defined part of any galaxy.  For our
+purposes, it is more important to have high signal-to-noise
+measurements of galaxy colors than to measure exactly the same regions
+of galaxies at low and high redshift.  We extinction-correct both
+types of magnitudes using the dust maps of \citet{schlegel98a}.
 
@@
 
-For the purposes of applying the method described in Section
-\ref{sedfit} to SDSS data, we take a subsample of the data consisting
-of around 30,000 objects in the range $0.0<z<0.5$. The sample includes
-both the main sample and the LRGs; we sample these spectra such that
-we get an approximately even distribution within our range of
-redshifts. In addition, we add results from galaxies in several plates
+For the purposes of using the SDSS data to constrain our template set,
+we take a subsample of the data consisting of around 30,000 objects in
+the range $0.0<z<0.5$. The sample includes both the main sample and
+the LRGs, and is designed such that there is an approximately even
+distribution over redshift within our range of redshifts. In addition,
+we add results from galaxies in several spectroscopic plates
 (totalling about 1,000 objects) which were selected by a photometric
 redshift algorithm (Csabai {\it et al.}, in preparation) to be at
-around $z\sim 0.3$--$0.4$. These objects are invaluable for tying down
-the blue end of the templates. Finally, we exclude galaxies in the
+around $z\sim 0.3$--$0.4$, and subsequently observed
+spectroscopically. These objects are invaluable for tying down the
+blue end of the templates.  Finally, we exclude galaxies in the
 redshift range $0.28 <z<0.32$ from the fit for the templates, for
@@
 band. We ignore these objects entirely when fitting for the
-templates. However, we still want to determine a best-fit SED for each
-object, once the templates are determined. Thus, after the templates
-have been determined, we constrain all the SED fits to all galaxies
-using the method described in Section \ref{direct}. As a constraint,
-we use the covariance matrix and mean value (essentially zero, by
-design) of the coefficients of the well-measured set of galaxies used
-to determine the templates; we multiply all the elements of the
-covariance matrix by a factor of four to weaken our constraint (which
-is appropriate, since the actual distribution of coefficients is not
-well described by a single gaussian).  Thus, we account for the
-missing information by simply requiring that the object SED have
-``reasonable'' properties. The constraint does not affect the results
-for galaxies which have well-measured magnitudes in all bands.
+templates.
 
@@
 (0.05 mags, 0.02 mags, 0.02 mags, 0.02 mags, and 0.03 mags for
-$ugriz$, respectively). Not accounting for this extra source of errors
-can cause the fits to be ill-behaved.
+$ugriz$, respectively). The choice of these values is based on a
+qualitative sense of the photometric errors present in the data. Not
+accounting for this extra source of errors can cause the fits to be
+ill-behaved.
 
 For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
-defined by linear combinations of ten Bruzual-Charlot instantaneous
-burst models with ages ranging from $3 \times 10^7$ to $2\times
-10^{10}$ years, five with metallicity $Z=0.02$ and five with
-$Z=0.004$, all assuming a Salpeter Initial Mass Function. Since we
-cannot recover information about the SED below a certain wavelength
-scale, we smooth most of the wavelength regime of the templates using
-a Gaussian with a standard deviation of $\sigma = 300$\AA; for the
-region of the spectrum which contains the sharpest gradients, around
-4000\AA, we smooth only with $\sigma = 150$\AA. None of our results
-change dramatically if we vary our smoothing procedure or if we add
-reddened templates to our allowed SED space.
+defined by ten Bruzual-Charlot instantaneous burst models with ages
+ranging from $3 \times 10^7$ to $2\times 10^{10}$ years, five with
+metallicity $Z=0.02$ and five with $Z=0.004$, all assuming a Salpeter
+Initial Mass Function. Since we cannot recover information about the
+SED below a certain wavelength resolution, we smooth most of the
+wavelength regime of the templates using a Gaussian with a standard
+deviation of $\sigma = 300$\AA; for the region of the spectrum which
+contains the sharpest gradients, around 4000\AA, we smooth only with
+$\sigma = 150$\AA. None of our results change dramatically if we vary
+our smoothing procedure or if we add reddened templates to our allowed
+SED space.
 
@@
 well. However, the fourth template is necessary to recover the
-$u$-band flux to any better than about 15\%. In addition, since one of
-the applications of these SED determinations is the distribution of
-galaxy colors in fixed frame magnitudes, we don't want to artificially
-reduce the dimensionality of the color space to only two.
+$u$-band flux to better than about 15\%. In addition, since one of the
+applications of these SED determinations is the distribution of galaxy
+colors in fixed frame magnitudes, we don't want to artificially reduce
+the dimensionality of the color space to only two.
 
@@
 choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA\ and
-$\lambda_{\mathrm{max}}=7500$\AA, since this range is included for
-almost all galaxies in the sample. We refer here and in other papers
-to the flux and luminosity in this range as the ``visual flux'' $f_v$
-and the ``visual luminosity'' $L_v$.
+$\lambda_{\mathrm{max}}=7500$\AA, since this rest frame range is
+observed for almost all galaxies in the sample. We refer here and in
+other papers to the flux and luminosity in this range as the ``visual
+flux'' $f_v$ and the ``visual luminosity'' $L_v$.
+
+Once we have finished fitting for the templates, we still want to
+determine a best-fit SED for each object.  Thus, after the templates
+have been determined, we fit for the coefficients $a_i$ for each
+galaxy, including those excluded from the template-fitting procedure
+because of large errors. In order to accommodate the presence of
+errors, we constrain all the SED fits to all galaxies using the method
+described in Section \ref{direct}. As a constraint, we use the
+covariance matrix and mean value (essentially zero, by design) of the
+coefficients of the well-measured set of galaxies used to determine
+the templates; we multiply all the elements of the covariance matrix
+by a factor of four to weaken our constraint (which is appropriate,
+since the actual distribution of coefficients is not well described by
+a single gaussian).  Thus, we account for any missing information by
+simply requiring that the object SED have ``reasonable''
+properties. The constraint does not affect the results for galaxies
+which have well-measured magnitudes in all bands.
 
@@
 spike at around 4000 \AA; that this direction is important is not
-surprising, since one of the most variable quantities of galaxy SEDs is
-the size of the 4000 \AA\ break. However, the gap between the $g$ and
-$r$ bands corresponds to the 4000 \AA\ break at around $z=0.3$. The
-existence of this spike in one direction in our four-dimensional space
-means that galaxies at $z=0.3$ can vary along this direction in order
-to better fit $u$, $i$, and $z$, with little effect on the $g-r$ color
-of the fit. This results in unphysical SED fits to galaxies near
-$z=0.3$.
+surprising, since one of the most variable quantities of galaxy SEDs
+is the size of the 4000 \AA\ break. However, the gap between the $g$
+and $r$ bands corresponds to the 4000 \AA\ break at around
+$z=0.3$. The existence of this spike in one direction in our
+four-dimensional space means that galaxies at $z=0.3$ can vary along
+this direction in order to better fit $u$, $i$, and $z$, with little
+effect on the $g-r$ color of the fit. This results in unphysical SED
+fits to galaxies near $z=0.3$.  One really only needs to worry about
+this effect in the SDSS when one is dealing with the LRGs, for
+galaxies in the range about $0.27 < z < 0.33$. The Main Sample results
+are not significantly affected at all (since few Main Sample galaxies
+are at $z>0.25$.  Furthermore, one can still observe a galaxy at
+$z=0.1$ and reliably infer what it would look like at $z=0.3$; it is
+only the reverse process which is difficult.
 
@@
 
-We have explored using a larger set of input spectrum, by including
-reddenned versions of all our templates, to test whether a more
+We have explored using a larger set of input spectra, by including
+reddened versions of all our templates, to test whether a more
 complete space of SEDs would cause our fit to choose a more realistic
 subspace. In addition, we have tried removing the increased resolution
-of the template SEDs around the 4000 \AA\ break. Neither of these tests
-yielded better results.
-
-Our view is that the ideal solution to this problem would be to
-minimize $\chi^2$ under the constraint that none of the original
-Bruzual-Charlot spectra defining our SED space contribute negatively
-to the SED fit. Such a solution would almost certainly have better
-properties than our linear approach here: it would almost certainly
-avoid the degeneracies, it would yield reasonable fit SED over a
-larger wavelength range, it would associate a reasonable
-star-formation history with each object (which could be used to
-evolution-correct the magnitudes).
+of the template SEDs around the 4000 \AA\ break. Neither of these
+tests yielded better results.
+
+Our opinion is that the ideal regularization which would solve this
+problem would be to minimize $\chi^2$ under the constraint that none
+of the original Bruzual-Charlot spectra defining our SED space
+contribute negatively to the SED fit. Such a solution would almost
+certainly have better properties than our linear approach here: it
+would almost certainly avoid the degeneracies, it would yield
+reasonable fit SED over a larger wavelength range, it would associate
+a reasonable star-formation history with each object (which could be
+used to evolution-correct the magnitudes).
 
@@
 section is sufficient to suppress the $g$/$r$ gap problem 
-for many purposes, as long as one $K$-corrected the magnitudes to
+for many purposes, as long as one $K$-corrects the magnitudes to
 $z=0.3$. In this way, one minimizes the $K$-corrections for the
 galaxies which one is least certain of.
-
-One really only needs to worry about this effect in the SDSS when you
-are dealing with the LRGs, for galaxies in the range about $0.27 < z <
-0.33$. The Main Sample results are not significantly affected at all.
-Furthermore, one can still observe a galaxy at $z=0.1$ and reliably
-infer what it would look like at $z=0.3$; it is only the reverse
-process which is difficult.
 
@@
 poor at the bluest and reddest edges of this diagram (the peak of the
-$z$-band response is at 8700~\AA restframe for a galaxy observed at
-$z=0$). Therefore, the odd behavior near the
-edges should not be taken too seriously. In addition, note the small
-spur extending from lower left to upper right through the right edge
-of the ($a_3/a_0$)-($a_2/a_0$) plane. Many of the objects in this spur are
-at around $z=0.3$, and this spur is the result of the degeracy due to
+$z$-band response is at 8700~\AA rest frame for a galaxy observed at
+$z=0$). Therefore, the odd behavior near the edges should not be taken
+too seriously. In addition, note the small spur extending from lower
+left to upper right through the right edge of the
+($a_3/a_0$)-($a_2/a_0$) plane. Many of the objects in this spur are at
+around $z=0.3$, and this spur is the result of the degeneracy due to
 the $g$/$r$ gap.
@@
 experiences a blueward shift of about 0.1 magnitudes between $z=0.1$
-and $z=0.45$, which is attributable to passive evolution.
+and $z=0.45$, which is attributable to passive evolution, though it
+could also be due to the LRG selection procedure.
 
@@
 example spectra the spectra of galaxies in the SDSS.  Since we cannot
-synthesize the observed $u$ and $z$ bands, we simply base them on the
-synthesized $g$ and $i$ band magnitudes plus the $u-g$ and $z-i$
-colors. (This procedure is unlikely to make the photometric
-estimations of the $K$-corrections {\it better}). Finally, we added
-2\% random errors to all of the measurements. Figure
-\ref{k_speck_plot.fitfib.0.1} compares the $K$-corrections from the
-synthesized photometry to that calculated from the spectrum
-itself: 
+synthesize the observed $u$ and $z$ bands from these spectra, we
+simply base them on the synthesized $g$ and $i$ band magnitudes plus
+the actual $u-g$ and $z-i$ colors. (This procedure is unlikely to make
+the photometric estimations of the $K$-corrections {\it
+better}). Finally, we added 2\% random errors to all of the
+measurements. Figure \ref{k_speck_plot.fitfib.0.1} compares the
+$K$-corrections from the synthesized photometry to that calculated
+from the spectrum itself:
 \begin{equation}
@@
 \end{equation}
-The agreement is very good, suggesting that our method can
-indeed recover the correct $K$-corrections based only on broad band
-magnitudes.
+Since all of the photometry used in this test is actually synthesized
+from the spectra, this is {\it not} a comparison of photometric and
+spectroscopic $K$-corrections; it is {\it only} a test of how well our
+method recovers $K$-corrections. The agreement in FIgure
+\ref{k_speck_plot.fitfib.0.1} is very good, suggesting that our method
+can indeed recover the correct $K$-corrections based only on broad
+band magnitudes.
 
@@
 the spectra to those of Figure \ref{k_kcorrect_plot}, finding that
-they are extremely similar. There is a trend with redshift in the
+they are quite similar. There is a trend with redshift in the
 $g$-band, which is due to the fact that on average the $g-r$ color is
@@
 we used the SDSS spectra as {\it example} spectra to validate the
-method; in this paragraph we have evaluated whether to rely on such
-spectra to calculate $K$-corrections).
+method as applied to synthesized photometry; in this paragraph we have
+evaluated whether to rely on fiber spectra to calculate
+$K$-corrections for the Petrosian photometry).
 
@@
 ($\sigma \sim 0.05$) in a comparison to SDSS galaxies with redshifts.
-Before this code can be used as a reliable photometric redshift
-indicator, more work has to be done (which the reader is invited to do
-using the released code) to identify galaxies which are likely to be
-redshift outliers. Another use of the three-dimensional space of
-galaxy SEDs is to create mock samples of galaxies observed in any
-bandpass at any redshift. We are working on quantifying the
-correlations between these coefficients and luminosity,
-surface-brighntess and radial profile in order to create mock catalogs
-which truly reflect the correlations between galaxy properties found
-in the data.
+However, we warn the reader that before this code can be used as a
+reliable photometric redshift indicator, more work has to be done
+(which the reader is invited to do using the released code) to
+identify galaxies which are likely to be redshift outliers and to
+handle galaxies at $z>0.5$ (which have photometric observations which
+are bluer in the rest frame than for any galaxies in our training
+set). Until this method is perfect and our templates are better
+constrained, we recommend the photometric redshift method and results
+of Csabai {\it et al.}, in preparation. Another use of the
+three-dimensional space of galaxy SEDs is to create mock samples of
+galaxies observed in any bandpass at any redshift. We are working on
+quantifying the correlations between these coefficients and
+luminosity, surface-brighntess and radial profile in order to create
+mock catalogs which truly reflect the correlations between galaxy
+properties found in the data.
 
@@
 
-MB is partially supported by {\bf NYU grants}.
-MB was supported at the beginning of this work by the DOE and NASA
-grant NAG 5-7092 at Fermilab. He is also grateful for the hospitality
-of the Department of Physics and Astronomy at the State University of
-New York at Stony Brook, who kindly provided computing facilities on
-his frequent visits there. 
-%MAS acknowledges the support of NSF grant AST-0071091.
+MB and DWH acknowledge NASA NAG5-11669 for partial support.  MB was
+supported at the beginning of this work by the DOE and NASA grant NAG
+5-7092 at Fermilab. He is also grateful for the hospitality of the
+Department of Physics and Astronomy at the State University of New
+York at Stony Brook, who kindly provided computing facilities on his
+frequent visits there.
 

### 2002-07-20 docs/paper/paper.tex
changed major error in overall zeropoint of K-corrections
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@114 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
-\newcommand{\band}[2]{\ensuremath{^{#1}\!{#2}}}
+\newcommand{\band}[2]{\ensuremath{^{{#1}}\!{#2}}}
 
@@
 \begin{equation}
+\label{kcorrecteqn}
 m = M + \mathrm{DM}(z) + K(z)
@@
 
-We here define a notation which we will use in subsequent papers for
-the effective rest-frame bandpass at any redshift in band $b$, namely:
+To characterize $K$-corrections, consider again Equation
+(\ref{kcorrecteqn}), which we here generalize to the case of relating
+an apparent magnitude observed in band $R$ to an absolute magnitude
+observed in band $Q$:
 \begin{equation}
-\band{z}{b}
+m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z).
 \end{equation}
-read, ``$b$-band at redshift $z$.'' For example, if we measure the
-flux of a galaxy at $z=0.22$ using a $g$-band filter, we have obtained
-the $\band{0.22}{g}$-band magnitude of the galaxy. The goal of
-``$K$-correction'' is to convert the $ugriz$ bands at redshift $z$
-for each galaxy to $ugriz$ bands at some fixed redshift $z_0$. While
-traditionally $K$-corrections are performed to $z_0=0$, this is not
-always desirable, particularly for surveys like the SDSS for which
-most of the galaxies are not observed near $z=0$.
-
-Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
-redshift $z$, one can use the method of the previous subsection to
-reconstruct the galaxy rest-frame SED.  With this reconstruction in
-hand, it is straightforward to project the SED onto any bands at any
-redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
-\begin{eqnarray}
-\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
-f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
-d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
-\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
-f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
-R(\nu)}\right]
-\end{eqnarray}
-The simple way to remember where the factors of $(1+z)$ appear in
-these equations is to remember that $K$-corrections are defined to be
-the change in magnitude with redshift for {\it fixed bolometric flux}.
-So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
-in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
-is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
-counteract the stretching of the spectrum.
-
-Given the reconstructed magnitudes, one can calculate the fixed frame
-magnitudes in one of two ways:
+The definition of an apparent AB magnitude is given by Equation
+(\ref{ABdef}). An absolute AB magnitude is defined to be the apparent
+AB magnitude which would be measured for a galaxy 10 pc distant, at
+rest.  In the restricted case of AB magnitudes, the resulting
+$K$-correction is given by:
+\begin{equation}
+K_{QR}(z) = -2.5 \log_{10} \left[
+\frac{
+\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
+}{
+\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) Q(\nu)
+}
+\right]
+\end{equation}
+A more general expression holds for other magnitude systems (such as
+the Vega-relative magnitude system) and for the case of
+$K$-corrections when $Q$ and $R$ are in different magnitude systems.
+
+In practice, we will be concerned here primarily with transforming the
+apparent magnitudes in band $R$ to a band $Q$ which is just the $R$
+band shifted by a factor $(1+z_0)$, which we will denote
+\begin{equation}
+\band{z_0}{R}(\nu) \equiv R(\nu/(1+z_0))  
+\end{equation}
+read, ``$R$-band shifted to $z_0$.''
+
+In the case of transforming $R$ to $\band{z_0}{R}$
+\begin{equation}
+K_{\band{z_0}{R}R}(z) = -2.5 \log_{10} \left[
+\frac{
+\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
+}{
+\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) R(\nu/(1+z_0))
+}
+\right]
+\end{equation}
+In the special case that $z=z_0$, one is transforming the magnitude of
+an observed galaxy at $z_0$ to its rest-frame magnitude in a band
+shifted by $z_0$. For that case
+\begin{equation}
+K_{\band{z_0}{R}R}(z_0) = -2.5 \log_{10} (1+z_0)
+\end{equation}
+independent of $f(\nu)$ (which is natural, since both magnitudes
+sample the same region of $f(\nu)$). The fact that one's uncertainties
+in $f(\nu)$ do not matter makes the use of the fixed frame magnitudes
+shifted to the redshift of the object in question useful, when it is
+possible.
+
+%Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
+%redshift $z$, one can use the method of the previous subsection to
+%reconstruct the galaxy rest-frame SED.  With this reconstruction in
+%hand, it is straightforward to project the SED onto any bands at any
+%redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
+%\begin{eqnarray}
+%\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
+%\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
+%f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
+%d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
+%\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
+%f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
+%R(\nu)}\right]
+%\end{eqnarray}
+%The simple way to remember where the factors of $(1+z)$ appear in
+%these equations is to remember that $K$-corrections are defined to be
+%the change in magnitude with redshift for {\it fixed bolometric flux}.
+%So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
+%in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
+%is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
+%counteract the stretching of the spectrum.
+
+%We here define a notation which we will use in subsequent papers for
+%the magnitude in a band $b$ which we have shifted blueward by a factor
+%$(1+z)$, namely:
+%\begin{equation}
+%\band{z}{b}
+%\end{equation}
+%read, ``$b$-band shifted to $z$.''  The goal of ``$K$-correction'' is
+%to convert the magnitudes of a galaxy at redshift $z$ observed in the
+%$ugriz$ bands at redshift $z$ to the magnitudes which would be
+%measured if the galaxy were at redshift at $z=0$ and observed through
+%the $ugriz$ bands shifted to some fixed redshift $z_0$. While
+%traditionally $K$-corrections are performed to $z_0=0$ (that is, the
+%bands are not shifted), this is not always desirable, particularly for
+%surveys like the SDSS for which most of the galaxies are not observed
+%near $z=0$.
+
+In practice, one can calculate the fixed frame magnitudes in one of
+two ways:
 \begin{enumerate}
-\item Simply use $\band{z_0}{m_{\mathrm{rec}}}$, the reconstructed
-magnitudes at redshift $z_0$, as the fixed frame magnitudes.
-\item Use the difference
-$\band{z_0}{m_{\mathrm{rec}}}-\band{z}{m_{\mathrm{rec}}}$ as an
-estimate of the correction to apply to the observed $\band{z}{m}$.
+\item Simply reconstruct $\band{z_0}{R}$ from the model
+$f_{\mathrm{rec}}(\nu)$ 
+\item Calculate the $K$-correction above and apply it to the observed
+$R$ magnitude
 \end{enumerate}
-If the reconstructions $\band{z}{m_{\mathrm{rec}}}$ are very similar
-to the actual observations $\band{z}{m}$ (as we will show is the case
-for our fits), there is very little difference between these two
-methods. However, if the reconstructions are poor, there will be large
-differences between the methods; in particular, method (1) will yield
-a color distribution of galaxies with artificially low
-dimensionality. Having estimated $\band{z_0}{m}$, you can then
-calculate the luminosity by simply applying to the calculated
-magnitude the cosmological distance modulus (that is, the luminosity
-distance-squared law, as tabulated by, {\it e.g.},  \citealt{hogg99a}).
+If the reconstructions of the observed magnitudes $R$ are very similar
+to the actual observations (as we will show is the case for our fits),
+there is very little difference between these two methods. However, if
+the reconstructions are poor, there will be large differences between
+the methods; in particular, method (1) will yield a color distribution
+of galaxies with artificially low dimensionality. Having estimated
+$\band{z_0}{m}$, you can then calculate the luminosity by simply
+applying to the calculated magnitude the cosmological distance modulus
+(that is, the luminosity distance-squared law, as tabulated by, {\it
+e.g.}, \citealt{hogg99a}).
 
@@
 
-MB and DWH acknowledge NASA NAG5-11669 for partial support.  MB was
-supported at the beginning of this work by the DOE and NASA grant NAG
-5-7092 at Fermilab. He is also grateful for the hospitality of the
-Department of Physics and Astronomy at the State University of New
-York at Stony Brook, who kindly provided computing facilities on his
-frequent visits there.
+Special thanks to Ivan Baldry for pointing out an error in this
+manuscript and in the distributed code.  MB and DWH acknowledge NASA
+NAG5-11669 for partial support.  MB was supported at the beginning of
+this work by the DOE and NASA grant NAG 5-7092 at Fermilab. He is also
+grateful for the hospitality of the Department of Physics and
+Astronomy at the State University of New York at Stony Brook, who
+kindly provided computing facilities on his frequent visits there.
 

### 2002-07-20 src/color_interp.c

@@ -62,8 +62,6 @@ static double scalefiltfunc(double lambda)
 
 static double tfunc(double lambda)
 {
-	unsigned long i,ip1;
-
 	return(ci_A*pow(lambda/(1.+ci_z),ci_alpha)/(1.+ci_z));
 } /* end tfunc */
 
@@ -76,8 +74,6 @@ static double tempfiltfunc(double lambda)
 
 static double bfunc(double lambda)
 {
-	unsigned long i,ip1;
-
 	if(lambda/(1.+ci_z)<4000.)
 		return(ci_break*pow(lambda/(1.+ci_z),2.)/(1.+ci_z));
 	else 
@@ -156,8 +152,8 @@ double balpha_fit(double alpha)
 
 int main(int argc, char **argv)
 {
-	double currz,cz,redden,alpha,A,val,scaleval;
-	int i,ii,j,k,l;
+	double alpha,A,val,scaleval;
+	int i,k;
 	char filename[255],bands[5];
 	FILE *fp;
 
@@ -338,6 +334,8 @@ int main(int argc, char **argv)
 	
 	printf("Done.\n");
 	fflush(stdout);
+	
+	return(0);
 	
 } /* end main */
 

### 2002-07-20 src/fit_coeffs.c

@@ -143,4 +143,6 @@ int main(int argc,
 	FREEVEC(galaxy_maggies);
 	FREEVEC(galaxy_invvar);
 	FREEVEC(coeffs);
+
+	return(0);
 } /* end main */

### 2002-07-20 src/fit_photoz.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -133,4 +134,6 @@ int main(int argc,
 	FREEVEC(galaxy_maggies);
 	FREEVEC(galaxy_invvar);
 	FREEVEC(coeffs);
+	
+	return(0);
 } /* end main */

### 2002-07-20 src/idl_k_create_r.c

@@ -16,7 +16,7 @@ IDL_LONG idl_k_create_r
 	IDL_LONG nk,nb,nl,nz,*filter_n,maxn;
 	double *rmatrix,*bmatrix,*lambda,*zvals,*filter_lambda,*filter_pass;
 	
-	IDL_LONG i,j;
+	IDL_LONG i;
 	IDL_LONG retval=1;
 
 	/* 0. allocate pointers from IDL */

### 2002-07-20 src/idl_k_fit_templates.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <math.h>
 #include <stdlib.h>
 #include <kcorrect.h>

### 2002-07-20 src/idl_k_reconstruct_maggies.c

@@ -13,7 +13,7 @@ IDL_LONG idl_k_reconstruct_maggies
   (int      argc,
    void *   argv[])
 {
-   IDL_LONG nt,nz,nk,nb,ngalaxy,*galaxy_clip,initialized_dmatrix;
+   IDL_LONG nt,nz,nk,nb,ngalaxy;
    double *ematrix,*zvals,*rmatrix,*amatrix,*reconstruct_maggies;
 	 double *galaxy_z,*band_shift;
 

### 2002-07-20 src/k_create_r.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -94,8 +95,6 @@ IDL_LONG k_create_r(double *rmatrix,
 {
 	double lammin,lammax,scale;
 	IDL_LONG i,l,k,indxoff;
-	char filename[255];
-	FILE *fp;
 
 	/* make local copies of bmatrix */
 	cr_bmatrix=(double *) malloc(nb*nl*sizeof(double));

### 2002-07-20 src/k_fileopen.c

@@ -1,5 +1,6 @@
 #include <string.h>
 #include <stdio.h>
+#include <stdlib.h>
 
 /*
  * k_fileopen.c : contains routine to open a file 

### 2002-07-20 src/k_fit_coeffs.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>

### 2002-07-20 src/k_fit_photoz.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -24,17 +25,18 @@ static double *pz_coeffs=NULL;
 static double *pz_reconstruct_maggies=NULL;
 static double *pz_galaxy_maggies=NULL;
 static double *pz_galaxy_invvar=NULL;
-static double *pz_coeffspnorm=NULL;
-static double pz_z,pz_linepos;
-static IDL_LONG pz_nt,pz_nz,pz_nk,pz_nb,pz_puse,pz_np,pz_p;
+static IDL_LONG pz_nt,pz_nz,pz_nk,pz_nb;
 
 double pz_calc_chi2(double z) 
 {
-	double *reconstruct_maggies,chi2;
+	double chi2;
+	double band_shift;
 	IDL_LONG k;
 	
+	band_shift=0.;
 	k_reconstruct_maggies(pz_ematrix,pz_nt,pz_zvals,pz_nz,pz_rmatrix,
-											 pz_nk,pz_nb,pz_coeffs,&z,pz_reconstruct_maggies,1);
+												pz_nk,pz_nb,pz_coeffs,&z,&band_shift, 
+												pz_reconstruct_maggies,1);
 	chi2=0.;
 	for(k=0;k<pz_nk;k++)
 		chi2+=(pz_reconstruct_maggies[k]-pz_galaxy_maggies[k])
@@ -46,8 +48,7 @@ double pz_calc_chi2(double z)
 /* calculates chi2 */
 double pz_fit_coeffs(double z) 
 {
-	double chi2,az,bz,cz;
-	IDL_LONG i,k;
+	double chi2;
 	
 	/* fit coeffs; no direct constraints are put on the coeffs in this case */
 	k_fit_coeffs(pz_ematrix,pz_nt,pz_zvals,pz_nz,pz_rmatrix,
@@ -74,8 +75,8 @@ IDL_LONG k_fit_photoz(double *ematrix,    /* eigentemplates */
 											double *galaxy_z,
 											IDL_LONG ngalaxy)
 {
-	double *zgrid,*chi2,chi2min,az,bz,cz,sl;
-	IDL_LONG i,j,k,b,p,nzsteps,jmin;
+	double *zgrid,*chi2,chi2min,az,bz,cz;
+	IDL_LONG i,j,k,nzsteps,jmin;
 
 	/* allocate memory */
 	pz_nk=nk;

### 2002-07-20 src/k_load_ascii_table.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -59,4 +60,6 @@ IDL_LONG k_load_ascii_table(double **table,
 	for(i=0;i<nelem;i++) 
 		fscanf(fp,"%lf",&((*table)[i]));
 	fclose(fp);
+
+	return(1);
 } /* end k_load_ascii_table */

### 2002-07-20 src/k_polint.c

@@ -1,5 +1,6 @@
 #include <string.h>
 #include <stdio.h>
+#include <stdlib.h>
 #include <math.h>
 #include <kcorrect.h>
 #define NRANSI

### 2002-07-20 src/k_qromo.c

@@ -1,5 +1,6 @@
 #include <string.h>
 #include <stdio.h>
+#include <stdlib.h>
 #include <math.h>
 #include <kcorrect.h>
 #define EPS 1.0e-4

### 2002-07-20 src/k_write_ascii_table.c

@@ -1,4 +1,5 @@
 #include <stdio.h>
+#include <stdlib.h>
 #include <string.h>
 #include <math.h>
 #include <kcorrect.h>
@@ -46,4 +47,5 @@ IDL_LONG k_write_ascii_table(double *table,
 		fprintf(fp,"%e\n",table[i]);
 	fclose(fp);
 
+	return(1);
 } /* end k_write_ascii_table */

### 2002-07-20 src/reconstruct_maggies.c

@@ -8,7 +8,27 @@
  * reconstruct_maggies.c
  *
  * Calculates the reconstructed flux, given the coefficients and the template
- * information
+ * information; one can specify two parameters which characterize the 
+ * magnitude:
+ *
+ * band_z : the shift applied to the band pass
+ * at_z   : the redshift at which to observe the galaxy
+ *
+ * To use reconstruct_maggies as part of a system to produce absolute
+ * magnitudes, use at_z=0. and set band_z to the shifted bandpass you
+ * want to K-correct to. Eg. to K-correct to ^{0.2}b (in the notation
+ * of the paper) call it like:
+ *
+ *  cat coeffs.dat | reconstruct_maggies 0.2 0. > maggies.dat
+ *
+ * where we specify at_z = 0. because this guarantees that we 
+ * observe the galaxy at z=0. (appropriate for K-correction defn).
+ *
+ * To reconstruct the observed magnitudes at the observed redshift:
+ *
+ *  cat coeffs.dat | reconstruct_maggies > maggies.dat
+ *
+ * (Assuming coeffs.dat is actually output of fit_coeffs)
  *
  * Mike Blanton
  * 1/2002 */
@@ -28,6 +48,7 @@ static double *filter_pass=NULL;
 static IDL_LONG *filter_n=NULL;
 static IDL_LONG maxn;
 static double *galaxy_z=NULL;
+static double *band_shift=NULL;
 static double *coeffs=NULL;
 
 int main(int argc,
@@ -47,14 +68,14 @@ int main(int argc,
 
 	/* read arguments */
 	if(argc<0) {
-		fprintf(stderr,"Usage: cat <coeff file> | reconstruct_maggies [at_z] [band_z] [version [version path]]\n");
+		fprintf(stderr,"Usage: cat <coeff file> | reconstruct_maggies [band_z] [at_z] [version [version path]]\n");
 		exit(1);
 	} /* end if */
 	i=1;
 	if(argc>=2) 
-		at_z=atof(argv[i]); i++; 
-	if(argc>=3) 
 		band_z=atof(argv[i]); i++; 
+	if(argc>=3) 
+		at_z=atof(argv[i]); i++; 
 	if(argc>=4) 
[…]

### 2002-09-03 docs/paper/paper.tex
paper update
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@145 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-\newcommand{\kversion}{{\tt v1\_10}}
+\newcommand{\kversion}{{\tt v1\_11}}
 
@@
 bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
-relationship between the apparent magnitude $m$ and the absolute
-magnitude $M$: 
+relationship between the apparent magnitude $m_R$ of band $R$ and the
+absolute magnitude $M$ of (in general different) band $Q$:
 \begin{equation}
 \label{kcorrecteqn}
-m = M + \mathrm{DM}(z) + K(z)
+m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z)
 \end{equation}
-Sometimes a single function has been applied regardless of the galaxy
-type, though more recently it has become standard to use a discrete
-set of $K(z)$ functions depending on galaxy type (based either on
-morphology or spectral features). As described in \cite{oke68a} and
-later papers, the function is based on the projection of an assumed
-galaxy SED redshifted to $z$ onto the measured instrument response as
-a function of wavelength (including the effects of the atmosphere, the
-reflectivity of the mirrors, the filter transmission, and the response
-of the CCD device or photographic emulsion). For example, for the
-Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
-\citet{fukugita95a} have both presented results for the
-$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
-(called $k(z)$ by \citealt{frei94a}) for galaxies of various
-morphological types, based on galaxy spectrophotometry.\footnote{Note
-that the description on p. 57 of \citet{binney98a} of the meaning of
-$k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
-exactly equivalent to our $K(z)$.}
+The traditional definition of the $K$-correction takes $Q=R$; however,
+we note that in practice many surveys do perform $K$-corrections from
+one bandpass to another when comparing high redshift and low redshift
+observations.
+
+Sometimes a single function $K(z)$ has been applied regardless of the
+galaxy type, though more recently it has become standard to use a
+discrete set of $K(z)$ functions depending on galaxy type (based
+either on morphology or spectral features). As described in
+\cite{oke68a} and later papers, the function is based on the
+projection of an assumed galaxy SED redshifted to $z$ onto the
+measured instrument response as a function of wavelength (including
+the effects of the atmosphere, the reflectivity of the mirrors, the
+filter transmission, and the response of the CCD device or
+photographic emulsion). For example, for the Sloan Digital Sky Survey
+(SDSS) filter system, \citet{frei94a} and \citet{fukugita95a} have
+both presented results for the $K$-corrections of galaxies. Each of
+these papers tabulates $K(z)$ (called $k(z)$ by \citealt{frei94a}) for
+galaxies of various morphological types, based on galaxy
+spectrophotometry.\footnote{Note that the description on p. 57 of
+\citet{binney98a} of the meaning of $k(z)$ in \citet{frei94a} is
+incorrect; in fact, their $k(z)$ is exactly equivalent to our $K(z)$.}
 
@@
 To characterize $K$-corrections, consider again Equation
-(\ref{kcorrecteqn}), which we here generalize to the case of relating
-an apparent magnitude observed in band $R$ to an absolute magnitude
-observed in band $Q$:
+(\ref{kcorrecteqn}): 
 \begin{equation}
@@
 A more general expression holds for other magnitude systems (such as
-the Vega-relative magnitude system) and for the case of
-$K$-corrections when $Q$ and $R$ are in different magnitude systems.
+the Vega-relative magnitude system).
 
@@
 shifted to the redshift of the object in question useful, when it is
-possible.
+possible. We note in passing that \citet{oke68a} state that the
+$K$-correction ``would disappear if intensity measurements of
+redshifted galaxies were made with a detector whose spectral
+acceptance band was shifted by $1+z$ at all wavelengths;'' here we
+have just shown this statement to be untrue for an AB system (though
+one could define a magnitude system in which this were true).
 
@@
 model of the galaxy SED, we do not correct to bandpasses with simpler
-shapes (for example, top hats). The reason is that we want the effective
-bandpasses to actually have been observed for some galaxies in the
-sample, so that the $K$-correction for those galaxies are formally and
-actually zero. This property is desirable because, as noted above,
-there are inherent uncertainties in the $K$-corrections. Nevertheless,
-we note that synthesizing top hat magnitudes may be appropriate in
-certain situations, as they have the virtue (similarly to the $AB$
-magnitude system) that they are very easy to comprehend and synthesize
-from theory.
+shapes (for example, top hats). The reason is that we want the
+effective bandpasses to actually have been observed for some galaxies
+in the sample, so that the $K$-correction for those galaxies are
+independent of the galaxy SED.  This property is desirable because, as
+noted above, there are inherent uncertainties in the $K$-corrections
+due to our lack of knowledge of the SEDs. Nevertheless, we note that
+synthesizing top hat magnitudes may be appropriate in certain
+situations, as they have the virtue (similarly to the $AB$ magnitude
+system) that they are very easy to comprehend and synthesize from
+theory.
 %For example, if we wanted to convert our results into a
@@
 Special thanks to Ivan Baldry for pointing out an error in this
-manuscript and in the distributed code.  MB and DWH acknowledge NASA
-NAG5-11669 for partial support.  MB was supported at the beginning of
-this work by the DOE and NASA grant NAG 5-7092 at Fermilab. He is also
-grateful for the hospitality of the Department of Physics and
-Astronomy at the State University of New York at Stony Brook, who
-kindly provided computing facilities on his frequent visits there.
+manuscript and in the distributed code for versions {\tt v1\_10} and
+previous.  MB and DWH acknowledge NASA NAG5-11669 for partial support.
+MB was supported at the beginning of this work by the DOE and NASA
+grant NAG 5-7092 at Fermilab. He is also grateful for the hospitality
+of the Department of Physics and Astronomy at the State University of
+New York at Stony Brook, who kindly provided computing facilities on
+his frequent visits there.
 

### 2002-09-04 docs/paper/paper.tex
all
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@148 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 objective, it is not {\it always} necessary or appropriate. Because
-$K$-corrections are inherently uncertain (the broad band magnitudes
-do not uniquely determine the SED) they should be avoided or
-minimized when possible. For example, if one were calculating the
-evolution of clustering of red and blue galaxies separately, it would
-perhaps be wise to perform the separation not on the $K$-corrected
-colors of the galaxies but on the median color of $M_\ast$
-galaxies as a function of redshift. Similarly, in situations where
-$K$-corrections cannot be ignored, such as the calculation of the
-evolution of the luminosity function, their effect should be minimized
-by, for example, correcting to the median redshift of galaxies in the
-sample.
+$K$-corrections are inherently uncertain (the broad band magnitudes do
+not uniquely determine the SED) they should be avoided or minimized
+when possible. For example, if one were calculating the evolution of
+clustering of red and blue galaxies separately, it would perhaps be
+wise to perform the separation not on the $K$-corrected colors of the
+galaxies but on the median color of $M_\ast$ galaxies as a function of
+redshift. Similarly, in situations where $K$-corrections cannot be
+ignored, such as the calculation of the evolution of the luminosity
+function, their effect should be minimized by, for example, correcting
+to a band shifted to the median redshift of galaxies in the sample.
 

### 2003-03-30 docs/paper/paper.tex
tiny
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@183 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 (\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
-22.5$. The ultimate goal is to spectroscopically observe Main Sample
-galaxies 900,000 galaxies (down to $r_{\mathrm{lim}}\approx 17.77$;
+22.5$. The ultimate goal is to spectroscopically observe 900,000 Main
+Sample galaxies (down to $r_{\mathrm{lim}}\approx 17.77$;
 \citealt{strauss02a}), 100,000 Luminous Red Galaxies (LRGs;
-\citealt{eisenstein01a}), and 100,000 QSOs (\citealt{fan99a}; Newberg
-{\it et al.}, in preparation).  This spectroscopic follow up uses two
-digital spectrographs on the same telescope as the imaging
-camera. Many of the details of the galaxy survey are described in a
-description of the galaxy target selection paper
-(\citealt{strauss02a}). Other aspects of the survey are described in
-the Early Data Release (EDR; \citealt{stoughton02a}).
+\citealt{eisenstein01a}), and 100,000 QSOs (\citealt{fan99a};
+\citealt{richards02a}.  This spectroscopic follow up uses two digital
+spectrographs on the same telescope as the imaging camera. Many of the
+details of the galaxy survey are described in a description of the
+galaxy target selection paper (\citealt{strauss02a}). Other aspects of
+the survey are described in the Early Data Release (EDR;
+\citealt{stoughton02a}).
 

### 2004-01-17 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@292 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2003-03-30 docs/paper/paper.tex
@@
 
-%% You can insert a short comment on the title page using the command below.
-
 \slugcomment{To be submitted to \aj}
 
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-%% The left head contains a list of authors,
-%% usually a maximum of three (otherwise use et al.).  The right
-%% head is a modified title of up to roughly 44 characters.  Running heads
-%% will not print in the manuscript style.
+\shortauthors{Blanton {\it et al.} (2000)}
+\shorttitle{$K$-corrections and photometric redshifts}
 
-\shortauthors{Blanton {\it et al.} (2000)}
-\shorttitle{Modeling Galaxy SEDs}
-
-%% This paper uses runs 752/756, rerun 4
-%% It uses redshifts from version 3 of spectro1d
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% paper itself with \begin{document}.
+% TODO:
+% finish final templates
+% fix kphotoz banding
+% fit SFHs to final spectra to get stellar masses, mean ages, 
+% mean metallicities
 
@@
 
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire.
+\title{$K$-corrections and photometric redshifts in SDSS and 2MASS$^1$}
 
-\title{Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$}
-
-%% Use \author, \affil, and the \and command to format
-%% author and affiliation information.
-%% Note that \email has replaced the old \authoremail command
-%% from AASTeX v4.0. You can use \email to mark an email address
-%% anywhere in the paper, not just in the front matter.
-%% As in the title, you can use \\ to force line breaks.
-
-%\author{Michael Blanton}
-%\affil{NASA/Fermilab Astrophysics Center\\
-%Fermi National Accelerator Laboratory, Batavia, IL 60510-0500}
-%\author{\and lots and lots of other, probably more important people}
-%\affil{Some Institution\\
-%Somewhere, Some City, Some State or Province}
-%\email{blanton@fnal.gov}
-
-% Authorship determined by those I was directly involved with
-% in performing this work as well as those responsible for the photo-z
-% plates and those who have characterized the filter curves (since
-% these might be publicly distributed). 
 \author{
-Michael R. Blanton\altaffilmark{\ref{NYU}},
+Michael R. Blanton\altaffilmark{\ref{NYU}}
 %Tamas Budavari\altaffilmark{\ref{JHU}},
 %Andrew J. Connolly\altaffilmark{\ref{Pitt}},
-J.~Brinkmann\altaffilmark{\ref{APO}},
-Istv\'an Csabai\altaffilmark{\ref{JHU}},
-Mamoru Doi\altaffilmark{\ref{Tokyo}},
-Daniel Eisenstein\altaffilmark{\ref{Arizona}},
-Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
-James E. Gunn\altaffilmark{\ref{Princeton}},
-David W. Hogg\altaffilmark{\ref{NYU}}, and
-David J. Schlegel\altaffilmark{\ref{Princeton}}
+%J.~Brinkmann\altaffilmark{\ref{APO}},
+%Istv\'an Csabai\altaffilmark{\ref{JHU}},
+%Mamoru Doi\altaffilmark{\ref{Tokyo}},
+%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
+%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
+%James E. Gunn\altaffilmark{\ref{Princeton}},
+%David W. Hogg\altaffilmark{\ref{NYU}}, and
+%David J. Schlegel\altaffilmark{\ref{Princeton}}
 %Julianne Dalcanton\altaffilmark{\ref{UW}},
@@
 \label{NYU}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Department of Physics and Astronomy, The Johns Hopkins University,
-Baltimore, MD 21218
-\label{JHU}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Department of Physics and Astronomy, The Johns Hopkins University,
+%Baltimore, MD 21218
+%\label{JHU}}
 %\altaffiltext{\theaddress}{
@@
 %\label{Pitt}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Department of Astronomy and Research Center for 
-the Early Universe,
-School of Science, University of Tokyo,
-Tokyo 113-0033, Japan
-\label{Tokyo}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Steward Observatory, 
-933 N. Cherry Ave., Tucson, AZ
-85721
-\label{Arizona}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Princeton University Observatory, Princeton,
-NJ 08544
-\label{Princeton}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Department of Astronomy and Research Center for 
+%the Early Universe,
+%School of Science, University of Tokyo,
+%Tokyo 113-0033, Japan
+%\label{Tokyo}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Steward Observatory, 
+%933 N. Cherry Ave., Tucson, AZ
+%85721
+%\label{Arizona}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Princeton University Observatory, Princeton,
+%NJ 08544
+%\label{Princeton}}
 %\addtocounter{address}{1}
@@
 %\label{Ohio}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Apache Point Observatory,
-2001 Apache Point Road,
-P.O. Box 59, Sunspot, NM 88349-0059
-\label{APO}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Apache Point Observatory,
+%2001 Apache Point Road,
+%P.O. Box 59, Sunspot, NM 88349-0059
+%\label{APO}}
 %\altaffiltext{\theaddress}{
@@
 %\label{Berkeley}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Institute for Cosmic Ray Research, University of
-Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
-\label{CosmicRay}}
-\altaffiltext{\theaddress}{
-\stepcounter{address}
-Institute for Advanced Study, Olden Lane,
-Princeton, NJ 08540
-\label{IAS}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Institute for Cosmic Ray Research, University of
+%Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
+%\label{CosmicRay}}
+%\altaffiltext{\theaddress}{
+%\stepcounter{address}
+%Institute for Advanced Study, Olden Lane,
+%Princeton, NJ 08540
+%\label{IAS}}
 %\altaffiltext{\theaddress}{
@@
 
-%% Mark off your abstract in the ``abstract'' environment. In the manuscript
-%% style, abstract will output a Received/Accepted line after the
-%% title and affiliation information. No date will appear since the author
-%% does not have this information. The dates will be filled in by the
-%% editorial office after submission.
 \begin{abstract}
-Broad-band measurements of flux for galaxies at different redshifts
-measure different regions of the rest-frame galaxy spectrum. Certain
-astronomical questions, such as the evolution of the luminosity
-function of galaxies, require transforming these inherently
-redshift-dependent magnitudes into redshift-independent quantities. To
-prepare to address these astronomical questions, investigated in
-detail in subsequent papers, we fit spectral energy distributions
-(SEDs) to broad band photometric observations, in the context of the
-optical observations of the Sloan Digital Sky Survey (SDSS). Linear
-combinations of four spectral templates can reproduce the five SDSS
-magnitudes of all galaxies to the precision of the
-photometry. Expressed in the appropriate coordinate system, the
-locus of the coefficients multiplying the templates is planar,
-and in fact nearly linear. The resulting reconstructed SEDs can be
-used to recover fixed frame magnitudes over a range of redshifts. This
-process yields consistent results, in the sense that within each
-sample the intrinsic colors of similar type galaxies are nearly
-constant with redshift. We compare our results to simpler
-interpolation methods and galaxy spectrophotometry from the SDSS.  The
-software that generates these results is publicly available and easily
-adapted to handle a wide range of galaxy observations.
+We present a method to develop templates for calculating
+$K$-corrections and photometric redshifts suitable for ultraviolet,
+optical, and infrared observations. Since we base the templates on
+stellar population synthesis results, the results are intepretable in
+terms of approximate stellar masses and star-formation histories. We
+present templates fit with this method to the Sloan Digital Sky Survey
+and Two-Micron All Sky Survey photometry.  The software that generates
+these results is publicly available and easily adapted to handle a
+wide range of galaxy observations.
 \end{abstract}
@@
 
-%
-% Introduction and motivation
-%
-
 \section{Motivation}
@@
 
-In the future, observations of galaxies (and indeed of any
-astronomical sources) will be performed using devices which combine
-high spatial resolution and high spectral resolution. By that time,
-the phenomena of broad band filters and the quaint terminology
-surrounding their usage --- magnitudes, $K$-corrections, color terms,
-{\it etc.} --- will have long since been forgotten. However, until
-such time as wide field spatially-resolved spectroscopy is cheaply
-available, virtually all observations of galaxies will be made through
-broad band filters, and special care has to be taken to recover
-knowledge of galaxy spectral energy distributions (SEDs) from these
-observations.  Reconstructing galaxy SEDs is nontrivial because SEDs
-of galaxies contain important information on scales considerably
-smaller than the width of typical broad band filters. Furthermore,
-SEDs of distant galaxies are redshifted such that the more distant the
-galaxy, the further the observed bandpass is blueshifted relative to
-the rest-frame spectral energy distribution of the object observed.
-This paper focuses on a solution to these problems which (in the
-optical wavelength regime) is fast, robust, and consistent over a
-large range of redshifts.
-
-In the past, people have accounted for these effects using
-``$K$-corrections'' applied to the observed magnitudes. In these
-analyses, a function $K(z)$ is added to the standard cosmological
-bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
-relationship between the apparent magnitude $m_R$ of band $R$ and the
-absolute magnitude $M$ of (in general different) band $Q$:
-\begin{equation}
-\label{kcorrecteqn}
-m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z)
-\end{equation}
-The traditional definition of the $K$-correction takes $Q=R$; however,
-we note that in practice many surveys do perform $K$-corrections from
-one bandpass to another when comparing high redshift and low redshift
-observations.
-
-Sometimes a single function $K(z)$ has been applied regardless of the
-galaxy type, though more recently it has become standard to use a
-discrete set of $K(z)$ functions depending on galaxy type (based
-either on morphology or spectral features). As described in
-\cite{oke68a} and later papers, the function is based on the
-projection of an assumed galaxy SED redshifted to $z$ onto the
-measured instrument response as a function of wavelength (including
-the effects of the atmosphere, the reflectivity of the mirrors, the
-filter transmission, and the response of the CCD device or
-photographic emulsion). For example, for the Sloan Digital Sky Survey
-(SDSS) filter system, \citet{frei94a} and \citet{fukugita95a} have
-both presented results for the $K$-corrections of galaxies. Each of
-these papers tabulates $K(z)$ (called $k(z)$ by \citealt{frei94a}) for
-galaxies of various morphological types, based on galaxy
-spectrophotometry.\footnote{Note that the description on p. 57 of
-\citet{binney98a} of the meaning of $k(z)$ in \citet{frei94a} is
-incorrect; in fact, their $k(z)$ is exactly equivalent to our $K(z)$.}
-
-However, galaxies do not all have the same SED, nor are they selected
-from some discrete set of SEDs. For this reason, these schemes for
-applying the $K$-corrections can fail to be self-consistent: the SED
-assumed for the $K$-corrections can be significantly inconsistent with
-the observed galaxy colors! As we demand more precision from our
-astronomical data analysis in new, large multi-band surveys, and in
-particular as we try to quantify the evolution of galaxies, we must
-take a more sophisticated approach to approximating fixed frame
-observations of galaxies.
-
-Our approach here is to use a method for inferring the underlying SEDs
-of a set of galaxies at a range of redshifts by requiring that their
-SEDs all be drawn from a similar population. For each galaxy, we will
-recover a model SED, which can be used to synthesize the galaxy's
-magnitude in any bandpass. Although the operation we are performing on
-the magnitudes is not a ``$K$-correction'' in the historical sense of
-the term, we will refer to it as such in this and subsequent
-papers. Our approach is equivalent (nearly identical) to the
-photometric redshift estimation methods of \citet{csabai00a},
-except we use slightly different coordinate systems for our SED
-template space. In fact, the software includes a fast and relatively
-accurate photometric redshift estimator based on their method.
-
-We implemented this system and are publishing it in order to lay the
-groundwork for upcoming papers which will rely heavily on the
-reliability of the fixed frame magnitudes determined here.  Because
-the SDSS is one of the largest astronomical surveys to date, our
-method will be useful to a large number of other investigators, both
-inside and outside the SDSS collaboration.  Thus, this paper also
-serves the purpose of describing the release of a piece of
-software. The computer software we distribute builds itself into a C
-shared object library, around which we have written both stand-alone C
-programs and IDL routines. We have made the source code available
-publicly, through a web page and through a public CVS
-repository. Improvements or ports to other languages implemented by
-users may be incorporated into the code upon request. The conditions
-of use for the code are that this paper be cited in any resulting
-refereed journal article and that the version of the code used is
-specified in any such paper. The version of the code used to make the
-figures for this paper is {\tt kcorrect \kversion}.
-
-We refer throughout this paper to $AB$ magnitudes, first defined by
-\citet{oke83a} to measure the ratio of the number of photons included
-in the signal of the detector relative to that number for a flat
-spectrum source with $g(\nu)=3.631 \times 10^{-20}$ ergs cm$^{-2}$
-s$^{-1}$ Hz$^{-1}$. For a source with a spectrum $f(\nu)$ the $AB$
-magnitude should be (for a perfectly calibrated $AB$ system)
-\begin{eqnarray}
-\label{ABdef}
-m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
-{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
-\cr
-&=& -48.60 - 2.5 \log_{10}\left[
-\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
-{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
-\end{eqnarray}
-where $R(\lambda)$ is the fraction of photons entering the Earth's
-atmosphere which are included in the signal as a function of
-wavelength (a unitless quantity). Note that $R(\lambda)$ can be
-defined even for devices which do not count photons directly (such as
-bolometers). This equation is written such that $f(\lambda)$ is in
-units of ergs cm$^{-2}$ s$^{-1}$ \AA$^{-1}$ and $f(\nu)$ is in units
-of ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$, while $\lambda$ is expressed in
-\AA\ and $\nu$ is expressed in Hz. The normalizations defined here
-mean that an object with $f(\nu) = g(\nu) = 3631\mathrm{~Jy} = 3.631
-\times 10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB
-magnitudes equal to zero.  The $\lambda^{-1}$ appears in the integrand
-of the denominator of the first equation because
-$g(\lambda)=c/\lambda^2$ for a ``flat spectrum'' source with
-$g(\nu)=1$.  The difference in the zeropoints of the two equations
-simply corresponds to the factor of the speed of light $c$ (expressed
-in \AA\ s$^{-1}$) in that expression for $g(\lambda)$.
-
-Section \ref{sedfit} describes our method of fitting galaxy SEDs to
-broad band photometry and how to calculate $K$-corrections. Section
-\ref{data} applies the method to galaxies in the SDSS, showing that
-the fits are robust.  Section \ref{conclusions} concludes and
-discusses future development of the method described here.
-
-\section{Calculating Fixed Frame Galaxy Magnitudes}
+\section{Fitting SEDs to broad-band magnitudes}
 \label{sedfit}
@@
 First, we describe how we reconstruct galaxy SEDs from broad band
-magnitude measurements. Second, we describe how we convert these SEDs
-into estimates of fixed frame galaxy magnitudes.
+magnitude measurements.
 
-\subsection{Fitting SEDs to Galaxy Broad-band Magnitudes}
+A broad-band flux $F_k$ in filter $k$ is a measurement of the
+projection of the source spectrum $f(\lambda)$ on the response
+$R_k(\lambda)$ (where typically this function includes the effects of
+the atmosphere, telescope, filter $k$, and the CCD itself)
+\begin{equation}
+F_k = \int_0^\infty d\lambda f(\lambda) R_k(\lambda).
+\end{equation}
+Now, let us model $f(\lambda)$ as a sum of many template spectra:
+\begin{equation}
+f(\lambda) = \sum_j a_j v_j(\lambda)
+\end{equation}
+In order to fit the $a_j$ given the observations in the bands $k$ for
+a single object, we minimize
+\begin{eqnarray}
+\label{chi2}
+\chi^2 &=& \sum_{k} \left[ \frac{F_{k,\mathrm{obs}} - F(k)}{\sigma_k}
+\right]^2 \cr
+&=& \sum_{ik} \left[ \frac{F_{k,\mathrm{obs}} -\sum_j a_j
+  v_j(\lambda)}{\sigma_k} \right]^2
+\end{eqnarray}
+over the $a_j$, where $\sigma_k$ is the standard deviation of the
+error in band $k$. Since we will use $v_j(\lambda)$ which are the light
+emitted from stars and ionized gas, we make the reasonable assumption
+that these only {\it add} light to the spectrum, not subtract it, and
+require nonnegativity: that $a_j\ge 0$ for all $j$. (We discuss dust
+below).  In order to perform this nonnegative minimization of Equation
+\ref{chi2}, we use the method of \citet{sha02a}.
 
-Our task is to recover a model for the galaxy SED from broad band
-photometric measurements. Since a set of broad band galaxy fluxes does
-not correspond uniquely to a particular SED, and because galaxy SEDs
-are known to have significant structure over wavelength ranges small
-compared to our bandpass, the inverse problem of reconstructing SEDs
-from broad band fluxes is ill-posed. The task is made yet more
-difficult by the fact that the separation of the filters are large
-compared to their widths; the SEDs are thus not well-sampled in the
-Nyquist sense of that term. However, we are not completely ignorant
-about the forms which galaxy SEDs take, so we can attempt to use what
-we know about galaxy SEDs to appropriately regularize our fits.  The
-method described here for doing so follows closely the method of
-estimating photometric redshifts used by \citet{csabai00a}.  It is
-designed to take advantage of what we already understand about galaxy
-spectral energy distributions.
-%In addition, as implemented here, the method is
-%equivalent to a fit to the star-formation history of the galaxies
-%using a spectrophotometric model; we will investigate the implications
-%of the model fits in a separate paper and concentrate here on the
-%empirical aspects of the results.
+At this point, the reader probably realizes that the number of
+templates $j$ should not be much higher than the number of bands $k$,
+or else the problem becomes very degenerate. In fact, for the case of
+unconstrained minimization of Equation \ref{chi2} the problem is
+formally degenerate if $j>k$. The nonnegativity constraint adds enough
+information that this is no longer the case, but still one cannot get
+strong constraints on a huge number of parameters given a small number
+of measurements. Nevertheless, below we {\it will} use a huge number
+of $j$ given the number of $k$, simply for the purposes of finding the
+subspace of possible galaxy spectra which actually conform to galaxy
+observations. We do not mean to imply that we can determine each of
+the resulting parameters at high signal-to-noise.
 
-Begin with an SED space defined by $N_b$ template galaxy SEDs
-$\vv{v}_i(\lambda)$ (for example from \citealt{bruzual93a}), where $N_b$ is
-large. In principle, this should be a complete set of galaxy SEDs if
-we want {\it all} galaxy SEDs to be in the space spanned by the
-$\vv{v}_i(\lambda)$; in practice, this property is not necessary.
-Given a set of $\vv{v}_i(\lambda)$, one can find an orthonormal set of
-spectra $\vv{b}_i(\lambda)$ spanning the same space.  To be more
-precise, we define the dot-product in the space of galaxy spectra to
-be:
-\begin{equation}
-\vv{v}_i \cdot \vv{v}_j =
-\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
-v_i(\lambda) v_j(\lambda),
-\end{equation}
-where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
-wavelength range used to define orthogonality. These limits are
-defined not over the full range covered by the $\vv{v}_i(\lambda)$
-spectra, but instead over a smaller range which is within observable
-wavelengths over most of the redshift range of the sample.\footnote{
-If the template spectra were identical in the subrange chosen and only
-differed outside of it, our choice would cause computational problems,
-but in fact the spectra are all independent in the restricted, as well
-as the full, wavelength range.}  We then define the basis of the SED
-space such that the $N_b$ template SEDs, $\vv{v}_i$, are linear
-combinations of $N_b$ basis spectra $\vv{b}_i$ and that $\vv{b}_i
-\cdot \vv{b}_j = \delta_{ij}$ (where $\delta_{ij}$ represents the
-Kronecker delta). These conditions do not fully specify the
-$\vv{b}_i$, which naturally may be rotated arbitrarily within our
-$N_b$-dimensional subspace of spectral space.
+For our model templates we use a set of spectra from the stellar
+evolution synthesis results of PEGASE2 (\citealt{fioc97a}). We use 15
+bursts of varying ages. The top panel of Figure \ref{sfh} shows the
+star-formation history of each of these bursts. For each age we create
+five templates with five different metallicities ($Z=0.03$, $0.02$,
+$0.008$, $0.004$, and $0.001$). The bottom panel of Figure \ref{sfh}
+shows the template for the $Z=0.008$ case at every other age, starting
+with the oldest. The bottom panel of For each age and metallicity, we
+create three templates with three different extinctions ($\tau_V=0$,
+$1$, and $3$) using the Milky Way extinction law using the ``clumpy,''
+``dusty'' model of \citet{witt00a}. This procedure results in 225
+templates and spans a very large range of possible spectra. Our
+procedure is designed to take this large dimensionality space and
+determine a low dimensionality subspace in which galaxies reside.
 
-In our case the number of observed bands is less than $N_b$, so we
-cannot determine an individual galaxy's position in this
-$N_b$-dimensional space from its broad band colors alone.  However, as
-outlined by \citet{csabai00a}, one can find a lower dimensional space
-defined by $N_t$ vectors $\vv{e}_k$ which best fits the {\it set} of
-galaxies, as follows. The SED of galaxy $i$ may be reconstructed from a
-linear combination of the basis spectra in this lower dimensional
-space:
-\begin{equation}
-\label{lincomb}
-f_{\mathrm{rec},i}(\lambda) d\lambda = 
-\sum_k a_{i,k} \sum_j e_{k,j} b_j(\lambda) d\lambda,
-\end{equation}
-where $a_k$ are the components of the galaxy projected in the
-low-dimensional space. From this SED it is easy to determine the
-reconstructed flux
-$F_{\mathrm{rec},il}$ in each survey bandpass by projecting
-the model SED onto the bandpass $l$: 
-\begin{equation}
-F_{\mathrm{rec},il} = \int_0^\infty d\lambda
-f_{\mathrm{rec},i}(\lambda) S_l(\lambda),
-\end{equation}
-where $S_l(\lambda)$ is the response of the instrument in band $l$.  We define
-\begin{equation}
-\chi^2 = \sum_i \sum_l
-\frac{(F_{\mathrm{obs},il}-F_{\mathrm{rec},il})^2}{\sigma_{il}^2},
-\end{equation}
-where $\sigma_{il}$ is the estimated error in the observed flux
-$F_{\mathrm{obs},il}$ in bandpass $l$ for galaxy $i$. Taking
-derivatives of $\chi^2$ with respect to $a_{i,k}$ and $e_{k,j}$
-results in a set of equations bilinear in $a_{i,k}$ and $e_{k,j}$.
-\citet{csabai00a} describe how to iteratively solve this bilinear
-equation by first fixing $e_{k,j}$ and fitting for $a_{i,k}$, then
-fixing $a_{i,k}$ and fitting for $e_{k,j}$, and iterating that
-procedure.
+We fit these models to a set of galaxies for which we know the
+redshifts. The majority of the redshifts are from the SDSS, but a
+small number of the redshifts are from CNOC2 ({\bf ref}). We consider
+only redshifts with 1$^\circ$.25 of the J2000 Equator, between
+$-60<\alpha<100$. We match the locations of these spectra to the SDSS
+photometric catalog ({\bf describe}) in all of the observations in
+this region and to the 2MASS Extended Source Catalog (XSC). In
+passing, we note that there are large errors (many arcseconds) in the
+catalog astrometry for CNOC2, and we needed to take care to account
+for these errors. For spectra which had matches in multiple runs of
+the SDSS, we take the (unweighted) mean of their photometry to reduce
+errors {\bf how many is this done for?}. We select a random set of
+these galaxies in such a way that the number per redshift is nearly
+constant across redshift. For each galaxy, we consider its
+extinction-corrected model photometry. Figure \ref{gmrz} shows the
+observed $g-r$ colors and redshifts of the resulting set of
+galaxies. Note that only for the lower redshift galaxies do we have
+matches to the 2MASS XSC.
 
-%We put one additional constraint on our SEDs. To regularize the edges
-%of the SEDs, we define two additional bands blueward and redward of
-%$u$ and $z$, respectively, which we will call $a_1$ and $a_2$. The
-%bandpasses are simply Gaussians centered at 2700\AA\ and 10200\AA\
-%with standard deviations of 250\AA\ and 350\AA, respectively. From
-%studying synthetic galaxy SEDs redshifted between $z=0$ and $z=0.2$,
-%it is clear that the $u-a_1$ color varies not much more than 1.5
-%magnitudes around the value 1.5, and that the $z-a_2$ color usually
-%varies only by about 0.2 magnitudes around 0.1. Thus, given the $u$
-%and $z$ magnitude, we can put a very weak constraint on the $a_1$ and
-%$a_2$ magnitudes.  Similarly, when one of the SDSS bands is
-%unobserved, we replace it based on an adjacent band and the mean color
-%for galaxies in the sample, with a large error bar. These weak
-%constraints prevent the inferred galaxy SEDs from becoming too
-%unreasonable in the regions of the spectrum without data.
+Figure \ref{kfit} shows the residuals between the observed and fit
+colors after minimizing Equation \ref{chi2} for all the galaxies in
+our sample. Interestingly, even this large set of parameters cannot
+perfectly describe the trends of color with redshift. There are
+deviations at the percent level or so in $ugriz$. The $J$-band 2MASS
+observations at moderate redshift appear to be about $0.1$ magnitudes
+off of the models; this discrepancy is probably related to the similar
+discrepancies found by \citet{blanton03d} and \citet{baldry03a} when
+comparing the luminosity density of the universe to models.  Our
+attitude in this paper will be to trust the data in preference to the
+models and so below we describe how we deal with these discrepancies.
 
-We define the ``total flux'' $f_k$ of each template as the flux in the
-range $\lambda_{\mathrm{min}}<\lambda<\lambda_{\mathrm{max}}$; then an
-estimate of the flux in this range is $F_{{t}}=\Sigma_k a_k f_k$.
-Galaxies of some fixed $F_t$ clearly lie on a plane in the component
-space $a_k$. Therefore, a natural choice for the orientation of the
-axes $\vv{e}_k$ is to let $\vv{e}_0$ be perpendicular to the planes of
-constant $F_t$; in this manner, the coefficient $a_0$ in Equation
-(\ref{lincomb}) is directly proportional to $F_t$ and is thus purely a
-measure of the object's flux (in the wavelength range
-$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$), while the
-parameters $a_i/a_0$ (where $i>0$) primarily measure the shape of the
-galaxy SED in the wavelength range $\lambda_{\mathrm{min}} <\lambda <
-\lambda_{\mathrm{max}}$. $F_t$ can be trivially related to $L_t$, the
-total luminosity in this range, by the inverse square of the
-luminosity distance (from, for example, \citealt{hogg99a}).  As it
-turns out, the parameters $a_i/a_0$ tend to be distributed
-approximately in an ellipsoid, with not much curvature within the
-space defined by the template SEDs. Thus, we transform the axes such
-that $a_i/a_0 = 0$ is the mean of the distribution and rotate the axes
-$\vv{e}_i$ (for $i>0$) such that they are aligned with the principal
-axes of the ellipsoid. Furthermore, we normalize the vectors
-$\vv{e}_i$ such that the zeroth component $a_0=F_t$ is output by the
-code in units of ergs cm$^{-2}$ s$^{-1}$ (within the range
-$\lambda_{\mathrm{min}} <\lambda < \lambda_{\mathrm{max}}$).
+Each many-parameter fit to the broad-band fluxes of each galaxy yields
+a complete spectrum from which to derive $K$-corrections. It also
+provides a value of $\chi^2$; in principle, for galaxies without
+redshifts we can find the assumed redshift with minimum $\chi^2$ in
+order to find the photometric redshift. In practice, the fit takes too
+long to perform to be useful for either of these purposes. Thus, we
+describe here our method for reducing the dimensionality to a set of
+three templates which adequately describe the data.
 
-In this way, we can reconstruct SEDs based on the broad band
-magnitudes of galaxies. From these reconstructed SEDs one can estimate
-$K$-corrections, develop a measure of galaxy spectral type, or
-synthesize other measurements of galaxy flux.  The templates
-determined during the procedure can, of course, be used to estimate
-photometric redshifts, as \citet{csabai00a} describes.
+Our procedure has two steps. First, we use principle component
+analysis on the set of 225-coefficent vectors describing all the fits
+in order to find a 10-dimensional subspace which still describes the
+data. Second, within this 10-dimensional subspace we group the
+galaxies into a set of ten groups, average the spectra within each
+group, and pick the triple among these groups which best describes the
+data. We adopted this procedure because we did not know of a
+principled way in which to reduce the dimensionality without
+sacrificing nonnegativity; however, it turns out that such methods do
+exist, and we recommend those methods for future investigators ({\bf
+refs from sam}) while retaining this method for the moment because it
+appears to work for our purposes.
 
-\subsection{Applying Direct Constraints on $a_i/a_0$}
-\label{direct}
+Because the galaxies all have very different bolometric fluxes, there
+is a strong variation in the coefficient values simply due to the
+total flux differing by large factors. Therefore, before performing
+principal component analysis, we scale all the coefficients of each
+galaxy by a single number (per galaxy) such that the galaxies all lie
+exactly on a plane in coefficient space. We choose the plane to be the
+plane on which the sum of the templates has a constant flux between
+1500 \AA\ and 30000 \AA. After scaling the galaxies to this plane, we
+perform principal component analysis to obtain the five most principal
+axes, using the efficient method of \citet{roweis99a}. There is almost
+no difference between the full fit shown above and that using the
+first five eigencoefficients.
 
-Sometimes we will want to apply direct constraints on the coefficients
-$a_i/a_0$ of the galaxy we are fitting. For example, in this paper we
-fit our templates to a set of galaxies with high quality photometry in
-all bands. However, we would like to use these templates to fit for
-the SEDs of other galaxies, which might have much worse photometry in
-one or more bands. Since the errors in the photometry can be
-catastrophic (that is, non-gaussian), when we fit the SEDs of the
-lower signal-to-noise galaxies it makes sense to at least weakly
-constrain these galaxies to have reasonable SEDs, in the sense that
-the $a_i/a_0$ ought not to be too different than those for
-well-measured galaxies. We can assume a gaussian prior probability
-distribution with a mean $\avg{a_{c,j}/a_{c,0}}$, a variance matrix
-$V_{c,jl}$, and a weight $A_c$.  We calculate a preliminary value of
-${\hat{a}}_{c,0}$, assuming $a_i = 0$. Then we constrain our fit by
-adding a term to $\chi^2$ as follows:
-\begin{equation}
-\sum_j \sum_l A_c V^{-1}_{jl}
-\left(\frac{a_l}{{\hat{a}}_{c,0}}-\Avg{\frac{a_{c,l}}{a_{c,0}}}\right)
-\left(\frac{a_j}{{\hat{a}}_{c,0}}-\Avg{\frac{a_{c,j}}{a_{c,0}}}\right)
-\end{equation}
-The resulting $\chi^2$ can still be minimized linearly, so we have
-sacrificed very little in speed.  Note that with such a constraint,
-one can calculate $K$-corrections even for galaxies with only one or
-two bands measured, and still obtain sensible results.
+On the other hand, the five eigencomponents natural consist partly of
+positive star-formation components and partly of negative
+star-formation components. This means that we cannot simply sum them
+nonnegatively to get a nonnegative star-formation history. In order to
+get a set of components that consists purely of all-positive
+star-formation histories, we consider the distribution of galaxies in
+the five-component space. We perform a $K$-means grouping on this set
+of points, which produces ten groups. This is simply a way of very
+roughly mapping the locus on which the galaxies reside (it is
+significantly nonlinear within the five-dimensional space, so we
+cannot simply take fewer principal components). We then average the
+spectra in each group. This gives ten spectra who roughly span the
+space of galaxy types. Finally, we pick from among those ten the
+triple which best describes the galaxies, in the sense of minimizing
+$\chi^2$ in Equation \ref{chi2}.
 
-\subsection{Calculating $K$-corrections}
+Figure \ref{kfit3} shows the color residuals after fitting with the
+final three templates.
 
-To characterize $K$-corrections, consider again Equation
-(\ref{kcorrecteqn}): 
-\begin{equation}
-m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z).
-\end{equation}
-The definition of an apparent AB magnitude is given by Equation
-(\ref{ABdef}). An absolute AB magnitude is defined to be the apparent
-AB magnitude which would be measured for a galaxy 10 pc distant, at
-rest.  In the restricted case of AB magnitudes, the resulting
-$K$-correction is given by:
-\begin{equation}
-K_{QR}(z) = -2.5 \log_{10} \left[
-\frac{
-\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
-}{
-\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) Q(\nu)
-}
-\right]
-\end{equation}
-A more general expression holds for other magnitude systems (such as
-the Vega-relative magnitude system).
+[ new model: 3 PEGASE2 spectra; Figure show spectra, Figure stellar
+  mass to light ratios ]
 
-In practice, we will be concerned here primarily with transforming the
-apparent magnitudes in band $R$ to a band $Q$ which is just the $R$
-band shifted by a factor $(1+z_0)$, which we will denote
-\begin{equation}
-\band{z_0}{R}(\nu) \equiv R(\nu/(1+z_0))  
-\end{equation}
-read, ``$R$-band shifted to $z_0$.''
+[ to deal with that for K-correction and photo-z, tweak ]
 
-In the case of transforming $R$ to $\band{z_0}{R}$
-\begin{equation}
-K_{\band{z_0}{R}R}(z) = -2.5 \log_{10} \left[
-\frac{
-\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
-}{
-\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) R(\nu/(1+z_0))
-}
-\right]
-\end{equation}
-In the special case that $z=z_0$, one is transforming the magnitude of
-an observed galaxy at $z_0$ to its rest-frame magnitude in a band
-shifted by $z_0$. For that case
-\begin{equation}
-K_{\band{z_0}{R}R}(z_0) = -2.5 \log_{10} (1+z_0)
-\end{equation}
-independent of $f(\nu)$ (which is natural, since both magnitudes
-sample the same region of $f(\nu)$). The fact that one's uncertainties
-in $f(\nu)$ do not matter makes the use of the fixed frame magnitudes
-shifted to the redshift of the object in question useful, when it is
-possible. We note in passing that \citet{oke68a} state that the
-$K$-correction ``would disappear if intensity measurements of
-redshifted galaxies were made with a detector whose spectral
-acceptance band was shifted by $1+z$ at all wavelengths;'' here we
-have just shown this statement to be untrue for an AB system (though
-one could define a magnitude system in which this were true).
+We recognize that this procedure is rather convoluted and a bit of a
+hack. As noted above, we have adopted it simply as one way of
+automatically getting a small set of templates that explain galaxy
+broad-band fluxes as a sum of star-formation histories. There are
+methods for directly performing principal component analysis subject
+to the conditions that the components and the coefficients are all
+positive, and we recommend the investigation of such methods to the
+reader ({\bf ref}).
 
-%Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
-%redshift $z$, one can use the method of the previous subsection to
-%reconstruct the galaxy rest-frame SED.  With this reconstruction in
-%hand, it is straightforward to project the SED onto any bands at any
-%redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
-%\begin{eqnarray}
-%\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
-%\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
-%f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
-%d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
-%\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
-%f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
-%R(\nu)}\right]
-%\end{eqnarray}
-%The simple way to remember where the factors of $(1+z)$ appear in
-%these equations is to remember that $K$-corrections are defined to be
-%the change in magnitude with redshift for {\it fixed bolometric flux}.
-%So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
-%in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
-%is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
-%counteract the stretching of the spectrum.
+\section{Determining $K$-corrections} 
+\label{kcorrect}
 
-%We here define a notation which we will use in subsequent papers for
-%the magnitude in a band $b$ which we have shifted blueward by a factor
-%$(1+z)$, namely:
-%\begin{equation}
-%\band{z}{b}
-%\end{equation}
-%read, ``$b$-band shifted to $z$.''  The goal of ``$K$-correction'' is
-%to convert the magnitudes of a galaxy at redshift $z$ observed in the
-%$ugriz$ bands at redshift $z$ to the magnitudes which would be
-%measured if the galaxy were at redshift at $z=0$ and observed through
-%the $ugriz$ bands shifted to some fixed redshift $z_0$. While
-%traditionally $K$-corrections are performed to $z_0=0$ (that is, the
-%bands are not shifted), this is not always desirable, particularly for
-%surveys like the SDSS for which most of the galaxies are not observed
-%near $z=0$.
+Given a model spectrum for the galaxy, the determination of the
+$K$-correction is straightforward. Here we repeat the
 
-In practice, one can calculate the fixed frame magnitudes in one of
-two ways:
-\begin{enumerate}
-\item Simply reconstruct $\band{z_0}{R}$ from the model
-$f_{\mathrm{rec}}(\nu)$ 
-\item Calculate the $K$-correction above and apply it to the observed
-$R$ magnitude
-\end{enumerate}
-If the reconstructions of the observed magnitudes $R$ are very similar
-to the actual observations (as we will show is the case for our fits),
-there is very little difference between these two methods. However, if
-the reconstructions are poor, there will be large differences between
-the methods; in particular, method (1) will yield a color distribution
-of galaxies with artificially low dimensionality. Having estimated
-$\band{z_0}{m}$, you can then calculate the luminosity by simply
-applying to the calculated magnitude the cosmological distance modulus
-(that is, the luminosity distance-squared law, as tabulated by, {\it
-e.g.}, \citealt{hogg99a}).
+[ Figure demonstrate K-corrections ]
 
-We want to emphasize here that, while $K$-correcting to a fixed frame
-bandpass is {\it sometimes} necessary in order to achieve a scientific
-objective, it is not {\it always} necessary or appropriate. Because
-$K$-corrections are inherently uncertain (the broad band magnitudes do
-not uniquely determine the SED) they should be avoided or minimized
-when possible. For example, if one were calculating the evolution of
-clustering of red and blue galaxies separately, it would perhaps be
-wise to perform the separation not on the $K$-corrected colors of the
-galaxies but on the median color of $M_\ast$ galaxies as a function of
-redshift. Similarly, in situations where $K$-corrections cannot be
-ignored, such as the calculation of the evolution of the luminosity
-function, their effect should be minimized by, for example, correcting
-to a band shifted to the median redshift of galaxies in the sample.
+[ Figure show K-corrections to z=0 ]
 
-Typically one's measurements will be difficult to connect to solar
-luminosities in $\band{z}{b}$ (unless $z=0$), simply because nobody
-has projected the appropriate stellar spectrophotometry onto these
-bandpasses. However, it is our position that stars are better
-understood than galaxies, and that it is therefore simpler in the end
-to stay as close as possible to the system in which the galaxies are
-observed. In any case, astronomy is quickly reaching a level of
-precision for which the exact nature of the bandpasses used has to be
-known and considered in most analyses of observational data.
+[ note band-shifting ]
 
-The reader may ask why, if we are basing our $K$-corrections on a full
-model of the galaxy SED, we do not correct to bandpasses with simpler
-shapes (for example, top hats). The reason is that we want the
-effective bandpasses to actually have been observed for some galaxies
-in the sample, so that the $K$-correction for those galaxies are
-independent of the galaxy SED.  This property is desirable because, as
-noted above, there are inherent uncertainties in the $K$-corrections
-due to our lack of knowledge of the SEDs. Nevertheless, we note that
-synthesizing top hat magnitudes may be appropriate in certain
-situations, as they have the virtue (similarly to the $AB$ magnitude
-system) that they are very easy to comprehend and synthesize from
-theory.
-%For example, if we wanted to convert our results into a
-%form easily comprehended by the hypothetical future astronomers
-%discussed in the first paragraph of the introduction, it would be best
-%to calculate fluxes within wide top-hat filters so that the
-%difficulties of understanding the magnitude system were minimized.
+[ Figure test: against a bunch of spectra ]
 
-\subsection{Public Access to the Code}
+\section{Determining photometric redshifts}
+\label{kphotoz}
 
-The version of the $K$-correction code (\kversion) implementing the
-calculations described here, along with eigentemplates and filter
-curves, is publicly available on the World Wide Web at {\tt
-http://physics.nyu.edu/\~\ mb144/kcorrect}.  The whole of the code can
-be used through the Research Systems, Incorporated, IDL language;
-everything except for the template-fitting also exists in stand-alone
-C programs (which call the same routines, guaranteeing consistency).
+[ describe simple fitting procedure ]
 
-\section{Application to SDSS Data}
-\label{data}
+[ Figure show training set results ]
 
-In this section we describe how we applied the above method to the
-SDSS data set.
+[ Figure show results on deeper sdss data ]
 
-\subsection{SDSS Spectroscopic Data}
+\section{Public access to the code}
 
-The SDSS (\citealt{york00a}) is producing imaging and spectroscopic
-surveys over $\pi$ steradians in the Northern Galactic Cap. A
-dedicated 2.5m telescope (Siegmund {\it et al.}, in preparation) at
-Apache Point Observatory, Sunspot, New Mexico, images the sky in five
-bands between 3000 and 10000 \AA\ ($u$, $g$, $r$, $i$, $z$;
-\citealt{fukugita96a}) using a drift-scanning, mosaic CCD camera
-(\citealt{gunn98a}), detecting objects to a flux limit of $r'\sim
-22.5$. The ultimate goal is to spectroscopically observe 900,000 Main
-Sample galaxies (down to $r_{\mathrm{lim}}\approx 17.77$;
-\citealt{strauss02a}), 100,000 Luminous Red Galaxies (LRGs;
-\citealt{eisenstein01a}), and 100,000 QSOs (\citealt{fan99a};
-\citealt{richards02a}.  This spectroscopic follow up uses two digital
-spectrographs on the same telescope as the imaging camera. Many of the
-details of the galaxy survey are described in a description of the
-galaxy target selection paper (\citealt{strauss02a}). Other aspects of
-the survey are described in the Early Data Release (EDR;
-\citealt{stoughton02a}).
-
-The results of the photometric pipeline for all of the galaxies were
-extracted from the SDSS Operational Database. The photometry used here
-for the bulk of these objects was the same as that used when the
-objects were targeted. However, for those objects which were in the
-EDR photometric catalog, we used the better calibrations and
-photometry from the EDR. The versions of the SDSS pipeline PHOTO used
-for the reductions of these data ranged from {\tt v5\_0} to {\tt
-v5\_2}. The treatment of relatively small galaxies, which account for
-most of our sample, did not change substantially throughout these
-versions. As described in \citet{smith02a}, magnitudes are calibrated to a
-standard star network approximately in the $AB$ system, described in
-Section \ref{motivation}.
-
-The response functions $R(\lambda)$ of the filter-CCD combinations
-have been measured using a monochrometer by Mamoru Doi.  Using a model
-for the atmospheric transmission and the reflectivity of the primary
-and secondary mirrors, one can then model the response of the entire
-system. For each bandpass in the SDSS, there are six different CCDs;
-it has been shown that the differences are small between these CCD and
-filter responses for each bandpass.  The resulting set of filter
-curves is shown in Figure \ref{response_sdss}, in comparison to a
-model of a galaxy spectral energy distribution observed at $z=0$ (a 4
-Gyr old instantaneous burst from the models of \citealt{bruzual93a}).
-
-The results of the spectroscopic observations are treated as follows.
-We extract one-dimensional spectra from the two-dimension images using
-a pipeline ({\tt specBS v4\_8}) created specifically for the SDSS
-instrumentation (\citealt{schlegel02a}), which also fits for the
-redshift of each spectrum. The official SDSS redshifts are obtained
-from a different pipeline (\citealt{subbarao02a}). The two independent
-versions provide a consistency check on the redshift determination,
-and for galaxies they agree on more than 99\% of the
-objects.\footnote{The disagreement is more significant for objects
-with unusual spectra, such as certain types of stars and QSOs.}
-
-We use two types of magnitudes determined by the SDSS. First, the SDSS
-Petrosian magnitudes, a modified form of the magnitude described by
-\citet{petrosian76a}, as described in \citet{strauss02a}.  Petrosian
-magnitudes are circular aperture magnitudes with an aperture size
-which is determined by the shape of the radial profile (not its
-amplitude). The resulting apertures are empirically nearly constant in
-metric size as a function of redshift; in addition, all the bands use
-the same aperture, so the measured SED corresponds (to within the
-effects of seeing) to the SED of an identifiable region of the
-galaxy. However, for faint objects, the Petrosian magnitudes tend to
-become noisy. Thus, for galaxies with $z>0.25$ we instead use the
-higher signal-to-noise ``model magnitudes.'' Model magnitudes are
-calculated in all bands using a single weighted aperture convolved
-with the point spread function; thus, the measured colors again
-correspond to an identifiable region of the galaxy, though a different
-one than would be measured by the Petrosian magnitude. The weighted
-aperture is the better fitting model (pure exponential or pure de
-Vaucouleurs) to the galaxy image in the $r$-band. Thus, the model
-magnitudes weight the centers of galaxies more strongly than do
-Petrosian magnitudes. In \citet{stoughton02a} we are explicitly warned
-not to mix Petrosian and model magnitudes in a single analysis, since
-they measure galaxies in very different ways. Nevertheless, for the
-purposes of contraining galaxy template SEDs, it is perfectly
-acceptable to use {\it any} well-defined part of any galaxy.  For our
-purposes, it is more important to have high signal-to-noise
-measurements of galaxy colors than to measure exactly the same regions
-of galaxies at low and high redshift.  We extinction-correct both
-types of magnitudes using the dust maps of \citet{schlegel98a}.
-
-\subsection{Fitting SEDs to SDSS Data}
-
-For the purposes of using the SDSS data to constrain our template set,
-we take a subsample of the data consisting of around 30,000 objects in
-the range $0.0<z<0.5$. The sample includes both the main sample and
-the LRGs, and is designed such that there is an approximately even
-distribution over redshift within our range of redshifts. In addition,
-we add results from galaxies in several spectroscopic plates
-(totalling about 1,000 objects) which were selected by a photometric
-redshift algorithm (Csabai {\it et al.}, in preparation) to be at
-around $z\sim 0.3$--$0.4$, and subsequently observed
-spectroscopically. These objects are invaluable for tying down the
-blue end of the templates.  Finally, we exclude galaxies in the
-redshift range $0.28 <z<0.32$ from the fit for the templates, for
-reasons which we will explain more fully below (nevertheless, we still
-can and do use the resulting templates to analyze galaxies in this
-range).
-
-Some of the objects have missing or poorly constrained data. For
-example, the $u$- or $z$-band fluxes for some objects are swamped by
-the photon noise of the sky. We identify such cases as magnitude
-errors greater than 2.0 or magnitudes fainter than 24.0 in any
-band. We ignore these objects entirely when fitting for the
-templates.
-
-In addition, the photometric errors for most objects in the
-spectroscopic galaxy sample of the SDSS are not dominated by the
-estimated errors on the photometry listed in the catalog. Instead, the
-errors are dominated by local calibration errors and other systematic
-effects, which are poorly known. To account for these errors, we add
-extra error terms in quadrature with the errors listed in the catalog
-(0.05 mags, 0.02 mags, 0.02 mags, 0.02 mags, and 0.03 mags for
-$ugriz$, respectively). The choice of these values is based on a
-qualitative sense of the photometric errors present in the data. Not
-accounting for this extra source of errors can cause the fits to be
-ill-behaved.
-
-For the SED space (our $\vv{v}_i(\lambda)$), we use the subspace
-defined by ten Bruzual-Charlot instantaneous burst models with ages
-ranging from $3 \times 10^7$ to $2\times 10^{10}$ years, five with
-metallicity $Z=0.02$ and five with $Z=0.004$, all assuming a Salpeter
-Initial Mass Function. Since we cannot recover information about the
-SED below a certain wavelength resolution, we smooth most of the
-wavelength regime of the templates using a Gaussian with a standard
-deviation of $\sigma = 300$\AA; for the region of the spectrum which
-contains the sharpest gradients, around 4000\AA, we smooth only with
-$\sigma = 150$\AA. None of our results change dramatically if we vary
-our smoothing procedure or if we add reddened templates to our allowed
-SED space.
-
-We choose to fit for $N_t = 4$ eigentemplates, the maximum one can use
-and still allow freedom to fit for the templates themselves.  Simply
-using five templates (which obviously reproduces all the magnitudes
-exactly) tends to yield unphysical trends of galaxy SED versus
-redshift ({\it cf.}  Figure \ref{main_colors_plot} below). Using three
-templates does nearly as well as four templates in the sense that the
-resulting templates reproduce the $griz$ magnitudes nearly as
-well. However, the fourth template is necessary to recover the
-$u$-band flux to better than about 15\%. In addition, since one of the
-applications of these SED determinations is the distribution of galaxy
-colors in fixed frame magnitudes, we don't want to artificially reduce
-the dimensionality of the color space to only two.
-
-One more choice needs to be made, the wavelength regime over which to
-orthogonalize the templates and over which to calculate the flux. We
-choose the range defined by $\lambda_{\mathrm{min}}=3500$\AA\ and
-$\lambda_{\mathrm{max}}=7500$\AA, since this rest frame range is
-observed for almost all galaxies in the sample. We refer here and in
-other papers to the flux and luminosity in this range as the ``visual
-flux'' $f_v$ and the ``visual luminosity'' $L_v$.
-
-Once we have finished fitting for the templates, we still want to
-determine a best-fit SED for each object.  Thus, after the templates
-have been determined, we fit for the coefficients $a_i$ for each
-galaxy, including those excluded from the template-fitting procedure
-because of large errors. In order to accommodate the presence of
-errors, we constrain all the SED fits to all galaxies using the method
-described in Section \ref{direct}. As a constraint, we use the
-covariance matrix and mean value (essentially zero, by design) of the
-coefficients of the well-measured set of galaxies used to determine
-the templates; we multiply all the elements of the covariance matrix
-by a factor of four to weaken our constraint (which is appropriate,
-since the actual distribution of coefficients is not well described by
-a single gaussian).  Thus, we account for any missing information by
-simply requiring that the object SED have ``reasonable''
-properties. The constraint does not affect the results for galaxies
-which have well-measured magnitudes in all bands.
-
-\subsection{The $g$/$r$ Gap}
-\label{grgap}
-
-In Figure \ref{response_sdss}, there is clearly a gap between the $g$
-and $r$ bands. Linear fits to the data allow considerable freedom in
-the resulting reconstructed SEDs. As it happens, one of the directions
-in our best fit coefficient space $a_i/a_0$ corresponds to a large
-spike at around 4000 \AA; that this direction is important is not
-surprising, since one of the most variable quantities of galaxy SEDs
-is the size of the 4000 \AA\ break. However, the gap between the $g$
-and $r$ bands corresponds to the 4000 \AA\ break at around
-$z=0.3$. The existence of this spike in one direction in our
-four-dimensional space means that galaxies at $z=0.3$ can vary along
-this direction in order to better fit $u$, $i$, and $z$, with little
-effect on the $g-r$ color of the fit. This results in unphysical SED
-fits to galaxies near $z=0.3$.  One really only needs to worry about
-this effect in the SDSS when one is dealing with the LRGs, for
-galaxies in the range about $0.27 < z < 0.33$. The Main Sample results
-are not significantly affected at all (since few Main Sample galaxies
-are at $z>0.25$.  Furthermore, one can still observe a galaxy at
-$z=0.1$ and reliably infer what it would look like at $z=0.3$; it is
-only the reverse process which is difficult.
-
-To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
-corresponding to the direction represented by the vector $ \vv{e}_2 +
-0.7 \vv{e}_3$, for our best fit templates. We overplot the
-\band{0.3}{g} and \band{0.3}{r} bands. This direction has a strong
-peak in the gap between the two bands at $z=0.3$.  This difficulty is
-why, in Section \ref{data}, we exclude the regime around $z\sim 0.3$;
-otherwise the templates became distracted by the degeneracies in this
-range.
-
-We have explored using a larger set of input spectra, by including
-reddened versions of all our templates, to test whether a more
-complete space of SEDs would cause our fit to choose a more realistic
-subspace. In addition, we have tried removing the increased resolution
-of the template SEDs around the 4000 \AA\ break. Neither of these
-tests yielded better results.
-
-Our opinion is that the ideal regularization which would solve this
-problem would be to minimize $\chi^2$ under the constraint that none
-of the original Bruzual-Charlot spectra defining our SED space
-contribute negatively to the SED fit. Such a solution would almost
-certainly have better properties than our linear approach here: it
-would almost certainly avoid the degeneracies, it would yield
-reasonable fit SED over a larger wavelength range, it would associate
-a reasonable star-formation history with each object (which could be
-used to evolution-correct the magnitudes).
-
-However, we have not tackled this approach yet. As it happens, the
-direct constraint on the coefficients described in the previous
-section is sufficient to suppress the $g$/$r$ gap problem 
-for many purposes, as long as one $K$-corrects the magnitudes to
-$z=0.3$. In this way, one minimizes the $K$-corrections for the
-galaxies which one is least certain of.
-
-%It is clear from Figure \ref{lrg_colors_plot} that $z=0.3$ is a
-%special redshift regime where our fits become unphysical. In fact, it
-%is easy to demonstrate that the diagonal spur (going from the upper
-%left to the lower right) around the red dot in the upper right panel
-%of Figure \ref{k_coeffdist_plot} is due almost completely to galaxies
-%at $z\sim 0.3$. The spur arises because the 4000 \AA\ break lies in
-%the gap between $g$ and $r$ at $z=0.3$, visible in Figure
-%\ref{response_sdss}. Our algorithm has strong gradients in the
-%spectrum around the 4000\AA\ break, which is necessary to properly
-%model galaxy spectra. In the presence of these sharp gradients, the
-%gap in the bands causes certain combinations of the template spectra
-%to be degenerate at around $z=0.3$.
-
-%The question arises of what to do about this problem. One solution is
-%to simply work around it. It is clear we cannot use the fits presented
-%here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
-%reliably --- their coefficients $a_k$ are unusual and probably
-%unphysical. However, nothing prevents one from $K$-correcting galaxies
-%at other redshifts {\it to} $z=0.3$. If you are considering the LRG
-%sample, the median redshift is close to $z=0.3$ anyway, so this is a
-%viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).
-
-%A second solution is to place extra constraints on the fit which
-%prevent the spur in the coefficients. One can imagine fitting the
-%distribution in the $(a_1/a_0)$-$(a_3/a_0)$ plane with a simple model,
-%such as a multigaussian fit, and using the resulting fit as a prior in
-%one's fit (with gaussians, the extra constraints would be linear and
-%thus easily accounted for). One could design these constraints to be
-%strong enough (at least near the redshifts in question) to prevent the
-%spur. However, the disadvantage of this solution is that for some
-%scientific projects the change in the galaxy SED is exactly what one
-%is trying to measure. In these cases, it may be inappropriate to be
-%putting explicit {\it a priori} constraints on exactly what one is
-%trying to measure.
-
-%A third solution, not adopted here, is to place stronger constraints
-%in the SED fit on the form of the 4000 \AA\ break. One obvious way of
-%doing this is to use the spectra to constrain the size of the break
-%(making it a weak constraint to prevent biases). This constraint could
-%be applied in a linear fashion to retain the speed of our method ---
-%except for the fact that you would have to read in the spectra to
-%apply the $K$-correction. For simplicity, and because the aforemention
-%work-around is so easy to obey, we have not implemented this solution
-%here.
-
-\subsection{Results}
-
-%The resulting four templates are shown in Figure \ref{k_espec_plot} in
-%the range $2000<\lambda<12000$\AA, the region which is constrained by
-%our observations. 
-The top two panels of Figure \ref{k_coeffdist_plot} show the pairwise
-joint distributions of $a_1/a_0$, $a_2/a_0$, and $a_3/a_0$ for a
-random subset of about 10,000 of the galaxies in the SDSS sample (not
-only the ones which we used to fit the templates). In the bottom
-panel, three spectra taken from a one-dimensional sequence along the
-galaxy locus are shown, showing that the spectra become progressively
-bluer along that sequence. Note that constraints on the SED become
-poor at the bluest and reddest edges of this diagram (the peak of the
-$z$-band response is at 8700~\AA rest frame for a galaxy observed at
-$z=0$). Therefore, the odd behavior near the edges should not be taken
-too seriously. In addition, note the small spur extending from lower
-left to upper right through the right edge of the
-($a_3/a_0$)-($a_2/a_0$) plane. Many of the objects in this spur are at
-around $z=0.3$, and this spur is the result of the degeneracy due to
-the $g$/$r$ gap.
-
-These reconstructed spectra do an excellent job of reproducing the
-observed galaxy fluxes. Figure \ref{k_model_plot} shows the
-differences between the observed and reconstructed fluxes as a
-function of redshift. There are no systematic trends with redshift,
-and the standard deviations of the differences between the observed
-and the reconstructed fluxes (shown on the right bottom corner of each
-panel) are of order the photometric errors in the sample. In
-particular, for our 30,000 galaxies (and thus approximately 30,000
-degrees of freedom) we find $\chi^2 \sim 55,000$, indicating that our
-fit is reasonable.
-
-An important test of the consistency of the fits is to check that for
-a fixed type of galaxy, the distribution of the fixed frame colors
-depends only weakly on redshift. Because the galaxies shown in Figures
-\ref{k_coeffdist_plot} and \ref{k_model_plot} are inhomogeneously
-selected --- some are Main Sample galaxies, some are LRGs, and some
-are selected from the photometric redshift plates --- we will split
-our sample into two well-defined sets of objects to perform this
-test. First, we choose a set of Main Sample galaxies in the luminosity
-range $-21.5 < M_{\band{0.1}{r}} < -21.2$. Figure
-\ref{main_colors_plot.z} shows the observed-frame colors of all the
-galaxies as a function of redshift.  The right-hand panels show the
-distribution for redshifts $0.05<z<0.10$ (solid histogram) and
-$0.10<z<0.17$ (dotted histogram). A trend exists in all colors, most
-strongly in $\band{z}{(g-r)}$. Figure \ref{main_colors_plot} shows the
-$K$-corrected colors $\band{0.1}{(u-g)}$, $\band{0.1}{(g-r)}$,
-$\band{0.1}{(r-i)}$, and $\band{0.1}{(i-z)}$ in the same manner.  The
-plots show a general consistency in the fixed frame color distribution
-of these objects with redshift.  Small changes are discernible in the
-distributions, mostly attributable to the increased errors at higher
-redshift. Note that for $z<0.1$ the $\band{0.1}{(u-g)}$ color depends
-on an extrapolation of the SED in the blue, while for $z>0.1$ the
-$\band{0.1}{(i-z)}$ color depends on an extrapolation of the SED in
-the red.
-
-Second, we choose a set of Cut I LRGs (see \citealt{eisenstein01a} for
-details) with $-22.8 < M_{\band{0.3}{r}} < -22.5$. Figure
-\ref{lrg_colors_plot.z} shows the observed colors as a function of
-redshift for these objects. Again, there is a strong redshift
-dependence. Figure \ref{lrg_colors_plot} shows the colors
-$K$-corrected to $z=0.3$, which appear approximately constant with
-redshift.
-%Again, for $z<0.3$ the $\band{{0.3}}{(u-g)}$ color relies
-%on an extrapolation, which is probably responsible for the scatter in
-%that color at low redshifts. Otherwise, the distribution of colors is
-%remarkably constant. 
-$\band{0.3}{(g-r)}$ (which is very similar to $\band{0.0}{(u-g)}$)
-experiences a blueward shift of about 0.1 magnitudes between $z=0.1$
-and $z=0.45$, which is attributable to passive evolution, though it
-could also be due to the LRG selection procedure.
-
-In short, these fits to galaxy SEDs provide estimates which reproduce
-the galaxy photometry nearly to the level of the errors in the
-photometry itself, seem physically reasonable, and are consistent over
-the range of redshifts we consider ($0.0 < z < 0.5$).
-
-\subsection{$K$-corrections in the SDSS}
-
-We show, in Figure \ref{k_kcorrect_plot}, the resulting $K$-corrections to
-$z=0.3$ inferred from this method, for all five SDSS bands. Note that
-the $K$-corrections are largest (and thus most uncertain) in
-$\band{0.3}{u}$ and $\band{0.3}{g}$. Remember that the $K$-corrections
-to $\band{0.3}{u}$ are extrapolations for $z<0.3$ and that the
-$K$-corrections to $\band{0.3}{z}$ are extrapolations for $z>0.3$, so
-those results should not be taken too seriously. (Although the
-$K$-corrections given are not {\it too} unreasonable).
-
-An important test of our method is to synthesize broad band photometry
-from spectra, and then try to recover the $K$-corrections in a case
-where we know the spectrum completely. For this purpose, we use as
-example spectra the spectra of galaxies in the SDSS.  Since we cannot
-synthesize the observed $u$ and $z$ bands from these spectra, we
-simply base them on the synthesized $g$ and $i$ band magnitudes plus
-the actual $u-g$ and $z-i$ colors. (This procedure is unlikely to make
-the photometric estimations of the $K$-corrections {\it
-better}). Finally, we added 2\% random errors to all of the
-measurements. Figure \ref{k_speck_plot.fitfib.0.1} compares the
-$K$-corrections from the synthesized photometry to that calculated
-from the spectrum itself:
-\begin{equation}
-\Delta K = K_{\mathrm{spec}} - K_{\mathrm{photo}}
-\end{equation}
-Since all of the photometry used in this test is actually synthesized
-from the spectra, this is {\it not} a comparison of photometric and
-spectroscopic $K$-corrections; it is {\it only} a test of how well our
-method recovers $K$-corrections. The agreement in FIgure
-\ref{k_speck_plot.fitfib.0.1} is very good, suggesting that our method
-can indeed recover the correct $K$-corrections based only on broad
-band magnitudes.
-
-To test how robust these $K$-correction results are to our model
-assumptions, we compare them to other methods of calculating fixed
-frame galaxy magnitudes.  An extremely simple method is to calculate
-the flux in any desired bandpass by fitting a power-law slope and
-amplitude to the fluxes in the two adjacent bandpasses (extrapolating
-when necessary). Figure \ref{compareci} shows the differences in
-the $K$-corrections inferred from this method and those inferred from
-the method of Figure \ref{k_kcorrect_plot}, as a function of
-redshift. $r$, $i$, and $z$ are all reasonably similar in either
-method; $u$ and $g$, however, have distinct trends with redshift,
-mostly due to the non-power law nature of the template galaxy spectra
-(and probably actual galaxy spectra) in this regime. To show this
-fact, we perform a similar power-law fit, only this time including a
-break in the spectrum at 4000 \AA. We use the $u$-$g$ color to fit the
-break, assuming that the slope blueward of 4000 \AA\ is
-$f(\lambda)\propto \lambda^{2}$.  Figure \ref{comparecibreak} shows
-the results of this fit; the redshift trend in $g$ is greatly reduced,
-as is the trend in $u$, but a large amount of scatter remains in the
-$u$ band. This results from the fact that you can {\it either} fit the
-slope of the SED below the 4000 \AA\ break {\it or} the size of the
-4000 \AA\ break itself; it is not possible with this data to constrain
-both in an individual spectrum, which is a limitation of our
-determination of fixed frame $u$-band magnitudes.
-
-Finally, it is possible to use the galaxy spectra obtained with the
-spectrograph to estimate the $K$-correction for each object.  However,
-this procedure requires trusting the spectrophotometry over a wide
-wavelength range. In addition, the region of the galaxy which the
-spectra cover is not completely well defined. The small scale
-wavelength features are set by the fiber aperture ($3''$ in diameter)
-while large scale wavelength features are constrained by a short
-``smear exposure'' which covers a larger area. The smear exposure
-corrections are clearly the correct approach for point sources, but
-for galaxies it causes the small scale and large scale features of the
-spectrum to be determined by somewhat different regions of the galaxy.
-
-Nevertheless, in Figure \ref{k_speck_plot.0.1}, we compare the
-$K$-corrections to $z=0.1$ of Main Sample galaxies calculated based on
-the spectra to those of Figure \ref{k_kcorrect_plot}, finding that
-they are quite similar. There is a trend with redshift in the
-$g$-band, which is due to the fact that on average the $g-r$ color is
-about 0.1 magnitudes redder in the fiber aperture than in the
-Petrosian aperture. The implication is that the $K$-corrections are
-stronger for the aperture which corresponds to the spectrum than for
-the Petrosian aperture. This indicates that for work which requires
-precision, it is best to avoid using the $K$-corrections determined
-from the spectra. (Note that in the second paragraph of this section
-we used the SDSS spectra as {\it example} spectra to validate the
-method as applied to synthesized photometry; in this paragraph we have
-evaluated whether to rely on fiber spectra to calculate
-$K$-corrections for the Petrosian photometry).
-
-For completeness, Figure \ref{k_speck_plot.0.3} compares the
-spectroscopic and photometric $K$-corrections to $z=0.3$ of LRGs. The
-\band{0.3}{r} and \band{0.3}{i} bands agree very well. However, there
-is considerable scatter in the \band{0.3}{g} band (nearly equivalent
-to the \band{0.0}{u} band), at around 20\%. Because the LRGs tend to
-have small color gradients, the issues of fiber sizes and smear
-exposures are not important in this context.
-
-\section{Conclusions and Future Work}
+\section{Summary}
 \label{conclusions}
 
-We have presented a method and an implementation for estimating galaxy
-SEDs for the purpose of calculating fixed frame galaxy magnitudes over
-a range of redshifts. We have demonstrated that it gives sensible and
-consistent results. We will be using this method in future papers
-which will describe the joint distribution of luminosities and colors
-of galaxies, as well as the evolution of the luminosity function of
-galaxies. Furthermore, we plan to incorporate observations of objects
-in bands other than the SDSS bands to further describe the nature of
-galaxy SEDs.
+[ summarize ]
 
-We have not discussed in detail the question of error analysis. There
-is no answer to this question which is simple, general, {\it and}
-realistic. The photometric errors can of course be propagated to the
-errors on the reconstructed magnitudes, but these estimates do not
-describe the errors associated with the restriction to a
-three-dimensional space of SED shapes. An idea of the level of errors
-can be gleaned from the comparison of our method to other methods of
-$K$-correction. Clearly, $r$, $i$ and $z$ can be reconstructed to the
-level of the photometry. Judging from Figure
-\ref{k_speck_plot.fitfib.0.1}, the $g$ band can be reconstructed to
-about 5\% accuracy. The $u$ band is more difficult to evaluate.
-Judging from the poor reconstructions of the $\band{0.3}{g}$ band
-(nearly equal to the $\band{0.0}{u}$ band, the reconstructions may be
-as bad as 20\% accuracy.
-
-Topics we have not discussed here are dust- and evolution-correction
-of the magnitudes. Some of the scatter in the three-dimensional space
-describing the shape of the galaxy SED is probably due to dust. It may
-be possible to evaluate the effects of reddening in this space and, by
-assuming that galaxies corrected for internal reddening live in an
-even lower dimensional space, perform a reddening correction. We will
-be investigating this question in the near future. One can perform a
-similar exercise with evolution estimates, relating the position of a
-galaxy along the $a_1$ axis in Figure \ref{k_coeffdist_plot} to a
-particular star-formation history, and estimating the evolution of the
-object from that history.
-
-The distribution of galaxies in the three-dimensional space of galaxy
-SEDs described here may be useful for other purposes, as well. For
-example, you can calculate photometric redshifts using the method
-described here. {\tt kcorrect \kversion} has a simple (and very fast)
-photometric redshift estimator with very tight core of residuals
-($\sigma \sim 0.05$) in a comparison to SDSS galaxies with redshifts.
-However, we warn the reader that before this code can be used as a
-reliable photometric redshift indicator, more work has to be done
-(which the reader is invited to do using the released code) to
-identify galaxies which are likely to be redshift outliers and to
-handle galaxies at $z>0.5$ (which have photometric observations which
-are bluer in the rest frame than for any galaxies in our training
-set). Until this method is perfect and our templates are better
-constrained, we recommend the photometric redshift method and results
-of Csabai {\it et al.}, in preparation. Another use of the
-three-dimensional space of galaxy SEDs is to create mock samples of
-galaxies observed in any bandpass at any redshift. We are working on
-quantifying the correlations between these coefficients and
-luminosity, surface-brighntess and radial profile in order to create
-mock catalogs which truly reflect the correlations between galaxy
-properties found in the data.
+[ this does not describe what constraints photometry puts on the SFH ]
 
@@
 
-Special thanks to Ivan Baldry for pointing out an error in this
-manuscript and in the distributed code for versions {\tt v1\_10} and
-previous.  MB and DWH acknowledge NASA NAG5-11669 for partial support.
-MB was supported at the beginning of this work by the DOE and NASA
-grant NAG 5-7092 at Fermilab. He is also grateful for the hospitality
-of the Department of Physics and Astronomy at the State University of
-New York at Stony Brook, who kindly provided computing facilities on
-his frequent visits there.
+Sam Roweis
 
@@
 
-\begin{thebibliography}{DUM}
-\input{bib_paper.tex}
-\end{thebibliography}
+\bibliographystyle{../../../nyu-astro/tex/apj}
+\bibliography{../../../nyu-astro/tex/apj-jour,../../../nyu-astro/tex/ccpp}
+
+\appendix
+
+\section{Approximate conversions between common magnitude systems}
 
@@
 
-\include{figures}
+\include{kcorrect_figures}
 

### 2004-01-17 docs/paper/kcorrect_figures.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@292 75dff457-8751-0410-ab1a-cd701654ce6b

\clearpage

\setcounter{thefigs}{0}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{k_sfh.ps}
\caption{\label{sfh} The top panel are the star-formation histories of
the 15 bursts we use in our model. The bottom panel shows (for no dust
and for $Z=0.008$) the PEGASE2 spectrum predicted for these bursts. }
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{gmrz.ps}
\caption{\label{gmrz} }
\end{figure}

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{kfit.ps}
\caption{\label{kfit} The color residuals between the observed values
  and those of the best fit to the nonnegative combination of 225
  templates ($\delta(b-r)\equiv [b-r]_{\mathrm{obs}} -
  [b-r]_{\mathrm{fit}}$). The greyscale is proportional to the
  distribution of color residual at a given redshift. The lines are
  0.16, 0.45, and 0.84 fractional quantiles. Note that we have zoomed
  into $g-r$, $r-i$, and $i-z$ in order to see the residuals
  better. At high redshift many galaxies have abnormally low $u$-band
  fluxes relative to their $r$-band flux. At low redshift there are a
  few galaxies whose near infrared $z$-band magnitudes are
  lower. There is a general offset of nearly 0.1 magnitudes in
  $J-r$. }
\end{figure}

### 2004-02-17 src/k_fit_photoz.c
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@295 75dff457-8751-0410-ab1a-cd701654ce6b

@@ -12,7 +12,7 @@
  * Mike Blanton
  * 5/2003 */
 
-#define ZRES 0.02
+#define ZRES 0.10
 #define TOL 0.02
 #define MAXITER 3000
 #define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
@@ -65,16 +65,16 @@ IDL_LONG k_fit_photoz(float *photoz,
 											float *chi2,
 											IDL_LONG verbose)
 {
-	float *zgrid,*chi2grid,chi2min,az,bz,cz;
+	float *lzgrid, *lchi2grid, *zgrid,*chi2grid,chi2min,az,bz,cz;
   float z_0,z_1,z_2,chi2_0,chi2_1,chi2_2;
-	IDL_LONG i,j,k,nzsteps,jmin;
+	IDL_LONG i,j,k,nzsteps,lnzsteps,jmin;
 
   pz_applyprior=1;
 	pz_nk=nk;
 	pz_nv=nv;
 	pz_nz=nz;
 	pz_nprior=nprior;
-	nzsteps=(IDL_LONG) floor((zvals[nz-1]-zvals[0])/ZRES);
+	nzsteps=(IDL_LONG) floor((zvals[nz-1]-zvals[0])/ZRES)+1;
 	chi2grid=(float *) malloc(nzsteps*sizeof(float));
 	pz_maggies=(float *) malloc(nk*sizeof(float));
 	pz_maggies_ivar=(float *) malloc(nk*sizeof(float));
@@ -94,6 +94,8 @@ IDL_LONG k_fit_photoz(float *photoz,
 	zgrid=(float *) malloc(nzsteps*sizeof(float));
 	for(j=0;j<nzsteps;j++) 
 		zgrid[j]=zvals[0]+(zvals[nz-1]-zvals[0])*(float)j/(float)(nzsteps-1);
+  lchi2grid=(float *) malloc(10000*sizeof(float));
+  lzgrid=(float *) malloc(10000*sizeof(float));
 	for(i=0;i<ngalaxy;i++) {
 
     pz_ncheck=0;
@@ -133,8 +135,38 @@ IDL_LONG k_fit_photoz(float *photoz,
 		} /* end if..else */
 		chi2min=k_brent(az,bz,cz,pz_fit_coeffs,TOL,&(photoz[i]));
 #else 
-    photoz[i]=zgrid[jmin];
-#endif
[…]

### 2004-10-30 src/k_choldc.c
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@316 75dff457-8751-0410-ab1a-cd701654ce6b

@@ -12,7 +12,7 @@ void k_choldc(float *a, int n, float p[])
 			for (sum=a[i*n+j],k=i-1;k>=0;k--) sum -= a[i*n+k]*a[j*n+k];
 			if (i == j) {
 				if (sum <= 0.0) {
-					printf("ERROR: choldc failed (sum==%e)\n",sum);
+					fprintf(stderr,"ERROR: choldc failed (sum==%e)\n",sum);
 					return;
 				}
 				p[i]=sqrt(sum);

### 2005-03-15 docs/paper/kcorrect.tex
some writing
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@344 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-\title{$K$-corrections and photometric redshifts in SDSS and 2MASS$^1$}
+\title{$K$-corrections and photometric redshifts in GALEX, SDSS,
+2MASS, DEEP, and Spitzer data$^1$}
 
@@
 terms of approximate stellar masses and star-formation histories. We
-present templates fit with this method to the Sloan Digital Sky Survey
-and Two-Micron All Sky Survey photometry.  The software that generates
-these results is publicly available and easily adapted to handle a
-wide range of galaxy observations.
-\end{abstract}
+present templates fit with this method to data from GALEX, Sloan
+Digital Sky Survey spectroscopy and photometry, the Two-Micron All Sky
+Survey, the Deep Extragalactic Evolutionary Probe, and the Spitzer
+Space Telescope. In addition, we provide software for using such data
+to estimate $K$-corrections, photometric redshifts, evolution
+corrections, stellar masses, star formation histories, and
+spectroscopic continua suitable for use in emission line measurements.
+This software is suitable for analyzing many other data sets.
+\end{abstract}code
 

### 2005-05-05 docs/paper/fitting.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@384 75dff457-8751-0410-ab1a-cd701654ce6b

\documentclass[10pt]{aastex}

\newcommand{\vv}[1]{{\bf #1}}
\newcommand{\df}{\delta}
\newcommand{\dfft}{{\tilde{\delta}}}
\newcommand{\betaft}{{\tilde{\beta}}}
\newcommand{\erf}{{\mathrm{erf}}}
\newcommand{\erfc}{{\mathrm{erfc}}}
\newcommand{\Step}{{\mathrm{Step}}}
\newcommand{\ee}[1]{\times 10^{#1}}
\newcommand{\avg}[1]{{\langle{#1}\rangle}}
\newcommand{\Avg}[1]{{\left\langle{#1}\right\rangle}}
\def\simless{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'074$}}\mathchar"7218\,$}}} % < or of order
\def\simgreat{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'076$}}\mathchar"7218\,$}}} % > or of order
\newcommand{\iras}{{\sl IRAS\/}}
\newcommand{\petroratio}{{{\mathcal{R}}_P}}
\newcommand{\petroradius}{{{r}_P}}
\newcommand{\petronumber}{{{N}_P}}
\newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
\newcommand{\band}[2]{\ensuremath{^{{#1}}\!{#2}}}

\newcommand{\kversion}{{\tt v1\_11}}

\setlength{\footnotesep}{9.6pt}

\newcounter{thefigs}
\newcommand{\fignum}{\arabic{thefigs}}

\newcounter{thetabs}
\newcommand{\tabnum}{\arabic{thetabs}}

\newcounter{address}

\slugcomment{To be submitted to \aj}

\shortauthors{Blanton {\it et al.} (2005)}
\shorttitle{Template star-formation histories}

\begin{document}

\title{ Using heterogeneous data to constrain star-formation histories
of galaxies}

\author{
Michael R. Blanton\altaffilmark{\ref{NYU}} and 
Sam Roweis\altaffilmark{\ref{NYU}}}

\setcounter{address}{1}
\altaffiltext{\theaddress}{
\stepcounter{address}
New York University, Department of Physics, 4 Washington Place, New
York, NY 10003
\label{NYU}}
\clearpage

\begin{abstract}
Astronomers have a huge set of well-calibrated photometric and
spectroscopic data on galaxies from recent large surveys at all
redshifts, such as (at low redshift) the GALaxy Evolution eXplorer in
the ultraviolet, the Sloan Digital Sky Survey in the optical, and the
Two-Micron All Sky Survey in the near infrared, and (at higher
redshifts) the Deep Extragalactic Evolutionary Probe 2 and the Great
Observatories Origins Deep Survey.  However, this data comes in a
rather heterogeneous form --- sometimes in the form of spectra,
sometimes in the form of broad-band photometry. In addition,
broad-band photometry suffers from two forms of heterogeneity: system
responses differ among different surveys, and galaxies are observed at
many different redshifts (and the system responses therefore differ
among all galaxies in the rest frame).  How can we explore this very
heterogeneous set of data?  Here we present an implementation of an
algorithm for fitting star-formation histories to these data. Our
approach is to restrict the problem by trying to explain the
observations of each galaxy as a nonnegative linear combination of a
small set of star-formation history templates.  The algorithm has
almost arbitrary freedom in choosing the set of templates, but it must
explain every galaxy in terms of a nonnegative sum of the same set.
Our algorithm can handle whatever spectral or broad-band information
one has on each galaxy.  The results of the algorithm have scientific
applications --- such as exploring the star-formation histories of
galaxies, how they vary with environment, how they correlate with
other properties, and how they change with redshift. In addition, they
have technical applications --- such as $K$-corrections, photometric
redshifts, and continuum-fitting of spectra. This algorithm and its
ilk are likely to become more and more useful as astronomers continue
to combine large sets of inhomogeneous data.
\end{abstract}

\keywords{galaxies: fundamental parameters --- galaxies: photometry
--- galaxies: statistics}

\section{Motivation}
\label{motivation}

Astronomers have observed the broad-band colors and spectra of a huge
number of galaxies in the Universe, in many different passbands.  For
example, at low redshift the Galaxy Evolution Explorer (GALEX; {\bf
ref}) observes in the ultraviolet (UV), the Sloan Digital Sky Survey
(SDSS; \citealt{york00a}) observes spectroscopically and
photometrically in the optical, and the Two-Micron All Sky Survey
(2MASS; \citealt{strutskie97a}) observes in the near infrared
(NIR). At higher redshifts, the Deep Extragalactic Evolutionary Probe
2 (DEEP2; {\bf ref} and the Great Observatories Origins Deep Survey
(GOODS; {\bf ref}) observe in the rest-frame UV. These and other data
sets provide a huge set of information about galaxy colors and spectra
which we can use to help understand their star-formation histories.

However, this data comes in rather heterogeneous forms. Some of the
information is spectrophotometric, and some of it is broad-band
photometry. Broad-band photometry differs among different instruments:
for example, the SDSS uses the $ugriz$ bands to observe in the
optical, whereas DEEP2 uses the $BRI$ bands. In addition, these bands
probe different parts of the rest-frame galaxy spectrum at different
redshifts. So the constraints these observations apply to our fits of
the star-formation histories of these objects vary considerably from
one object to another.

On the other hand, from the spectroscopic observations we know the
galaxy spectra reside in a low-dimensional locus. Principal Component
Analysis (PCA) of galaxy spectra, introduced by \citet{connolly94a} and
applied many times since then ({\bf more refs}; \citealt{yip05a}),
demonstrates that most of the variance in the distribution of galaxies
in spectral space can be explained using a few templates. This means,
that in the linear space of all possible spectra, galaxies exist in
only a small local. Therefore, even with very heterogeneous data, we
should be able to determine the properties of this locus.

Here we present an approach to combining the heterogeneous data
described above in order to determine the properties of the locus of
galaxy spectra. Rather than taking the model-free approach used by
PCA, we here restrict the space of possible spectra to those predicted
from the high resolution stellar population synthesis model of
\citet{bruzual03a}. This approach both constrains the problem
appropriately and yields a natural theoretical interpretation of the
results in terms of star-formation histories.

In a nutshell, our algorithm does the following. Given the
observations (and uncertainties of those observations) available for
each galaxy, it finds the nonnegative linear combination of $N$
template star-formation histories which best predict those
observations in the $\chi^2$ sense. Given the entire set of galaxy
observations available, it also fits for those template star-formation
histories. The technical name for this algorithm is Nonnegative Matrix
Factorization (NMF).

This approach is similar to PCA in that it finds the small spectral
subspace in which galaxies exist, and can in some ways be thought of
as ``nonnegative PCA.''  However, our method has several advantages
over standard PCA approach. First, it has a natural physical
interpretation associated with that spectral subspace, which is the
corresponding subspace of all possible star-formation
histories. Second, it naturally handles data uncertainties and missing
data which allows it to ignore variation which is due purely to
statistical errors.  Third, it handles the complications of observing
galaxy spectra photometrically using broad-band filters.

The results of this work yield, first of all, a set of templates which
are useful for any number of technical tasks, for example
$K$-corrections, evolution corrections, photometric redshifts, and
continuum subtraction of galaxy spectra. Second, the fits to
individual galaxies provide a theoretical interpretation of their
emission, in terms of their stellar masses, dust content, and
star-formation histories.

In Section \ref{algorithm} we present and test the NMF algorithm
itself. All of the software we use is available publicly in the {\tt
idlutils}, {\tt idlspec2d} and {\tt kcorrect} packages.  In Section
\ref{data} we present the data we use, most of which is publicly
available. In Section \ref{results}, we present the templates we have
found for our data set. We summarize in Section
\ref{summary}. Separate papers describe the application of these
results to the scientific and technical questions described above.

\section{Method}
\label{algorithm}

Our method is fundamentally very simple. We take $N$ star-formation
history templates and all of our observations of our galaxies.  We
then find the templates and (for each galaxy) their linear combination
by minimizing the $\chi^2$ difference between the actual observations
and the model predictions for those observations. In practice there
are some complications and details to consider. In Section \ref{model}
we describe how we express the star-formation histories of each
galaxy.  In Section \ref{spectra} we explain the models we use to
convert these star-formation histories into spectra. In Section
\ref{observe} we describe how a particular star-formation history can
be transformed into all possible observations for that galaxy. In
Section \ref{nmf} we describe our minimization procedure.

\subsection{Model for the star-formation histories}
\label{model}

Our model for the star-formation history of each galaxy $i$ is a linear
combination of $N$ templates, as follows:
\begin{equation}
G_i(t, m, d) = \sum_{i=0}^{N-1} a_{ik} T_k(t, m, \tau_V).
\end{equation}
We restrict to nonnegative star-formation rates ($a_{ik} \ge
0$). $T_k(t, m, \tau_V)$ is the star-formation history, described as
the stellar mass (in solar masses) formed $t$ years ago, with
metallicity $m$, and currently obscured by $\tau_V$ magnitudes in the
$V$ band.

Each template is in turn expressed as a linear combination of a set of
instantaneous bursts:
\begin{equation}
T_k(t, m, d) = 
\sum_{l=0}^{N_t-1} 
\sum_{n=0}^{N_m-1} 
\sum_{p=0}^{N_\tau-1} 
t_{lmn} 
\delta^D(t-t_l)
\delta^D(m-m_n)
\delta^D(\tau_V-\tau_{V,p})
\end{equation}

\subsection{Predicted spectra from the star-formation histories}
\label{spectra}

Note that we do not {\it have} to limit our spectral library to those
predicted by star-formation histories. In fact, we can add arbitrary
spectra in this step to account for emission lines, polycyclic
aromatic hydrocarbon (PAH) emission, active galactic nuclei, or any
other constituent of galaxies.

In fact, as an example of this, we here include nine emission lines:
OII 3728 \AA; H$\beta$ 4863 \AA; OIII 4960, 5008 \AA; NII 6550, 6585
\AA; H$\alpha$ 6565 \AA; SII 6718, 6733 (all vacuum wavelengths). We
include each of these lines as a template with a Gaussian centered on
the line center with a dispersion corresponding to 300 km s^{-1}.

\subsection{Observations of the predicted spectra}
\label{observe}

\subsection{Finding the best-fit templates}
\label{nmf}

\section{Data}
\label{data}

\section{Results}
\label{results}

\section{Summary}
\label{summary}

\acknowledgments

Funding for the creation and distribution of the SDSS Archive has been
provided by the Alfred P. Sloan Foundation, the Participating
Institutions, the National Aeronautics and Space Administration, the
National Science Foundation, the U.S. Department of Energy, the
Japanese Monbukagakusho, and the Max Planck Society. The SDSS Web site
is {\tt http://www.sdss.org/}.

The SDSS is managed by the Astrophysical Research Consortium (ARC) for
the Participating Institutions. The Participating Institutions are The
University of Chicago, Fermilab, the Institute for Advanced Study, the
Japan Participation Group, The Johns Hopkins University, Los Alamos
National Laboratory, the Max-Planck-Institute for Astronomy (MPIA),
the Max-Planck-Institute for Astrophysics (MPA), New Mexico State
University, Princeton University, the United States Naval Observatory,
and the University of Washington.

\begin{thebibliography}{DUM}
\input{bib_paper.tex}
\end{thebibliography}

\newpage

%\include{tables}

\include{figures}

\end{document}

### 2005-05-26 docs/paper/fitting.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@387 75dff457-8751-0410-ab1a-cd701654ce6b
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@402 75dff457-8751-0410-ab1a-cd701654ce6b

@@
-\documentclass[10pt]{aastex}
+\documentclass[10pt,preprint]{aastex}
 
@@
 
-\title{ Using heterogeneous data to constrain star-formation histories
-of galaxies}
+\title{ Creating spectroscopic templates using heterogeneous data}
 
@@
 Observatories Origins Deep Survey.  However, this data comes in a
-rather heterogeneous form --- sometimes in the form of spectra,
-sometimes in the form of broad-band photometry. In addition,
-broad-band photometry suffers from two forms of heterogeneity: system
-responses differ among different surveys, and galaxies are observed at
-many different redshifts (and the system responses therefore differ
-among all galaxies in the rest frame).  How can we explore this very
-heterogeneous set of data?  Here we present an implementation of an
-algorithm for fitting star-formation histories to these data. Our
-approach is to restrict the problem by trying to explain the
-observations of each galaxy as a nonnegative linear combination of a
-small set of star-formation history templates.  The algorithm has
-almost arbitrary freedom in choosing the set of templates, but it must
-explain every galaxy in terms of a nonnegative sum of the same set.
-Our algorithm can handle whatever spectral or broad-band information
-one has on each galaxy.  The results of the algorithm have scientific
-applications --- such as exploring the star-formation histories of
-galaxies, how they vary with environment, how they correlate with
-other properties, and how they change with redshift. In addition, they
-have technical applications --- such as $K$-corrections, photometric
-redshifts, and continuum-fitting of spectra. This algorithm and its
-ilk are likely to become more and more useful as astronomers continue
-to combine large sets of inhomogeneous data.
+rather heterogeneous form --- sometimes in the form of
+spectrophotometry, sometimes in the form of broad-band photometry. In
+addition, broad-band photometry suffers from two forms of
+heterogeneity: system responses differ among different surveys, and
+galaxies are observed at many different redshifts (and the system
+responses therefore differ among all galaxies in the rest frame).  How
+can we explore this very heterogeneous set of data?  Here we present
+an implementation of an algorithm for fitting star-formation histories
+to these data. Our approach is to restrict the problem by trying to
+explain the observations of each galaxy as a nonnegative linear
+combination of a small set of star-formation history templates.  The
+algorithm has almost arbitrary freedom in choosing the set of
+templates, but it must explain every galaxy in terms of a nonnegative
+sum of the same small set.  Our algorithm can handle whatever spectral
+or broad-band information one has on each galaxy.  The results of the
+algorithm have scientific applications --- such as exploring the
+star-formation histories of galaxies, how they vary with environment,
+how they correlate with other properties, and how they change with
+redshift. In addition, they have technical applications --- such as
+$K$-corrections, photometric redshifts, and continuum-fitting of
+spectra. This algorithm and its ilk are likely to become more and more
+useful as astronomers continue to combine large sets of inhomogeneous
+data.
 \end{abstract}
@@
 galaxy spectra reside in a low-dimensional locus. Principal Component
-Analysis (PCA) of galaxy spectra, introduced by \citet{connolly94a} and
-applied many times since then ({\bf more refs}; \citealt{yip05a}),
+Analysis (PCA) of galaxy spectra, introduced by \citet{connolly94a}
+and applied many times since then ({\bf more refs}; \citealt{yip05a}),
 demonstrates that most of the variance in the distribution of galaxies
-in spectral space can be explained using a few templates. This means,
-that in the linear space of all possible spectra, galaxies exist in
-only a small local. Therefore, even with very heterogeneous data, we
+in spectral space can be explained using a few templates. This means
+that, in the linear space of all possible spectra, galaxies exist in
+only a small locus. Therefore, even with very heterogeneous data, we
 should be able to determine the properties of this locus.
@@
 from the high resolution stellar population synthesis model of
-\citet{bruzual03a}. This approach both constrains the problem
-appropriately and yields a natural theoretical interpretation of the
-results in terms of star-formation histories.
+\citet{bruzual03a} and the emission line models of {\bf mappings
+ref}.  This approach both constrains the problem appropriately and
+yields a natural theoretical interpretation of the results in terms of
+star-formation histories.
 
@@
 as ``nonnegative PCA.''  However, our method has several advantages
-over standard PCA approach. First, it has a natural physical
+over the standard PCA approach. First, it has a natural physical
 interpretation associated with that spectral subspace, which is the
@@
 histories. Second, it naturally handles data uncertainties and missing
-data which allows it to ignore variation which is due purely to
+data, which allows it to ignore that variation which is due purely to
 statistical errors.  Third, it handles the complications of observing
@@
 \begin{equation}
+\label{stellar}
 G_i(t, m, d) = \sum_{i=0}^{N-1} a_{ik} T_k(t, m, \tau_V).
 \end{equation}
-We restrict to nonnegative star-formation rates ($a_{ik} \ge
-0$). $T_k(t, m, \tau_V)$ is the star-formation history, described as
-the stellar mass (in solar masses) formed $t$ years ago, with
-metallicity $m$, and currently obscured by $\tau_V$ magnitudes in the
-$V$ band.
+$G_i(t, m, \tau_V)$ is the star-formation history of each galaxy $i$,
+described as the stellar mass (in solar masses) formed $t$ years ago,
+with metallicity $m$, and currently obscured by $\tau_V$ magnitudes in
+the $V$ band. $T_k(t, m, \tau_V)$ is the star-formation history of one
+of $N$ templates out of which we allow ourselves to build each
+galaxy. We restrict to nonnegative combinations of these templates
+($a_{ik} \ge 0$).
 
@@
 \sum_{p=0}^{N_\tau-1} 
-t_{lmn} 
+t_{k,lmn} 
 \delta^D(t-t_l)
@@
 \end{equation}
+That is, for each template $k$ a certain mass in stars $t_{k,lmn}$
+(expressed in solar masses) were formed time $t_l$ ago with
+metallicity $m_n$, and are at the time of observation obscured by
+$\tau_V$ magnitudes of dust in the $V$ band.
 
-\subsection{Predicted spectra from the star-formation histories}
+As described in Section \ref{spectra}, these star-formation histories
+can be converted into spectra using population synthesis models.
+
+\subsection{Model for the ionized gas}
+
+In the optical, the other important contribution to the flux other
+than the stellar light is emission from ionized gas. Depending on how
+it is ionized and what its metallicity is, the recombination of this
+gas can yield rather different spectra.
+
+Thus, we treat the gas as a sum of the same templates:
+\begin{equation}
+\label{emission}
+E_i(m, q) = \sum_{i=0}^{N-1} a_{ik} S_k(m, q). 
+\end{equation}
+$E_i(m, q)$ is the amount of gas in each galaxy with metallicity $m$
+and ionization parameter $q$.  $S_k(m, q)$ is the amount of such gas
+in each of our $N$ templates.  We constrain the $a_{ik}$ in this
+equation to be the same as that in Equation \ref{stellar} for the
+star-formation histories. That is, each star formation history
+templates $T_k{t, m, \tau_V}$ will be associated with the
+corresponding distribution of ionized gas $S_k(m,q)$.
+
+Again, each template is actually a sum of delta functions:
+\begin{equation}
+S_k(m, q) = 
+\sum_{r=0}^{N_m-1} 
+\sum_{s=0}^{N_q-1} 
+t_{k,rs}'
+\delta^D(m-m_r)
+\delta^D(q-q_s)
+\end{equation}
+where for each metallicity $m_r$ and ionization parameter $q_s$ we 
+have a prediction for the emergent line fluxes, which we describe in
+Section \ref{spectra}.
+
+\subsection{Predicted spectra from the star-formation histories and
+ionized gas}
 \label{spectra}
@@
 Note that we do not {\it have} to limit our spectral library to those
-predicted by star-formation histories. In fact, we can add arbitrary
-spectra in this step to account for emission lines, polycyclic
-aromatic hydrocarbon (PAH) emission, active galactic nuclei, or any
-other constituent of galaxies.
-
-In fact, as an example of this, we here include nine emission lines:
-OII 3728 \AA; H$\beta$ 4863 \AA; OIII 4960, 5008 \AA; NII 6550, 6585
-\AA; H$\alpha$ 6565 \AA; SII 6718, 6733 (all vacuum wavelengths). We
-include each of these lines as a template with a Gaussian centered on
-the line center with a dispersion corresponding to 300 km s^{-1}.
+predicted by star-formation histories or the emission line models. In
+fact, we can add arbitrary spectra in this step to account for
+emission lines, polycyclic aromatic hydrocarbon (PAH) emission in the
+mid-infrared, active galactic nuclei in the ultraviolet, or any other
+constituent of galaxies.
 
@@
 
-\subsection{Finding the best-fit templates}
+\subsection{Summary of the problem we want to solve}
+\label{problem}
+
+\section{Non-negative matrix factorization to find the best templates}
 \label{nmf}
@@
 
-%\include{tables}
+%\include{fitting_tables}
 
-\include{figures}
+\include{fitting_figures}
 

### 2005-07-07 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@406 75dff457-8751-0410-ab1a-cd701654ce6b
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@411 75dff457-8751-0410-ab1a-cd701654ce6b
changes
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@450 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2004-01-17 docs/paper/kcorrect.tex
@@
 \shortauthors{Blanton {\it et al.} (2000)}
-\shorttitle{$K$-corrections and photometric redshifts}
-
-% TODO:
-% finish final templates
-% fix kphotoz banding
-% fit SFHs to final spectra to get stellar masses, mean ages, 
-% mean metallicities
+\shorttitle{$K$-corrections and filter transformations}
 
@@
 
-\title{$K$-corrections and photometric redshifts in SDSS and 2MASS$^1$}
+\title{$K$-corrections and filter transformations \\
+from the restframe
+ultraviolet to the near infrared}
 
 \author{
-Michael R. Blanton\altaffilmark{\ref{NYU}}
+Michael R. Blanton\altaffilmark{\ref{NYU}} and 
+Samuel Roweis
 %Tamas Budavari\altaffilmark{\ref{JHU}},
@@
 \begin{abstract}
-We present a method to develop templates for calculating
-$K$-corrections and photometric redshifts suitable for ultraviolet,
-optical, and infrared observations. Since we base the templates on
-stellar population synthesis results, the results are intepretable in
-terms of approximate stellar masses and star-formation histories. We
-present templates fit with this method to the Sloan Digital Sky Survey
-and Two-Micron All Sky Survey photometry.  The software that generates
-these results is publicly available and easily adapted to handle a
-wide range of galaxy observations.
-\end{abstract}
+We present a method of developing model-based spectroscopic templates
+to explain observed galaxy fluxes. Such templates are useful for
+calculating $K$-corrections, calculating photometric redshifts, and
+converting among observations of galaxies at various wavelengths. In
+principle, we could apply this method to galaxy data at any redshift
+and in any set of wavelengths. The implementation we present here is
+suitable for ultraviolet, optical, and near-infrared
+observations. Since we base the templates on stellar population
+synthesis results, the results are intepretable in terms of
+approximate stellar masses and star-formation histories. We present
+templates fit with this method to data from GALEX, Sloan Digital Sky
+Survey spectroscopy and photometry, the Two-Micron All Sky Survey, the
+Deep Extragalactic Evolutionary Probe and Great Observatories Origins
+Deep Survey.  In addition, we present software for using such data to
+estimate $K$-corrections, photometric redshifts, evolution
+corrections, stellar masses, and recent star-formation rates. This
+software is suitable for analyzing many other data sets as well.
+\end{abstract}code
 
@@
 
-\section{Fitting SEDs to broad-band magnitudes}
-\label{sedfit}
+New surveys at a low and high redshift have provided us with estimates
+of galaxy spectral energy distributions (SEDs) for an enormous number
+of galaxies. When comparing populations of galaxies at different
+redshifts in these surveys, we need to use comparable measurements of
+the galaxy SEDs. However, different surveys use different bandpasses
+and the restframe wavelengths of these bandpasses necessarily vary
+with redshift. We need to be able to handle this heterogeneity in
+order to make sensible comparisons among all of these new surveys.
 
-First, we describe how we reconstruct galaxy SEDs from broad band
-magnitude measurements.
+In this paper, we present a method for doing so, by calculating
+$K$-corrections between observed and desired bandpasses.  The
+$K$-correction between a bandpass $R$ used to observe a galaxy at
+redshift $z$ and the desired bandpass $Q$ is defined by the equation
+(\cite{oke68a, hogg02a}):
+\begin{equation}
+\label{kcorrecteqn}
+m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z) - 5 \log_10 h 
+\end{equation}
+where $\mathrm{DM}(z) = 25 - 5\log_{10} (d_L / (h^{-1}{\mathrm{~Mpc}}))$ is the
+bolometric distance modulus calculated from the luminosity distance
+$d_L$, and $M_Q$ is the absolute magnitude. The absolute magnitude is
+defined as the apparent magnitude an object would have if were
+observed 10 pc away, in bandpass $Q$, at rest.  The traditional
+definition of the $K$-correction takes $Q=R$. however, we note that in
+practice many surveys do perform $K$-corrections from one observed
+bandpass $R$ to another bandpass $Q$ in the rest frame. This practice
+is particularly common when dealing with high redshift observations.
+In addition to $K$-corrections, this method also provides an
+interpretation of the data in terms of a physical model which
+describes the stellar mass and star-formation history of each galaxy.
 
-A broad-band flux $F_k$ in filter $k$ is a measurement of the
-projection of the source spectrum $f(\lambda)$ on the response
-$R_k(\lambda)$ (where typically this function includes the effects of
-the atmosphere, telescope, filter $k$, and the CCD itself)
+The templates are designed to work well for a wide range of data sets.
+We constrain them with photometry and spectroscopy of the Sloan
+Digital Sky Survey (SDSS; \citealt{york00a}) the Galaxy Evolution
+Explorer (GALEX; {\bf ref})) in the ultraviolet, and the Two-Micron
+All Sky Survey (2MASS; \citealt{strutskie97a}) in the near infrared
+(NIR). In addition, at higher redshifts, we use constraints from the
+Deep Extragalactic Evolutionary Probe 2 (DEEP2; {\bf ref} and the
+Great Observatories Origins Deep Survey (GOODS; {\bf ref}).  These and
+other data sets provide a huge set of information about galaxy colors
+and spectra which we can use to help understand their star-formation
+histories.
+
+We note that the methods used here are generally useful. Using these
+methods to develop templates has the advantage that one can easily
+search a large space of models in order to find the few templates
+necessary in order to explain a given set of data.
+
+We also release the templates in electronic form as well as an
+implementation of the methods used to fit the templates to data. This
+software {\tt kcorrect v4\_1} is distributed on the World Wide
+Web{\footnote {\tt http://cosmo.nyu.edu/blanton/kcorrect/}}. It
+consists of a core C library which performs most of the complex and
+computationally intensive tasks, plus an IDL library that provides a
+high-level interface.  This software is an update of two major earlier
+releases ({\tt v1\_16}, \citet{blanton03b}; {\tt v3\_2}). The
+improvement over the latest version is twofold. First, we have
+improved the templates such that they successfully fit galaxies in the
+restframe UV, as observed by GALEX, DEEP2, and GOODS. Second, because
+the templates are now completely model-based, the fits have a physical
+interpretation in terms of a star-formation history. The IDL library
+also has a number of new useful functions.
+
+\section{Finding the templates}
+\label{algorithm}
+
+\subsection{Overview}
+
+From the spectroscopic observations we know that galaxy spectra reside
+in a low-dimensional locus. Principal Component Analysis (PCA) of
+galaxy spectra, introduced by \citet{connolly94a} and applied many
+times since then ({\bf more refs}; \citealt{yip05a}), demonstrates
+that most of the variance in the distribution of galaxies in spectral
+space can be explained using a few templates. This means that, in the
+linear space of all possible spectra, galaxies exist in only a small
+locus. Therefore, even with very heterogeneous data, we should be able
+to determine the properties of this locus.
+
+Here we present an approach to combining the heterogeneous data
+described above in order to determine the properties of the locus of
+galaxy spectra. Rather than taking the model-free approach used by
+PCA, we here restrict the space of possible spectra to those predicted
+from the high resolution stellar population synthesis model of
+\citet{bruzual03a} and the emission line models of {\bf mappings
+ref}.  This approach both constrains the problem appropriately and
+yields a natural theoretical interpretation of the results in terms of
+star-formation histories.
+
+In a nutshell, our algorithm does the following. Given the
+observations (and uncertainties of those observations) available for
+each galaxy, it finds the nonnegative linear combination of $N$
+template star-formation histories which best predict those
+observations in the $\chi^2$ sense. Given the entire set of galaxy
+observations available, it also fits for those template star-formation
+histories. The technical name for this algorithm is Nonnegative Matrix
+Factorization (NMF).
+
+This approach is similar to PCA in that it finds the small spectral
+subspace in which galaxies exist, and can in some ways be thought of
+as ``nonnegative PCA.''  However, our method has several advantages
+over the standard PCA approach. First, it has a natural physical
+interpretation associated with that spectral subspace, which is the
+corresponding subspace of all possible star-formation
+histories. Second, it naturally handles data uncertainties and missing
+data, which allows it to ignore that variation which is due purely to
+statistical errors.  Third, it handles the complications of observing
+galaxy spectra photometrically using broad-band filters.
+
+In this section, we describe how to take the astrophysical
+models and the astronomical observations, and to pose the problem in
+the form needed for the purposes of NMF. In the Appendix we describe
+how to solve to NMF problem.
+
+\subsection{The models}
+
+We begin with a basis set of 485 spectral templates. Of these, 450 are
+a set of instantaneous bursts from \citet{bruzual03a}. There are 6
+metallicities (0.005, 0.02, 0.2, 0.4, 1, and 2.5 times solar). For
+each metallicity there are 25 ages (between 1 Myr and 13.75 Gyr,
+spaced almost logarithmically in age).  For each age and metallicity
+there are 3 choices of dust model: (1) no dust extinction; (2) $\tau_V
+= 3$ dust with Milky Way type extinction (as input into the models of
+\citealt{witt00a} with a ``homogeneous'' distribution and ``shell''
+geometry); (3) $\tau_V = 3$ dust with SMC type extinction (from the
+same models). We smooth each template to $300$ km s$^{-1}$ velocity
+dispersion.
+
+The remaining 35 templates are from MAPPINGS-III {\bf ref} models of
+emission from ionized gas. We choose the predictions for an 8 Myr old
+continuous star-formation history with 5 possible metalliciticites
+(0.5, 0.2, 0.4, 1.0, and 2. times solar) and 7 possible ionization
+parameters ($q= 5\times 10^6$, $10^7$, $2\times 10^7$, $4\times 10^7$,
+$8 \times 10^7$, $1.5\times 10^8$, and $3\times 10^8$ cm s$^{-1}$). We
+take the spectra (given as a set of delta functions) and smooth them
+to $300$ km s$^{-1}$ velocity dispersion.
+
+Let us refer to these 485 basis templates as $M_{\lambda,
+j}(\lambda)$, expressed in units of ergs s$^{-1}$ \AA$^{-1}$.
+
+We seek five templates $F_{\lambda, i}(\lambda)$ which are built from
+nonnegative combinations of the original basis set of $N$ templates:
 \begin{equation}
-F_k = \int_0^\infty d\lambda f(\lambda) R_k(\lambda).
+F_{\lambda, i}(\lambda) = \sum_j b_{ij} M_{\lambda, j}(\lambda),
 \end{equation}
-Now, let us model $f(\lambda)$ as a sum of many template spectra:
+in units of ergs s$^{-1}$ \AA$^{-1}$.  For each galaxy $k$
+we want our model ${\hat{F}_\lambda}(\lambda)$ for their spectrum to
+be a nonegative sum of these five templates:
 \begin{equation}
-f(\lambda) = \sum_j a_j v_j(\lambda)
+{\hat{F}}_{\lambda,_k}(\lambda) = \sum_i a_{ki} F_{\lambda,i}(\lambda).
 \end{equation}
-In order to fit the $a_j$ given the observations in the bands $k$ for
-a single object, we minimize
+
+\subsection{The data}
+
+The training set consists of:
+\begin{enumerate}
+\item SDSS spectroscopic data in the observed range $3800 < \lambda <
+9000$ \AA, for 400 Luminous Red Galaxies between $0.15 < z < 0.5$ (LRGs;
+\citealt{eisenstein02a}) and 1600 Main sample galaxies between $0.001
+< z < 0.4$ (\citealt{strauss02a})
+\item SDSS photometric data on an independent set of LRGs ($griz$
+photometry only) and Main sample galaxies (using the full $ugriz$
+photometry). We use 2000 LRGs and 7000 Main sample galaxies.  For
+these galaxies we include 2MASS $JHK_s$ data ({\bf ref}) where
+available.
+\item GALEX DR1 far UV ($\sim 1500$ \AA) and near UV ($\sim 2300$ \AA)
+photometry for objects with SDSS redshifts and $ugriz$ photometry (4000
+galaxies; {\bf ref}).
+\item The $BRI$ photometry for high redshift galaxies in the DEEP2 
+DR1 release between $0.6 < z < 1.5$ (4000
+galaxies; {\bf ref}).
+\item The $BVizJHK_s$ photometry for GOODS galaxies between $0.5 < z < 2$ (1000
+galaxies; {\bf ref}).
+\end{enumerate}
+The SDSS, 2MASS, and GALEX galaxies were selected from and the matches
+were obtained by the New York University Value-Added Galaxy Catalog
+(NYU-VAGC;
+\citealt{blanton05b}).
+
+The SDSS data processing consists of astrometry \citep{pier03a};
+source identification, deblending and photometry \citep{lupton01a};
+photometricity determination \citep{hogg01a}; calibration
+\citep{fukugita96a,smith02a}; spectroscopic target selection
+\citep{eisenstein01a,strauss02a,richards02a}; spectroscopic fiber
+placement \citep{blanton03a}; and spectroscopic data reduction.  We
+recalibrated our photometry using the ``ubercalibration'' procedure
+described in \citet{blanton05b}.  Descriptions of these pipelines also
+exist in \citet{stoughton02a}.  An automated pipeline called {\tt
+idlspec2d} (in this case {\tt v4\_9}) measures the redshifts and
+classifies the reduced spectra (Schlegel et al., in preparation).
+
+Note that there are 20 different broadband photometric filters listed
+above ($B$ in DEEP2 is different than the $B$ used by GOODS). The {\tt
+kcorrect} product described in Section \ref{code} contains a
+tabulation of the response functions for all of these filters.
+
+\subsection{Comparing data and models}
+
+The data consist of spectra and bandband photometric measurements of
+galaxies at a number of redshifts, and we have to relate the models to
+these measurements.
+
+For the spectra, we take the observed spectrum
+$f_{\lambda,k}(\lambda)$ (in ergs cm$^{-2}$ s$^{-1}$ \AA$^{-1}$) of
+each galaxy $k$ at redshift $z$ (corresponding to a luminosity
+distance $d_L(z)$ in cm; see \citealt{hogg99a}) and calculate the
+restframe luminosity per unit wavelength:
+\begin{equation}
+{{F}}_{\lambda, k}(\lambda) = 
+{{f}}_{\lambda, k}[\lambda (1+z)](1+z) (4 \pi d_L^2),
+\end{equation}
+That is, the spectrum is shifted due to the redshift, while the
+integral of the numerator over wavelength (the total luminosity) is
+constant with redshift, and the total flux is related to the total
+luminosity by the inverse square law.  In addition, we smooth each
+spectrum by such an amount that, given its estimated velocity
+dispersion, its total velocity dispserion after smoothing is 300 km
+s$^{-1}$.
+
+Note that if we discretize the spectra to wavelengths $\lambda_l$, the
+relationship between the predicted SED for a galaxy $k$ and the basis
+set $M_{jl}$, becomes simply:
+\begin{equation}
+\hat{F}_{kl} = \sum_{ij} a_{ki} b_{ij} M_{jl}
+\end{equation}
+
+If one has a spectrum for a galaxy the
+expression for the contribution to $\chi^2$ from each wavelength
+$\lambda_l$ in the spectrum is then quite simple:
 \begin{eqnarray}
+\label{chi2spec}
+\chi_{kl}^2 &=& \left[\frac{F_{\lambda,k}(\lambda_l) -
+{\hat{F}}_{\lambda,k}(\lambda_l)} 
+{\sigma^2_k(\lambda_l)}\right]^2 \cr
+&=&
+\left[\frac{F_{\lambda,k}(\lambda_l) -
+\sum_{ij} a_{ki} b_{ij} M_{jl}
+}{\sigma^2_k(\lambda_l)}\right]^2. 
+\end{eqnarray}
+
+Comparing observed broadband flux measurements is a bit more
+complicated, because it is a projection of the spectrum onto the
+broadband filter $p$ at the observed redshift $z$. Instead of
+adjusting the observed fluxes as we could so easily do for the spectra
+above, for the photometry we express the models in terms of predicted
+broadband fluxes.
+
+Here we express the flux for each galaxy $k$ in units of AB maggies
+$\mu_p$, which are defined as what one would measure in bandpass $p$
+relative to the AB standard source. For example, if we transform our
+spectral energy density basis function $M_{\lambda, j}(\lambda)$ to a flux
+density $m_{\lambda, j}(\lambda)$, we can calculate the contribution of
+that basis function to the predicted maggies as:
+\begin{equation}
+\label{maggies}
+\eta_{jp} = 
+\frac{\int_0^{\infty} d\lambda \lambda R_p(\lambda) {m}_{\lambda, j}(\lambda)}
+{\int_0^\infty d\lambda \lambda R_p(\lambda) f_{\lambda,
+\mathrm{AB}}(\lambda) }
+\end{equation}
+Here, the response function $R_p(\lambda)$ is proportional to the
+contribution to the detector signal of a photon with wavelength
+$\lambda$ entering the Earth's atmosphere (or entering the telescope
+for a space telescope). The AB standard source is $f_{\lambda,
+\mathrm{AB}}(\lambda) d\lambda = f_{\nu,
+\mathrm{AB}} (\nu) d\nu$ and $f_{\nu, \mathrm{AB}}=3631$ Jy $= 3.631 \times
+10^{-20}$ erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$. Of course maggies $\mu$
+are related to magnitudes $m$ as:
+\begin{equation}
+m = -2.5 \log_{10} \mu, 
+\end{equation}
+such that the AB standard source would (if it existed) have $\mu=1$
+and $m=0$ for all bandpasses.
+
+In this context it is worth noting that many authors
+(e.g. \citealt{bessell90a}) tabulate the contribution to the detector
+signal per {\it unit of energy} in photons of wavelength $\lambda$
+instead of per {\it photon} with wavelength $\lambda$.  We will refer here
+to that quantity as $R_p'(\lambda)$, though in the literature it is
+often referred to without a prime (and without any explicit
+definition!). Clearly $R_p(\lambda) \propto R_p'(\lambda)/\lambda$,
+since the higher the frequency of the photon, at a fixed response per
+unit energy there is a higher response per photon. With this
+substitution, one can reexpress Equation
+\ref{maggies} appropriately in terms of $R_p'(\lambda)$.  
+Generally, though not universally, authors tabulate $R_p'(\lambda)$
+for bandpasses whose standards were originally calibrated using
+energy-counting devices rather than photon-counting devices. However,
+from the point of view of the analysis of the observations it is
+irrelevant what the devices used for the standards and the
+observations are, as long as one calculates $R_p(\lambda)$ and uses
+Equation \ref{maggies}.
+
+The prediction for the broadband fluxes from Equation
+\ref{maggies} is only for a specific redshift $z$. It turns out to
+simplify our mathematics to calculate the projection of each basis
+function $j$ onto each filter $p$ for a grid of redshifts (in this
+case spaced by 0.005 between redshifts 0 and 2). Thus, below we will
+take $p$ to index all the filters at all such redshifts.
+
+Just as before, we can now write down the relationship between the
+predicted broadband flux and the basis set $\mu_{jp}$, in the bandpass
+and redshift corresponding to the index $p$, for a galaxy $k$:
+\begin{equation}
+\hat{\mu}_{kp} = \sum_{ij} a_{ki} b_{ij} \eta_{jp}
+\end{equation}
+
+The contribution to the total $\chi^2$ of a broadband flux is
+therefore just:
+\begin{eqnarray}
+\label{chi2photo}
+\chi^2_{kp} &=& \left[\frac{\mu_{kp} - \hat\mu_{kp}}
+{\sigma_{kp}}\right]^2 \cr
+&=&
+\left[\frac{\mu_{kp} -
+\sum_{ij} a_{ki} b_{ij} \eta_{jp}
+}{\sigma^2_{kp}}\right]^2. 
+\end{eqnarray}
+For a given galaxy we do not have every filter, and we only have an
+observation of each filter at a single redshift. We pick the closest
+redshift on the redshift grid, and use the measured filters at that
+redshift for our expression of $\chi^2$, and set $1/\sigma_{kp}^2 = 0$
+for the rest of the values of the index $p$ so that we zero-weight
+those predictions.
+
+$\eta_{jp}$ and $M_{jl}$ are totally fixed, and we combine them into a
+single matrix $M_{jn}$ and the indices $p$ and $l$ into a single index
+$n$. We can similarly combine our observations $\mu_{kp}$ and
+$F_{\lambda, k}(\lambda_l)$ and their uncertainties $\sigma_{kp}$ and 
+$\sigma_{k}(\lambda_l)$ into vectors $x_{kn}$ and $\sigma_{kn}$.
+Then we can combine the Equations \ref{chi2spec} and \ref{chi2photo}
+into a single equation:
+\begin{equation}
 \label{chi2}
-\chi^2 &=& \sum_{k} \left[ \frac{F_{k,\mathrm{obs}} - F(k)}{\sigma_k}
-\right]^2 \cr
-&=& \sum_{ik} \left[ \frac{F_{k,\mathrm{obs}} -\sum_j a_j
-  v_j(\lambda)}{\sigma_k} \right]^2
-\end{eqnarray}
-over the $a_j$, where $\sigma_k$ is the standard deviation of the
-error in band $k$. Since we will use $v_j(\lambda)$ which are the light
-emitted from stars and ionized gas, we make the reasonable assumption
-that these only {\it add} light to the spectrum, not subtract it, and
-require nonnegativity: that $a_j\ge 0$ for all $j$. (We discuss dust
-below).  In order to perform this nonnegative minimization of Equation
-\ref{chi2}, we use the method of \citet{sha02a}.
+\chi^2 = \sum_{kn} \left[ 
+\frac{x_{kn} - \sum_{ij} a_{ki} b_{ij} M_{jn}}
+{\sigma_{kn}} \right]^2
+\end{equation}
+There is a simple method that we describe in the Appendix called
+nonnegative matrix factorization (NMF) to iterate to the nonnegative
+$a_{ki}$ and $b_{ij}$ which minimize Equation \ref{chi2}. The basic
+method is implemented in a public piece of code named {\tt
+NMF\_SPARSE} in the {\tt idlutils} distribution of IDL
+utilities.\footnote{\tt http://skymaps.info}. As the name implies, our
+implementation takes advantage of the fact that many of the matrix
+operations are on very sparse matrices (for example, for each galaxy
+with photometric data there are {\it no} spectroscopic data points and
+{\it only} photometric data at a single redshift).
 
-At this point, the reader probably realizes that the number of
-templates $j$ should not be much higher than the number of bands $k$,
-or else the problem becomes very degenerate. In fact, for the case of
-unconstrained minimization of Equation \ref{chi2} the problem is
-formally degenerate if $j>k$. The nonnegativity constraint adds enough
-information that this is no longer the case, but still one cannot get
-strong constraints on a huge number of parameters given a small number
-of measurements. Nevertheless, below we {\it will} use a huge number
-of $j$ given the number of $k$, simply for the purposes of finding the
-subspace of possible galaxy spectra which actually conform to galaxy
-observations. We do not mean to imply that we can determine each of
-the resulting parameters at high signal-to-noise.
+Once we have fit for $b_{ij}$ using the training set, we can minimize
+Equation \ref{chi2} for any other galaxy using any nonnegative least
+squares algorithm (since the minimization of Equation \ref{chi2} has a
+linear form in that case).  When we do so here we use the beautifully
+simple iterative method of \citet{sha02a}.
 
-For our model templates we use a set of spectra from the stellar
-evolution synthesis results of PEGASE2 (\citealt{fioc97a}). We use 15
-bursts of varying ages. The top panel of Figure \ref{sfh} shows the
-star-formation history of each of these bursts. For each age we create
-five templates with five different metallicities ($Z=0.03$, $0.02$,
-$0.008$, $0.004$, and $0.001$). The bottom panel of Figure \ref{sfh}
-shows the template for the $Z=0.008$ case at every other age, starting
-with the oldest. The bottom panel of For each age and metallicity, we
-create three templates with three different extinctions ($\tau_V=0$,
-$1$, and $3$) using the Milky Way extinction law using the ``clumpy,''
-``dusty'' model of \citet{witt00a}. This procedure results in 225
-templates and spans a very large range of possible spectra. Our
-procedure is designed to take this large dimensionality space and
-determine a low dimensionality subspace in which galaxies reside.
+\section{Results}
 
-We fit these models to a set of galaxies for which we know the
-redshifts. The majority of the redshifts are from the SDSS, but a
-small number of the redshifts are from CNOC2 ({\bf ref}). We consider
-only redshifts with 1$^\circ$.25 of the J2000 Equator, between
-$-60<\alpha<100$. We match the locations of these spectra to the SDSS
-photometric catalog ({\bf describe}) in all of the observations in
-this region and to the 2MASS Extended Source Catalog (XSC). In
-passing, we note that there are large errors (many arcseconds) in the
-catalog astrometry for CNOC2, and we needed to take care to account
-for these errors. For spectra which had matches in multiple runs of
-the SDSS, we take the (unweighted) mean of their photometry to reduce
-errors {\bf how many is this done for?}. We select a random set of
-these galaxies in such a way that the number per redshift is nearly
-constant across redshift. For each galaxy, we consider its
-extinction-corrected model photometry. Figure \ref{gmrz} shows the
-observed $g-r$ colors and redshifts of the resulting set of
-galaxies. Note that only for the lower redshift galaxies do we have
-matches to the 2MASS XSC.
+\subsection{An example: the Luminous Red Galaxy templates}
 
-Figure \ref{kfit} shows the residuals between the observed and fit
-colors after minimizing Equation \ref{chi2} for all the galaxies in
-our sample. Interestingly, even this large set of parameters cannot
-perfectly describe the trends of color with redshift. There are
-deviations at the percent level or so in $ugriz$. The $J$-band 2MASS
-observations at moderate redshift appear to be about $0.1$ magnitudes
-off of the models; this discrepancy is probably related to the similar
-discrepancies found by \citet{blanton03d} and \citet{baldry03a} when
-comparing the luminosity density of the universe to models.  Our
-attitude in this paper will be to trust the data in preference to the
-models and so below we describe how we deal with these discrepancies.
+We begin with the simplest case, which is fitting a {\it single}
+template to the photometric data of the LRG sample
+(\citealt{eisenstein02a}). For many of these galaxies, which extend to
+$r<19.5$ and are intrinsically red, the $u$-band flux is extremely
+poorly measured, so we ignore the $u$-band for all LRGs.
 
-Each many-parameter fit to the broad-band fluxes of each galaxy yields
-a complete spectrum from which to derive $K$-corrections. It also
-provides a value of $\chi^2$; in principle, for galaxies without
-redshifts we can find the assumed redshift with minimum $\chi^2$ in
-order to find the photometric redshift. In practice, the fit takes too
-long to perform to be useful for either of these purposes. Thus, we
-describe here our method for reducing the dimensionality to a set of
-three templates which adequately describe the data.
+Figure \ref{spec_lrg} shows the spectrum of the best-fit LRG.  This
+template is constrained by data between about 2000 and 10000 \AA;
+outside that range it is an extrapolation.
 
-Our procedure has two steps. First, we use principle component
-analysis on the set of 225-coefficent vectors describing all the fits
-in order to find a 10-dimensional subspace which still describes the
-data. Second, within this 10-dimensional subspace we group the
-galaxies into a set of ten groups, average the spectra within each
-group, and pick the triple among these groups which best describes the
-data. We adopted this procedure because we did not know of a
-principled way in which to reduce the dimensionality without
-sacrificing nonnegativity; however, it turns out that such methods do
-exist, and we recommend those methods for future investigators ({\bf
-refs from sam}) while retaining this method for the moment because it
-appears to work for our purposes.
+Figure \ref{sfh_lrg} shows the star-formation history corresponding to
+this best-fit LRG spectrum. The top panel shows the star-formation per
+unit time as a function of look-back time.  The thick, straight
 
-Because the galaxies all have very different bolometric fluxes, there
-is a strong variation in the coefficient values simply due to the
-total flux differing by large factors. Therefore, before performing
-principal component analysis, we scale all the coefficients of each
-galaxy by a single number (per galaxy) such that the galaxies all lie
-exactly on a plane in coefficient space. We choose the plane to be the
-plane on which the sum of the templates has a constant flux between
-1500 \AA\ and 30000 \AA. After scaling the galaxies to this plane, we
-perform principal component analysis to obtain the five most principal
-axes, using the efficient method of \citet{roweis99a}. There is almost
-no difference between the full fit shown above and that using the
-first five eigencoefficients.
+Figure \ref{lrg_colors} shows the LRG colors as a function of
+redshift, with the best fit template color overplotted as the smooth,
+thick line. Thus, the best fit is a good fit to the data
 
-On the other hand, the five eigencomponents natural consist partly of
-positive star-formation components and partly of negative
-star-formation components. This means that we cannot simply sum them
-nonnegatively to get a nonnegative star-formation history. In order to
-get a set of components that consists purely of all-positive
-star-formation histories, we consider the distribution of galaxies in
-the five-component space. We perform a $K$-means grouping on this set
-of points, which produces ten groups. This is simply a way of very
-roughly mapping the locus on which the galaxies reside (it is
-significantly nonlinear within the five-dimensional space, so we
-cannot simply take fewer principal components). We then average the
-spectra in each group. This gives ten spectra who roughly span the
-space of galaxy types. Finally, we pick from among those ten the
-triple which best describes the galaxies, in the sense of minimizing
-$\chi^2$ in Equation \ref{chi2}.
+Note that in this example we have fit a nonevolving template, which is
+inappropriate over this range of redshifts --- even if it is a good
+fit to the colors! In principle, we can adjust the methods used here
+for the case of evolving templates, but we will not do so here.
 
-Figure \ref{kfit3} shows the color residuals after fitting with the
-final three templates.
+\subsection{Description of the templates}
 
-[ new model: 3 PEGASE2 spectra; Figure show spectra, Figure stellar
-  mass to light ratios ]
+Figure \ref{spec_templates} shows the spectrum from the ultraviolet
+through the near infrared for the resulting templates. Note that there
+is a very old template, a very young template, and several
+intermediate templates, including one which is close to that of an A
+star.
 
-[ to deal with that for K-correction and photo-z, tweak ]
+Figure \ref{sfh_templates} shows the star-formation histories and
+metallicities associated with the templates. We have not imposed any
+smoothness criterion on these fits, which explains the ragged
+appearance of these histories. We do not believe the details of these
+fits, since neighboring ages or metallicities are highly degenerate
+with one another.
 
-We recognize that this procedure is rather convoluted and a bit of a
-hack. As noted above, we have adopted it simply as one way of
-automatically getting a small set of templates that explain galaxy
-broad-band fluxes as a sum of star-formation histories. There are
-methods for directly performing principal component analysis subject
-to the conditions that the components and the coefficients are all
-positive, and we recommend the investigation of such methods to the
-reader ({\bf ref}).
+{\bf how are numbers released}
+
+\subsection{Explanatory power of templates}
+
+These templates explain the photometric data rather well. 
+[SDSS]
+
+[SDSS + 2MASS]
+
+[SDSS + GALEX]
+
+[DEEP2]
+
+[GOODS]
+
+In addition, they provide fairly good fits to the SDSS spectroscopic
+data. However, in detail they do not fit the SDSS data.
+
+\subsection{Predictive power of templates}
+
+[2MASS]
+
+[GALEX]
+
+[ H alpha color relationship ]
 
@@
 
-[ Figure test: against a bunch of spectra ]
+[ Figure test: against projection of a bunch of spectra ]
 
-\section{Determining photometric redshifts}
-\label{kphotoz}
+\section{Physical interpretation of the models}
 
-[ describe simple fitting procedure ]
+[ solar magnitudes ]
 
-[ Figure show training set results ]
+[ Figure mass-to-light ratio ]
 
-[ Figure show results on deeper sdss data ]
+[ Figure recent SF ]
+
+\section{Linear relationships between common magnitude systems}
+
+[ vega to ab ]
+
+[ conversions ]
 
 \section{Public access to the code}
+\label{code}
 
@@
 
-Sam Roweis
+This publication makes use of data products from the Two Micron All
+Sky Survey, which is a joint project of the University of
+Massachusetts and the Infrared Processing and Analysis
+Center/California Institute of Technology, funded by the National
+Aeronautics and Space Administration and the National Science
+Foundation.
+
+The Galaxy Evolution Explorer (GALEX) is a NASA Small Explorer. The
+mission was developed in cooperation with the Centre National d'Etudes
+Spatiales of France and the Korean Ministry of Science and Technology.
+
+DEEP2 is a collaboration between UC Santa Cruz and UC Berkeley.
+Funding for the DEEP2 survey has been provided by NSF grant
+AST-0071048 and AST-0071198.
 
@@
 
-\section{Approximate conversions between common magnitude systems}
+\section{Nonnegative matrix factorization (NMF)}
+
+\section{$K$-corrections to Shifted bandpasses}
+
+{\bf NMF}
 

### 2005-07-07 src/k_fit_photoz.c

@@ -12,8 +12,8 @@
  * Mike Blanton
  * 5/2003 */
 
-#define ZRES 0.10
-#define TOL 0.02
+#define ZRES 0.05
+#define TOL 0.005
 #define MAXITER 3000
 #define FREEVEC(a) {if((a)!=NULL) free((char *) (a)); (a)=NULL;}
 

### 2005-07-26 docs/paper/kcorrect.tex
writing paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@455 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \newcommand{\band}[2]{\ensuremath{^{{#1}}\!{#2}}}
-
-\newcommand{\kversion}{{\tt v1\_11}}
 
@@
 to explain observed galaxy fluxes. Such templates are useful for
-calculating $K$-corrections, calculating photometric redshifts, and
-converting among observations of galaxies at various wavelengths. In
-principle, we could apply this method to galaxy data at any redshift
-and in any set of wavelengths. The implementation we present here is
-suitable for ultraviolet, optical, and near-infrared
-observations. Since we base the templates on stellar population
-synthesis results, the results are intepretable in terms of
-approximate stellar masses and star-formation histories. We present
-templates fit with this method to data from GALEX, Sloan Digital Sky
-Survey spectroscopy and photometry, the Two-Micron All Sky Survey, the
-Deep Extragalactic Evolutionary Probe and Great Observatories Origins
-Deep Survey.  In addition, we present software for using such data to
-estimate $K$-corrections, photometric redshifts, evolution
+calculating $K$-corrections and converting among observations of
+galaxies at various wavelengths. In principle, we could apply this
+method to galaxy data at any redshift and in any set of
+wavelengths. The implementation we present here is suitable for
+ultraviolet, optical, and near-infrared observations. Since we base
+the templates on stellar population synthesis results, the results are
+intepretable in terms of approximate stellar masses and star-formation
+histories. We present templates fit with this method to data from
+GALEX, Sloan Digital Sky Survey spectroscopy and photometry, the
+Two-Micron All Sky Survey, the Deep Extragalactic Evolutionary Probe
+and Great Observatories Origins Deep Survey.  In addition, we present
+software for using such data to estimate $K$-corrections, evolution
 corrections, stellar masses, and recent star-formation rates. This
 software is suitable for analyzing many other data sets as well.
-\end{abstract}code
+\end{abstract}
 
@@
 redshift $z$ and the desired bandpass $Q$ is defined by the equation
-(\cite{oke68a, hogg02a}):
+(\citealt{oke68a, hogg02a}):
 \begin{equation}
 \label{kcorrecteqn}
-m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z) - 5 \log_10 h 
+m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z) - 5 \log_{10} h 
 \end{equation}
@@
 Explorer (GALEX; {\bf ref})) in the ultraviolet, and the Two-Micron
-All Sky Survey (2MASS; \citealt{strutskie97a}) in the near infrared
+All Sky Survey (2MASS; \citealt{skrutskie97a}) in the near infrared
 (NIR). In addition, at higher redshifts, we use constraints from the
-Deep Extragalactic Evolutionary Probe 2 (DEEP2; {\bf ref} and the
-Great Observatories Origins Deep Survey (GOODS; {\bf ref}).  These and
-other data sets provide a huge set of information about galaxy colors
-and spectra which we can use to help understand their star-formation
-histories.
+Deep Extragalactic Evolutionary Probe 2 (DEEP2; \citealt{davis03a,
+faber03a}) and the Great Observatories Origins Deep Survey (GOODS;
+\citealt{giavalisco04a}).  These and other data sets provide a huge set of
+information about galaxy colors and spectra which we can use to help
+understand their star-formation histories.
 
@@
 high-level interface.  This software is an update of two major earlier
-releases ({\tt v1\_16}, \citet{blanton03b}; {\tt v3\_2}). The
+releases ({\tt v1\_16}, \citealt{blanton03b}; {\tt v3\_2}). The
 improvement over the latest version is twofold. First, we have
@@
 also has a number of new useful functions.
+
+In Section \ref{algorithm}, we describe how to find suitable
+templates, given a set of models and a set of data. We also describe
+the data and models used here. In Section \ref{results}, we show the
+results: the best-fit templates and how well they fit the data.
 
@@
 models and the astronomical observations, and to pose the problem in
-the form needed for the purposes of NMF. In the Appendix we describe
+the form needed for the purposes of NMF. In Appendix \ref{nmf} we describe
 how to solve to NMF problem.
@@
 recalibrated our photometry using the ``ubercalibration'' procedure
-described in \citet{blanton05b}.  Descriptions of these pipelines also
+described in \citet{blanton05a}.  Descriptions of these pipelines also
 exist in \citet{stoughton02a}.  An automated pipeline called {\tt
@@
 above, for the photometry we express the models in terms of predicted
-broadband fluxes.
+broadband fluxes at each redshift.
 
@@
 \label{maggies}
-\eta_{jp} = 
+\mu_{jp} = 
 \frac{\int_0^{\infty} d\lambda \lambda R_p(\lambda) {m}_{\lambda, j}(\lambda)}
@@
 \end{equation}
-There is a simple method that we describe in the Appendix called
+There is a simple method that we describe in Appendix \ref{nmf} called
 nonnegative matrix factorization (NMF) to iterate to the nonnegative
@@
 \section{Results}
+\label{results}
 
@@
 Figure \ref{spec_lrg} shows the spectrum of the best-fit LRG.  This
-template is constrained by data between about 2000 and 10000 \AA;
+template is constrained by data between about 2000 \AA\ and 10000 \AA;
 outside that range it is an extrapolation.
@@
 this best-fit LRG spectrum. The top panel shows the star-formation per
-unit time as a function of look-back time.  The thick, straight
+unit time as a function of look-back time.  The thick, straight line
+is the power law:
+\begin{equation}
+\mathrm{SFR} = \left(\frac{t}{t_0})^2,
+\end{equation}
+where $t$ is the look-back time, for $t<13.75$ $h^{-1}$ Gyr, and $t_0=$
+{\bf number}. The bottom panel shows the mean metallicity of the
+forming stars as a function of time.
 
@@
 redshift, with the best fit template color overplotted as the smooth,
-thick line. Thus, the best fit is a good fit to the data
+thick line. In the code described in Section \ref{code}, we used the
+routine {\tt sdss\_kcorrect} (with the optional flag {\tt /lrg}) to
+perform these fits. The best fit is a good fit to the data.
 
@@
 
-\subsection{Description of the templates}
+In Appendix \ref{format} we describe the form in which we release 
+the star-formation histories and spectra associated with this
+template.
+
+\subsection{Templates for the full data set}
 
@@
 
-{\bf how are numbers released}
+In all the results below, we use these five templates, unaltered. That
+is, when we speak below of ``fitting'' the templates, we mean we fix
+$b_{ij}$ to the five templates shown in this section and fit only for
+$a_{ki}$. In general below, the tests we perform are not to the
+training set but to independent sets of galaxies.
+
+In Appendix \ref{format} we describe the form in which we release the
+star-formation histories and spectra associated with these templates.
 
@@
 
-These templates explain the photometric data rather well. 
-[SDSS]
+These templates explain the photometric data rather well. For example,
+consider Figure \ref{fullfits}, showing the color residuals of the
+observations with respect to the best fits, when fitting to galaxies
+with GALEX, SDSS, and 2MASS data. In this context we define the color
+residuals as (for example):
+\begin{equation}
+\Delta [u-g] = [u_{\mathrm{obs}}- g_{\mathrm{obs}}] -
+[u_{\mathrm{model}}- g_{\mathrm{model}}] .
+\end{equation}
+In the code described in Section \ref{code}, we used the routine {\tt
+gst\_kcorrect} to perform these fits.  The templates do just as well
+on higher redshift data. For example, Figure \ref{goods} shows the
+color residuals for GOODS data.
 
-[SDSS + 2MASS]
-
-[SDSS + GALEX]
-
-[DEEP2]
-
-[GOODS]
-
-In addition, they provide fairly good fits to the SDSS spectroscopic
-data. However, in detail they do not fit the SDSS data.
+{\bf comparison to spectra}
 
@@
 
-[2MASS]
+Perhaps more interestingly, the templates do a good job of predicting
+{\it missing} data. That is, we can ask the question: if we use the
+templates to fit only to some bands but leave out others, how well do
+the best fits {\it predict} the bands left out?
 
-[GALEX]
+Consider Figures \ref{twomass_resid} and \ref{twomass_predicted}. The
+former shows the color residuals of {\it fitting} to both SDSS and
+2MASS data for each galaxy. The latter shows the color residuals when
+we fit {\it only} to the SDSS data and do not include 2MASS in the fit
+at all. Without any input from 2MASS the templates do a very good job
+of predicting the 2MASS fluxes, with a scatter of 20--30\% {\bf real
+number} --- not far from the uncertainties in the 2MASS fluxes
+themselves. Of course, it is not surprising that it is easy to predict
+the 2MASS fluxes --- stellar spectra are very simple in the NIR, so
+having the $ugriz$-band fluxes from SDSS is very information regarding
+the redder bandpasses.
+
+However, we will note here that this result flies in the face of a
+persistent insistence that NIR observations are necessary to measure
+stellar masses of galaxies. If we can predict the NIR observations
+themselves, clearly they cannot be adding much to our knowledge of the
+underlying stellar mass. This fact changes one's decisions about what
+data is best to use for calculating the stellar mass function.  Basing
+it on 2MASS data only improves slightly the stellar mass estimate of
+each galaxy, while restricting (and thus biasing) the sample
+significantly at the lowest luminosities and surface brightnesses
+relative to the SDSS.
+
+In addition, consider Figures \ref{galex_resid} and
+\ref{galex_predicted}. The former shows the color residuals of {\it
+fitting} to both SDSS and GALEX data for each galaxy. The latter shows
+the color residuals when we fit {\it only} to the SDSS data and do not
+include GALEX in the fit at all. Without any input from GALEX the
+templates do a significantly worse job at predicting the GALEX
+fluxes. The scatter becomes about $\sigma \sim 0.5$ mag {\bf real
+number}. As is common knowledge, it is more difficult to predict the
+UV fluxes, because dust and recent star-formation are so variable
+among galaxies. However, the median residuals based on are our
+templates still near zero.
 
@@
 Given a model spectrum for the galaxy, the determination of the
-$K$-correction is straightforward. Here we repeat the
+$K$-correction is straightforward. Here we give the relevant formulae
+for the $K$-corrections, leaving the derivation to \citet{hogg02a}.
+Then, we show the typical $K$-corrections for the data we have fit
+to. Finally, we test our method on actual spectra projected on
+bandpasses to make sure that it works properly.
 
-[ Figure demonstrate K-corrections ]
+The $K$-correction between a bandpass $R$ used to observe a galaxy at
+redshift $z$ and the desired bandpass $Q$ is defined by the equation
+(\citealt{oke68a, hogg02a}):
+\begin{equation}
+\label{kcorrecteqn}
+m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z) - 5 \log_{10} h 
+\end{equation}
+where $\mathrm{DM}(z) = 25 - 5\log_{10} (d_L /
+(h^{-1}{\mathrm{~Mpc}}))$ is the bolometric distance modulus
+calculated from the luminosity distance $d_L$, and $M_Q$ is the
+absolute magnitude. The absolute magnitude is defined as the apparent
+magnitude an object would have if were observed 10 pc away, in
+bandpass $Q$, at rest.  The traditional definition of the
+$K$-correction takes $Q=R$. however, we note that in practice many
+surveys do perform $K$-corrections from one observed bandpass $R$ to
+another bandpass $Q$ in the rest frame. This practice is particularly
+common when dealing with high redshift observations.  In addition to
+$K$-corrections, this method also provides an interpretation of the
+data in terms of a physical model which describes the stellar mass and
+star-formation history of each galaxy.
+
+{\bf kcorrection equations + figure}
+
+For example, consider Figures \ref{galex_kcorrect},
+\ref{sdss_kcorrect} and \ref{twomass_kcorrect}. For a randomly
+selected set of galaxies these figures show the $K$-corrections from
+the observed frame bandpasses of GALEX, SDSS, and 2MASS respectively,
+to the rest-frame bandpasses of each survey.
 
@@
 \section{Nonnegative matrix factorization (NMF)}
+\label{nmf}
 
-\section{$K$-corrections to Shifted bandpasses}
+\section{Format for templates}
+\label{format}
 
-{\bf NMF}
+\section{$K$-corrections to shifted bandpasses}
+\label{bandshift}
 

### 2005-07-27 docs/paper/kcorrect.tex
change fitting defaults
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@470 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \newcommand{\band}[2]{\ensuremath{^{{#1}}\!{#2}}}
+\newcommand{\lambdaobs}{\lambda_o}
+\newcommand{\lambdaemit}{\lambda_e}
 
@@
 \begin{equation}
-\mathrm{SFR} = \left(\frac{t}{t_0})^2,
+\mathrm{SFR} = \left(\frac{t}{t_0}\right)^2,
 \end{equation}
@@
 
-{\bf kcorrection equations + figure}
+Equation (\ref{eq:definition}) holds if the $K$-correction $K_{QR}$ is
+\begin{equation}
+\label{eq:wavelengthL}
+K_{QR} = -2.5\,\log_{10}\left[\frac{1}{[1+z]}\,
+  \frac{\displaystyle
+  \int\mathrm{d}\lambdaobs\,\lambdaobs\,L_{\lambda}\!\left(\frac{\lambdaobs}{1+z}\right)\,R(\lambdaobs)\,
+    \int\mathrm{d}\lambdaemit\,\lambdaemit\,
+    g^Q_{\lambda}(\lambdaemit)\,Q(\lambdaemit)}
+       {\displaystyle
+  \int\mathrm{d}\lambdaobs\,\lambdaobs\,g^R_{\lambda}(\lambdaobs)\,R(\lambdaobs)\,
+    \int\mathrm{d}\lambdaemit\,\lambdaemit\,
+    L_{\lambda}(\lambdaemit)\,Q(\lambdaemit)}
+\right] \;\;\;.
+\end{equation}
+Here, $R(\lambda)$ and $Q(\lambda)$ represent the response of the
+instrument per unit photon entering the Earth's atmosphere (or the
+telescope aperture for a space instrument).  $g^R_\lambda$ is the flux
+density per unit wavelength (e.g.~ ergs s$^{-1}$ cm$^{-2}$ \AA$^{-1}$)
+for the standard source for band $R$, and similarly for band $Q$. For
+example, if the magnitudes are AB relative, then these represent the
+AB standard source, while if they are AB relative then they represent
+the spectrum of Vega.
+
+A particularly common special case is when $R=Q$:
+\begin{equation}
+\label{eq:specialL}
+K_{QR} = -2.5\,\log_{10}\left[\frac{1}{[1+z]}\,
+  \frac{\displaystyle
+  \int\mathrm{d}\lambdaobs\,\lambdaobs\,L_{\lambda}\!\left(\frac{\lambdaobs}{1+z}\right)\,R(\lambdaobs)\,}
+       {\displaystyle
+    \int\mathrm{d}\lambdaemit\,\lambdaemit\,
+    L_{\lambda}(\lambdaemit)\,R(\lambdaemit)}
+\right] \;\;\;.
+\end{equation}
 

### 2005-08-05 docs/paper/kcorrect.tex
add late
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@472 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 DR1 release between $0.6 < z < 1.5$ (4000
-galaxies; {\bf ref}).
+galaxies; \citealt{davis03a, faber03a}).
 \item The $BVizJHK_s$ photometry for GOODS galaxies between $0.5 < z < 2$ (1000
-galaxies; {\bf ref}).
+galaxies; \citealt{giavalisco04a}).
 \end{enumerate}
@@
 the observed frame bandpasses of GALEX, SDSS, and 2MASS respectively,
-to the rest-frame bandpasses of each survey.
+to the same bandpasses in the rest frame.
 
-[ Figure show K-corrections to z=0 ]
+Often, it makes sense to $K$-correct to a different set of band passes
+than the rest frame. For example, the SDSS Main sample is mostly
+around $z=0.1$, so it does not always make sense to $K$-correct from
+the observed frame $r$ band to the rest frame $r$ band. Doing so means
+applying large and uncertain corrections to all galaxies. If precision
+is not important, this uncertainty might be worth the simplification
+it brings. However, if one is (say) interested in the evolution of
+galaxies, the clustering of galaxies, or something else that requires
+precision, it makes more sense to avoid introducing unnecessary
+uncertainty.
 
-[ note band-shifting ]
+Our solution to this problem is to correct the magnitude to
+blue-shifted bandpasses that correspond to the observed bandpass at
+some intermediate redshift. We denote these bandpasses \band{z}{b},
+where the blue-shift in this case is by a factor $1+z$. For example,
+in the case of SDSS galaxies it makes sense to correct to the
+\band{0.3}{r} bandpass --- the $r$ band blue-shifted to $z=0.1$. This
+bandpass has the property that the $K$-correction is independent of
+the SED of the galaxy. If the magnitudes are AB relative, then the
+$K$-correction to \band{z}{b} for a galaxy at redshift $z$ is simply
+$-2.5\log_{10} (1+z)$.
+
+Figure \ref{main_kcorrect} shows the $K$-corrections from the
+$[ugriz]$ bands to the $\band{0.1}{[ugriz]}$ bands as a function of
+redshift for SDSS galaxies. Note that the $K$-corrections converge at
+$z=0.1$, so that around that redshift (where most galaxies are) the
+uncertainties in the $K$-corrections are minimized. Compare this
+simplicity to Figure \ref{sdss_kcorrect}, where the $K$-corrections at
+redshift $z=0.1$ in the $g$-band range between $0.$ and $0.2$ (and in
+the $u$-band from $0.$ and $0.5$). Most galaxies are near redshift
+$z=0.1$, and obviously the $K$-corrections are not perfect --- why
+introduce a large uncertainty for most galaxies when it can easily be
+avoided?
 

### 2005-08-23 docs/paper/kcorrect.tex
doc paper work
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@475 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 In the code described in Section \ref{code}, we used the routine {\tt
-gst\_kcorrect} to perform these fits.  The templates do just as well
-on higher redshift data. For example, Figure \ref{goods} shows the
-color residuals for GOODS data.
+gst\_kcorrect} to perform these fits.  In Figure \ref{fullfits} the
+thick dashed lines show the estimated 1$\sigma$ uncertainties in the
+colors from the photometric catalogs. Relative to the uncertainties,
+there are no significant biases or redshift trends in these fits. Note
+that the galaxies used in this test are distinct from the training set
+on which we fit the five templates.
+
+The templates do just as well on higher redshift data. For example,
+Figure \ref{goods} shows the color residuals for GOODS data.
 
@@
 redshift $z=0.1$ in the $g$-band range between $0.$ and $0.2$ (and in
-the $u$-band from $0.$ and $0.5$). Most galaxies are near redshift
-$z=0.1$, and obviously the $K$-corrections are not perfect --- why
-introduce a large uncertainty for most galaxies when it can easily be
-avoided?
+the $u$-band from $0.$ and $0.5$). Most SDSS galaxies are near
+redshift $z=0.1$, and obviously the $K$-corrections are not perfect
+--- why introduce a large uncertainty for most galaxies when it can
+easily be avoided?
 

### 2005-08-23 docs/paper/kcorrect_figures.tex
doc paper work
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@475 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \figurenum{\fignum}
-\plotone{k_sfh.ps}
-\caption{\label{sfh} The top panel are the star-formation histories of
-the 15 bursts we use in our model. The bottom panel shows (for no dust
-and for $Z=0.008$) the PEGASE2 spectrum predicted for these bursts. }
+\plotone{spec_lrg.ps}
+\caption{\label{spec_lrg} Best fit LRG spectral template
+	(normalization is for a 1 $M_\odot$ galaxy located 10 pc away, or
+	equivalently a $10^12$ $M_\odot$ galaxy located 10 Mpc away).
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{gmrz.ps}
-\caption{\label{gmrz} }
+\plotone{sfh_lrg.ps}
+\caption{\label{sfh_lrg} Star-formation history corresponding to LRG
+	spectral template of Figure \ref{spec_lrg}. Top panel shows the
+	number of stars formed per logarithmic time interval ($t$ is
+	expressed in years, curve is normalized for a $10^{12}$ $M_\odot$
+	galaxy). Almost all of the stars are formed in the first couple of
+	billion years --- note that the recent ``spike'' is represents a
+	tiny fraction ($\sim 10^{-8}$) of the total number of stars. Bottom
+	panel shows the mean metallicity of the population as a function of
+	time.  }
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{kfit.ps}
-\caption{\label{kfit} The color residuals between the observed values
-  and those of the best fit to the nonnegative combination of 225
-  templates ($\delta(b-r)\equiv [b-r]_{\mathrm{obs}} -
-  [b-r]_{\mathrm{fit}}$). The greyscale is proportional to the
-  distribution of color residual at a given redshift. The lines are
-  0.16, 0.45, and 0.84 fractional quantiles. Note that we have zoomed
-  into $g-r$, $r-i$, and $i-z$ in order to see the residuals
-  better. At high redshift many galaxies have abnormally low $u$-band
-  fluxes relative to their $r$-band flux. At low redshift there are a
-  few galaxies whose near infrared $z$-band magnitudes are
-  lower. There is a general offset of nearly 0.1 magnitudes in
-  $J-r$. }
+\plotone{lrg_colors.ps}
+\caption{\label{lrg_colors} SDSS colors of LRGs as a function of
+	redshift. The greyscale is the conditional distribution of color
+	within each redshift bin. The thin lines are the 10\%, 25\%, 50\%,
+	75\%, and 90\% quantiles of the distribution. The thick line
+	is the prediction of the model. The $u$ band is not included in the
+	fit, and the $u$ magnitudes of most LRGs are poorly known. The other
+	colors fit the models reasonably well. This model, remember, is
+	given incredible freedom, meaning that the above agreement is the
+	best one can do with the stellar population synthesis code of
+	\citet{bruzual03a}. }
 \end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{fullfits.ps}
+\caption{\label{fullfits} Color residuals (defined explicitly in the
+	text) of GALEX, SDSS, and 2MASS observations relative to our best
+	fit 5-template model.  The greyscale is the conditional distribution
+	of the color residual given the redshift.  The thin lines are the
+	10\%, 25\%, 50\%, 75\%, and 90\% quantiles of the distribution. The
+	thick dashed lines show the estimated 1$\sigma$ uncertainties in the
+	colors from the photometric catalogs. Relative to the uncertainties,
+	there are no significant biases or redshift trends in these fits. }
+\end{figure}

### 2005-08-24 docs/paper/kcorrect.tex
wrok on paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@493 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-The templates do just as well on higher redshift data. For example,
-Figure \ref{goods} shows the color residuals for GOODS data.
-
-{\bf comparison to spectra}
+The templates do well on higher redshift data, as well. For example,
+Figure \ref{goods} shows the color residuals for GOODS data. Here we
+find that there are somewhat larger residuals with respect to the
+uncertainties (again shown as the thick dashed lines). We can compare
+this result to that in Figure \ref{goods_special}, which we get by
+fitting to {\it only} the GOODS data (this alternate set of five
+templates is also distributed in the manner that Appendix \ref{format}
+describes). From Figure \ref{goods_special}, we conclude that {\it
+some} of the errors are irreducible, and require either a larger set
+of templates for high redshift galaxies (a possibility we find a
+priori unlikely), or that there are simply errors in the catalogs or
+the input filter curves. For example, in the $z$ and $V$ bands we find
+significant trends between redshifts $z=0.5$ and $1.5$ (echoed in
+Figure \ref{goods}. More significantly, in the $H$ band there is at
+least a 10\% offset in the magnitudes across all redshifts, which is
+almost certainly a catalog error.  {\bf GOODS people: this is your
+challenge! what is going on?} Other errors (such as a large scatter in
+the residuals in the $B$ band and some redshift dependence in the $J$
+band) are clearly due to the fact that the templates are not primarily
+designed for GOODS data. All in all, we recommend using the special
+templates for $K$-corrections within the GOODS data set.
+
+We can also test how well these templates recover actual spectra given
+only photometry. To do so, we take eight random SDSS spectra (chosen
+to span color space), project them onto the $g$, $r$ and $i$
+bandpasses, and then fit the five templates to just these three
+fluxes.  Figure \ref{specfit} compares the reconstructed model spectra
+to the original spectrum. The two agree extremely well. Naturally,
+this success depends heavily on the quality of the original stellar
+spectra used in the population synthesis code of \citet{bruzual03a}.
 

### 2005-08-24 docs/paper/kcorrect_figures.tex
wrok on paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@493 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{goods.ps}
+\caption{\label{goods} Color residuals in fit using the standard five
+templates to GOODS data, compared to the typical uncertainties (thick
+dashed lines). Note that the fits always do poorly on the $H$ band,
+which we believe to be a catalog error. }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{goods_special.ps}
+\caption{\label{goods_special} Same as Figure \ref{goods}, but fitting
+using five templates specially designed for GOODS. These templates
+have smaller residuals in many respects but still fail to fit the $H$
+band data. }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{specfit.ps}
+\caption{\label{specfit} Best fit model spectra based on the five
+template fit to $g$, $r$ and $i$ fluxes, compared to the original
+SDSS spectra from which we computed those fluxes.  The models and the
+original spectra agree very well.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{twomass_resid.ps}
+\caption{\label{twomass_resid} Similar to Figure \ref{fullfits} but
+for galaxies observed in both SDSS and 2MASS and only using SDSS and
+2MASS bands. }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{twomass_resid.ps}
+\caption{\label{twomass_resid} Similar to Figure \ref{fullfits} but
+for galaxies observed in both SDSS and 2MASS and only using SDSS and
+2MASS bands. }
+\end{figure}

### 2005-08-25 docs/paper/kcorrect.tex
work on paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@498 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \begin{equation}
-\label{kcorrecteqn}
 m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z) - 5 \log_{10} h 
@@
 9000$ \AA, for 400 Luminous Red Galaxies between $0.15 < z < 0.5$ (LRGs;
-\citealt{eisenstein02a}) and 1600 Main sample galaxies between $0.001
+\citealt{eisenstein01a}) and 1600 Main sample galaxies between $0.001
 < z < 0.4$ (\citealt{strauss02a})
@@
 (NYU-VAGC;
-\citealt{blanton05b}).
+\citealt{blanton05a}).
 
@@
 template to the photometric data of the LRG sample
-(\citealt{eisenstein02a}). For many of these galaxies, which extend to
+(\citealt{eisenstein01a}). For many of these galaxies, which extend to
 $r<19.5$ and are intrinsically red, the $u$-band flux is extremely
@@
 
-Figure \ref{spec_templates} shows the spectrum from the ultraviolet
-through the near infrared for the resulting templates. Note that there
-is a very old template, a very young template, and several
-intermediate templates, including one which is close to that of an A
-star.
-
-Figure \ref{sfh_templates} shows the star-formation histories and
-metallicities associated with the templates. We have not imposed any
-smoothness criterion on these fits, which explains the ragged
-appearance of these histories. We do not believe the details of these
-fits, since neighboring ages or metallicities are highly degenerate
-with one another.
+The right-most column of Figure \ref{sfh_templates} shows the spectrum
+from the ultraviolet through the near infrared for the resulting
+templates. Note that there is a very old template, a very young
+template, and several intermediate templates, including one which is
+close to that of an A star.
+
+The left and middle columns of Figure \ref{sfh_templates} show the
+star-formation histories and metallicities associated with the
+templates. We have not imposed any smoothness criterion on these fits,
+which explains the ragged appearance of these histories. We do not
+believe the details of these fits, since neighboring ages or
+metallicities are highly degenerate with one another.
 
@@
 the 2MASS fluxes --- stellar spectra are very simple in the NIR, so
-having the $ugriz$-band fluxes from SDSS is very information regarding
-the redder bandpasses.
+having the $ugriz$-band fluxes from SDSS yields much information
+regarding the redder bandpasses.
 
@@
 templates do a significantly worse job at predicting the GALEX
-fluxes. The scatter becomes about $\sigma \sim 0.5$ mag {\bf real
-number}. As is common knowledge, it is more difficult to predict the
-UV fluxes, because dust and recent star-formation are so variable
-among galaxies. However, the median residuals based on are our
-templates still near zero.
-
-[ H alpha color relationship ]
+fluxes. The scatter becomes about $\sigma \sim 0.5$ mag.  As is common
+knowledge, it is more difficult to predict the UV fluxes, because dust
+and recent star-formation are so variable among galaxies. However, the
+median residuals based on are our templates still near zero.
 
@@
 
-Equation (\ref{eq:definition}) holds if the $K$-correction $K_{QR}$ is
+Equation (\ref{kcorrecteqn}) holds if the $K$-correction $K_{QR}$ is
 \begin{equation}
@@
 
-[ Figure test: against projection of a bunch of spectra ]
-
 \section{Physical interpretation of the models}
 
+In addition to $K$-corrections, these template fits also provide
+physical interpretations of the galaxies, since they correspond to
+actual stellar population synthesis models. For example, 
+Figure \ref{mtol} shows the mass-to-light ratios of galaxies as a
+function of color.
+
+[ Figure mass-to-light ratio; Figure compare to whoever ]
+
+[ Figure recent SF ]
+
+\section{Linear relationships between common magnitude systems}
+
+Finally, there some practical relationships between different band
+systems and conventions.
+
 [ solar magnitudes ]
-
-[ Figure mass-to-light ratio ]
-
-[ Figure recent SF ]
-
-\section{Linear relationships between common magnitude systems}
 

### 2005-08-25 docs/paper/kcorrect_figures.tex
work on paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@498 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 	(normalization is for a 1 $M_\odot$ galaxy located 10 pc away, or
-	equivalently a $10^12$ $M_\odot$ galaxy located 10 Mpc away).
+	equivalently a $10^12$ $M_\odot$ galaxy located 10 Mpc away).}
 \end{figure}
@@
 \caption{\label{lrg_colors} SDSS colors of LRGs as a function of
+	redshift. The greyscale is the conditional distribution of color
+	within each redshift bin. The thin lines are the 10\%, 25\%, 50\%,
+	75\%, and 90\% quantiles of the distribution. The thick line
+	is the prediction of the model. The $u$ band is not included in the
+	fit, and the $u$ magnitudes of most LRGs are poorly known. The other
+	colors fit the models reasonably well. This model, remember, is
+	given incredible freedom, meaning that the above agreement is the
+	best one can do with the stellar population synthesis code of
+	\citet{bruzual03a}. }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{sfh_templates.ps}
+\caption{\label{sfh_templates} SDSS colors of LRGs as a function of
 	redshift. The greyscale is the conditional distribution of color
@@
 for galaxies observed in both SDSS and 2MASS and only using SDSS and
-2MASS bands. }
+2MASS bands. The fits are to the SDSS and 2MASS data together.}
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{twomass_resid.ps}
-\caption{\label{twomass_resid} Similar to Figure \ref{fullfits} but
-for galaxies observed in both SDSS and 2MASS and only using SDSS and
-2MASS bands. }
+\plotone{twomass_predicted.ps}
+\caption{\label{twomass_predicted} Similar to Figure
+\ref{twomass_resid} but now the fits are {\it only} to the SDSS
+bands. The residuals in the 2MASS bands remain very small, indicating
+that the 2MASS measurements do not add a lot of information about
+these galaxies.}
 \end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{galex_resid.ps}
+\caption{\label{galex_resid} Similar to Figure \ref{twomass_resid} but
+for galaxies observed in both SDSS and GALEX and only using SDSS and
+GALEX bands. The fits are to the SDSS and GALEX data together.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{galex_predicted.ps}
+\caption{\label{galex_predicted} Similar to Figure
+\ref{galex_resid} but now the fits are {\it only} to the SDSS
+bands. }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{galex_kcorrect.ps}
+\caption{\label{galex_kcorrect} $K$-corrections as a function of
+redshift in the GALEX near (N) and far (F) UV bands. }
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{twomass_kcorrect.ps}
+\caption{\label{twomass_kcorrect} $K$-corrections as a function of
+redshift in the 2MASS $J$, $H$ and $K_s$ bands.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{sdss_kcorrect.ps}
+\caption{\label{sdss_kcorrect} $K$-corrections as a function of
+redshift in the SDSS $u$, $g$, $r$, $i$, and $z$ bands.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{sdss_kcorrect.ps}
+\caption{\label{sdss_kcorrect} $K$-corrections as a function of
+redshift in the SDSS $u$, $g$, $r$, $i$, and $z$ bands.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{main_kcorrect.ps}
+\caption{\label{main_kcorrect} Same as Figure \ref{sdss_kcorrect}, but
+$K$-correcting to the \band{0.1}{u}, \band{0.1}{g}, \band{0.1}{r},
+\band{0.1}{i}, and \band{0.1}{z} bands.}  
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{mtol.ps}
+\caption{\label{mtol} Mass-to-light ratios of galaxies (in solar
+units) as a function of galaxy color. The solid line satisfies the
+relationship $\log_{10}(M/L) = 1.65 (g-r) - 0.54$, which has 
+the same slope as that found by \citet{bell01a}, with a higher
+normalization by 0.3 dex.}
+\end{figure}

### 2005-08-30 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@499 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 physical interpretations of the galaxies, since they correspond to
-actual stellar population synthesis models. For example, 
-Figure \ref{mtol} shows the mass-to-light ratios of galaxies as a
-function of color.
-
-[ Figure mass-to-light ratio; Figure compare to whoever ]
-
-[ Figure recent SF ]
+actual stellar population synthesis models. Indeed, the code outputs
+three parameters relevant to the star-formation history: the current
+mass of stars in the galaxy; the stellar mass-to-light ratio of the
+galaxy; and the fraction of the total amount of star-formation that
+has occurred recently.
+
+\citet{kauffmann03a} have also calculated stellar masses for SDSS
+galaxies. Roughly, they have calculated the $z$-band stellar
+mass-to-light ratios of models which best fit the H$\delta$ absorption
+and D4000 measurements of the spectra, and then applied those
+mass-to-light ratios to the observed $z$-band luminosities of the
+imaging. Figure \ref{mass_to_garching} shows a galaxy-by-galaxy
+comparison of their stellar masses $M_{\ast,s}$ relative to ours
+$M_\ast$, as a function of $M_\ast$ and of intrinsic $g-r$ color. The
+two sets of masses are very similar to each other, with a scatter of
+only 0.1 dex and trends of less than 0.2 dex with stellar mass. As we
+did, \citet{kauffmann03a} used the library of \citet{bruzual03a} with
+the \citet{chabrier03} initial mass function, which makes some of this
+agreement unsurprising.  To be explicit, the masses in Figure
+\ref{mass_to_garching} correspond to the current mass in stars
+remaining in the galaxy. This mass can differ from the total
+star-formation rate integrated over time, due to the fact that stars
+can die --- and in practice for our fits the total integrated
+star-formation rate is usually about twice as high as the total
+current mass in stars.
+
+Similarly, Figure \ref{mtol} shows the mass-to-light ratios of SDSS
+galaxies in the $V$ band as a function of color in $B-V$. Here, we
+have calculated the $B$ and $V$ band fluxes based on the template fits
+to each galaxy, as explained in Section \ref{sdss2bessell} below. We
+give the mass-to-light ratio in solar units. The solid line is the
+relationship that \citet{bell01a} give based on their fits to spiral
+galaxies. For blue galaxies ($B-V < 0.8$) this line is indeed a good
+fit to our results. For the typical red galaxy, the \citet{bell01a}
+relation predicts a larger stellar mass than we do.
+
+Finally, we can also calculate measures of the recent star-formation
+rate from these fits. Since our fits are to broad-band photometry, and
+not based on emission line measurements, we cannot expect to have
+detailed information about the very recent star-formation
+rate. However, we can try to measure what fraction of the total
+star-formation has occurred in, say, the past 1 Gyr:
+\begin{equation}
+b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
+{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
+\end{equation}
+Note that here $b_G$ is defined in terms of the total integrated
+star-formation history (unlike the stellar masses referred to above)
+as is standard for the $b$ parameter (\citealt{kennicutt98a}). Figure
+\ref{umr_bg} shows the relationship between the rest-frame $u-r$ color
+for SDSS galaxies and $b_G$. For galaxies with $u-r < 2.5$ (redder
+than which galaxies are mostly reddened, edge-on disks) these two
+parameters follow a simple relationship:
+\begin{equation}
+\log_{10} b_g = -0.55 (u-r) - 0.25.
+\end{equation}
 
 \section{Linear relationships between common magnitude systems}
-
-Finally, there some practical relationships between different band
-systems and conventions.
-
-[ solar magnitudes ]
-
-[ vega to ab ]
+\label{sdss2bessell}
+
+The software and templates that we present here are also useful for
+determining simple conversions between bandpasses performing other
+common tasks, such as calculating the absolute magnitude of the Sun in
+various band systems, and calculating the conversion between Vega and
+AB magnitudes. In this section, we present these tools.
+
+For example, Table \ref{solarmagnitudes} lists the effective
+wavelengths, conversion from Vega to AB magnitudes, and absolute
+magnitudes of the Sun in a number of filters: the GALEX filters, the
+Bessell filters, the SDSS filters, and the 2MASS filters. The listed
+numbers are the results of running the IDL functions {\tt
+k\_lambda\_eff}, {\tt k\_vega2ab}, and {\tt k\_solar\_magnitudes},
+respectively, so the reader can calculate the same thing easily on any
+given filter.  The effective wavelengths listed for each filter use
+the definition
+\begin{equation}
+\label{lden}
+\lambda_{\mathrm{eff}} = \exp\left[ 
+\frac{\int d(\ln\lambda) R(\lambda) \ln\lambda}
+{\int d(\ln\lambda) R(\lambda)} \right],
+\end{equation}
+following \citet{fukugita96a} and \citet{schneider83a}.  For the
+spectrum of Vega in the conversion of AB to Vega magnitudes, we use
+the \citet{kurucz91a} theoretical Vega spectrum (normalized at 5000
+\AA\ to match the \citet{hayes85a} spectrophotometry of Vega). In {\tt
+k\_vega2ab}, the user has the option to use the spectrum of
+\citet{hayes85a} instead.  We use the solar spectrum of
+\citet{kurucz91a} for the calculation of the solar absolute magnitude.
+
+For comparison among results from different surveys, one also wants to
+be able to convert the magnitudes in one set of bandpasses to
+magnitudes in another set of bandpasses. An example of doing so is
+given in our piece of code {\tt sdss2bessell}, which takes SDSS input
+magnitudes and outputs absolute magnitudes in the Bessell $U$, $B$,
+$V$, $R$ and $I$ bandpasses.
+
+However, often one wants to just make quick and dirty comparisons,
+without going through the trouble of running any software. For these
+purposes, we provide Table \ref{lineareqs}, which provides the linear
+relationships among the same bandpasses as those listed in Table
+\ref{solarmagnitudes}. Note that Table \ref{lineareqs} refers to AB
+magnitudes {\it throughout}; use Table \ref{solarmagnitudes} to get
+the relationships to Vega magnitudes.
+
+[ sdss2bessell ]
 
@@
 
-%\include{tables}
+\include{kcorrect_tables}
 

### 2005-08-30 docs/paper/kcorrect_figures.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@499 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \figurenum{\fignum}
+\plotone{mass_to_garching.ps}
+\caption{\label{mass_to_garching} Our galaxy stellar mass estimates
+	$M_\ast$ compared to those of \citet{kauffmann03a} $M_{s,\ast}$, as
+	a function of stellar mass (top panel) and of color (bottom
+	panel). The greyscale is the conditional distribution
+	$M_{s,\ast}/M_\ast$ on each quantity . The lines are the 10\%, 25\%,
+	50\%, 75\% and 90\% quantiles.}
+\end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
 \plotone{mtol.ps}
-\caption{\label{mtol} Mass-to-light ratios of galaxies (in solar
-units) as a function of galaxy color. The solid line satisfies the
-relationship $\log_{10}(M/L) = 1.65 (g-r) - 0.54$, which has 
-the same slope as that found by \citet{bell01a}, with a higher
-normalization by 0.3 dex.}
+\caption{\label{mtol} Mass-to-light ratios of galaxies in the $V$ band
+	(in solar units) as a function of galaxy $B-V$ color. The solid line
+	satisfies the relationship $\log_{10}(M/L_V) = 1.40 (g-r) - 0.73$,
+	given by \citet{bell01a} for their sample of spiral galaxies. 
+	Their estimates and ours agree for $B-V < 0.8$, where spiral
+	galaxies dominate the galaxy population.}
 \end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{umr_bg.ps}
+\caption{\label{umr_bg} Fraction of the total star-formation that has
+	occurred in the previous billion years $b_G$, as a function of restframe
+	$u-r$ color, for SDSS galaxies. The greyscale is the conditional
+	distribution of $b_G$ on $u-r$. The lines are the 10\%, 25\%, 50\%,
+	75\% and 90\% quantiles. For $u-r < 2.5$, the median relationship
+	follows the simple form listed in the figure. Galaxies with $u-r >
+	2.5$ are often highly reddened star-forming galaxies.}
+\end{figure}

### 2005-08-30 docs/paper/kcorrect_tables.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@499 75dff457-8751-0410-ab1a-cd701654ce6b

\begin{deluxetable}{rrrrr}
\tablewidth{0pt}
\tablecolumns{5}
\tablecaption{\label{solarmagnitudes} Properties of various filters}
\tablehead{ Band & $\lambda_{\mathrm{eff}}$ (\AA) & $m_{\mathrm{AB}} -
	m_{\mathrm{Vega}}$ & $M_\odot$ (AB) &
	$M_\odot$ (Vega)}
\startdata
\input{solarmagnitudes.tex}
\enddata
\tablecomments{Uses model solar spectrum and model Vega spectrum of
	\citet{kurucz91a}. Effective wavelength is defined in the text.}
\end{deluxetable}

\begin{deluxetable}{cc}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
\input{lineareqs.tex}
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}

### 2005-09-07 docs/paper/kcorrect.tex
update docs and paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@504 75dff457-8751-0410-ab1a-cd701654ce6b
update paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@520 75dff457-8751-0410-ab1a-cd701654ce6b
paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@523 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 Michael R. Blanton\altaffilmark{\ref{NYU}} and 
-Samuel Roweis
+Sam Roweis
 %Tamas Budavari\altaffilmark{\ref{JHU}},
@@
 in a low-dimensional locus. Principal Component Analysis (PCA) of
-galaxy spectra, introduced by \citet{connolly94a} and applied many
-times since then ({\bf more refs}; \citealt{yip05a}), demonstrates
-that most of the variance in the distribution of galaxies in spectral
-space can be explained using a few templates. This means that, in the
-linear space of all possible spectra, galaxies exist in only a small
-locus. Therefore, even with very heterogeneous data, we should be able
-to determine the properties of this locus.
+galaxy spectra, introduced by \citet{connolly95b} and applied many
+times since then, demonstrates that most of the variance in the
+distribution of galaxies in spectral space can be explained using a
+few templates. This means that, in the linear space of all possible
+spectra, galaxies exist in only a small locus. Therefore, even with
+very heterogeneous data, we should be able to determine the properties
+of this locus.
 
@@
 \subsection{The models}
+\label{models}
 
@@
 above ($B$ in DEEP2 is different than the $B$ used by GOODS). The {\tt
-kcorrect} product described in Section \ref{code} contains a
+kcorrect} product described in Section \ref{summary} contains a
 tabulation of the response functions for all of these filters.
@@
 redshift, with the best fit template color overplotted as the smooth,
-thick line. In the code described in Section \ref{code}, we used the
+thick line. In the code described in Section \ref{summary}, we used the
 routine {\tt sdss\_kcorrect} (with the optional flag {\tt /lrg}) to
@@
 \end{equation}
-In the code described in Section \ref{code}, we used the routine {\tt
+In the code described in Section \ref{summary}, we used the routine {\tt
 gst\_kcorrect} to perform these fits.  In Figure \ref{fullfits} the
@@
 \label{eq:specialL}
-K_{QR} = -2.5\,\log_{10}\left[\frac{1}{[1+z]}\,
+K_R(z) = -2.5\,\log_{10}\left[\frac{1}{[1+z]}\,
   \frac{\displaystyle
@@
 did, \citet{kauffmann03a} used the library of \citet{bruzual03a} with
-the \citet{chabrier03} initial mass function, which makes some of this
+the \citet{chabrier03a} initial mass function, which makes some of this
 agreement unsurprising.  To be explicit, the masses in Figure
@@
 give the mass-to-light ratio in solar units. The solid line is the
-relationship that \citet{bell01a} give based on their fits to spiral
+relationship that \citet{bell01b} give based on their fits to spiral
 galaxies. For blue galaxies ($B-V < 0.8$) this line is indeed a good
-fit to our results. For the typical red galaxy, the \citet{bell01a}
+fit to our results. For the typical red galaxy, the \citet{bell01b}
 relation predicts a larger stellar mass than we do.
@@
 \begin{equation}
-\log_{10} b_g = -0.55 (u-r) - 0.25.
+\log_{10} b_g = -0.55 (u-r). 
 \end{equation}
@@
 relationships among the same bandpasses as those listed in Table
-\ref{solarmagnitudes}. Note that Table \ref{lineareqs} refers to AB
+\ref{solarmagnitudes}. In fact, these relationships are usually good
+to 0.05 mag or better. Note that Table \ref{lineareqs} refers to AB
 magnitudes {\it throughout}; use Table \ref{solarmagnitudes} to get
@@
 
-[ sdss2bessell ]
-
-[ conversions ]
-
-\section{Public access to the code}
-\label{code}
-
 \section{Summary}
-\label{conclusions}
-
-[ summarize ]
-
-[ this does not describe what constraints photometry puts on the SFH ]
+\label{summary}
+
+We have here described a method for fitting templates to
+nearly-arbitrary sets of spectra and broad-band fluxes. The basic
+concept is that we can recover a small set of templates (based on
+models), nonnegative linear combinations of which can explain a much
+larger set of inhomogeneous observations.  We have applied this method
+to SDSS, GALEX, 2MASS, DEEP2, and GOODS data to come up with a set of
+templates which are effective for describing all of these data sets.
+
+Nonnegative matrix factorization (NMF) is an effective way to reduce
+the dimensionality of any large data set, and has that in common with
+PCA.  It is, in some sense, a ``nonnegative PCA.''  However, at least
+in the current context the method has several advantages over
+PCA. First, it has a natural physical interpretation associated with
+that spectral subspace, which is the corresponding subspace of all
+possible star-formation histories. Second, it naturally handles data
+uncertainties and missing data, which allows it to ignore that
+variation which is due purely to statistical errors.  Third, it
+handles the complications of observing galaxy spectra photometrically
+using broad-band filters.
+
+We note here that the general NMF method does not {\tt depend} on
+using a model --- our ``model'' could just have been a set of top hat
+functions or Gaussians on a wavelength grid. In some situations, such
+as datasets for which there are not well-developed theoretical models,
+this approach could be more appropriate.
+
+We have released our results in the form of a templates and a code
+base called {\tt kcorrect} which fits those templates to many types of
+data (SDSS photometric and spectroscopic data, GALEX data, GOODS data,
+DEEP2 data, and 2MASS data). Furthermore, this code returns
+$K$-corrections and a physical interpretation of the photometry. All
+of the plots in this paper were created using code in the
+repository. The code is available from a web page maintained by one of
+the authors\footnote{{\tt
+http://cosmo.nyu.edu/blanton/kcorrect}}. This web page is kept updated
+on improvements in the code and new developments. It consists of a
+{\tt C} language library, stand-alone {\tt C} programs, and IDL
+language wrappers around the {\tt C} library. Thus, one can use the
+basic templates and fitting code using only a machine with a {\tt C}
+compiler. However, there is significant functionality which is
+programmed in the (unfortunately, proprietary) IDL language. These
+routines depend on the {\tt idlutils} library\footnote{{\tt
+http://skymaps.info}}.
 
 \acknowledgments
+
+The authors would like to thank Alison Coil, David W. Hogg, David
+Schiminovich,
 
@@
 
+[sam sam sam sam]
+
 \section{Format for templates}
@@
 
-\section{$K$-corrections to shifted bandpasses}
-\label{bandshift}
+We have fit several different sets of templates that we release
+with the code, which we denote:
+\begin{enumerate}
+\item {\tt default}, the default set of five templates, 
+\item {\tt lrg1}, the single template fit to LRGs, and 
+\item {\tt goods}, the five template fit to just the GOODS data.
+\end{enumerate}
+The information about the default set of templates is contained in the
+file {\tt data/templates/k\_nmf\_derived.default.fits} in the {\tt
+kcorrect} project.  This file has 25 Header and Data Units (HDUs),
+each listed in Table \ref{derived} and described in more detail in the
+paragraphs below.  There are similar files for the {\tt lrg1} and {\tt
+goods} template sets.
+
+As described in Section \ref{models}, there are $N_{\mathrm{basis}} =
+485$ basis spectra, consisting of 450 different instantaneous burst
+stellar populations ($N_{\mathrm{age}} = 25$ ages, $N_{\mathrm{mets}}
+= 6$ metallicities, and $N_{\mathrm{dust}}=3$ dust properties) plus 35
+different emission line models. The {\tt templates} HDU has the
+coefficients of the five templates in this basis space.
+
+For each of the five templates, we have the template spectrum {\tt
+spec} in units of ergs s$^{-1}$ cm$^{-2}$ \AA$^{-1}$ per solar mass,
+as it would be observed at 10 pc distance. The wavelength grid is in
+the {\tt lambda} HDU. As noted above, the models are smoothed at 300
+km s$^{-1}$ resolution; the {\tt rawspec} HDUs have the original
+(unsmoothed) models from \citet{bruzual03a}. In addition, we give
+versions without the emission lines (with the {\tt \_nl} suffix) and
+without the dust extinction applied (with the {\tt \_nd} suffix). The
+actual emission-line-only spectra for each template are in the {\tt
+lspec} HDU. Finally, the multiplicative dust extinction factor for
+each template spectrum is given in the {\tt extinction} HDU.
+
+We include the star-formation rate as a function of time in the {\tt
+sfr} HDU (the age grid used is in the {\tt ages} HDU). The time
+differential used to quantify this rate is in the {\tt dage} HDU; to
+get the total number of stars formed at each age, multiply {\tt sfr}
+by {\tt dage}. The average metallicity of the stars formed as a
+function of time is in the {\tt metallicity} HDU.
+
+The total mass formed in each template (in solar masses) is in the
+{\tt mass} HDU --- this is actually just unity for each template. The
+total surviving stellar mass for each template is in the {\tt mremain}
+HDU. The metallicity in the surviving stars is given in the {\tt mets}
+HDU.  The total mass in stars formed in the last 300 Myrs is in the
+{\tt m300} HDU, and the total mass formed in the last 1 Gyr is in the
+{\tt m1000} HDU.
+
+The properties of the 450 stellar population basis vectors are given
+in the four HDU: {\tt basis\_ages}, {\tt basis\_mets}, {\tt
+basis\_dusts}, and {\tt basis\_mremain}. Respectively, these give the
+ages, metallicities, dust properties, and fraction of original stellar
+mass remaining for each instantaneous burst in the grid. The dust 
+properties are given in terms of a structure with four elements which
+refer to the properties in the multiple scattering model of \citet{witt00a}:
+\begin{enumerate}
+\item {\tt GEOMETRY}: the geometry of the dust (e.g. ``shell'') 
+\item {\tt DUST}: the dust extinction curve (e.g. ``MW'' for Milky
+	Way extinction and ``SMC'' for Small Magellanic Cloud type
+	extinction) 
+\item {\tt STRUCTURE}: structure of the dust distribution (``h'' for
+	homogeneous, ``c'' for clumpy)
+\item {\tt TAUV}: the amount of dust (the total optical depth in the
+	$V$ band)
+\end{enumerate}
 

### 2005-09-07 docs/paper/kcorrect_figures.tex
paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@523 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 	(normalization is for a 1 $M_\odot$ galaxy located 10 pc away, or
-	equivalently a $10^12$ $M_\odot$ galaxy located 10 Mpc away).}
+	equivalently a $10^{12}$ $M_\odot$ galaxy located 10 Mpc away).}
 \end{figure}
@@
 \figurenum{\fignum}
-\plotone{sdss_kcorrect.ps}
-\caption{\label{sdss_kcorrect} $K$-corrections as a function of
-redshift in the SDSS $u$, $g$, $r$, $i$, and $z$ bands.}
-\end{figure}
-
-\clearpage
-\stepcounter{thefigs}
-\begin{figure}
-\figurenum{\fignum}
 \plotone{main_kcorrect.ps}
@@
 	satisfies the relationship $\log_{10}(M/L_V) = 1.40 (g-r) - 0.73$,
-	given by \citet{bell01a} for their sample of spiral galaxies. 
+	given by \citet{bell01b} for their sample of spiral galaxies. 
 	Their estimates and ours agree for $B-V < 0.8$, where spiral

### 2005-09-07 docs/paper/kcorrect_tables.tex
update paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@520 75dff457-8751-0410-ab1a-cd701654ce6b
paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@523 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 
-\begin{deluxetable}{cc}
+\begin{deluxetable}{lr}
 \tablewidth{0pt}
@@
 \end{deluxetable}
+
+\begin{deluxetable}{llcl}
+\tablewidth{0pt}
+\tablecolumns{4}
+\tablecaption{\label{derived} HDU of {\tt k\_nmf\_derived.default.fits}}
+\tablehead{ Number & Name & Dimensions & Description}
+\startdata
+0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
+in basis space for each template \cr
+1 & {\tt spec} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template \cr
+2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template, lines removed \cr
+3 & {\tt spec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template, without dust \cr
+4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template, without lines or dust \cr
+5 & {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
+each template \cr
+6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
+each template, lines removed \cr
+7 & {\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
+each template, without dust \cr
+8 & {\tt rawspec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ &
+Unsmoothed spectrum of each template, without lines or dust \cr
+9 & {\tt lspec} & $N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr
+10 & {\tt extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as
+a function of wavelength \cr
+11 & {\tt lambda} & $N_{\mathrm{spec}}$ & Wavelength grid (\AA) \cr
+12 & {\tt sfr} & $N_{\mathrm{age}} \times N_t$ & Star-formation rate
+($M_\odot$ per year) as a function of time \cr
+13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
+metallicity as a function of time \cr
+14 & {\tt ages} & $N_{\mathrm{age}}$ & Age grid (yrs) \cr
+15 & {\tt dage} & $N_{\mathrm{age}}$ & Age bin size (yrs);
+multiply {\tt sfr} for total star-formation in each bin \cr
+16 & {\tt mass} & $N_t$ & total mass formed in each template \cr
+17 & {\tt mremain} & $N_t$ & total current stellar mass in each template \cr
+18 & {\tt mets} & $N_t$ & average metallicity of current stars in each
+template \cr
+19 & {\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in each template \cr
+20 & {\tt m1000} & $N_t$ & total stellar mass formed in last 1 Gyr in each template \cr
+21 & {\tt basis\_ages} &
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+ages of each stellar population basis vector  \cr
+22 & {\tt basis\_mets} & 
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+metallicity of each stellar population basis vector  \cr
+23 & {\tt basis\_dusts} & 
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+dust properties of each stellar population basis vector \cr
+24 & {\tt basis\_mremain} & 
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+fraction of original stellar mass surviving for age of each basis vector  \cr
+\enddata 
+\tablecomments{All quantities are in floating point (four byte)
+precision, except {\tt dusts} which is an array of structures, whose
+format is described in text.}
+\end{deluxetable}

### 2005-09-15 docs/paper/kcorrect_tables.tex
redit
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@524 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2005-08-30 docs/paper/kcorrect_tables.tex
@@
 \end{deluxetable}
+
+\begin{deluxetable}{llcl}
+\tablewidth{0pt}
+\tablecolumns{4}
+\tablecaption{\label{derived} HDU of {\tt k\_nmf\_derived.default.fits}}
+\tablehead{ Number & Name & Dimensions & Description}
+\startdata
+0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
+in basis space for each template \cr
+1 & {\tt spec} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template \cr
+2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template, lines removed \cr
+3 & {\tt spec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template, without dust \cr
+4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
+each template, without lines or dust \cr
+5 & {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
+each template \cr
+6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
+each template, lines removed \cr
+7 & {\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
+each template, without dust \cr
+8 & {\tt rawspec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ &
+Unsmoothed spectrum of each template, without lines or dust \cr
+9 & {\tt lspec} & $N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr
+10 & {\tt extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as
+a function of wavelength \cr
+11 & {\tt lambda} & $N_{\mathrm{spec}}$ & Wavelength grid (\AA) \cr
+12 & {\tt sfr} & $N_{\mathrm{age}} \times N_t$ & Star-formation rate
+($M_\odot$ per year) as a function of time \cr
+13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
+metallicity as a function of time \cr
+14 & {\tt ages} & $N_{\mathrm{age}}$ & Age grid (yrs) \cr
+15 & {\tt dage} & $N_{\mathrm{age}}$ & Age bin size (yrs);
+{\tt sfr}$\times${\tt dage} $=$ total star-formation in each bin \cr
+16 & {\tt mass} & $N_t$ & total mass formed in each template \cr
+17 & {\tt mremain} & $N_t$ & total current stellar mass in each template \cr
+18 & {\tt mets} & $N_t$ & average metallicity of current stars in each
+template \cr
+19 & {\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in each template \cr
+20 & {\tt m1000} & $N_t$ & total stellar mass formed in last 1 Gyr in each template \cr
+21 & {\tt basis\_ages} &
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+ages of each stellar population basis vector  \cr
+22 & {\tt basis\_mets} & 
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+metallicity of each stellar population basis vector  \cr
+23 & {\tt basis\_dusts} & 
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+dust properties of each stellar population basis vector \cr
+24 & {\tt basis\_mremain} & 
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+fraction of original stellar mass surviving for each basis vector  \cr
+\enddata 
+\tablecomments{All quantities are in floating point (four byte)
+precision, except {\tt dusts} which is an array of structures, whose
+format is described in text.}
+\end{deluxetable}

### 2006-02-16 docs/paper/kcorrect.tex
redit
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@524 75dff457-8751-0410-ab1a-cd701654ce6b
update paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@541 75dff457-8751-0410-ab1a-cd701654ce6b
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@549 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 wavelengths. The implementation we present here is suitable for
-ultraviolet, optical, and near-infrared observations. Since we base
-the templates on stellar population synthesis results, the results are
-intepretable in terms of approximate stellar masses and star-formation
-histories. We present templates fit with this method to data from
-GALEX, Sloan Digital Sky Survey spectroscopy and photometry, the
-Two-Micron All Sky Survey, the Deep Extragalactic Evolutionary Probe
-and Great Observatories Origins Deep Survey.  In addition, we present
-software for using such data to estimate $K$-corrections, evolution
-corrections, stellar masses, and recent star-formation rates. This
-software is suitable for analyzing many other data sets as well.
+ultraviolet, optical, and near-infrared observations in the redshift
+range $0 < z < 1.5$. Since we base the templates on stellar population
+synthesis results, the results are intepretable in terms of
+approximate stellar masses and star-formation histories. We present
+templates fit with this method to data from GALEX, Sloan Digital Sky
+Survey spectroscopy and photometry, the Two-Micron All Sky Survey, the
+Deep Extragalactic Evolutionary Probe and Great Observatories Origins
+Deep Survey.  In addition, we present software for using such data to
+estimate $K$-corrections, stellar masses, and recent star-formation
+rates. This software is suitable for analyzing many other data sets as
+well.
 \end{abstract}
@@
 
-New surveys at a low and high redshift have provided us with estimates
+New surveys at low and high redshift have provided us with estimates
 of galaxy spectral energy distributions (SEDs) for an enormous number
@@
 observed 10 pc away, in bandpass $Q$, at rest.  The traditional
-definition of the $K$-correction takes $Q=R$. however, we note that in
+definition of the $K$-correction takes $Q=R$. However, we note that in
 practice many surveys do perform $K$-corrections from one observed
@@
 Digital Sky Survey (SDSS; \citealt{york00a}) the Galaxy Evolution
-Explorer (GALEX; {\bf ref})) in the ultraviolet, and the Two-Micron
-All Sky Survey (2MASS; \citealt{skrutskie97a}) in the near infrared
-(NIR). In addition, at higher redshifts, we use constraints from the
-Deep Extragalactic Evolutionary Probe 2 (DEEP2; \citealt{davis03a,
-faber03a}) and the Great Observatories Origins Deep Survey (GOODS;
+Explorer (GALEX; \citealt{martin05a}) in the ultraviolet, and the
+Two-Micron All Sky Survey (2MASS; \citealt{skrutskie97a}) in the near
+infrared (NIR). In addition, at higher redshifts, we use constraints
+from the Deep Extragalactic Evolutionary Probe 2 (DEEP2;
+\citealt{davis03a, faber03a}) and the Great Observatories Origins Deep
+Survey (GOODS;
 \citealt{giavalisco04a}).  These and other data sets provide a huge set of
@@
 high-level interface.  This software is an update of two major earlier
-releases ({\tt v1\_16}, \citealt{blanton03b}; {\tt v3\_2}). The
+releases ({\tt v1\_16}; \citealt{blanton03b}; {\tt v3\_2}). The
 improvement over the latest version is twofold. First, we have
@@
 results: the best-fit templates and how well they fit the data.
+
+Where necessary, we have assumed cosmological parameters $\Omega_0 =
+0.3$, $\Omega_\Lambda = 0.7$, and $H_0 = 100$ $h$ km s$^{-1}$
+Mpc$^{-1}$ (with $h=1$). All magnitudes are (unless otherwise noted)
+AB-relative.  Except where noted, I will always be referring to the
+version of the {\tt kcorrect} software labeled {\tt
+v4\_1\_4}.
 
@@
 from the high resolution stellar population synthesis model of
-\citet{bruzual03a} and the emission line models of {\bf mappings
-ref}.  This approach both constrains the problem appropriately and
-yields a natural theoretical interpretation of the results in terms of
+\citet{bruzual03a} and the nebular emission line models of \citet{kewley01a}.
+This approach both constrains the problem appropriately and yields a
+natural theoretical interpretation of the results in terms of
 star-formation histories.
@@
 histories. The technical name for this algorithm is Nonnegative Matrix
-Factorization (NMF).
+Factorization (NMF). Note that this problem is not the same as the
+well-known nonnegative least squares problem, which would fit for the
+best-fit nonnegative linear combination of templates, but not for the
+form of the templates themselves.
 
@@
 statistical errors.  Third, it handles the complications of observing
-galaxy spectra photometrically using broad-band filters.
-
-In this section, we describe how to take the astrophysical
-models and the astronomical observations, and to pose the problem in
-the form needed for the purposes of NMF. In Appendix \ref{nmf} we describe
+galaxy spectra photometrically using broad-band filters of galaxies at
+varying redshifts, whereas standard PCA requires a constant grid of
+observations in rest-frame wavelength space.
+
+In this section, we describe how to take the astrophysical models and
+the astronomical observations, and to pose the problem in the form
+needed for the purposes of NMF. Section \ref{models} describes the
+basis set of models, Section \ref{In Appendix \ref{nmf} we describe
 how to solve to NMF problem.
@@
 We begin with a basis set of 485 spectral templates. Of these, 450 are
-a set of instantaneous bursts from \citet{bruzual03a}. There are 6
-metallicities (0.005, 0.02, 0.2, 0.4, 1, and 2.5 times solar). For
-each metallicity there are 25 ages (between 1 Myr and 13.75 Gyr,
-spaced almost logarithmically in age).  For each age and metallicity
-there are 3 choices of dust model: (1) no dust extinction; (2) $\tau_V
-= 3$ dust with Milky Way type extinction (as input into the models of
+a set of instantaneous bursts from \citet{bruzual03a}, using the
+\citet{chabrier03a} stellar initial mass function and the Padova1994
+isochrones. We select 6 metallicities (0.005, 0.02, 0.2, 0.4, 1, and
+2.5 times solar). For each metallicity we select 25 ages (between 1
+Myr and 13.75 Gyr, spaced almost logarithmically in age).  For each
+age and metallicity we make 3 choices of dust model: (1) no dust
+extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction (as
+input into the models of
 \citealt{witt00a} with a ``homogeneous'' distribution and ``shell''
@@
 same models). We smooth each template to $300$ km s$^{-1}$ velocity
-dispersion.
-
-The remaining 35 templates are from MAPPINGS-III {\bf ref} models of
-emission from ionized gas. We choose the predictions for an 8 Myr old
-continuous star-formation history with 5 possible metalliciticites
-(0.5, 0.2, 0.4, 1.0, and 2. times solar) and 7 possible ionization
-parameters ($q= 5\times 10^6$, $10^7$, $2\times 10^7$, $4\times 10^7$,
-$8 \times 10^7$, $1.5\times 10^8$, and $3\times 10^8$ cm s$^{-1}$). We
-take the spectra (given as a set of delta functions) and smooth them
-to $300$ km s$^{-1}$ velocity dispersion.
+dispersion (we will smooth the SDSS spectra in the training set to the
+same resolution).
+
+The remaining 35 templates are from MAPPINGS-III (\citealt{kewley01a})
+models of emission from ionized gas. We choose the predictions for an
+8 Myr old continuous star-formation history with 5 possible
+metalliciticites (0.5, 0.2, 0.4, 1.0, and 2. times solar) and 7
+possible ionization parameters ($q= 5\times 10^6$, $10^7$, $2\times
+10^7$, $4\times 10^7$, $8 \times 10^7$, $1.5\times 10^8$, and $3\times
+10^8$ cm s$^{-1}$). We take the spectra (given as a set of delta
+functions) and smooth them to $300$ km s$^{-1}$ velocity
+dispersion. The one alteration we make to all of these templates is to
+remove Ly$\alpha$, because it is generally much larger than observed
+in real galaxies.
 
@@
 
-We seek five templates $F_{\lambda, i}(\lambda)$ which are built from
-nonnegative combinations of the original basis set of $N$ templates:
+We seek to reduce this full basis space to a subspace in which
+galaxies actually exist. In particular, we seek five templates
+$F_{\lambda, i}(\lambda)$ built from nonnegative combinations of the
+original basis set of $N$ templates:
 \begin{equation}
@@
 \end{equation}
-in units of ergs s$^{-1}$ \AA$^{-1}$.  For each galaxy $k$
-we want our model ${\hat{F}_\lambda}(\lambda)$ for their spectrum to
-be a nonegative sum of these five templates:
+in units of ergs s$^{-1}$ \AA$^{-1}$. In principle, we could seek any
+number of templates; from our experiments, we have found that five
+turns out to be a number large enough to explain the data we use
+here. For each galaxy $k$ we want our model
+${\hat{F}_\lambda}(\lambda)$ for their spectrum to be a nonegative sum
+of these five templates:
 \begin{equation}
@@
 photometry only) and Main sample galaxies (using the full $ugriz$
-photometry). We use 2000 LRGs and 7000 Main sample galaxies.  For
-these galaxies we include 2MASS $JHK_s$ data ({\bf ref}) where
-available.
+photometry) in the same redshift ranges. We use 2000 LRGs and 7000
+Main sample galaxies.  For these galaxies we include 2MASS $JHK_s$
+extended source catalog data (\citealt{jarrett00a}) where available.
 \item GALEX DR1 far UV ($\sim 1500$ \AA) and near UV ($\sim 2300$ \AA)
-photometry for objects with SDSS redshifts and $ugriz$ photometry (4000
-galaxies; {\bf ref}).
+photometry for Main sample SDSS galaxies with redshifts and $ugriz$
+photometry (4000 galaxies; \citealt{martin05a}).
 \item The $BRI$ photometry for high redshift galaxies in the DEEP2 
@@
 idlspec2d} (in this case {\tt v4\_9}) measures the redshifts and
-classifies the reduced spectra (Schlegel et al., in preparation).
+classifies the reduced spectra (Schlegel et al., in preparation). We
+use the Petrosian magnitudes for the SDSS data (except where noted
+below for the LRG-only sample).
+
+For the GALEX and GOODS data we use ``auto'' magnitudes, which are the
+Kron-like magnitudes from SExtractor (\citealt{bertin96a}).  For 2MASS
+we use the ``extrapolated'' magnitudes from the Extended Source
+Catalog (\citealt{jarrett00a}).
 
@@
 spectrum by such an amount that, given its estimated velocity
-dispersion, its total velocity dispserion after smoothing is 300 km
-s$^{-1}$.
+dispersion, its total velocity dispersion after smoothing is 300 km
+s$^{-1}$ (there are a small number of galaxies with larger velocity
+dispersions, which we leave unchanged).
 
@@
 
-$\eta_{jp}$ and $M_{jl}$ are totally fixed, and we combine them into a
-single matrix $M_{jn}$ and the indices $p$ and $l$ into a single index
-$n$. We can similarly combine our observations $\mu_{kp}$ and
-$F_{\lambda, k}(\lambda_l)$ and their uncertainties $\sigma_{kp}$ and 
-$\sigma_{k}(\lambda_l)$ into vectors $x_{kn}$ and $\sigma_{kn}$.
-Then we can combine the Equations \ref{chi2spec} and \ref{chi2photo}
-into a single equation:
+The matrices $\eta_{jp}$ (related to the broad-band fluxes) and
+$M_{jl}$ (related to the spectra) are totally fixed, and we combine
+them into a single matrix $M_{jn}$ and the indices $p$ and $l$ into a
+single index $n$, to handle both the broad-band fluxes and spectra
+simultaneously. We can similarly combine our observations $\mu_{kp}$
+and $F_{\lambda, k}(\lambda_l)$ and their uncertainties $\sigma_{kp}$
+and $\sigma_{k}(\lambda_l)$ into vectors $x_{kn}$ and $\sigma_{kn}$.
+Then we can combine the Equations \ref{chi2spec} and
+\ref{chi2photo} into a single equation:
 \begin{equation}
@@
 
+{\bf nonconvexity; not reproducible}
+
 Once we have fit for $b_{ij}$ using the training set, we can minimize
@@
 this best-fit LRG spectrum. The top panel shows the star-formation per
-unit time as a function of look-back time.  The thick, straight line
-is the power law:
-\begin{equation}
-\mathrm{SFR} = \left(\frac{t}{t_0}\right)^2,
-\end{equation}
-where $t$ is the look-back time, for $t<13.75$ $h^{-1}$ Gyr, and $t_0=$
-{\bf number}. The bottom panel shows the mean metallicity of the
-forming stars as a function of time.
+unit time as a function of look-back time.  The bottom panel shows the
+mean metallicity of the forming stars as a function of time.  When
+considering these plots and those below, do remember that although
+these results are our best-fit, the parameters are highly degenerate
+(especially in, for example, neighboring age bins).
 
@@
 templates. We have not imposed any smoothness criterion on these fits,
-which explains the ragged appearance of these histories. We do not
-believe the details of these fits, since neighboring ages or
-metallicities are highly degenerate with one another.
+which explains the ragged appearance of these histories.  The details
+of these fits are weakly constrained due to degeneracies within and
+among the templates, but we show them here for completeness.
 
@@
 $b_{ij}$ to the five templates shown in this section and fit only for
-$a_{ki}$. In general below, the tests we perform are not to the
-training set but to independent sets of galaxies.
+$a_{ki}$. Since we have determined these templates to be appropriate
+for the data sets we include in the training set, they end up
+providing good fits to most other galaxies from the parent data sets.
+In general below, the tests we perform are not to the training set but
+to independent sets of galaxies.
 
@@
 significant trends between redshifts $z=0.5$ and $1.5$ (echoed in
-Figure \ref{goods}. More significantly, in the $H$ band there is at
+Figure \ref{goods}). More significantly, in the $H$ band there is at
 least a 10\% offset in the magnitudes across all redshifts, which is
-almost certainly a catalog error.  {\bf GOODS people: this is your
-challenge! what is going on?} Other errors (such as a large scatter in
-the residuals in the $B$ band and some redshift dependence in the $J$
-band) are clearly due to the fact that the templates are not primarily
-designed for GOODS data. All in all, we recommend using the special
-templates for $K$-corrections within the GOODS data set.
+almost certainly a catalog error. Other errors (such as a large
+scatter in the residuals in the $B$ band and some redshift dependence
+in the $J$ band) are clearly due to the fact that the templates are
+not primarily designed for GOODS data. All in all, we recommend using
+the special templates for $K$-corrections within the GOODS data set.
 
@@
 fluxes.  Figure \ref{specfit} compares the reconstructed model spectra
-to the original spectrum. The two agree extremely well. Naturally,
-this success depends heavily on the quality of the original stellar
-spectra used in the population synthesis code of \citet{bruzual03a}.
+to the original spectrum. The two agree well in general, though some
+features (like the emission lines) are not reproduced well. 
+Naturally, this success is mostly due to the quality of the original
+stellar spectra used in the population synthesis code of
+\citet{bruzual03a}. These results encourage us that by using
+broad-band data we can infer $K$-corrections of galaxies.
 
@@
 at all. Without any input from 2MASS the templates do a very good job
-of predicting the 2MASS fluxes, with a scatter of 20--30\% {\bf real
-number} --- not far from the uncertainties in the 2MASS fluxes
-themselves. Of course, it is not surprising that it is easy to predict
-the 2MASS fluxes --- stellar spectra are very simple in the NIR, so
-having the $ugriz$-band fluxes from SDSS yields much information
-regarding the redder bandpasses.
+of predicting the 2MASS fluxes, with a scatter of 20--30\% --- not far
+from the uncertainties in the 2MASS fluxes themselves. Of course, it
+is not surprising that it is easy to predict the 2MASS fluxes ---
+stellar spectra are very simple in the NIR, so having the $ugriz$-band
+fluxes from SDSS yields much information regarding the redder
+bandpasses.
 
@@
 significantly at the lowest luminosities and surface brightnesses
-relative to the SDSS.
+relative to the SDSS. Of course, if at high redshift galaxies have
+a very different locus of their SEDs, then perhaps this statement is
+less true for them.
 
@@
 bandpass $Q$, at rest.  The traditional definition of the
-$K$-correction takes $Q=R$. however, we note that in practice many
+$K$-correction takes $Q=R$. However, we note that in practice many
 surveys do perform $K$-corrections from one observed bandpass $R$ to
@@
 in the case of SDSS galaxies it makes sense to correct to the
-\band{0.3}{r} bandpass --- the $r$ band blue-shifted to $z=0.1$. This
+\band{0.1}{r} bandpass --- the $r$ band blue-shifted to $z=0.1$. This
 bandpass has the property that the $K$-correction is independent of
@@
 two sets of masses are very similar to each other, with a scatter of
-only 0.1 dex and trends of less than 0.2 dex with stellar mass. As we
-did, \citet{kauffmann03a} used the library of \citet{bruzual03a} with
-the \citet{chabrier03a} initial mass function, which makes some of this
-agreement unsurprising.  To be explicit, the masses in Figure
-\ref{mass_to_garching} correspond to the current mass in stars
-remaining in the galaxy. This mass can differ from the total
-star-formation rate integrated over time, due to the fact that stars
-can die --- and in practice for our fits the total integrated
-star-formation rate is usually about twice as high as the total
-current mass in stars.
+only 0.1 dex and trends of less than 0.2 dex with stellar mass.  To be
+explicit, the masses in Figure \ref{mass_to_garching} correspond to
+the current mass in stars remaining in the galaxy. This mass can
+differ from the total star-formation rate integrated over time, due to
+the fact that stars can die --- and in practice for our fits the total
+integrated star-formation rate is usually about twice as high as the
+total current mass in stars.
 
@@
 \end{equation}
-Note that here $b_G$ is defined in terms of the total integrated
+where the definition here is meant to be similar to the birthrate
+parameter $b$ in \citet{kennicutt94a} and references therein.  Note
+that here $b_G$ is defined in terms of the total integrated
 star-formation history (unlike the stellar masses referred to above)
-as is standard for the $b$ parameter (\citealt{kennicutt98a}). Figure
-\ref{umr_bg} shows the relationship between the rest-frame $u-r$ color
-for SDSS galaxies and $b_G$. For galaxies with $u-r < 2.5$ (redder
-than which galaxies are mostly reddened, edge-on disks) these two 
-parameters follow a simple relationship:
+as is standard for the $b$ parameter. Figure \ref{umr_bg} shows the
+relationship between the rest-frame $u-r$ color for SDSS galaxies and
+$b_G$. For galaxies with $u-r < 2.5$ (redder than which galaxies are
+mostly reddened, edge-on disks) these two parameters follow a simple
+relationship:
 \begin{equation}
@@
 \end{equation}
+Note, however, the large scatter (about 0.3 dex at 1$\sigma$).
 
@@
 
-The authors would like to thank Alison Coil, David W. Hogg, David
-Schiminovich,
+The authors would like to thank Alison Coil, Michael Cooper, David
+W. Hogg, John Moustakas, Leonidas Moustakas, and David Schiminovich
+for their detailed comments and feedback. {\bf funding}
 

### 2006-02-16 docs/paper/kcorrect_figures.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@549 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \plotone{spec_lrg.ps}
-\caption{\label{spec_lrg} Best fit LRG spectral template
-	(normalization is for a 1 $M_\odot$ galaxy located 10 pc away, or
-	equivalently a $10^{12}$ $M_\odot$ galaxy located 10 Mpc away).}
+\caption{\label{spec_lrg} Best fit LRG spectral template in the rest
+frame (normalization is for a 1 $M_\odot$ galaxy located 10 pc away,
+or equivalently a $10^{12}$ $M_\odot$ galaxy located 10 Mpc away).}
 \end{figure}
@@
 	panel shows the mean metallicity of the population as a function of
-	time.  }
+	time. Note that the details of these functions are rather poorly
+	constrained. }
 \end{figure}
@@
 \plotone{sfh_templates.ps}
-\caption{\label{sfh_templates} SDSS colors of LRGs as a function of
-	redshift. The greyscale is the conditional distribution of color
-	within each redshift bin. The thin lines are the 10\%, 25\%, 50\%,
-	75\%, and 90\% quantiles of the distribution. The thick line
-	is the prediction of the model. The $u$ band is not included in the
-	fit, and the $u$ magnitudes of most LRGs are poorly known. The other
-	colors fit the models reasonably well. This model, remember, is
-	given incredible freedom, meaning that the above agreement is the
-	best one can do with the stellar population synthesis code of
-	\citet{bruzual03a}. }
+\caption{\label{sfh_templates} Similar to Figure \ref{sfh_lrg}, but
+for the five global templates. Again, there are many degeneracies in
+the fit parameters (though there are fewer in the actual spectra), so
+these figures need to be interpreted appropriately.}
 \end{figure}

### 2006-02-16 docs/paper/kcorrect_tables.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@549 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2005-08-30 docs/paper/kcorrect_tables.tex
@@
 \tablecomments{Uses model solar spectrum and model Vega spectrum of
-	\citet{kurucz91a}. Effective wavelength is defined in the text.}
+	\citet{kurucz91a}. Effective wavelength is defined in the text. The
+	$UBRVI$ filters are those of \citet{bessell90a}. The $ugriz$ filters
+	are those determined by Mamoru Doi, Daniel Eisenstein, and James
+	Gunn and available on the SDSS DR4 web site.\footnote{{\tt
+	http://www.sdss.org/dr4/}} The $JHK_s$ filters are those from
+	\citet{cohen03a}. }
 \end{deluxetable}
 
-\begin{deluxetable}{cc}
+\begin{deluxetable}{lr}
 \tablewidth{0pt}
@@
 \end{deluxetable}
+
+\begin{deluxetable}{llcl}
+\tablewidth{0pt}
+\tablecolumns{4}
+\tablecaption{\label{derived} Header and Data Units (HDUs) of {\tt k\_nmf\_derived.default.fits}}
+\tablehead{ Number & Name & Dimensions & Description}
+\startdata
+0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
+in basis space for each template \cr 1 & {\tt spec} &
+$N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of each template
+\cr 2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed
+spectrum of each template, lines removed \cr 3 & {\tt spec\_nd} &
+$N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of each template,
+without dust \cr 4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times
+N_t$ & Smoothed spectrum of each template, without lines or dust \cr 5
+& {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum
+of each template \cr 6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times
+N_t$ & Unsmoothed spectrum of each template, lines removed \cr 7 &
+{\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed
+spectrum of each template, without dust \cr 8 & {\tt rawspec\_nl\_nd}
+& $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of each
+template, without lines or dust \cr 9 & {\tt lspec} &
+$N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr 10 & {\tt
+extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as a
+function of wavelength \cr 11 & {\tt lambda} & $N_{\mathrm{spec}}$ &
+Wavelength grid (\AA) \cr 12 & {\tt sfr} & $N_{\mathrm{age}} \times
+N_t$ & Star-formation rate ($M_\odot$ per year) as a function of time
+\cr 13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
+metallicity as a function of time \cr 14 & {\tt ages} &
+$N_{\mathrm{age}}$ & Age grid (yrs) \cr 15 & {\tt dage} &
+$N_{\mathrm{age}}$ & Age bin size (yrs); {\tt sfr}$\times${\tt dage}
+$=$ total star-formation in each bin \cr 16 & {\tt mass} & $N_t$ &
+total mass formed in each template \cr 17 & {\tt mremain} & $N_t$ &
+total current stellar mass in each template \cr 18 & {\tt mets} &
+$N_t$ & average metallicity of current stars in each template \cr 19 &
+{\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in
+each template \cr 20 & {\tt m1000} & $N_t$ & total stellar mass formed
+in last 1 Gyr in each template \cr 21 & {\tt basis\_ages} &
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+ages of each stellar population basis vector \cr 22 & {\tt
+basis\_mets} & $N_{\mathrm{age}} \times N_{\mathrm{mets}} \times
+N_{\mathrm{dust}}$ & metallicity of each stellar population basis
+vector \cr 23 & {\tt basis\_dusts} & $N_{\mathrm{age}} \times
+N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ & dust properties of each
+stellar population basis vector \cr 24 & {\tt basis\_mremain} &
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+fraction of original stellar mass surviving for each basis vector \cr
+\enddata 
+\tablecomments{All quantities are in floating point (four byte)
+precision, except {\tt dusts} which is an array of structures, whose
+format is described in text.}
+\end{deluxetable}

### 2006-04-19 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@550 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \documentclass[10pt,preprint]{aastex}
+\usepackage{amsmath}
+\usepackage{epsfig}
 
@@
 \newcommand{\lambdaemit}{\lambda_e}
+
+\def\beq#1\eeq{\begin{equation}#1\end{equation}}
+\def\beqa#1\eeqa{\begin{align}#1\end{align}}
+\def\beqq#1\eeqq{\begin{equation*}#1\end{equation*}}
+\def\beqqa#1\eeqqa{\begin{align*}#1\end{align*}}
+\def\bseq#1\eseq{\begin{subequations}#1\end{subequations}}
+
+\newcommand{\T}{^{\scriptscriptstyle \top}}
+\newcommand{\iT}{^{\scriptscriptstyle -\top}}
+\newcommand{\inv}{^{-1}}
+
+\newcommand{\ie}{i.e.}
+
+\newcommand{\Log}{\mathrm{Log}}
+\newcommand{\trace}{\mathrm{Tr}}
+\newcommand{\detM}{\det(M)}
+
+\newcommand{\XX}{X}
+\newcommand{\XXt}{\tilde{X}}
+\newcommand{\XXh}{\hat{X}}
+\newcommand{\Xij}{X_{ij}}
+\newcommand{\Xijh}{\hat{X}_{ij}}
+\renewcommand{\AA}{A}
+\newcommand{\AAt}{\tilde{A}}
+\newcommand{\WW}{W}
+\newcommand{\WWt}{\tilde{W}}
+\newcommand{\Wik}{W_{ik}}
+\newcommand{\HH}{H}
+\newcommand{\HHt}{\tilde{H}}
+\newcommand{\Hkj}{H_{kj}}
+\newcommand{\BB}{B}
+\newcommand{\BBt}{\tilde{B}}
+\newcommand{\MM}{M}
+\newcommand{\MMt}{\tilde{M}}
+\newcommand{\Aik}{A_{ik}}
+\newcommand{\Bkr}{B_{kr}}
+\newcommand{\Mrj}{M_{rj}}
+\newcommand{\sij}{\sigma_{ij}}
 
@@
 needed for the purposes of NMF. Section \ref{models} describes the
-basis set of models, Section \ref{In Appendix \ref{nmf} we describe
+basis set of models, Section {\bf redo} In Appendix \ref{nmf} we describe
 how to solve to NMF problem.
@@
 
-[sam sam sam sam]
+\subsection{Standard NMF}
+The standard non-negative matrix factorization (NMF) problem, as
+originally posed by Lee and Seung\cite{??}, is to approximate a data
+matrix $\XX$ (of size $n\times d$) as the outer product of two
+rank-$k$ non-negative matrices $\WW$ and $\HH$. For example, in this
+paper, $\XX$ represents the fluxes at $n$ different wavelengths for
+$d$ different galaxies, $\HH$ represents a set of templates to build
+each galaxy from, and $\WW$ represents, for each galaxy, the weights
+to give each template. This factorization is similar in spirit to
+singular value decomposition (SVD) and principal components analysis
+(PCA) but crucially, in NMF, $\XX$, $\WW$ and $\HH$ are all
+constrained to have non-negative entries. This changes the problem
+fundamentally, since no ``cancellation'' of positive and negative
+basis functions or coefficients is possible -- all approximation
+interactions are strictly additive. The goal, as with SVD or PCA, is
+to minmize the squared approximation error (Frobenius norm):
+\beqa
+\chi^2 &= \|\XX-\WW\HH\|^2\\
+%&= \|\XX-\XXh\|^2\\
+&= \sum_{ij} \left(\Xij - \sum_k \Wik\Hkj \right)^2
+\eeqa
+In their seminal papers\cite{??}, Lee and Seung show that this approximation
+error in non-increasing under the following very simple multiplicative
+update rules: 
+\beqa
+\Wik &\leftarrow \Wik \frac{[\XX\HH\T]_{ik}}{[\WW\HH\HH\T]_{ik}}\\
+\Hkj &\leftarrow \Hkj \frac{[\WW\T\XX]_{kj}}{[\WW\T\WW\HH]_{kj}}
+\eeqa
+$[M]_{ab}$ denotes the $(ab)$ element of the matrix valued expression
+$M$. $\HH\T$ is the transpose of matrix $\HH$.
+
+These rules are remarkable because, although they make finite (not
+infinitesimal) adjustments to the elements of the approximation
+matrices, they have no step-size parameters and are always guaranteed
+to reduce the error (or leave it invariant once they have converged).
+$\chi^2$ is in general a ``non-convex'' function of $\WW$ and $\HH$,
+meaning we cannot guarantee there is only one local
+minimum. Therefore, the procedure does not necessarily find the global
+optimum. But in practice, even with random initialization, these rules
+tend to converge to good solutions when the data matrix $\XX$ is
+highly structured. {\bf explain!}
+
+\subsection{NMF in the space of coefficients for a known basis}
+The factorization problem we face in this paper is a slight variation on the
+basic NMF setup described above. We wish to approximate $\XX$ by the
+product $\AA\BB\MM$, 
+where $\MM$ is a given (fixed) basis matrix of size $r \times d$
+and the optimization is over the matrices $\AA$ (of size
+$n\times k$) and $\BB$ (of size $k\times r$). All matrices
+$\XX,\AA,\BB\,\MM$ have non-negative entries. This can be thought of
+as performing non-negative matrix factorization on the coefficients of
+an approximation, given a fixed (non-negative) basis, described by the
+columns of $\MM$. Once again, the objective we wish to minimize is the
+squared approximation error: 
+\beqa
+\chi^2 &= \|\XX-\AA\BB\MM\|^2\\
+&= \sum_{ij} \left(\Xij - \sum_{kr} \Aik\Bkr\Mrj \right)^2
+\eeqa
+In the case of this paper, this generalization allows us to express
+the templates in terms of star-formation histories, but to compare the
+predicted fluxes for the star-formation histories through the $\MM$
+matrix to the observed fluxes $\XX$.
+
+Of course, in the very special case where $\MM$ is an invertible
+matrix, this problem can be transformed into the original NMF problem above by
+right multiplying the data matrix by $\MM\inv$. However, in most
+situations, including ours, $\MM$ has many fewer rows than columns, and
+as such is far from invertible. Fortunately, however, it is possible
+to derive multiplicative updates for this extended problem 
+which minimze the error directly, even when $\MM$ is not invertible.
+
+First, by thinking of the tensor product $\BB\MM$ as a single
+non-negative matrix $\HH$, we can trivially derive a multiplicative
+update equation for the elements of $\AA$ by using the $\WW$ update
+provided above: 
+\beq
+\Aik \leftarrow \Aik 
+\frac{[\XX\MM\T\BB\T]_{ik}}{[\AA\BB\MM\MM\T\BB\T]_{ik}}
+\eeq
+By Lee and Seung's original proof, this update is guaranteed (for any
+non-negative matrices $\XX$,$\MM$ and $\BB$) not to increase the
+approximation error.
+
+Our main algorithmic contribution is to derive a similar update equation
+for the elements of $\BB$:
+\beq
+\Bkr \leftarrow \Bkr
+\frac{[\AA\T\XX\MM\T]_{kr}}{[\AA\T\AA\BB\MM\MM\T]_{kr}}
+\eeq
+Following the technique outlined in \cite{??}, it can still be shown
+that the error is nonincreasing under the application of this update.
+
+\subsection{Nonuniform uncertainties}
+When different entries $\Xij$ of the data matrix have different
+uncertainties, the natural objective function is the weighted approximation
+error (which also corresponds to the negative log likelihood under a
+Gaussian noise assumption):
+\beq
+\chi^2 = \sum_{ij} \left( \frac{\Xij - 
+\sum_k \Wik\Hkj}{\sij}\right)^2
+\eeq
+
+This case can easily be handled since during the update for $\WW$ the
+elements of $\HH$ are fixed and vice versa and the updates are
+guaranteed not to increase the cost for any nonnegative matrices.
+%In particular, define $\XXt_{ij}=\Xij/\sij$,
+%$\WWt(j)_{ik}=\Wik/\sij$ and 
+%$\HHt(i)_{kj}=\Hkj/\sij$. 
+In particular, when updating $\WW$, we can rewrite the cost function as
+\beq
+\chi^2 = \sum_{ij} \left( \frac{\Xij}{\sij} - 
+\sum_k \Wik\frac{\Hkj}{\sij}\right)^2
+%= \sum_{ij} \left(\XXt_{ij} - 
+%\sum_k \Wik\HHt(i)_{kj}\right)^2
+\eeq
+yielding updates of the form:
+\beq
+\Wik \leftarrow \Wik 
+\left( \sum_j \frac{\Xij\Hkj}{\sij^2} \right) / 
+\left( \sum_{mj} \frac{W_{im}H_{mj}\Hkj}{\sij^2}\right)
+%\Wik \leftarrow \Wik \frac{[\XXt\HHt(i)\T]_{ik}}{[\WW\HHt(i)\HHt(i)\T]_{ik}}
+\eeq
+
+and similarly, when updating $\HH$, we can rewrite the cost function as
+\beq
+\chi^2 = \sum_{ij} \left( \frac{\Xij}{\sij} - 
+\sum_k \frac{\Wik}{\sij}\Hkj\right)^2
+%= \sum_{ij} \left(\XXt_{ij} - 
+%\sum_k \WWt(j)_{ik}\Hkj\right)^2
+\eeq
+yielding updates of the form:
+\beq
+\Hkj \leftarrow \Hkj
+\left( \sum_i \frac{\Wik\Xij}{\sij^2}\right) /
+\left( \sum_{mi} \frac{\Wik{}W_{im}H_{mj}}{\sij^2}\right)
+%\Hkj \leftarrow \Hkj \frac{[\WWt(j)\T\XX]_{kj}}{[\WWt(j)\T\WWt(j)\HH]_{kj}}
+\eeq
+
+This argument can be equally applied to our extended model, yeilding
+the final update equations which are actually implemented in the
+kcorrect code:
+\centerline{\fbox{\parbox{4.3in}{
+\beqa 
+\Aik &\leftarrow \Aik 
+\left( \sum_{rj} \frac{\Xij\Mrj\Bkr}{\sij^2}\right) / 
+\left( \sum_{mnrj} \frac{A_{im}B_{mr}M_{rj}M_{jn}B_{kn}}{\sij^2}\right) \\
+\Bkr &\leftarrow \Bkr
+\left( \sum_{ij} \frac{\Aik\Xij\Mrj}{\sij^2}\right) / 
+\left( \sum_{mnij} \frac{A_{ik}A_{im}B_{mn}M_{nj}M_{rj}}{\sij^2}\right)
+\eeqa
+}}}
+
+%\centerline{\fbox{\parbox{2.6in}{
+%\beqa
+%\Aik &\leftarrow \Aik 
+%\frac{[\XXt\MMt(i)\T\BB\T]_{ik}}{[\AA\BB\MMt(i)\MMt(i)\T\BB\T]_{ik}}\\
+%\Bkr &\leftarrow \Bkr
+%\frac{[\AAt(?)\T\AAt(?)\BB\MM\MM\T]_{kr}}{[\AAt(?)\T\XXt\MM\T]_{kr}}
+%\eeqa
+%}~~}}\\[.2in]
+%pwhere $\XXt_{ij}=\Xij/\sij$, $\MMt(i)_{rj} = \Mrj/\sij$ and
+%$\AAt(?)_{ik}=\Aik/??$. 
+%}}}
 

### 2006-05-16 docs/paper/kcorrect.tex
adjust
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@552 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \newcommand{\XXh}{\hat{X}}
-\newcommand{\Xij}{X_{ij}}
-\newcommand{\Xijh}{\hat{X}_{ij}}
+\newcommand{\Xkn}{X_{kn}}
+\newcommand{\Xknh}{\hat{X}_{kn}}
 \renewcommand{\AA}{A}
@@
 \newcommand{\WWt}{\tilde{W}}
-\newcommand{\Wik}{W_{ik}}
+\newcommand{\Wki}{W_{ki}}
 \newcommand{\HH}{H}
 \newcommand{\HHt}{\tilde{H}}
-\newcommand{\Hkj}{H_{kj}}
+\newcommand{\Hin}{H_{in}}
 \newcommand{\BB}{B}
@@
 \newcommand{\MMt}{\tilde{M}}
-\newcommand{\Aik}{A_{ik}}
-\newcommand{\Bkr}{B_{kr}}
-\newcommand{\Mrj}{M_{rj}}
-\newcommand{\sij}{\sigma_{ij}}
+\newcommand{\Aki}{A_{ki}}
+\newcommand{\Bij}{B_{ij}}
+\newcommand{\Mjn}{M_{jn}}
+\newcommand{\skn}{\sigma_{kn}}
 
@@
 Michael R. Blanton\altaffilmark{\ref{NYU}} and 
-Sam Roweis
+Sam Roweis\altaffilmark{\ref{UToronto}}
 %Tamas Budavari\altaffilmark{\ref{JHU}},
@@
 
-\altaffiltext{1}{Based on observations obtained with the
-Sloan Digital Sky Survey} 
-\setcounter{address}{2}
+\setcounter{address}{0}
 \altaffiltext{\theaddress}{
 \stepcounter{address}
-New York University, Department of Physics, 4 Washington Place, New
+New York University, Center for Cosmology and Particle Physics, 4 Washington Place, New
 York, NY 10003
 \label{NYU}}
+\setcounter{address}{1}
+\altaffiltext{\theaddress}{
+\stepcounter{address}
+University of Toronto, Dept. of Computer Science, 6 King's College
+Rd., Toronto, Ontario M5S 3G4, CANADA
+\label{UToronto}}
 %\altaffiltext{\theaddress}{
@@
 
-The templates are designed to work well for a wide range of data sets.
-We constrain them with photometry and spectroscopy of the Sloan
-Digital Sky Survey (SDSS; \citealt{york00a}) the Galaxy Evolution
-Explorer (GALEX; \citealt{martin05a}) in the ultraviolet, and the
-Two-Micron All Sky Survey (2MASS; \citealt{skrutskie97a}) in the near
-infrared (NIR). In addition, at higher redshifts, we use constraints
-from the Deep Extragalactic Evolutionary Probe 2 (DEEP2;
-\citealt{davis03a, faber03a}) and the Great Observatories Origins Deep
-Survey (GOODS;
-\citealt{giavalisco04a}).  These and other data sets provide a huge set of
-information about galaxy colors and spectra which we can use to help
-understand their star-formation histories.
-
-We note that the methods used here are generally useful. Using these
-methods to develop templates has the advantage that one can easily
-search a large space of models in order to find the few templates
-necessary in order to explain a given set of data.
+The method is designed to work well for a wide range of data sets.  It
+uses photometry and spectroscopy of the Sloan Digital Sky Survey
+(SDSS; \citealt{york00a}) the Galaxy Evolution Explorer (GALEX;
+\citealt{martin05a}) in the ultraviolet, and the Two-Micron All Sky
+Survey (2MASS; \citealt{skrutskie97a}) in the near infrared (NIR). In
+addition, at higher redshifts, we use constraints from the Deep
+Extragalactic Evolutionary Probe 2 (DEEP2; \citealt{davis03a,
+faber03a}) and the Great Observatories Origins Deep Survey (GOODS;
+\citealt{giavalisco04a}).  These and other data sets provide a huge
+set of information about galaxy colors and spectra which we can use to
+help understand their star-formation histories.
+
+We note that some of the algorithmic techniques used here may have
+applications in other areas of astrophysics. First, nonnegative matrix
+factorization (NMF; \citealt{lee00a}), and the extensions to it we
+describe here, is a particularly useful variant on principal component
+analysis. Second, the nonnegative least squares algorithm of
+\citet{sha02a} has the virtue of being extremely simple to implement.
+These methods may find other applications in image analysis, spectroscopic
+analysis, and model fitting in astrophysics.
 
@@
 releases ({\tt v1\_16}; \citealt{blanton03b}; {\tt v3\_2}). The
-improvement over the latest version is twofold. First, we have
+improvement over the previous version is twofold. First, we have
 improved the templates such that they successfully fit galaxies in the
@@
 the data and models used here. In Section \ref{results}, we show the
-results: the best-fit templates and how well they fit the data.
+results: the best-fit templates and how well they fit the data.  In
+Section \ref{kcorrect}, we describe how to convert our results into an
+estimate of the $K$-correction. In Section \ref{physical}, we describe
+the physical interpretation of the templates and of fits to data.  In
+Section \ref{sdss2bessell}, we present some simplified, linear
+transformations between various bandpasses. In Section \ref{summary},
+we summarize our results. Appendix \ref{nmf} describes the NMF
+algorithm.  Appendix \ref{format} describes the electronic format of
+our results.
 
@@
 0.3$, $\Omega_\Lambda = 0.7$, and $H_0 = 100$ $h$ km s$^{-1}$
-Mpc$^{-1}$ (with $h=1$). All magnitudes are (unless otherwise noted)
-AB-relative.  Except where noted, I will always be referring to the
-version of the {\tt kcorrect} software labeled {\tt
-v4\_1\_4}.
+Mpc$^{-1}$ (with $h=1$), unless otherwise noted. All magnitudes are
+(unless otherwise noted) AB-relative.  Except where noted, I will
+always be referring to the version of the {\tt kcorrect} software
+labeled {\tt v4\_1\_4}.
 
@@
 From the spectroscopic observations we know that galaxy spectra reside
-in a low-dimensional locus. Principal Component Analysis (PCA) of
+in a low-dimensional subspace. Principal Component Analysis (PCA) of
 galaxy spectra, introduced by \citet{connolly95b} and applied many
@@
 few templates. This means that, in the linear space of all possible
-spectra, galaxies exist in only a small locus. Therefore, even with
+spectra, galaxies exist in only a small subspace. Therefore, even with
 very heterogeneous data, we should be able to determine the properties
-of this locus.
-
-Here we present an approach to combining the heterogeneous data
-described above in order to determine the properties of the locus of
-galaxy spectra. Rather than taking the model-free approach used by
-PCA, we here restrict the space of possible spectra to those predicted
-from the high resolution stellar population synthesis model of
-\citet{bruzual03a} and the nebular emission line models of \citet{kewley01a}.
-This approach both constrains the problem appropriately and yields a
-natural theoretical interpretation of the results in terms of
-star-formation histories.
+of this subspace.
+
+Here we present an approach to combining heterogeneous data in order
+to determine the properties of the subspace of galaxy spectra. Rather
+than taking the model-free approach used by PCA, we here restrict the
+space of possible spectra to those predicted from the high resolution
+stellar population synthesis model of \citet{bruzual03a} and the
+nebular emission line models of \citet{kewley01a}.  This approach both
+constrains the problem appropriately and yields a natural theoretical
+interpretation of the results in terms of star-formation histories.
 
@@
 observations in the $\chi^2$ sense. Given the entire set of galaxy
-observations available, it also fits for those template star-formation
-histories. The technical name for this algorithm is Nonnegative Matrix
-Factorization (NMF). Note that this problem is not the same as the
-well-known nonnegative least squares problem, which would fit for the
-best-fit nonnegative linear combination of templates, but not for the
-form of the templates themselves.
+observations available, it also fits for the $N$ template
+star-formation histories. The technical name for this algorithm is
+Nonnegative Matrix Factorization (NMF; \citealt{lee99a, lee00a}), and
+we describe it in detail in Appendix \ref{nmf}. Note that this problem
+is not the same as the well-known nonnegative least squares problem,
+which would fit for the best-fit nonnegative linear combination of
+templates, but not for the form of the templates themselves.
 
@@
 as ``nonnegative PCA.''  However, our method has several advantages
-over the standard PCA approach. First, it has a natural physical
-interpretation associated with that spectral subspace, which is the
-corresponding subspace of all possible star-formation
-histories. Second, it naturally handles data uncertainties and missing
-data, which allows it to ignore that variation which is due purely to
-statistical errors.  Third, it handles the complications of observing
-galaxy spectra photometrically using broad-band filters of galaxies at
-varying redshifts, whereas standard PCA requires a constant grid of
-observations in rest-frame wavelength space.
-
-In this section, we describe how to take the astrophysical models and
-the astronomical observations, and to pose the problem in the form
-needed for the purposes of NMF. Section \ref{models} describes the
-basis set of models, Section {\bf redo} In Appendix \ref{nmf} we describe
-how to solve to NMF problem.
+over the standard PCA approach. First, the results yield, along with a
+subspace in the space of all possible spectra, a natural
+interpretation, which is the corresponding subspace in the space of
+all possible star-formation histories. Second, it naturally handles
+data uncertainties and missing data, which allows it to ignore that
+variation which is due purely to statistical errors.  Third, it
+handles the complications of observing galaxy spectra photometrically
+using broad-band filters of galaxies at varying redshifts, whereas
+standard PCA requires a constant grid of observations in rest-frame
+wavelength space.
+
+In the following subsections, we introduce the basis set of models we
+use here (\S\ref{models}), the data we fit to (\S\ref{data}), and how
+we combine the two to set up the NMF problem (\S\ref{setupnmf}).  In
+Appendix \ref{nmf} we describe how to actually solve to NMF problem.
 
@@
 
-Let us refer to these 485 basis templates as $M_{\lambda,
-j}(\lambda)$, expressed in units of ergs s$^{-1}$ \AA$^{-1}$.
+Let us refer to these 485 basis templates as $M_{j}(\lambda)$,
+expressed in units of ergs s$^{-1}$ \AA$^{-1}$.
 
@@
 galaxies actually exist. In particular, we seek five templates
-$F_{\lambda, i}(\lambda)$ built from nonnegative combinations of the
+$F_{i}(\lambda)$ built from nonnegative combinations of the
 original basis set of $N$ templates:
 \begin{equation}
-F_{\lambda, i}(\lambda) = \sum_j b_{ij} M_{\lambda, j}(\lambda),
+F_{i}(\lambda) = \sum_j b_{ij} M_{j}(\lambda),
 \end{equation}
@@
 \begin{equation}
-{\hat{F}}_{\lambda,_k}(\lambda) = \sum_i a_{ki} F_{\lambda,i}(\lambda).
+{\hat{F}}_{k}(\lambda) = \sum_i a_{ki} F_{i}(\lambda).
 \end{equation}
@@
 \subsection{The data}
+\label{data}
 
@@
 \subsection{Comparing data and models}
+\label{setupnmf}
 
@@
 For the spectra, we take the observed spectrum
-$f_{\lambda,k}(\lambda)$ (in ergs cm$^{-2}$ s$^{-1}$ \AA$^{-1}$) of
+$f_{k}(\lambda)$ (in ergs cm$^{-2}$ s$^{-1}$ \AA$^{-1}$) of
 each galaxy $k$ at redshift $z$ (corresponding to a luminosity
@@
 \begin{equation}
-{{F}}_{\lambda, k}(\lambda) = 
-{{f}}_{\lambda, k}[\lambda (1+z)](1+z) (4 \pi d_L^2),
+{{F}}_{k}(\lambda) = 
+{{f}}_{k}[\lambda (1+z)](1+z) (4 \pi d_L^2),
 \end{equation}
@@
 \label{chi2spec}
-\chi_{kl}^2 &=& \left[\frac{F_{\lambda,k}(\lambda_l) -
-{\hat{F}}_{\lambda,k}(\lambda_l)} 
+\chi_{kl}^2 &=& \left[\frac{F_{k}(\lambda_l) -
+{\hat{F}}_{k}(\lambda_l)} 
 {\sigma^2_k(\lambda_l)}\right]^2 \cr
 &=&
-\left[\frac{F_{\lambda,k}(\lambda_l) -
+\left[\frac{F_{k}(\lambda_l) -
 \sum_{ij} a_{ki} b_{ij} M_{jl}
@@
 relative to the AB standard source. For example, if we transform our
-spectral energy density basis function $M_{\lambda, j}(\lambda)$ to a flux
-density $m_{\lambda, j}(\lambda)$, we can calculate the contribution of
+spectral energy density basis function $M_{j}(\lambda)$ to a flux
+density $m_{j}(\lambda)$, we can calculate the contribution of
 that basis function to the predicted maggies as:
@@
 \mu_{jp} = 
-\frac{\int_0^{\infty} d\lambda \lambda R_p(\lambda) {m}_{\lambda, j}(\lambda)}
-{\int_0^\infty d\lambda \lambda R_p(\lambda) f_{\lambda,
-\mathrm{AB}}(\lambda) }
+\frac{\int_0^{\infty} d\lambda \lambda R_p(\lambda) {m}_{j}(\lambda)}
+{\int_0^\infty d\lambda \lambda R_p(\lambda) f_{\mathrm{AB}}(\lambda) }
 \end{equation}
@@
 $\lambda$ entering the Earth's atmosphere (or entering the telescope
-for a space telescope). The AB standard source is $f_{\lambda,
-\mathrm{AB}}(\lambda) d\lambda = f_{\nu,
-\mathrm{AB}} (\nu) d\nu$ and $f_{\nu, \mathrm{AB}}=3631$ Jy $= 3.631 \times
+for a space telescope). The AB standard source is $f_{
+\mathrm{AB}}(\lambda) d\lambda = f_{
+\mathrm{AB}} (\nu) d\nu$ and $f_{\mathrm{AB}}(\nu)=3631$ Jy $= 3.631 \times
 10^{-20}$ erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$. Of course maggies $\mu$
@@
 instead of per {\it photon} with wavelength $\lambda$.  We will refer here
-to that quantity as $R_p'(\lambda)$, though in the literature it is
+to the former quantity as $R_p'(\lambda)$, though in the literature it is
 often referred to without a prime (and without any explicit
@@
 simultaneously. We can similarly combine our observations $\mu_{kp}$
-and $F_{\lambda, k}(\lambda_l)$ and their uncertainties $\sigma_{kp}$
+and $F_{k}(\lambda_l)$ and their uncertainties $\sigma_{kp}$
 and $\sigma_{k}(\lambda_l)$ into vectors $x_{kn}$ and $\sigma_{kn}$.
@@
 nonnegative matrix factorization (NMF) to iterate to the nonnegative
-$a_{ki}$ and $b_{ij}$ which minimize Equation \ref{chi2}. The basic
-method is implemented in a public piece of code named {\tt
+$a_{ki}$ and $b_{ij}$ which locally minimize Equation \ref{chi2}. The
+basic method is implemented in a public piece of code named {\tt
 NMF\_SPARSE} in the {\tt idlutils} distribution of IDL
@@
 
-{\bf nonconvexity; not reproducible}
+As we describe in Appendix \ref{nmf}, the NMF problem is not convex:
+that is, there are multiple local minima. Our method finds one of the
+local minima, but is not guaranteed to find the global minimum. Thus,
+the reader who tries to reproduce our results using our methods or
+others may (depending on their initial conditions) find different
+local minima of $\chi^2$ than do we.
 
@@
 Equation \ref{chi2} for any other galaxy using any nonnegative least
-squares algorithm (since the minimization of Equation \ref{chi2} has a
-linear form in that case).  When we do so here we use the beautifully
-simple iterative method of \citet{sha02a}.
+squares algorithm to determine $a_{ki}$ (since the minimization of
+Equation \ref{chi2} has a linear form in that case).  When we do so
+here we use the beautifully simple iterative method of \citet{sha02a}.
 
@@
 these results are our best-fit, the parameters are highly degenerate
-(especially in, for example, neighboring age bins).
+(especially in, for example, neighboring age bins).  We show these
+merely to illustrate the general nature of the fits.
 
@@
 
-The right-most column of Figure \ref{sfh_templates} shows the spectrum
-from the ultraviolet through the near infrared for the resulting
+We next apply the method to the full data set and fit for five
+templates. From experiments, we found that we needed five to
+adequately explain the data. More than five templates would allow us
+to explain the data better, but only marginally so.  The right-most
+column of Figure \ref{sfh_templates} shows the spectrum from the
+ultraviolet through the near infrared for the resulting
 templates. Note that there is a very old template, a very young
@@
 this result to that in Figure \ref{goods_special}, which we get by
-fitting to {\it only} the GOODS data (this alternate set of five
-templates is also distributed in the manner that Appendix \ref{format}
-describes). From Figure \ref{goods_special}, we conclude that {\it
-some} of the errors are irreducible, and require either a larger set
-of templates for high redshift galaxies (a possibility we find a
-priori unlikely), or that there are simply errors in the catalogs or
-the input filter curves. For example, in the $z$ and $V$ bands we find
-significant trends between redshifts $z=0.5$ and $1.5$ (echoed in
-Figure \ref{goods}). More significantly, in the $H$ band there is at
-least a 10\% offset in the magnitudes across all redshifts, which is
-almost certainly a catalog error. Other errors (such as a large
-scatter in the residuals in the $B$ band and some redshift dependence
-in the $J$ band) are clearly due to the fact that the templates are
-not primarily designed for GOODS data. All in all, we recommend using
-the special templates for $K$-corrections within the GOODS data set.
+fitting a new set of templates to {\it only} the GOODS data (this
+alternate set of five templates is also distributed in the manner that
+Appendix \ref{format} describes). From Figure \ref{goods_special}, we
+conclude that {\it some} of the errors are irreducible, and require
+either a larger set of templates for high redshift galaxies (a
+possibility we find a priori unlikely), or that there are simply
+errors in the catalogs or the input filter curves. For example, in the
+$z$ and $V$ bands we find significant trends between redshifts $z=0.5$
+and $1.5$ (echoed in Figure \ref{goods}). More significantly, in the
+$H$ band there is at least a 10\% offset in the magnitudes across all
+redshifts, which is almost certainly a catalog or filter curve
+error. Other errors (such as a large scatter in the residuals in the
+$B$ band and some redshift dependence in the $J$ band) are clearly due
+to the fact that the templates are not primarily designed for GOODS
+data. All in all, we recommend using the special templates for
+$K$-corrections within the GOODS data set.
 
@@
 stellar spectra used in the population synthesis code of
-\citet{bruzual03a}. These results encourage us that by using
-broad-band data we can infer $K$-corrections of galaxies.
+\citet{bruzual03a}. We are not suggesting that $gri$ photometry
+contains as much information as the full spectra. However, these
+results encourage us that by using broad-band data we can infer
+$K$-corrections of galaxies.
 
@@
 Then, we show the typical $K$-corrections for the data we have fit
-to. Finally, we test our method on actual spectra projected on
-bandpasses to make sure that it works properly.
+to.
 
@@
 \section{Physical interpretation of the models}
+\label{physical}
 
@@
 imaging. Figure \ref{mass_to_garching} shows a galaxy-by-galaxy
-comparison of their stellar masses $M_{\ast,s}$ relative to ours
+comparison of their stellar masses $M_{\ast,s}$ relative to our
 $M_\ast$, as a function of $M_\ast$ and of intrinsic $g-r$ color. The
@@
 
+We do not want to overstate the validity of these physical
+interpretations. They are sufficiently good physical interpretations
+to explain broad band data, but are not uniquely so. Certainly more
+detailed spectroscopic analysis can provide better constraints on
+star-formation histories and stellar masses.
+
+One particular weakness for high redshift data is that the templates
+always assume that there is 14 Gyrs of star-formation history.  Thus,
+{\tt kcorrect} may greatly overestimate the stellar masses of galaxies
+at $z \sim 1$ (for example), where the actual star-formation history
+must last only 6 Gyrs.
+
 \section{Linear relationships between common magnitude systems}
@@
 
-We note here that the general NMF method does not {\tt depend} on
+We note here that the general NMF method does not {\it depend} on
 using a model --- our ``model'' could just have been a set of top hat
@@
 The authors would like to thank Alison Coil, Michael Cooper, David
-W. Hogg, John Moustakas, Leonidas Moustakas, and David Schiminovich
-for their detailed comments and feedback. {\bf funding}
+W. Hogg, John Moustakas, Leonidas Moustakas, David Schiminovich, Risa
+Wechsler, and Andrew West for their detailed comments and
+feedback. This work was funded by a GALEX Archival Program (38) and an
+HST Archival Research grant (AR-9912).
 
@@
 
-\bibliographystyle{../../../nyu-astro/tex/apj}
-\bibliography{../../../nyu-astro/tex/apj-jour,../../../nyu-astro/tex/ccpp}
+\bibliographystyle{/home/users/mb144/nyu-astro/tex/apj}
+\bibliography{/home/users/mb144/nyu-astro/tex/apj-jour,/home/users/mb144/nyu-astro/tex/ccpp}
 
@@
 The standard non-negative matrix factorization (NMF) problem, as
-originally posed by Lee and Seung\cite{??}, is to approximate a data
-matrix $\XX$ (of size $n\times d$) as the outer product of two
-rank-$k$ non-negative matrices $\WW$ and $\HH$. For example, in this
-paper, $\XX$ represents the fluxes at $n$ different wavelengths for
-$d$ different galaxies, $\HH$ represents a set of templates to build
+originally posed by \cite{lee00a}, is to approximate a data
+matrix $\XX$ (of size $N_k\times N_n$) as the outer product of two
+rank-$N_i$ non-negative matrices $\WW$ and $\HH$. For example, in this
+paper, $\XX$ represents the fluxes at $N_n$ different wavelengths for
+$N_k$ different galaxies, $\HH$ represents a set of $N_i$ templates to build
 each galaxy from, and $\WW$ represents, for each galaxy, the weights
@@
 %&= \|\XX-\XXh\|^2\\
-&= \sum_{ij} \left(\Xij - \sum_k \Wik\Hkj \right)^2
+&= \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
 \eeqa
-In their seminal papers\cite{??}, Lee and Seung show that this approximation
+In their seminal papers, \cite{lee00a} show that this approximation
 error in non-increasing under the following very simple multiplicative
@@
 \beqa
-\Wik &\leftarrow \Wik \frac{[\XX\HH\T]_{ik}}{[\WW\HH\HH\T]_{ik}}\\
-\Hkj &\leftarrow \Hkj \frac{[\WW\T\XX]_{kj}}{[\WW\T\WW\HH]_{kj}}
+\Wki &\leftarrow \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
+\Hin &\leftarrow \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
 \eeqa
@@
 optimum. But in practice, even with random initialization, these rules
-tend to converge to good solutions when the data matrix $\XX$ is
-highly structured. {\bf explain!}
+seem to converge to good solutions for real data (at least data like
+ours).
 
@@
 product $\AA\BB\MM$, 
-where $\MM$ is a given (fixed) basis matrix of size $r \times d$
+where $\MM$ is a given (fixed) basis matrix of size $N_j \times N_n$
 and the optimization is over the matrices $\AA$ (of size
-$n\times k$) and $\BB$ (of size $k\times r$). All matrices
+$N_k\times N_i$) and $\BB$ (of size $N_i\times N_j$). All matrices
 $\XX,\AA,\BB\,\MM$ have non-negative entries. This can be thought of
@@
 \chi^2 &= \|\XX-\AA\BB\MM\|^2\\
-&= \sum_{ij} \left(\Xij - \sum_{kr} \Aik\Bkr\Mrj \right)^2
+&= \sum_{kn} \left(\Xkn - \sum_{ij} \Aki\Bij\Mjn \right)^2
 \eeqa
+This equation is almost that posed in Equation \ref{chi2}.
 In the case of this paper, this generalization allows us to express
@@
 \beq
-\Aik \leftarrow \Aik 
-\frac{[\XX\MM\T\BB\T]_{ik}}{[\AA\BB\MM\MM\T\BB\T]_{ik}}
+\Aki \leftarrow \Aki 
+\frac{[\XX\MM\T\BB\T]_{ki}}{[\AA\BB\MM\MM\T\BB\T]_{ki}}
 \eeq
@@
 \beq
-\Bkr \leftarrow \Bkr
-\frac{[\AA\T\XX\MM\T]_{kr}}{[\AA\T\AA\BB\MM\MM\T]_{kr}}
+\Bij \leftarrow \Bij
+\frac{[\AA\T\XX\MM\T]_{ij}}{[\AA\T\AA\BB\MM\MM\T]_{ij}}
 \eeq
-Following the technique outlined in \cite{??}, it can still be shown
+Following the technique outlined in \cite{lee00a}, it can still be shown
 that the error is nonincreasing under the application of this update.
 
+{\bf proof??}
+
 \subsection{Nonuniform uncertainties}
-When different entries $\Xij$ of the data matrix have different
+When different entries $\Xkn$ of the data matrix have different
 uncertainties, the natural objective function is the weighted approximation
@@
 \beq
-\chi^2 = \sum_{ij} \left( \frac{\Xij - 
-\sum_k \Wik\Hkj}{\sij}\right)^2
+\chi^2 = \sum_{kn} \left( \frac{\Xkn - 
+\sum_i \Wki\Hin}{\skn}\right)^2
 \eeq
@@
 guaranteed not to increase the cost for any nonnegative matrices.
-%In particular, define $\XXt_{ij}=\Xij/\sij$,
-%$\WWt(j)_{ik}=\Wik/\sij$ and 
-%$\HHt(i)_{kj}=\Hkj/\sij$. 
+%In particular, define $\XXt_{ij}=\Xkn/\skn$,
+%$\WWt(j)_{ik}=\Wki/\skn$ and 
+%$\HHt(i)_{kj}=\Hin/\skn$. 
 In particular, when updating $\WW$, we can rewrite the cost function as
 \beq
-\chi^2 = \sum_{ij} \left( \frac{\Xij}{\sij} - 
-\sum_k \Wik\frac{\Hkj}{\sij}\right)^2
+\chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
+\sum_i \Wki\frac{\Hin}{\skn}\right)^2
 %= \sum_{ij} \left(\XXt_{ij} - 
-%\sum_k \Wik\HHt(i)_{kj}\right)^2
+%\sum_k \Wki\HHt(i)_{kj}\right)^2
 \eeq
@@
 \beq
-\Wik \leftarrow \Wik 
-\left( \sum_j \frac{\Xij\Hkj}{\sij^2} \right) / 
-\left( \sum_{mj} \frac{W_{im}H_{mj}\Hkj}{\sij^2}\right)
-%\Wik \leftarrow \Wik \frac{[\XXt\HHt(i)\T]_{ik}}{[\WW\HHt(i)\HHt(i)\T]_{ik}}
+\Wki \leftarrow \Wki 
+\left( \sum_n \frac{\Xkn\Hin}{\skn^2} \right) / 
+\left( \sum_{mn} \frac{W_{km}H_{mn}\Hin}{\skn^2}\right)
+%\Wki \leftarrow \Wki \frac{[\XXt\HHt(i)\T]_{ik}}{[\WW\HHt(i)\HHt(i)\T]_{ik}}
 \eeq
@@
 \beq
-\chi^2 = \sum_{ij} \left( \frac{\Xij}{\sij} - 
-\sum_k \frac{\Wik}{\sij}\Hkj\right)^2
+\chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
+\sum_i \frac{\Wki}{\skn}\Hin\right)^2
 %= \sum_{ij} \left(\XXt_{ij} - 
-%\sum_k \WWt(j)_{ik}\Hkj\right)^2
+%\sum_k \WWt(j)_{ik}\Hin\right)^2
 \eeq
@@
 \beq
-\Hkj \leftarrow \Hkj
-\left( \sum_i \frac{\Wik\Xij}{\sij^2}\right) /
-\left( \sum_{mi} \frac{\Wik{}W_{im}H_{mj}}{\sij^2}\right)
-%\Hkj \leftarrow \Hkj \frac{[\WWt(j)\T\XX]_{kj}}{[\WWt(j)\T\WWt(j)\HH]_{kj}}
+\Hin \leftarrow \Hin
+\left( \sum_k \frac{\Wki\Xkn}{\skn^2}\right) /
+\left( \sum_{mk} \frac{\Wki{}W_{km}H_{mn}}{\skn^2}\right)
+%\Hin \leftarrow \Hin \frac{[\WWt(j)\T\XX]_{kj}}{[\WWt(j)\T\WWt(j)\HH]_{kj}}
 \eeq
 
-This argument can be equally applied to our extended model, yeilding
+This argument can be equally applied to our extended model, yielding
 the final update equations which are actually implemented in the
 kcorrect code:
-\centerline{\fbox{\parbox{4.3in}{
+%%\centerline{\fbox{\parbox{4.3in}{
 \beqa 
-\Aik &\leftarrow \Aik 
-\left( \sum_{rj} \frac{\Xij\Mrj\Bkr}{\sij^2}\right) / 
-\left( \sum_{mnrj} \frac{A_{im}B_{mr}M_{rj}M_{jn}B_{kn}}{\sij^2}\right) \\
-\Bkr &\leftarrow \Bkr
-\left( \sum_{ij} \frac{\Aik\Xij\Mrj}{\sij^2}\right) / 
-\left( \sum_{mnij} \frac{A_{ik}A_{im}B_{mn}M_{nj}M_{rj}}{\sij^2}\right)
+\Aki &\leftarrow \Aki 
+\left( \sum_{jn} \frac{\Xkn\Mjn\Bij}{\skn^2}\right) / 
+\left( \sum_{mljn} \frac{A_{km}B_{mj}M_{jn}M_{nl}B_{il}}{\skn^2}\right) \\
+\Bij &\leftarrow \Bij
+\left( \sum_{kn} \frac{\Aki\Xkn\Mjn}{\skn^2}\right) / 
+\left( \sum_{mlkn} \frac{A_{ki}A_{km}B_{ml}M_{ln}M_{jn}}{\skn^2}\right)
 \eeqa
-}}}
+%%}}}
 
@@
 %\beqa
-%\Aik &\leftarrow \Aik 
+%\Aki &\leftarrow \Aki 
 %\frac{[\XXt\MMt(i)\T\BB\T]_{ik}}{[\AA\BB\MMt(i)\MMt(i)\T\BB\T]_{ik}}\\
-%\Bkr &\leftarrow \Bkr
+%\Bij &\leftarrow \Bij
 %\frac{[\AAt(?)\T\AAt(?)\BB\MM\MM\T]_{kr}}{[\AAt(?)\T\XXt\MM\T]_{kr}}
@@
 %}~~}}\\[.2in]
-%pwhere $\XXt_{ij}=\Xij/\sij$, $\MMt(i)_{rj} = \Mrj/\sij$ and
-%$\AAt(?)_{ik}=\Aik/??$. 
+%pwhere $\XXt_{ij}=\Xkn/\skn$, $\MMt(i)_{rj} = \Mjn/\skn$ and
+%$\AAt(?)_{ik}=\Aki/??$. 
 %}}}

### 2006-06-07 docs/paper/kcorrect-aj-v1/ms.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@553 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2006-05-16 docs/paper/kcorrect.tex
@@
-\documentclass[10pt,preprint]{aastex}
+\documentclass[12pt,preprint]{aastex}
 \usepackage{amsmath}
@@
 \newcommand{\Wki}{W_{ki}}
+\newcommand{\Wkl}{W_{k\ell}}
 \newcommand{\HH}{H}
@@
 \newcommand{\Hin}{H_{in}}
+\newcommand{\Hln}{H_{\ell{}n}}
 \newcommand{\BB}{B}
@@
 \newcommand{\Aki}{A_{ki}}
+\newcommand{\Akl}{A_{k\ell}}
 \newcommand{\Bij}{B_{ij}}
+\newcommand{\Bls}{B_{\ell{}s}}
 \newcommand{\Mjn}{M_{jn}}
+\newcommand{\Msn}{M_{sn}}
 \newcommand{\skn}{\sigma_{kn}}
+\newcommand{\PP}{P}
+\newcommand{\Pni}{P_{ni}}
+\newcommand{\ZZ}{Z}
+\newcommand{\Zki}{Z_{ki}}
+\newcommand{\Zkl}{Z_{k\ell}}
+\newcommand{\Zij}{Z_{ij}}
+\newcommand{\Zkj}{Z_{kj}}
+\newcommand{\Zks}{Z_{ks}}
+\newcommand{\zz}{z}
+\newcommand{\yy}{y}
+\newcommand{\yn}{y_n}
+\newcommand{\yi}{y_i}
+\newcommand{\zn}{z_n}
+\newcommand{\zi}{z_i}
+\newcommand{\betaij}{\beta_{ij}}
+\newcommand{\betalj}{\beta_{\ell{}j}}
 
@@
 \title{$K$-corrections and filter transformations \\
-from the restframe
-ultraviolet to the near infrared}
+in the ultraviolet, optical, and near infrared}
 
@@
 \begin{abstract}
-We present a method of developing model-based spectroscopic templates
-to explain observed galaxy fluxes. Such templates are useful for
-calculating $K$-corrections and converting among observations of
-galaxies at various wavelengths. In principle, we could apply this
-method to galaxy data at any redshift and in any set of
-wavelengths. The implementation we present here is suitable for
-ultraviolet, optical, and near-infrared observations in the redshift
-range $0 < z < 1.5$. Since we base the templates on stellar population
-synthesis results, the results are intepretable in terms of
-approximate stellar masses and star-formation histories. We present
-templates fit with this method to data from GALEX, Sloan Digital Sky
-Survey spectroscopy and photometry, the Two-Micron All Sky Survey, the
-Deep Extragalactic Evolutionary Probe and Great Observatories Origins
-Deep Survey.  In addition, we present software for using such data to
-estimate $K$-corrections, stellar masses, and recent star-formation
-rates. This software is suitable for analyzing many other data sets as
-well.
+Template fits to observed galaxy fluxes allow calculation of
+$K$-corrections and conversions among observations of galaxies at
+various wavelengths. We present a method for creating model-based
+template sets given a set of heterogeneous photometric and
+spectroscopic galaxy data. Our technique, non-negative matrix
+factorization, is akin to principle component analysis (PCA), except
+that it is constrained to produce nonnegative templates, it can use a
+basis set of models (rather than the delta function basis of PCA), and
+it naturally handles uncertainties, missing data, and heterogeneous
+data (including broad-band fluxes at various redshifts).  The
+particular implementation we present here is suitable for ultraviolet,
+optical, and near-infrared observations in the redshift range $0 < z <
+1.5$. Since we base our templates on stellar population synthesis
+models, the results are intepretable in terms of approximate stellar
+masses and star-formation histories. We present templates fit with
+this method to data from GALEX, Sloan Digital Sky Survey spectroscopy
+and photometry, the Two-Micron All Sky Survey, the Deep Extragalactic
+Evolutionary Probe and the Great Observatories Origins Deep Survey.
+In addition, we present software for using such data to estimate
+$K$-corrections and stellar masses.
 \end{abstract}
@@
 
-\bibliographystyle{/home/users/mb144/nyu-astro/tex/apj}
-\bibliography{/home/users/mb144/nyu-astro/tex/apj-jour,/home/users/mb144/nyu-astro/tex/ccpp}
+\begin{thebibliography}{39}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Bell} \& {de~Jong}(2001)}]{bell01b}
+{Bell}, E.~F. \& {de~Jong}, R.~S. 2001, \apj, 550, 212
+
+\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96a}
+{Bertin}, E. \& {Arnouts}, S. 1996, \aaps, 117, 393
+
+\bibitem[{{Bessell}(1990)}]{bessell90a}
+{Bessell}, M.~S. 1990, \pasp, 102, 1181
+
+\bibitem[{Blanton {et~al.}(2003)Blanton, Brinkmann, Csabai, Doi, Eisenstein,
+  Fukugita, Gunn, Hogg, \& Schlegel}]{blanton03b}
+Blanton, M.~R., Brinkmann, J., Csabai, I., Doi, M., Eisenstein, D.~J.,
+  Fukugita, M., Gunn, J.~E., Hogg, D.~W., \& Schlegel, D.~J. 2003, \aj, 125,
+  2348
+
+\bibitem[{{Blanton} {et~al.}(2003){Blanton}, {Lin}, {Lupton}, {Maley}, {Young},
+  {Zehavi}, \& {Loveday}}]{blanton03a}
+{Blanton}, M.~R., {Lin}, H., {Lupton}, R.~H., {Maley}, F.~M., {Young}, N.,
+  {Zehavi}, I., \& {Loveday}, J. 2003, \aj, 125, 2276
+
+\bibitem[{{Blanton} {et~al.}(2005)}]{blanton05a}
+{Blanton}, M.~R. {et~al.} 2005, \aj, 129, 2562
+
+\bibitem[{Bruzual \& Charlot(2003)}]{bruzual03a}
+Bruzual, A.~G. \& Charlot, S. 2003, \mnras, 344, 1000
+
+\bibitem[{{Chabrier}(2003)}]{chabrier03a}
+{Chabrier}, G. 2003, \pasp, 115, 763
+
+\bibitem[{{Cohen} {et~al.}(2003){Cohen}, {Wheaton}, \& {Megeath}}]{cohen03a}
+{Cohen}, M., {Wheaton}, W.~A., \& {Megeath}, S.~T. 2003, \aj, 126, 1090
+
+\bibitem[{{Connolly} {et~al.}(1995){Connolly}, {Szalay}, {Bershady}, {Kinney},
+  \& {Calzetti}}]{connolly95b}
+{Connolly}, A.~J., {Szalay}, A.~S., {Bershady}, M.~A., {Kinney}, A.~L., \&
+  {Calzetti}, D. 1995, \aj, 110, 1071
+
+\bibitem[{{Davis} {et~al.}(2003){Davis}, {Faber}, {Newman}, {Phillips},
+  {Ellis}, {Steidel}, {Conselice}, {Coil}, {Finkbeiner}, {Koo}, {Guhathakurta},
+  {Weiner}, {Schiavon}, {Willmer}, {Kaiser}, {Luppino}, {Wirth}, {Connolly},
+  {Eisenhardt}, {Cooper}, \& {Gerke}}]{davis03a}
+{Davis}, M., {Faber}, S.~M., {Newman}, J., {Phillips}, A.~C., {Ellis}, R.~S.,
+  {Steidel}, C.~C., {Conselice}, C., {Coil}, A.~L., {Finkbeiner}, D.~P., {Koo},
+  D.~C., {Guhathakurta}, P., {Weiner}, B., {Schiavon}, R., {Willmer}, C.,
+  {Kaiser}, N., {Luppino}, G.~A., {Wirth}, G., {Connolly}, A., {Eisenhardt},
+  P., {Cooper}, M., \& {Gerke}, B. 2003, in Discoveries and Research Prospects
+  from 6- to 10-Meter-Class Telescopes II. Edited by Guhathakurta, Puragra.
+  Proceedings of the SPIE, Volume 4834, pp. 161-172 (2003)., 161--172
+
+\bibitem[{Eisenstein {et~al.}(2001)}]{eisenstein01a}
+Eisenstein, D.~J. {et~al.} 2001, \aj, 122, 2267
+
+\bibitem[{{Faber} {et~al.}(2003){Faber}, {Phillips}, {Kibrick}, {Alcott},
+  {Allen}, {Burrous}, {Cantrall}, {Clarke}, {Coil}, {Cowley}, {Davis}, {Deich},
+  {Dietsch}, {Gilmore}, {Harper}, {Hilyard}, {Lewis}, {McVeigh}, {Newman},
+  {Osborne}, {Schiavon}, {Stover}, {Tucker}, {Wallace}, {Wei}, {Wirth}, \&
+  {Wright}}]{faber03a}
+{Faber}, S.~M., {Phillips}, A.~C., {Kibrick}, R.~I., {Alcott}, B., {Allen},
+  S.~L., {Burrous}, J., {Cantrall}, T., {Clarke}, D., {Coil}, A.~L., {Cowley},
+  D.~J., {Davis}, M., {Deich}, W.~T.~S., {Dietsch}, K., {Gilmore}, D.~K.,
+  {Harper}, C.~A., {Hilyard}, D.~F., {Lewis}, J.~P., {McVeigh}, M., {Newman},
+  J., {Osborne}, J., {Schiavon}, R., {Stover}, R.~J., {Tucker}, D., {Wallace},
+  V., {Wei}, M., {Wirth}, G., \& {Wright}, C.~A. 2003, in Instrument Design and
+  Performance for Optical/Infrared Ground-based Telescopes. Edited by Iye,
+  Masanori; Moorwood, Alan F. M. Proceedings of the SPIE, Volume 4841, pp.
+  1657-1669 (2003)., 1657--1669
+
+\bibitem[{Fukugita {et~al.}(1996)Fukugita, Ichikawa, Gunn, Doi, Shimasaku, \&
+  Schneider}]{fukugita96a}
+Fukugita, M., Ichikawa, T., Gunn, J.~E., Doi, M., Shimasaku, K., \& Schneider,
+  D.~P. 1996, \aj, 111, 1748
+
+\bibitem[{{Giavalisco} {et~al.}(2004)}]{giavalisco04a}
+{Giavalisco}, M. {et~al.} 2004, \apjl, 600, L93
+
+\bibitem[{{Hayes}(1985)}]{hayes85a}
+{Hayes}, D.~S. 1985, in IAU Symp. 111: Calibration of Fundamental Stellar
+  Quantities, Vol. 111, 225--249
+
+\bibitem[{{Hogg}(1999)}]{hogg99a}
+{Hogg}, D.~W. 1999, in ASP Conf. Ser. 193: The Hy-Redshift Universe: Galaxy
+  Formation and Evolution at High Redshift, 224--+
+
+\bibitem[{{Hogg} {et~al.}(2001){Hogg}, {Finkbeiner}, {Schlegel}, \&
+  {Gunn}}]{hogg01a}
+{Hogg}, D.~W., {Finkbeiner}, D.~P., {Schlegel}, D.~J., \& {Gunn}, J.~E. 2001,
+  \aj, 122, 2129
+
+\bibitem[{Hogg {et~al.}(2002)}]{hogg02a}
+Hogg, D.~W. {et~al.} 2002, \aj, 124, 646
+
+\bibitem[{{Jarrett} {et~al.}(2000){Jarrett}, {Chester}, {Cutri}, {Schneider},
+  {Skrutskie}, \& {Huchra}}]{jarrett00a}
+{Jarrett}, T.~H., {Chester}, T., {Cutri}, R., {Schneider}, S., {Skrutskie}, M.,
+  \& {Huchra}, J.~P. 2000, \aj, 119, 2498
+
+\bibitem[{{Kauffmann} {et~al.}(2003)}]{kauffmann03a}
+{Kauffmann}, G. {et~al.} 2003, \mnras, 341, 33
+
+\bibitem[{{Kennicutt} {et~al.}(1994){Kennicutt}, {Tamblyn}, \&
+  {Congdon}}]{kennicutt94a}
+{Kennicutt}, R.~C., {Tamblyn}, P., \& {Congdon}, C.~E. 1994, \apj, 435, 22
+
+\bibitem[{{Kewley} {et~al.}(2001){Kewley}, {Dopita}, {Sutherland}, {Heisler},
+  \& {Trevena}}]{kewley01a}
+{Kewley}, L.~J., {Dopita}, M.~A., {Sutherland}, R.~S., {Heisler}, C.~A., \&
+  {Trevena}, J. 2001, \apj, 556, 121
+
+\bibitem[{Kurucz(1991)}]{kurucz91a}
+Kurucz, R.~L. 1991, in Precision Photometry:\ Astrophysics of the Galaxy, L.
+  Davis Press, Inc., 27--44
+
+\bibitem[{Lee \& Seung(1999)}]{lee99a}
+Lee, D.~D. \& Seung, H.~S. 1999, Nature, 401, 788
+
+\bibitem[{Lee \& Seung(2000)}]{lee00a}
+Lee, D.~D. \& Seung, H.~S. 2000, in {NIPS}, 556--562
+
+\bibitem[{Lupton {et~al.}(2001)Lupton, Gunn, {Ivezi{\'c}}, Knapp, Kent, \&
+  Yasuda}]{lupton01a}
+Lupton, R.~H., Gunn, J.~E., {Ivezi{\'c}}, Z., Knapp, G.~R., Kent, S., \&
+  Yasuda, N. 2001, in ASP Conf.\ Ser.\ 238:\ Astronomical Data Analysis
+  Software and Systems X, Vol.~10, 269
+
+\bibitem[{Martin {et~al.}(2005)}]{martin05a}
+Martin, D.~C. {et~al.} 2005, \apjl, 619, L1
+
+\bibitem[{{Oke} \& {Sandage}(1968)}]{oke68a}
+{Oke}, J.~B. \& {Sandage}, A. 1968, \apj, 154, 21
+
+\bibitem[{Pier {et~al.}(2003)Pier, Munn, Hindsley, Hennessy, Kent, Lupton, \&
+  {Ivezi{\' c}}}]{pier03a}
+Pier, J.~R., Munn, J.~A., Hindsley, R.~B., Hennessy, G.~S., Kent, S.~M.,
+  Lupton, R.~H., \& {Ivezi{\' c}}, {\v Z}. 2003, \aj, 125, 1559
+
+\bibitem[{Richards {et~al.}(2002)}]{richards02a}
+Richards, G. {et~al.} 2002, \aj, 123, 2945
+
+\bibitem[{{Schneider} {et~al.}(1983){Schneider}, {Gunn}, \&
+  {Hoessel}}]{schneider83a}
+{Schneider}, D.~P., {Gunn}, J.~E., \& {Hoessel}, J.~G. 1983, \apj, 264, 337
+
+\bibitem[{Sha {et~al.}(2002)Sha, Saul, \& Lee}]{sha02a}
+Sha, F., Saul, L., \& Lee, D. 2002, University of Pennsylvania, Technical
+  Report MS-CIS-02-19
+
+\bibitem[{Skrutskie {et~al.}(1997)}]{skrutskie97a}
+Skrutskie, M.~F. {et~al.} 1997, in ASSL Vol. 210: The Impact of Large Scale
+  Near-IR Sky Surveys, 25
+
+\bibitem[{Smith {et~al.}(2002)Smith, Tucker, {et~al.}}]{smith02a}
+Smith, J.~A., Tucker, D.~L., {et~al.} 2002, \aj, 123, 2121
+
+\bibitem[{Stoughton {et~al.}(2002)}]{stoughton02a}
+Stoughton, C. {et~al.} 2002, \aj, 123, 485
+
+\bibitem[{Strauss {et~al.}(2002)}]{strauss02a}
+Strauss, M.~A. {et~al.} 2002, \aj, 124, 1810
+
+\bibitem[{{Witt} \& {Gordon}(2000)}]{witt00a}
+{Witt}, A.~N. \& {Gordon}, K.~D. 2000, \apj, 528, 799
+
+\bibitem[{York {et~al.}(2000)}]{york00a}
+York, D. {et~al.} 2000, \aj, 120, 1579
+
+\end{thebibliography}
 
@@
 to minmize the squared approximation error (Frobenius norm):
-\beqa
+\beqa \label{eq:nmfcost}
 \chi^2 &= \|\XX-\WW\HH\|^2\\
@@
 update rules: 
-\beqa
+\beqa \label{eq:wupdate}
 \Wki &\leftarrow \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
+\label{eq:hupdate}
 \Hin &\leftarrow \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
@@
 for the elements of $\BB$:
-\beq
+\beq \label{eq:bupdate}
 \Bij \leftarrow \Bij
@@
 
-{\bf proof??}
+The proof involves the use of an inequality lemma for symmetric
+non-negative matrices:
+
+{\bf Lemma:} 
+{\it For any symmetric matrix $\PP$ having non-negative entries $\Pni
+\geq 0$, any vector $\zz$ having non-negative entries $\zn \geq 0$,
+and any vector $\yy$,}
+\beq
+\sum_{ni} \yn \yi \Pni \leq \sum_{ni} \yn^2 \frac{\zi}{\zn} \Pni 
+\eeq
+%{\it If the vector $\yy$ has non-negative entries, equality can be
+%achieved by choosing $\zz=\yy$.}
+
+For the standard NMF problem, we can prove that the approximation
+error \eqref{eq:nmfcost} is non-increasing under \eqref{eq:wupdate} by
+using the above lemma to construct a function $\phi(\WW,\ZZ)$ which is
+an upper bound on the cost $\chi^2(\WW)$ for any non-negative matrix $\ZZ$:
+{\small 
+\beqa 
+\chi^2(\WW) &= \sum_{kni\ell} \Wki\Wkl \Hin\Hln
+- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
+\phi(\WW,\ZZ) &= \sum_{kni\ell} \Wki^2\frac{\Zkl}{\Zki} \Hin\Hln 
+- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
+\phi(\WW,\ZZ) &\geq \chi^2(\WW) \qquad \forall \: \Wki\geq 0,\: \Zki\geq 0
+\eeqa
+} with equality being achieved when $\ZZ=\WW$. Figure \ref{lemma}
+represents this definition schematically, showing the true $\chi^2$
+function and $\phi$. The trick of the method is to define $\phi$ such
+that it is equal to $\chi^2$ at $W$, greater than $\chi^2$ everywhere
+else, and is easily minimizable. Then one can use the minimum of
+$\phi$ as the update, and be guaranteed that one's updates do not
+increase $\chi^2$.
+
+The function $\phi$ can be analytically minimized with respect to
+its first argument:
+\beqa
+\phi(\WW^*,\ZZ) &\leq \phi(\WW,\ZZ) \qquad \forall \: \WW,\ZZ \\
+\Wki^* &= \Zki \frac{\sum_n \Xkn \Hin}{\sum_{n\ell} \Hin\Hln \Zkl}
+\eeqa
+If we set $\ZZ=\WW$, this becomes exactly \eqref{eq:wupdate} and
+now we can easily prove the validity of this update rule: 
+\beq
+\chi^2(\WW) = \phi(\WW,\WW) \geq \phi(\WW^*,\WW) \geq \chi^2(\WW^*)
+\eeq
+where the first inequality comes from the fact that $\WW^*$ minimizes
+$\phi(\cdot,\WW)$ with respect to its first argument
+and the second comes from the fact that $\phi(\WW^*,\cdot)$ is a bound
+on $\chi^2(\WW^*)$. The proof of validity for the update
+\eqref{eq:hupdate} is analogous by symmetry.
+
+To prove the validity of our update \eqref{eq:bupdate} for our new
+problem, we proceed in a similar fashion, using the lemma twice to
+construct consecutive upper bounds on the cost:
+{\small 
+\beqa 
+\chi^2(\BB) &= \sum_{kni\ell{}js} \Aki\Akl\Bij\Bls\Mjn\Msn\\ \nonumber
+&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
+\phi(\BB,\ZZ,\beta) &= \sum_{kni\ell{}js} \Aki\Akl\Bij^2
+\frac{\Zks\betalj}{\Zkj\betaij}\Mjn\Msn\\ \nonumber
+&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
+\phi(\BB,\ZZ,\beta) &\geq \chi^2(\BB) \qquad \forall 
+\: \Bij\geq 0,\: \Zkj\geq 0,\: \betaij\geq 0
+\eeqa
+} with equality being achieved when $\ZZ=\AA\BB$ and 
+$\beta=\BB$.
+
+Once again, the bound $\phi$ can be analytically minimized with respect to
+its first argument:
+\beqa
+\phi(\BB^*,\ZZ,\beta) &\leq \phi(\BB,\ZZ,\beta) \qquad 
+\forall \: \BB,\ZZ,\beta \\
+\Bij^* &= \betaij \frac{\sum_{kn} \Xkn\Aki\Mjn}
+{\sum_{kn\ell{}s} \Aki\Akl\frac{\Zks\betalj}{\Zkj}\Mjn\Msn}
+\eeqa
+If we set $\ZZ=\AA\BB$ and $\beta=\BB$, this becomes
+exactly \eqref{eq:bupdate} and 
+we can now prove the validity of this update rule: 
+\beq
+\chi^2(\BB) = \phi(\BB,\AA\BB,\BB) \geq \phi(\BB^*,\AA\BB,\BB) 
+\geq \chi^2(\BB^*) 
+\eeq
+where the first inequality comes from the fact that $\BB^*$ minimizes
+$\phi(\cdot,\AA\BB,\BB)$ with respect to its first argument
+and the second comes from the fact that $\phi(\BB^*,\cdot,\cdot)$ is a bound
+on $\chi^2(\BB^*)$. The only condition we require is that the elements
+of the matrix $\MM\MM\T$ are all non-negative.
 
@@
 
-\include{kcorrect_tables}
-
-\include{kcorrect_figures}
+\clearpage
+\begin{deluxetable}{rrrrr}
+\tablewidth{0pt}
+\tablecolumns{5}
+\tablecaption{\label{solarmagnitudes} Properties of various filters}
+\tablehead{ Band & $\lambda_{\mathrm{eff}}$ (\AA) & $m_{\mathrm{AB}} -
+	m_{\mathrm{Vega}}$ & $M_\odot$ (AB) &
+	$M_\odot$ (Vega)}
+\startdata
+$U$ & 3571. & 0.79 & 6.35 & 5.55\cr
+$B$ & 4344. & -0.09 & 5.36 & 5.45\cr
+$V$ & 5456. & 0.02 & 4.80 & 4.78\cr
+$R$ & 6442. & 0.21 & 4.61 & 4.41\cr
+$I$ & 7994. & 0.45 & 4.52 & 4.07\cr
+$u$ & 3546. & 0.91 & 6.38 & 5.47\cr
+$g$ & 4670. & -0.08 & 5.12 & 5.20\cr
+$r$ & 6156. & 0.16 & 4.64 & 4.49\cr
+$i$ & 7472. & 0.37 & 4.53 & 4.16\cr
+$z$ & 8917. & 0.54 & 4.51 & 3.97\cr
+$J$ & 12355. & 0.91 & 4.56 & 3.65\cr
+$H$ & 16458. & 1.39 & 4.71 & 3.32\cr
+$K_s$ & 21603. & 1.85 & 5.14 & 3.29\cr
+\band{0.1}{u} & 3224. & 1.25 & 6.78 & 5.53\cr
+\band{0.1}{g} & 4245. & -0.01 & 5.43 & 5.44\cr
+\band{0.1}{r} & 5597. & 0.04 & 4.76 & 4.71\cr
+\band{0.1}{i} & 6792. & 0.27 & 4.57 & 4.30\cr
+\band{0.1}{z} & 8107. & 0.46 & 4.52 & 4.05\cr
+\enddata
+\tablecomments{Uses model solar spectrum and model Vega spectrum of
+	\citet{kurucz91a}. Effective wavelength is defined in the text. The
+	$UBRVI$ filters are those of \citet{bessell90a}. The $ugriz$ filters
+	are those determined by Mamoru Doi, Daniel Eisenstein, and James
+	Gunn and available on the SDSS DR4 web site.\footnote{{\tt
+	http://www.sdss.org/dr4/}} The $JHK_s$ filters are those from
+	\citet{cohen03a}. }
+\end{deluxetable}
+
+\begin{deluxetable}{lr}
+\tablewidth{0pt}
+\tablecolumns{2}
+\tablecaption{\label{lineareqs} Conversions among various filters}
+\tablehead{ Equation & Color dispersion }
+\startdata
+$u = \band{0.1}{u} - 0.3310 - 0.3014 \left[ (\band{0.1}{u}-\band{0.1}{g}) - 1.2839 \right] $ & $\sigma\left[\band{0.1}{u}-\band{0.1}{g}\right] = 0.28$ \cr
+$g = \band{0.1}{g} - 0.3112 - 0.3530 \left[ (\band{0.1}{g}-\band{0.1}{r}) - 0.7187 \right] $ & $\sigma\left[\band{0.1}{g}-\band{0.1}{r}\right] = 0.20$ \cr
+$r = \band{0.1}{r} - 0.2026 - 0.3810 \left[ (\band{0.1}{r}-\band{0.1}{i}) - 0.3530 \right] $ & $\sigma\left[\band{0.1}{r}-\band{0.1}{i}\right] = 0.07$ \cr
+$i = \band{0.1}{i} - 0.1072 - 0.4273 \left[ (\band{0.1}{i}-\band{0.1}{z}) - 0.1914 \right] $ & $\sigma\left[\band{0.1}{i}-\band{0.1}{z}\right] = 0.13$ \cr
+$z = \band{0.1}{i} - 0.3161 - 1.2057 \left[ (\band{0.1}{i}-\band{0.1}{z}) - 0.1914 \right] $ & $\sigma\left[\band{0.1}{i}-\band{0.1}{z}\right] = 0.13$ \cr
+$u = U + 0.0682 + 0.0197 \left[ (U-B) - 0.9602 \right] $ & $\sigma\left[U-B\right] = 0.20$ \cr
+$g = B - 0.2354 - 0.3411 \left[ (B-V) - 0.5870 \right] $ & $\sigma\left[B-V\right] = 0.18$ \cr
+$r = V - 0.2585 - 0.5003 \left[ (V-R) - 0.3161 \right] $ & $\sigma\left[V-R\right] = 0.07$ \cr
+$i = R - 0.2000 - 0.4248 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
+$z = R - 0.4088 - 1.2495 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
+$\band{0.1}{u} = U + 0.3989 + 0.4135 \left[ (U-B) - 0.9602 \right] $ & $\sigma\left[U-B\right] = 0.20$ \cr
+$\band{0.1}{g} = U - 0.8845 - 0.9508 \left[ (U-B) - 0.9602 \right] $ & $\sigma\left[U-B\right] = 0.20$ \cr
+$\band{0.1}{g} = B + 0.0759 + 0.0620 \left[ (B-V) - 0.5870 \right] $ & $\sigma\left[B-V\right] = 0.18$ \cr
+$\band{0.1}{r} = B - 0.6429 - 1.0845 \left[ (B-V) - 0.5870 \right] $ & $\sigma\left[B-V\right] = 0.18$ \cr
+$\band{0.1}{r} = V - 0.0558 - 0.1803 \left[ (V-R) - 0.3161 \right] $ & $\sigma\left[V-R\right] = 0.07$ \cr
+$\band{0.1}{i} = R - 0.0927 + 0.0035 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
+$\band{0.1}{z} = R - 0.2841 - 1.0301 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
+$\band{0.1}{u} = u + 0.3310 + 0.3203 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
+$\band{0.1}{g} = u - 0.9528 - 0.7572 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
+$\band{0.1}{g} = g + 0.3113 + 0.4620 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
+$\band{0.1}{r} = g - 0.4075 - 0.8577 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
+$\band{0.1}{i} = r - 0.1504 - 0.3654 \left[ (r-i) - 0.2589 \right] $ & $\sigma\left[r-i\right] = 0.10$ \cr
+$\band{0.1}{z} = i - 0.0836 - 0.7518 \left[ (i-z) - 0.2083 \right] $ & $\sigma\left[i-z\right] = 0.10$ \cr
+$U = u - 0.0682 - 0.0140 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
+$B = u - 1.0286 - 0.7981 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
+$B = g + 0.2354 + 0.3915 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
+$V = g - 0.3516 - 0.7585 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
+$R = r - 0.0576 - 0.3718 \left[ (r-i) - 0.2589 \right] $ & $\sigma\left[r-i\right] = 0.10$ \cr
+$I = i - 0.0647 - 0.7177 \left[ (i-z) - 0.2083 \right] $ & $\sigma\left[i-z\right] = 0.10$ \cr
+$U = \band{0.1}{u} - 0.3992 - 0.3189 \left[ (\band{0.1}{u}-\band{0.1}{g}) - 1.2839 \right] $ & $\sigma\left[\band{0.1}{u}-\band{0.1}{g}\right] = 0.28$ \cr
+$B = \band{0.1}{g} - 0.0759 - 0.0545 \left[ (\band{0.1}{g}-\band{0.1}{r}) - 0.7187 \right] $ & $\sigma\left[\band{0.1}{g}-\band{0.1}{r}\right] = 0.20$ \cr
+$V = \band{0.1}{g} - 0.6628 - 0.9259 \left[ (\band{0.1}{g}-\band{0.1}{r}) - 0.7187 \right] $ & $\sigma\left[\band{0.1}{g}-\band{0.1}{r}\right] = 0.20$ \cr
+$R = \band{0.1}{r} - 0.2603 - 0.9162 \left[ (\band{0.1}{r}-\band{0.1}{i}) - 0.3530 \right] $ & $\sigma\left[\band{0.1}{r}-\band{0.1}{i}\right] = 0.07$ \cr
+$I = \band{0.1}{i} - 0.1725 - 0.9718 \left[ (\band{0.1}{i}-\band{0.1}{z}) - 0.1914 \right] $ & $\sigma\left[\band{0.1}{i}-\band{0.1}{z}\right] = 0.13$ \cr
+\enddata
+\tablecomments{Uses AB magnitudes throughout.}
+\end{deluxetable}
+
+\begin{deluxetable}{llcl}
+\tablewidth{0pt}
+\tablecolumns{4}
+\tablecaption{\label{derived} Header and Data Units (HDUs) of {\tt k\_nmf\_derived.default.fits}}
+\tablehead{ Number & Name & Dimensions & Description}
+\startdata
+0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
+in basis space for each template \cr 1 & {\tt spec} &
+$N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of each template
+\cr 2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed
+spectrum of each template, lines removed \cr 3 & {\tt spec\_nd} &
+$N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of each template,
+without dust \cr 4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times
+N_t$ & Smoothed spectrum of each template, without lines or dust \cr 5
+& {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum
+of each template \cr 6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times
+N_t$ & Unsmoothed spectrum of each template, lines removed \cr 7 &
+{\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed
+spectrum of each template, without dust \cr 8 & {\tt rawspec\_nl\_nd}
+& $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of each
+template, without lines or dust \cr 9 & {\tt lspec} &
+$N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr 10 & {\tt
+extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as a
+function of wavelength \cr 11 & {\tt lambda} & $N_{\mathrm{spec}}$ &
+Wavelength grid (\AA) \cr 12 & {\tt sfr} & $N_{\mathrm{age}} \times
+N_t$ & Star-formation rate ($M_\odot$ per year) as a function of time
+\cr 13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
+metallicity as a function of time \cr 14 & {\tt ages} &
+$N_{\mathrm{age}}$ & Age grid (yrs) \cr 15 & {\tt dage} &
+$N_{\mathrm{age}}$ & Age bin size (yrs); {\tt sfr}$\times${\tt dage}
+$=$ total star-formation in each bin \cr 16 & {\tt mass} & $N_t$ &
+total mass formed in each template \cr 17 & {\tt mremain} & $N_t$ &
+total current stellar mass in each template \cr 18 & {\tt mets} &
+$N_t$ & average metallicity of current stars in each template \cr 19 &
+{\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in
+each template \cr 20 & {\tt m1000} & $N_t$ & total stellar mass formed
+in last 1 Gyr in each template \cr 21 & {\tt basis\_ages} &
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+ages of each stellar population basis vector \cr 22 & {\tt
+basis\_mets} & $N_{\mathrm{age}} \times N_{\mathrm{mets}} \times
+N_{\mathrm{dust}}$ & metallicity of each stellar population basis
+vector \cr 23 & {\tt basis\_dusts} & $N_{\mathrm{age}} \times
+N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ & dust properties of each
+stellar population basis vector \cr 24 & {\tt basis\_mremain} &
+$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
+fraction of original stellar mass surviving for each basis vector \cr
+\enddata 
+\tablecomments{All quantities are in floating point (four byte)
+precision, except {\tt dusts} which is an array of structures, whose
+format is described in text.}
+\end{deluxetable}
+
+\clearpage
+
+\setcounter{thefigs}{0}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f1.eps}
+\caption{\label{spec_lrg} Best fit LRG spectral template in the rest
+frame (normalization is for a 1 $M_\odot$ galaxy located 10 pc away,
+or equivalently a $10^{12}$ $M_\odot$ galaxy located 10 Mpc away).}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f2.eps}
+\caption{\label{sfh_lrg} Star-formation history corresponding to LRG
+	spectral template of Figure \ref{spec_lrg}. Top panel shows the
+	number of stars formed per logarithmic time interval ($t$ is
+	expressed in years, curve is normalized for a $10^{12}$ $M_\odot$
+	galaxy). Almost all of the stars are formed in the first couple of
+	billion years --- note that the recent ``spike'' is represents a
+	tiny fraction ($\sim 10^{-8}$) of the total number of stars. Bottom
+	panel shows the mean metallicity of the population as a function of
+	time. Note that the details of these functions are rather poorly
+	constrained. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f3.eps}
+\caption{\label{lrg_colors} SDSS colors of LRGs as a function of
+	redshift. The greyscale is the conditional distribution of color
+	within each redshift bin. The thin lines are the 10\%, 25\%, 50\%,
+	75\%, and 90\% quantiles of the distribution. The thick line
+	is the prediction of the model. The $u$ band is not included in the
+	fit, and the $u$ magnitudes of most LRGs are poorly known. The other
+	colors fit the models reasonably well. This model, remember, is
+	given incredible freedom, meaning that the above agreement is the
+	best one can do with the stellar population synthesis code of
+	\citet{bruzual03a}. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f4.eps}
+\caption{\label{sfh_templates} Similar to Figure \ref{sfh_lrg}, but
+for the five global templates. Again, there are many degeneracies in
+the fit parameters (though there are fewer in the actual spectra), so
+these figures need to be interpreted appropriately.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f5.eps}
+\caption{\label{fullfits} Color residuals (defined explicitly in the
+	text) of GALEX, SDSS, and 2MASS observations relative to our best
+	fit 5-template model.  The greyscale is the conditional distribution
+	of the color residual given the redshift.  The thin lines are the
+	10\%, 25\%, 50\%, 75\%, and 90\% quantiles of the distribution. The
+	thick dashed lines show the estimated 1$\sigma$ uncertainties in the
+	colors from the photometric catalogs. Relative to the uncertainties,
+	there are no significant biases or redshift trends in these fits. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f6.eps}
+\caption{\label{goods} Color residuals in fit using the standard five
+templates to GOODS data, compared to the typical uncertainties (thick
+dashed lines). Note that the fits always do poorly on the $H$ band,
+which we believe to be a catalog error. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f7.eps}
+\caption{\label{goods_special} Same as Figure \ref{goods}, but fitting
+using five templates specially designed for GOODS. These templates
+have smaller residuals in many respects but still fail to fit the $H$
+band data. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f8.eps}
+\caption{\label{specfit} Best fit model spectra based on the five
+template fit to $g$, $r$ and $i$ fluxes, compared to the original
+SDSS spectra from which we computed those fluxes.  The models and the
+original spectra agree very well.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f9.eps}
+\caption{\label{twomass_resid} Similar to Figure \ref{fullfits} but
+for galaxies observed in both SDSS and 2MASS and only using SDSS and
+2MASS bands. The fits are to the SDSS and 2MASS data together.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f10.eps}
+\caption{\label{twomass_predicted} Similar to Figure
+\ref{twomass_resid} but now the fits are {\it only} to the SDSS
+bands. The residuals in the 2MASS bands remain very small, indicating
+that the 2MASS measurements do not add a lot of information about
+these galaxies.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f11.eps}
+\caption{\label{galex_resid} Similar to Figure \ref{twomass_resid} but
+for galaxies observed in both SDSS and GALEX and only using SDSS and
+GALEX bands. The fits are to the SDSS and GALEX data together.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f12.eps}
+\caption{\label{galex_predicted} Similar to Figure
+\ref{galex_resid} but now the fits are {\it only} to the SDSS
+bands. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f13.eps}
+\caption{\label{galex_kcorrect} $K$-corrections as a function of
+redshift in the GALEX near (N) and far (F) UV bands. }
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f14.eps}
+\caption{\label{twomass_kcorrect} $K$-corrections as a function of
+redshift in the 2MASS $J$, $H$ and $K_s$ bands.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f15.eps}
+\caption{\label{sdss_kcorrect} $K$-corrections as a function of
+redshift in the SDSS $u$, $g$, $r$, $i$, and $z$ bands.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f16.eps}
+\caption{\label{main_kcorrect} Same as Figure \ref{sdss_kcorrect}, but
+$K$-correcting to the \band{0.1}{u}, \band{0.1}{g}, \band{0.1}{r},
+\band{0.1}{i}, and \band{0.1}{z} bands.}  
+\end{figure}
+
+\clearpage 
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f17.eps}
+\caption{\label{mass_to_garching} Our galaxy stellar mass estimates
+	$M_\ast$ compared to those of \citet{kauffmann03a} $M_{s,\ast}$, as
+	a function of stellar mass (top panel) and of color (bottom
+	panel). The greyscale is the conditional distribution
+	$M_{s,\ast}/M_\ast$ on each quantity . The lines are the 10\%, 25\%,
+	50\%, 75\% and 90\% quantiles.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f18.eps}
+\caption{\label{mtol} Mass-to-light ratios of galaxies in the $V$ band
+	(in solar units) as a function of galaxy $B-V$ color. The solid line
+	satisfies the relationship $\log_{10}(M/L_V) = 1.40 (g-r) - 0.73$,
+	given by \citet{bell01b} for their sample of spiral galaxies. 
+	Their estimates and ours agree for $B-V < 0.8$, where spiral
+	galaxies dominate the galaxy population.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f19.eps}
+\caption{\label{umr_bg} Fraction of the total star-formation that has
+	occurred in the previous billion years $b_G$, as a function of restframe
+	$u-r$ color, for SDSS galaxies. The greyscale is the conditional
+	distribution of $b_G$ on $u-r$. The lines are the 10\%, 25\%, 50\%,
+	75\% and 90\% quantiles. For $u-r < 2.5$, the median relationship
+	follows the simple form listed in the figure. Galaxies with $u-r >
+	2.5$ are often highly reddened star-forming galaxies.}
+\end{figure}
+
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{f20.eps}
+\caption{\label{lemma} Example diagram of definition of $\phi$ in the
+	text. $\chi^2$ represents the actual $\chi^2$ while $\phi$ is
+	designed such that it exceeds $\chi^2$ except at the single point
+	$W$. By minimizing $\phi$ we can find a point $W^\ast$ that we know
+	to have an equal or better $\chi^2$ than that at $W$. }
+\end{figure}
 

### 2006-06-07 docs/paper/kcorrect_figures.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@553 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \end{figure}
+
+\clearpage
+\stepcounter{thefigs}
+\begin{figure}
+\figurenum{\fignum}
+\plotone{lemma.eps}
+\caption{\label{lemma} Example diagram of definition of $\phi$ in the
+	text. $\chi^2$ represents the actual $\chi^2$ while $\phi$ is
+	designed such that it exceeds $\chi^2$ except at the single point
+	$W$. By minimizing $\phi$ we can find a point $W^\ast$ that we know
+	to have an equal or better $\chi^2$ than that at $W$. }
+\end{figure}

### 2006-06-26 docs/paper/kcorrect.tex
typo
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@557 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \end{equation}
-where $\mathrm{DM}(z) = 25 - 5\log_{10} (d_L / (h^{-1}{\mathrm{~Mpc}}))$ is the
+where $\mathrm{DM}(z) = 25 + 5\log_{10} (d_L / (h^{-1}{\mathrm{~Mpc}}))$ is the
 bolometric distance modulus calculated from the luminosity distance
@@
 \end{equation}
-where $\mathrm{DM}(z) = 25 - 5\log_{10} (d_L /
+where $\mathrm{DM}(z) = 25 + 5\log_{10} (d_L /
 (h^{-1}{\mathrm{~Mpc}}))$ is the bolometric distance modulus

### 2006-08-01 docs/paper/kcorrect.tex
update docs
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@558 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 data uncertainties and missing data, which allows it to ignore that
-variation which is due purely to statistical errors.  Third, it
-handles the complications of observing galaxy spectra photometrically
-using broad-band filters of galaxies at varying redshifts, whereas
-standard PCA requires a constant grid of observations in rest-frame
-wavelength space.
+variation which is due purely to statistical errors. PCA, by
+constrast, does not use any information about the known uncertainties
+and is therefore left free to assign templates to express variations
+that are purely due to errors. Third, it handles the complications of
+observing galaxy spectra photometrically using broad-band filters of
+galaxies at varying redshifts, whereas standard PCA requires a
+constant grid of observations in rest-frame wavelength space.
 
@@
 \citet{chabrier03a} stellar initial mass function and the Padova1994
-isochrones. We select 6 metallicities (0.005, 0.02, 0.2, 0.4, 1, and
-2.5 times solar). For each metallicity we select 25 ages (between 1
-Myr and 13.75 Gyr, spaced almost logarithmically in age).  For each
-age and metallicity we make 3 choices of dust model: (1) no dust
-extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction (as
-input into the models of
-\citealt{witt00a} with a ``homogeneous'' distribution and ``shell''
-geometry); (3) $\tau_V = 3$ dust with SMC type extinction (from the
-same models). We smooth each template to $300$ km s$^{-1}$ velocity
-dispersion (we will smooth the SDSS spectra in the training set to the
-same resolution).
+isochrones. We use all 6 metallicities (mass fractions of elements
+heavier than He of $Z= 0.0001$, $0.0004$, $0.004$, $0.008$, $0.02$ and
+$0.05$).  For each metallicity we select 25 ages (between 1 Myr and
+13.75 Gyr, spaced almost logarithmically in age).  We smooth each
+template to $300$ km s$^{-1}$ velocity dispersion (we will smooth the
+SDSS spectra in the training set to the same resolution).
+
+There are limitations to all stellar population synthesis libraries,
+and our results are therefore restricted to be at best as good as our
+input models. There are some well-known areas of uncertainty in the
+stellar populations models in the UV and IR. In the UV, old
+populations sometimes contain extreme horizontal branch stars that
+have lost their envelopes and are thus extremely hot, that dominate
+the flux at $\lambda \sim 1500$ \AA. There is certainly no physical
+model that predicts when this happens, and no consensus about how
+these stars should be accounted for in isochrone synthesis (that is,
+what fraction of stars undergo this phase as a function of mass and
+metallicity). In the near-IR, the spectra of thermally-pulsating
+asymptotic giant branch stars may dominate the flux in certain
+intermediate age populations (2--3 Gyr). Again, the fraction of stars
+in this phase and what their spectra are is under dispute. {\bf what
+does BC03 do?}  We do not address these uncertainties here, though we
+note that for low redshift galaxies our results from BC03 do explain
+most galaxies' colors (though see below for our discussion of the high
+redshift GOODS sample). {\bf refs}
+
+For each age and metallicity we make 3 choices of dust model: (1) no
+dust extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction
+(as input into the models of \citealt{witt00a} with a ``homogeneous''
+distribution and ``shell'' geometry); (3) $\tau_V = 3$ dust with SMC
+type extinction (from the same models).
 
@@
 8 Myr old continuous star-formation history with 5 possible
-metalliciticites (0.5, 0.2, 0.4, 1.0, and 2. times solar) and 7
+metalliciticites ($Z=0.001$, 0.004, 0.008, 0.02, and 0.04) and 7
 possible ionization parameters ($q= 5\times 10^6$, $10^7$, $2\times
@@
 remove Ly$\alpha$, because it is generally much larger than observed
-in real galaxies.
+in real galaxies. We warn readers that galaxies exhibit many
+individual variations in equivalent widths of emission lines --- our
+final set of five templates will not have any hope of following these
+variations in detail.
 
 Let us refer to these 485 basis templates as $M_{j}(\lambda)$,
-expressed in units of ergs s$^{-1}$ \AA$^{-1}$.
-
-We seek to reduce this full basis space to a subspace in which
-galaxies actually exist. In particular, we seek five templates
-$F_{i}(\lambda)$ built from nonnegative combinations of the
-original basis set of $N$ templates:
+expressed in units of ergs s$^{-1}$ \AA$^{-1}$.  We seek to reduce
+this full basis space to a subspace in which galaxies actually
+exist. In particular, we seek five templates $F_{i}(\lambda)$ built
+from nonnegative combinations of the original basis set of $N$
+templates:
 \begin{equation}
@@
 exist in \citet{stoughton02a}.  An automated pipeline called {\tt
-idlspec2d} (in this case {\tt v4\_9}) measures the redshifts and
-classifies the reduced spectra (Schlegel et al., in preparation). We
-use the Petrosian magnitudes for the SDSS data (except where noted
-below for the LRG-only sample).
+idlspec2d}\footnote{Available on the World Wide Web at {\tt
+http://skymaps.info}} (in this case {\tt v4\_9}) measures the
+redshifts and classifies the reduced spectra (Schlegel et al., in
+preparation). We use the Petrosian magnitudes for the SDSS data
+(except where noted below for the LRG-only sample). Using the measured
+velocity dispersions of the spectra from {\tt idlspec2d}, we further
+smooth all spectra in the training sets such that each spectrum has a
+dispersion of 300 km s$^{-1}$, to ensure consistency of resolution
+among all the spectra and models.
 
@@
 We begin with the simplest case, which is fitting a {\it single}
-template to the photometric data of the LRG sample
+template to just the SDSS photometric data of the LRG sample
 (\citealt{eisenstein01a}). For many of these galaxies, which extend to
 $r<19.5$ and are intrinsically red, the $u$-band flux is extremely
-poorly measured, so we ignore the $u$-band for all LRGs.
+poorly measured, so we ignore the $u$-band for all LRGs and only fit
+to $g$, $r$, $i$ and $z$ fluxes using the SDSS ``model'' fluxes as
+described in \cite{stoughton01a} and elsewhere.
+In this case, we fit for the $a_{i0}$ and $b_{0j}$ coefficients 
+in Equation \ref{chi2}: we want just a single template, expressed 
+by the $b_{0j}$ coefficients, and a single scaling of its 
+flux for each galaxy, expressed by $a_{i0}$.
 
 Figure \ref{spec_lrg} shows the spectrum of the best-fit LRG.  This
-template is constrained by data between about 2000 \AA\ and 10000 \AA;
-outside that range it is an extrapolation.
+template is constrained by data between about 3000 \AA\ and 10000 \AA.
+Outside that range it is an extrapolation --- that is, it is simply
+the best fit model prediction, but there is no data to verify the
+model there.
 
 Figure \ref{sfh_lrg} shows the star-formation history corresponding to
-this best-fit LRG spectrum. The top panel shows the star-formation per
+this best-fit LRG spectrum. 
+The top panel shows the star-formation per
 unit time as a function of look-back time.  The bottom panel shows the
@@
 
-We next apply the method to the full data set and fit for five
-templates. From experiments, we found that we needed five to
-adequately explain the data. More than five templates would allow us
-to explain the data better, but only marginally so.  The right-most
-column of Figure \ref{sfh_templates} shows the spectrum from the
-ultraviolet through the near infrared for the resulting
-templates. Note that there is a very old template, a very young
-template, and several intermediate templates, including one which is
-close to that of an A star.
+We next apply the method to the full data set (that is, all five items
+listed in Section \ref{data}) and fit for five templates. From
+experiments, we found that we needed five to adequately explain the
+data. More than five templates would allow us to explain the data
+better, but only marginally so.  The right-most column of Figure
+\ref{sfh_templates} shows the spectrum from the ultraviolet through
+the near infrared for the resulting templates. Note that there is a
+very old template, a very young template, and several intermediate
+templates, including one which is close to that of an A star.
 
@@
 observations with respect to the best fits, when fitting to galaxies
-with GALEX, SDSS, and 2MASS data. In this context we define the color
+with GALEX, SDSS, and 2MASS data (using analogous sets of data to 
+those described in Section \ref{data}, but in detail a different
+set of galaxies). In this context we define the color
 residuals as (for example):
@@
 galaxy; and the fraction of the total amount of star-formation that
-has occurred recently.
+has occurred recently. We do not believe that fitting just to
+broad-band colors can ever give extremely accurate values for these
+quantities.  However, in this section
 
@@
 
+%The strength of the technique we have used here is that one can
+%quickly and reliably whittle down a large parameter space of models to
+%the few templates one needs to explain a particular set of data. We
+%should note that when fitting models, the most difficult part is
+%actually starting with an accurate set of models that are capable of
+%explaining the data at all. Here we have used the \citet{bruzual03a}
+%models, that do a reasonable but not perfect job of explaining the
+%available data.
+
 \acknowledgments
@@
 W. Hogg, John Moustakas, Leonidas Moustakas, David Schiminovich, Risa
-Wechsler, and Andrew West for their detailed comments and
+Wechsler, Andrew West, and the anonymous referee for their detailed comments and
 feedback. This work was funded by a GALEX Archival Program (38) and an

### 2006-08-01 docs/paper/kcorrect.tex
paper changes
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@560 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 what fraction of stars undergo this phase as a function of mass and
-metallicity). In the near-IR, the spectra of thermally-pulsating
-asymptotic giant branch stars may dominate the flux in certain
-intermediate age populations (2--3 Gyr). Again, the fraction of stars
-in this phase and what their spectra are is under dispute. {\bf what
-does BC03 do?}  We do not address these uncertainties here, though we
+metallicity). While the models of \cite{bruzual03a} do contain
+post-AGB and planetary nebula phases that provide FUV flux in old
+populations, they do not contain extreme horizontal branch stars or
+other candidates that might provide more of the ``UV-excess'' seen in
+elliptical galaxies (\citealt{greggio90a, brown00a}).  In the near-IR,
+the spectra of thermally-pulsating asymptotic giant branch (TP-AGB)
+stars may dominate the flux in certain intermediate age populations
+(2--3 Gyr). With respect to these phases, \citet{bruzual03a} use
+models that are calibrated to observations of TP-AGB stars
+nearby. \citet{maraston05a} show that a different choice of TB-AGB
+models can produce significantly higher flux in the NIR at
+intermediate ages. Thus, the choices for AGB and post-AGB evolution we
+make here by using the models \cite{bruzual03a} are not definitive,
+but we do not address these controversies here and they remain an
+uncertainty in the physical interpretation of galaxy fluxes.  We do
 note that for low redshift galaxies our results from BC03 do explain
 most galaxies' colors (though see below for our discussion of the high
-redshift GOODS sample). {\bf refs}
+redshift GOODS sample).
 
@@
 conclude that {\it some} of the errors are irreducible, and require
-either a larger set of templates for high redshift galaxies (a
-possibility we find a priori unlikely), or that there are simply
-errors in the catalogs or the input filter curves. For example, in the
-$z$ and $V$ bands we find significant trends between redshifts $z=0.5$
-and $1.5$ (echoed in Figure \ref{goods}). More significantly, in the
-$H$ band there is at least a 10\% offset in the magnitudes across all
-redshifts, which is almost certainly a catalog or filter curve
-error. Other errors (such as a large scatter in the residuals in the
-$B$ band and some redshift dependence in the $J$ band) are clearly due
-to the fact that the templates are not primarily designed for GOODS
-data. All in all, we recommend using the special templates for
-$K$-corrections within the GOODS data set.
+either a different set of templates for high redshift galaxies or that
+there are simply errors in the catalogs or the input filter
+curves. For example, in the $z$ and $V$ bands we find significant
+trends between redshifts $z=0.5$ and $1.5$ (echoed in Figure
+\ref{goods}). More significantly, in the $H$ band there is at least a
+10\% offset in the magnitudes across all redshifts, which is almost
+certainly a catalog or filter curve error. One possibility for these
+problems in the red are that intermediate age populations have flux in
+the near-IR from TP-AGB that our model omits; for example,
+\citet{maraston05a} has shown that such stars can significantly change
+the near-IR colors. We defer a fuller investigation of these
+discrepancies.  Other errors (such as a large scatter in the residuals
+in the $B$ band and some redshift dependence in the $J$ band) are
+clearly due to the fact that the templates are not primarily designed
+for GOODS data. All in all, we recommend using the special templates
+for $K$-corrections within the GOODS data set.
 
@@
 of predicting the 2MASS fluxes, with a scatter of 20--30\% --- not far
-from the uncertainties in the 2MASS fluxes themselves. Of course, it
-is not surprising that it is easy to predict the 2MASS fluxes ---
-stellar spectra are very simple in the NIR, so having the $ugriz$-band
-fluxes from SDSS yields much information regarding the redder
-bandpasses.
-
-However, we will note here that this result flies in the face of a
-persistent insistence that NIR observations are necessary to measure
-stellar masses of galaxies. If we can predict the NIR observations
-themselves, clearly they cannot be adding much to our knowledge of the
-underlying stellar mass. This fact changes one's decisions about what
-data is best to use for calculating the stellar mass function.  Basing
-it on 2MASS data only improves slightly the stellar mass estimate of
-each galaxy, while restricting (and thus biasing) the sample
-significantly at the lowest luminosities and surface brightnesses
-relative to the SDSS. Of course, if at high redshift galaxies have
-a very different locus of their SEDs, then perhaps this statement is
-less true for them.
+from the uncertainties in the 2MASS fluxes themselves.
+
+However, we will note here that this result flies in the face of the
+view of many that NIR observations are necessary to measure stellar
+masses of galaxies. If we can predict the NIR observations themselves,
+clearly they cannot be adding much to our knowledge of the underlying
+stellar mass. This fact changes one's decisions about what data is
+best to use for calculating the stellar mass function.  Basing it on
+2MASS data only improves slightly the stellar mass estimate of each
+galaxy, while restricting (and thus biasing) the sample significantly
+at the lowest luminosities and surface brightnesses relative to the
+SDSS. Of course, if at high redshift galaxies have a very different
+locus of their SEDs (for example, as perhaps is true of the GOODS
+galaxies in the NIR), then perhaps this statement is less true for
+them.
 
@@
 broad-band colors can ever give extremely accurate values for these
-quantities.  However, in this section
+quantities. In this section we demonstrate consistency between our
+results and those of other groups for the same galaxies.
 
@@
 \end{equation}
-Note, however, the large scatter (about 0.3 dex at 1$\sigma$).
+Note, however, the large scatter (about 0.3 dex at 1$\sigma$).  The
+red dots show the results of using the formula of \citet{hopkins03a}
+to infer the current star-formation rate from the $u$-band,
+integrating over 1 Gyr, scaling by the integrated stellar mass from
+the {\tt kcorrect} determination, and then multiplying by four (which
+is the minimum appropriate correction for dust extinction in $u$
+according to \citealt{hopkins03a}). Our results roughly scale as those
+of \citet{hopkins03a} do.
 
@@
 detailed spectroscopic analysis can provide better constraints on
-star-formation histories and stellar masses.
+star-formation histories and stellar masses.  However, the results
+from {\tt kcorrect} are roughly in agreement with what other authors
+find with comparable data.
 

### 2006-08-03 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@562 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 In addition, we present software for using such data to estimate
-$K$-corrections and stellar masses.
+$K$-corrections.
 \end{abstract}
@@
 factorization (NMF; \citealt{lee00a}), and the extensions to it we
-describe here, is a particularly useful variant on principal component
-analysis. Second, the nonnegative least squares algorithm of
+describe here, is a particularly useful variant on principal
+components
+analysis (PCA). Second, the nonnegative least squares algorithm of
 \citet{sha02a} has the virtue of being extremely simple to implement.
@@
 
-We also release the templates in electronic form as well as an
-implementation of the methods used to fit the templates to data. This
-software {\tt kcorrect v4\_1} is distributed on the World Wide
-Web{\footnote {\tt http://cosmo.nyu.edu/blanton/kcorrect/}}. It
-consists of a core C library which performs most of the complex and
-computationally intensive tasks, plus an IDL library that provides a
-high-level interface.  This software is an update of two major earlier
-releases ({\tt v1\_16}; \citealt{blanton03b}; {\tt v3\_2}). The
-improvement over the previous version is twofold. First, we have
-improved the templates such that they successfully fit galaxies in the
-restframe UV, as observed by GALEX, DEEP2, and GOODS. Second, because
-the templates are now completely model-based, the fits have a physical
+We also release the templates in electronic form and an implementation
+of the methods used to fit the templates to data. This software {\tt
+kcorrect v4\_1} is distributed on the World Wide Web{\footnote {\tt
+http://cosmo.nyu.edu/blanton/kcorrect/}}. It consists of a core C
+library which performs most of the complex and computationally
+intensive tasks, plus an IDL library that provides a high-level
+interface.  This software is an update of two major earlier releases
+({\tt v1\_16}; \citealt{blanton03b}; {\tt v3\_2}). The improvement
+over the previous version is twofold. First, we have improved the
+templates such that they successfully fit galaxies in the restframe
+UV, as observed by GALEX, DEEP2, and GOODS. Second, because the
+templates are now completely model-based, the fits have a physical
 interpretation in terms of a star-formation history. The IDL library
@@
 From the spectroscopic observations we know that galaxy spectra reside
-in a low-dimensional subspace. Principal Component Analysis (PCA) of
-galaxy spectra, introduced by \citet{connolly95b} and applied many
-times since then, demonstrates that most of the variance in the
-distribution of galaxies in spectral space can be explained using a
-few templates. This means that, in the linear space of all possible
-spectra, galaxies exist in only a small subspace. Therefore, even with
-very heterogeneous data, we should be able to determine the properties
-of this subspace.
+in a low-dimensional subspace of all possible spectra. Principal
+Component Analysis (PCA) of galaxy spectra, introduced by
+\citet{connolly95b} and applied many times since then, demonstrates
+that most of the variance in the distribution of galaxies in spectral
+space can be explained using a few templates. This means that, in the
+linear space of all possible spectra, galaxies exist in only a small
+subspace. Therefore, even with very heterogeneous data, we should be
+able to determine the properties of this subspace.
 
@@
 constrains the problem appropriately and yields a natural theoretical
-interpretation of the results in terms of star-formation histories.
+interpretation of the results in terms of star-formation
+histories. Generalizing this approach to use other models of emission
+from stars and gas is straightforward.
 
@@
 to $g$, $r$, $i$ and $z$ fluxes using the SDSS ``model'' fluxes as
-described in \cite{stoughton01a} and elsewhere.
+described in \cite{stoughton02a} and elsewhere.
 In this case, we fit for the $a_{i0}$ and $b_{0j}$ coefficients 
@@
 broad-band colors can ever give extremely accurate values for these
-quantities. In this section we demonstrate consistency between our
-results and those of other groups for the same galaxies.
+quantities, but to evaluate what {\tt kcorrect} does in this regard we
+here examine the consistency between our results and those of other
+groups for the same galaxies.
 
@@
 rate. However, we can try to measure what fraction of the total
-star-formation has occurred in, say, the past 1 Gyr:
+star-formation has occurred in, say, the past 300 Myr:
 \begin{equation}
-b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
+b_{300} = \frac{\int_0^{300\mathrm{~Myr}} dt\, \mathrm{SFR}(t)}
 {\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
@@
 parameter $b$ in \citet{kennicutt94a} and references therein.  Note
-that here $b_G$ is defined in terms of the total integrated
+that here $b_{300}$ is defined in terms of the total integrated
 star-formation history (unlike the stellar masses referred to above)
@@
 relationship between the rest-frame $u-r$ color for SDSS galaxies and
-$b_G$. For galaxies with $u-r < 2.5$ (redder than which galaxies are
-mostly reddened, edge-on disks) these two parameters follow a simple
-relationship:
+$b_{300}$. For our results, the median relationship is close to:
 \begin{equation}
-\log_{10} b_g = -0.55 (u-r). 
+\log_{10} b_{300} = 0.2 - 1.3 (u-r). 
 \end{equation}
 Note, however, the large scatter (about 0.3 dex at 1$\sigma$).  The
-red dots show the results of using the formula of \citet{hopkins03a}
+overlaid points show the results of using the formula of \citet{hopkins03a}
 to infer the current star-formation rate from the $u$-band,
-integrating over 1 Gyr, scaling by the integrated stellar mass from
-the {\tt kcorrect} determination, and then multiplying by four (which
-is the minimum appropriate correction for dust extinction in $u$
-according to \citealt{hopkins03a}). Our results roughly scale as those
-of \citet{hopkins03a} do.
+integrating over 300 Myr, scaling by the integrated stellar mass from
+the {\tt kcorrect} determination, and then correcting for the internal
+$u$-band dust extinction inferred by the {\tt kcorrect} fits (which
+turns out to be factors ranging from 3--10, similar to the factors
+used by \citealt{hopkins03a}).  Our results roughly scale as those of
+\citet{hopkins03a} do, but with lower star-formation rates by a factor
+of five or so. Since \citet{hopkins03a} calibrate their star-formation
+rates to a number of other indicators, we conclude that either {\tt
+kcorrect} probably underestimates the recent star-formation rate.
 
@@
 The standard non-negative matrix factorization (NMF) problem, as
-originally posed by \cite{lee00a}, is to approximate a data
-matrix $\XX$ (of size $N_k\times N_n$) as the outer product of two
-rank-$N_i$ non-negative matrices $\WW$ and $\HH$. For example, in this
-paper, $\XX$ represents the fluxes at $N_n$ different wavelengths for
-$N_k$ different galaxies, $\HH$ represents a set of $N_i$ templates to build
+originally posed by \cite{lee00a}, is to approximate a data matrix
+$\XX$ (of size $N_k\times N_n$) as the outer product of two rank-$N_i$
+non-negative matrices $\WW$ and $\HH$. For example, in this paper,
+$\XX$ represents the fluxes at $N_n$ different wavelengths for $N_k$
+different galaxies, $\HH$ represents a set of $N_i$ templates to build
 each galaxy from, and $\WW$ represents, for each galaxy, the weights
 to give each template. This factorization is similar in spirit to
-singular value decomposition (SVD) and principal components analysis
-(PCA) but crucially, in NMF, $\XX$, $\WW$ and $\HH$ are all
-constrained to have non-negative entries. This changes the problem
-fundamentally, since no ``cancellation'' of positive and negative
-basis functions or coefficients is possible -- all approximation
-interactions are strictly additive. The goal, as with SVD or PCA, is
-to minmize the squared approximation error (Frobenius norm):
-\beqa \label{eq:nmfcost}
+singular value decomposition (SVD) and PCA, but crucially, in NMF,
+$\XX$, $\WW$ and $\HH$ are all constrained to have non-negative
+entries. This changes the problem fundamentally, since no
+``cancellation'' of positive and negative basis functions or
+coefficients is possible -- all approximation interactions are
+strictly additive. The goal, as with SVD or PCA, is to minmize the
+squared approximation error (Frobenius norm): \beqa \label{eq:nmfcost}
 \chi^2 &= \|\XX-\WW\HH\|^2\\

### 2006-08-03 docs/paper/kcorrect_figures.tex
paper changes
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@560 75dff457-8751-0410-ab1a-cd701654ce6b
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@562 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \caption{\label{umr_bg} Fraction of the total star-formation that has
-	occurred in the previous billion years $b_G$, as a function of restframe
-	$u-r$ color, for SDSS galaxies. The greyscale is the conditional
-	distribution of $b_G$ on $u-r$. The lines are the 10\%, 25\%, 50\%,
-	75\% and 90\% quantiles. For $u-r < 2.5$, the median relationship
-	follows the simple form listed in the figure. Galaxies with $u-r >
-	2.5$ are often highly reddened star-forming galaxies.}
+	occurred in the previous 300 Myr $b_{300}$, as a function of
+	restframe $u-r$ color, for SDSS galaxies. The greyscale is the
+	conditional distribution of $b_{300}$ on $u-r$. The lines are the 10\%,
+	25\%, 50\%, 75\% and 90\% quantiles.  For comparison we show the
+	results of using the scaling of \citet{hopkins03a}.}
 \end{figure}

### 2006-08-17 docs/paper/kcorrect-aj-v2/ms.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@563 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2006-06-07 docs/paper/kcorrect-aj-v1/ms.tex
@@
 \citet{chabrier03a} stellar initial mass function and the Padova1994
-isochrones. We select 6 metallicities (0.005, 0.02, 0.2, 0.4, 1, and
-2.5 times solar). For each metallicity we select 25 ages (between 1
-Myr and 13.75 Gyr, spaced almost logarithmically in age).  For each
-age and metallicity we make 3 choices of dust model: (1) no dust
-extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction (as
-input into the models of
-\citealt{witt00a} with a ``homogeneous'' distribution and ``shell''
-geometry); (3) $\tau_V = 3$ dust with SMC type extinction (from the
-same models). We smooth each template to $300$ km s$^{-1}$ velocity
-dispersion (we will smooth the SDSS spectra in the training set to the
-same resolution).
+isochrones. We use all 6 metallicities (mass fractions of elements
+heavier than He of $Z= 0.0001$, $0.0004$, $0.004$, $0.008$, $0.02$ and
+$0.05$).  For each metallicity we select 25 ages (between 1 Myr and
+13.75 Gyr, spaced almost logarithmically in age).  We smooth each
+template to $300$ km s$^{-1}$ velocity dispersion (we will smooth the
+SDSS spectra in the training set to the same resolution).
+
+There are limitations to all stellar population synthesis libraries,
+and our results are therefore restricted to be at best as good as our
+input models. There are some well-known areas of uncertainty in the
+stellar populations models in the UV and IR. In the UV, old
+populations sometimes contain extreme horizontal branch stars that
+have lost their envelopes and are thus extremely hot, that dominate
+the flux at $\lambda \sim 1500$ \AA. There is certainly no physical
+model that predicts when this happens, and no consensus about how
+these stars should be accounted for in isochrone synthesis (that is,
+what fraction of stars undergo this phase as a function of mass and
+metallicity). While the models of \cite{bruzual03a} do contain
+post-AGB and planetary nebula phases that provide FUV flux in old
+populations, they do not contain extreme horizontal branch stars or
+other candidates that might provide more of the ``UV-excess'' seen in
+elliptical galaxies (\citealt{greggio90a, brown00a}).  In the near-IR,
+the spectra of thermally-pulsating asymptotic giant branch (TP-AGB)
+stars may dominate the flux in certain intermediate age populations
+(2--3 Gyr). With respect to these phases, \citet{bruzual03a} use
+models that are calibrated to observations of TP-AGB stars
+nearby. \citet{maraston05a} show that a different choice of TB-AGB
+models can produce significantly higher flux in the NIR at
+intermediate ages. Thus, the choices for AGB and post-AGB evolution we
+make here by using the models \cite{bruzual03a} are not definitive,
+but we do not address these controversies here and they remain an
+uncertainty in the physical interpretation of galaxy fluxes.  We do
+note that for low redshift galaxies our results from BC03 do explain
+most galaxies' colors (though see below for our discussion of the high
+redshift GOODS sample).
+
+For each age and metallicity we make 3 choices of dust model: (1) no
+dust extinction; (2) $\tau_V = 3$ dust with Milky Way type extinction
+(as input into the models of \citealt{witt00a} with a ``homogeneous''
+distribution and ``shell'' geometry); (3) $\tau_V = 3$ dust with SMC
+type extinction (from the same models).
 
@@
 We begin with the simplest case, which is fitting a {\it single}
-template to the photometric data of the LRG sample
+template to just the SDSS photometric data of the LRG sample
 (\citealt{eisenstein01a}). For many of these galaxies, which extend to
 $r<19.5$ and are intrinsically red, the $u$-band flux is extremely
-poorly measured, so we ignore the $u$-band for all LRGs.
+poorly measured, so we ignore the $u$-band for all LRGs and only fit
+to $g$, $r$, $i$ and $z$ fluxes using the SDSS ``model'' fluxes as
+described in \cite{stoughton02a} and elsewhere.
+In this case, we fit for the $a_{i0}$ and $b_{0j}$ coefficients 
+in Equation \ref{chi2}: we want just a single template, expressed 
+by the $b_{0j}$ coefficients, and a single scaling of its 
+flux for each galaxy, expressed by $a_{i0}$.
 
 Figure \ref{spec_lrg} shows the spectrum of the best-fit LRG.  This
-template is constrained by data between about 2000 \AA\ and 10000 \AA;
-outside that range it is an extrapolation.
+template is constrained by data between about 3000 \AA\ and 10000 \AA.
+Outside that range it is an extrapolation --- that is, it is simply
+the best fit model prediction, but there is no data to verify the
+model there.
 
 Figure \ref{sfh_lrg} shows the star-formation history corresponding to
-this best-fit LRG spectrum. The top panel shows the star-formation per
+this best-fit LRG spectrum. 
+The top panel shows the star-formation per
 unit time as a function of look-back time.  The bottom panel shows the
@@
 galaxy; and the fraction of the total amount of star-formation that
-has occurred recently.
+has occurred recently. We do not believe that fitting just to
+broad-band colors can ever give extremely accurate values for these
+quantities, but to evaluate what {\tt kcorrect} does in this regard we
+here examine the consistency between our results and those of other
+groups for the same galaxies.
 
@@
 relationship between the rest-frame $u-r$ color for SDSS galaxies and
-$b_G$. For galaxies with $u-r < 2.5$ (redder than which galaxies are
-mostly reddened, edge-on disks) these two parameters follow a simple
-relationship:
+$b_{300}$. For our results, the median relationship is close to:
 \begin{equation}
-\log_{10} b_g = -0.55 (u-r). 
+\log_{10} b_{300} = 0.2 - 1.3 (u-r). 
 \end{equation}
-Note, however, the large scatter (about 0.3 dex at 1$\sigma$).
+Note, however, the large scatter (about 0.3 dex at 1$\sigma$).  The
+overlaid points show the results of using the formula of \citet{hopkins03a}
+to infer the current star-formation rate from the $u$-band,
+integrating over 300 Myr, scaling by the integrated stellar mass from
+the {\tt kcorrect} determination, and then correcting for the internal
+$u$-band dust extinction inferred by the {\tt kcorrect} fits (which
+turns out to be factors ranging from 3--10, similar to the factors
+used by \citealt{hopkins03a}).  Our results roughly scale as those of
+\citet{hopkins03a} do, but with lower star-formation rates by a factor
+of five or so. Since \citet{hopkins03a} calibrate their star-formation
+rates to a number of other indicators, we conclude that either {\tt
+kcorrect} probably underestimates the recent star-formation rate.
 
@@
 
-\begin{thebibliography}{39}
+\begin{thebibliography}{43}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 {Blanton}, M.~R. {et~al.} 2005, \aj, 129, 2562
+
+\bibitem[{{Brown} {et~al.}(2000){Brown}, {Bowers}, {Kimble}, {Sweigart}, \&
+  {Ferguson}}]{brown00a}
+{Brown}, T.~M., {Bowers}, C.~W., {Kimble}, R.~A., {Sweigart}, A.~V., \&
+  {Ferguson}, H.~C. 2000, \apj, 532, 308
 
@@
 
+\bibitem[{{Greggio} \& {Renzini}(1990)}]{greggio90a}
+{Greggio}, L. \& {Renzini}, A. 1990, \apj, 364, 35
+
 \bibitem[{{Hayes}(1985)}]{hayes85a}
@@
 Hogg, D.~W. {et~al.} 2002, \aj, 124, 646
+
+\bibitem[{{Hopkins} {et~al.}(2003){Hopkins}, {Miller}, {Nichol}, {Connolly},
+  {Bernardi}, {G{\'o}mez}, {Goto}, {Tremonti}, {Brinkmann}, {Ivezi{\'c}}, \&
+  {Lamb}}]{hopkins03a}
+{Hopkins}, A.~M., {Miller}, C.~J., {Nichol}, R.~C., {Connolly}, A.~J.,
+  {Bernardi}, M., {G{\'o}mez}, P.~L., {Goto}, T., {Tremonti}, C.~A.,
+  {Brinkmann}, J., {Ivezi{\'c}}, {\v Z}., \& {Lamb}, D.~Q. 2003, \apj, 599, 971
 
@@
 
+\bibitem[{{Maraston}(2005)}]{maraston05a}
+{Maraston}, C. 2005, \mnras, 362, 799
+
 \bibitem[{Martin {et~al.}(2005)}]{martin05a}
@@
 
-\clearpage 
 \stepcounter{thefigs}

### 2006-08-21 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@565 75dff457-8751-0410-ab1a-cd701654ce6b

@@
 \documentclass[10pt,preprint]{aastex}
-\usepackage{amsmath}
-\usepackage{epsfig}
 
@@
 \newcommand{\lambdaemit}{\lambda_e}
-
-\def\beq#1\eeq{\begin{equation}#1\end{equation}}
-\def\beqa#1\eeqa{\begin{align}#1\end{align}}
-\def\beqq#1\eeqq{\begin{equation*}#1\end{equation*}}
-\def\beqqa#1\eeqqa{\begin{align*}#1\end{align*}}
-\def\bseq#1\eseq{\begin{subequations}#1\end{subequations}}
 
@@
 strictly additive. The goal, as with SVD or PCA, is to minmize the
-squared approximation error (Frobenius norm): \beqa \label{eq:nmfcost}
-\chi^2 &= \|\XX-\WW\HH\|^2\\
+squared approximation error (Frobenius norm): 
+\begin{eqnarray}
+\label{eq:nmfcost}
+\chi^2 &=& \|\XX-\WW\HH\|^2\\
 %&= \|\XX-\XXh\|^2\\
-&= \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
-\eeqa
+&=& \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
+\end{eqnarray}
 In their seminal papers, \cite{lee00a} show that this approximation
@@
 update rules: 
-\beqa \label{eq:wupdate}
-\Wki &\leftarrow \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
+\begin{eqnarray}
+ \label{eq:wupdate}
+\Wki &\leftarrow& \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
 \label{eq:hupdate}
-\Hin &\leftarrow \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
-\eeqa
+\Hin &\leftarrow& \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
+\end{eqnarray}
 $[M]_{ab}$ denotes the $(ab)$ element of the matrix valued expression
@@
 squared approximation error: 
-\beqa 
-\chi^2 &= \|\XX-\AA\BB\MM\|^2\\
-&= \sum_{kn} \left(\Xkn - \sum_{ij} \Aki\Bij\Mjn \right)^2
-\eeqa
+\begin{eqnarray} 
+\chi^2 &=& \|\XX-\AA\BB\MM\|^2\\
+&=& \sum_{kn} \left(\Xkn - \sum_{ij} \Aki\Bij\Mjn \right)^2
+\end{eqnarray}
 This equation is almost that posed in Equation \ref{chi2}.
@@
 provided above: 
-\beq
+\begin{equation}
 \Aki \leftarrow \Aki 
 \frac{[\XX\MM\T\BB\T]_{ki}}{[\AA\BB\MM\MM\T\BB\T]_{ki}}
-\eeq
+\end{equation}
 By Lee and Seung's original proof, this update is guaranteed (for any
@@
 for the elements of $\BB$:
-\beq \label{eq:bupdate}
+\begin{equation} \label{eq:bupdate}
 \Bij \leftarrow \Bij
 \frac{[\AA\T\XX\MM\T]_{ij}}{[\AA\T\AA\BB\MM\MM\T]_{ij}}
-\eeq
+\end{equation}
 Following the technique outlined in \cite{lee00a}, it can still be shown
@@
 and any vector $\yy$,}
-\beq
+\begin{equation}
 \sum_{ni} \yn \yi \Pni \leq \sum_{ni} \yn^2 \frac{\zi}{\zn} \Pni 
-\eeq
+\end{equation}
 %{\it If the vector $\yy$ has non-negative entries, equality can be
@@
 For the standard NMF problem, we can prove that the approximation
-error \eqref{eq:nmfcost} is non-increasing under \eqref{eq:wupdate} by
-using the above lemma to construct a function $\phi(\WW,\ZZ)$ which is
-an upper bound on the cost $\chi^2(\WW)$ for any non-negative matrix $\ZZ$:
-{\small 
-\beqa 
-\chi^2(\WW) &= \sum_{kni\ell} \Wki\Wkl \Hin\Hln
+error (Equation \ref{eq:nmfcost}) is non-increasing under Equation
+\ref{eq:wupdate} by using the above lemma to construct a function
+$\phi(\WW,\ZZ)$ which is an upper bound on the cost $\chi^2(\WW)$ for
+any non-negative matrix $\ZZ$: {\small
+\begin{eqnarray} 
+\chi^2(\WW) &=& \sum_{kni\ell} \Wki\Wkl \Hin\Hln
 - 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
-\phi(\WW,\ZZ) &= \sum_{kni\ell} \Wki^2\frac{\Zkl}{\Zki} \Hin\Hln 
+\phi(\WW,\ZZ) &=& \sum_{kni\ell} \Wki^2\frac{\Zkl}{\Zki} \Hin\Hln 
 - 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
-\phi(\WW,\ZZ) &\geq \chi^2(\WW) \qquad \forall \: \Wki\geq 0,\: \Zki\geq 0
-\eeqa
+\phi(\WW,\ZZ) 
+&&\geq \chi^2(\WW) \qquad \forall \: \Wki\geq 0,\: \Zki\geq 0
+\end{eqnarray}
 } with equality being achieved when $\ZZ=\WW$. Figure \ref{lemma}
@@
 its first argument:
-\beqa
-\phi(\WW^*,\ZZ) &\leq \phi(\WW,\ZZ) \qquad \forall \: \WW,\ZZ \\
-\Wki^* &= \Zki \frac{\sum_n \Xkn \Hin}{\sum_{n\ell} \Hin\Hln \Zkl}
-\eeqa
-If we set $\ZZ=\WW$, this becomes exactly \eqref{eq:wupdate} and
+\begin{eqnarray}
+\phi(\WW^*,\ZZ) &\leq& \phi(\WW,\ZZ) \qquad \forall \: \WW,\ZZ \\
+\Wki^* &=& \Zki \frac{\sum_n \Xkn \Hin}{\sum_{n\ell} \Hin\Hln \Zkl}
+\end{eqnarray}
+If we set $\ZZ=\WW$, this becomes exactly Equation \ref{eq:wupdate} and
 now we can easily prove the validity of this update rule: 
-\beq
+\begin{equation}
 \chi^2(\WW) = \phi(\WW,\WW) \geq \phi(\WW^*,\WW) \geq \chi^2(\WW^*)
-\eeq
+\end{equation}
 where the first inequality comes from the fact that $\WW^*$ minimizes
@@
 on $\chi^2(\WW^*)$. The proof of validity for the update
-\eqref{eq:hupdate} is analogous by symmetry.
-
-To prove the validity of our update \eqref{eq:bupdate} for our new
-problem, we proceed in a similar fashion, using the lemma twice to
-construct consecutive upper bounds on the cost:
-{\small 
-\beqa 
-\chi^2(\BB) &= \sum_{kni\ell{}js} \Aki\Akl\Bij\Bls\Mjn\Msn\\ \nonumber
-&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
-\phi(\BB,\ZZ,\beta) &= \sum_{kni\ell{}js} \Aki\Akl\Bij^2
+(Equation \ref{eq:hupdate}) is analogous by symmetry.
+
+To prove the validity of our update (Equation \ref{eq:bupdate}) for
+our new problem, we proceed in a similar fashion, using the lemma
+twice to construct consecutive upper bounds on the cost: {\small
+\begin{eqnarray} 
+\chi^2(\BB) &=& \sum_{kni\ell{}js} \Aki\Akl\Bij\Bls\Mjn\Msn\\ \nonumber
+&&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
+\phi(\BB,\ZZ,\beta) &=& \sum_{kni\ell{}js} \Aki\Akl\Bij^2
 \frac{\Zks\betalj}{\Zkj\betaij}\Mjn\Msn\\ \nonumber
-&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
-\phi(\BB,\ZZ,\beta) &\geq \chi^2(\BB) \qquad \forall 
+&&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
+\phi(\BB,\ZZ,\beta) &&\geq \chi^2(\BB) \qquad \forall 
 \: \Bij\geq 0,\: \Zkj\geq 0,\: \betaij\geq 0
-\eeqa
+\end{eqnarray}
 } with equality being achieved when $\ZZ=\AA\BB$ and 
@@
 its first argument:
-\beqa
-\phi(\BB^*,\ZZ,\beta) &\leq \phi(\BB,\ZZ,\beta) \qquad 
+\begin{eqnarray}
+\phi(\BB^*,\ZZ,\beta) &\leq& \phi(\BB,\ZZ,\beta) \qquad 
 \forall \: \BB,\ZZ,\beta \\
-\Bij^* &= \betaij \frac{\sum_{kn} \Xkn\Aki\Mjn}
+\Bij^* &=& \betaij \frac{\sum_{kn} \Xkn\Aki\Mjn}
 {\sum_{kn\ell{}s} \Aki\Akl\frac{\Zks\betalj}{\Zkj}\Mjn\Msn}
-\eeqa
+\end{eqnarray}
 If we set $\ZZ=\AA\BB$ and $\beta=\BB$, this becomes
-exactly \eqref{eq:bupdate} and 
+exactly Equation \ref{eq:bupdate} and 
 we can now prove the validity of this update rule: 
-\beq
+\begin{equation}
 \chi^2(\BB) = \phi(\BB,\AA\BB,\BB) \geq \phi(\BB^*,\AA\BB,\BB) 
 \geq \chi^2(\BB^*) 
-\eeq
+\end{equation}
 where the first inequality comes from the fact that $\BB^*$ minimizes
@@
 Gaussian noise assumption):
-\beq
+\begin{equation}
 \chi^2 = \sum_{kn} \left( \frac{\Xkn - 
 \sum_i \Wki\Hin}{\skn}\right)^2
-\eeq
+\end{equation}
 
@@
 In particular, when updating $\WW$, we can rewrite the cost function as
-\beq
+\begin{equation}
 \chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
@@
 %\sum_k \Wki\HHt(i)_{kj}\right)^2
-\eeq
+\end{equation}
 yielding updates of the form:
-\beq
+\begin{equation}
 \Wki \leftarrow \Wki 
@@
 %\Wki \leftarrow \Wki \frac{[\XXt\HHt(i)\T]_{ik}}{[\WW\HHt(i)\HHt(i)\T]_{ik}}
-\eeq
+\end{equation}
 
 and similarly, when updating $\HH$, we can rewrite the cost function as
-\beq
+\begin{equation}
 \chi^2 = \sum_{kn} \left( \frac{\Xkn}{\skn} - 
@@
 %\sum_k \WWt(j)_{ik}\Hin\right)^2
-\eeq
+\end{equation}
 yielding updates of the form:
-\beq
+\begin{equation}
 \Hin \leftarrow \Hin
@@
 %\Hin \leftarrow \Hin \frac{[\WWt(j)\T\XX]_{kj}}{[\WWt(j)\T\WWt(j)\HH]_{kj}}
-\eeq
+\end{equation}
 
@@
 %%\centerline{\fbox{\parbox{4.3in}{
-\beqa 
-\Aki &\leftarrow \Aki 
+\begin{eqnarray} 
+\Aki &\leftarrow& \Aki 
 \left( \sum_{jn} \frac{\Xkn\Mjn\Bij}{\skn^2}\right) / 
 \left( \sum_{mljn} \frac{A_{km}B_{mj}M_{jn}M_{nl}B_{il}}{\skn^2}\right) \\
-\Bij &\leftarrow \Bij
+\Bij &\leftarrow& \Bij
 \left( \sum_{kn} \frac{\Aki\Xkn\Mjn}{\skn^2}\right) / 
 \left( \sum_{mlkn} \frac{A_{ki}A_{km}B_{ml}M_{ln}M_{jn}}{\skn^2}\right)
-\eeqa
+\end{eqnarray}
 %%}}}

### 2006-10-04 docs/paper/kcorrect-aj-v3/ms.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@566 75dff457-8751-0410-ab1a-cd701654ce6b

--- 2006-08-17 docs/paper/kcorrect-aj-v2/ms.tex
@@
 \documentclass[12pt,preprint]{aastex}
-\usepackage{amsmath}
-\usepackage{epsfig}
 
@@
 other candidates that might provide more of the ``UV-excess'' seen in
-elliptical galaxies (\citealt{greggio90a, brown00a}).  In the near-IR,
-the spectra of thermally-pulsating asymptotic giant branch (TP-AGB)
-stars may dominate the flux in certain intermediate age populations
-(2--3 Gyr). With respect to these phases, \citet{bruzual03a} use
-models that are calibrated to observations of TP-AGB stars
-nearby. \citet{maraston05a} show that a different choice of TB-AGB
-models can produce significantly higher flux in the NIR at
-intermediate ages. Thus, the choices for AGB and post-AGB evolution we
-make here by using the models \cite{bruzual03a} are not definitive,
-but we do not address these controversies here and they remain an
-uncertainty in the physical interpretation of galaxy fluxes.  We do
-note that for low redshift galaxies our results from BC03 do explain
-most galaxies' colors (though see below for our discussion of the high
-redshift GOODS sample).
+elliptical galaxies (\citealt{greggio90a, brown00a}).
+
+In the near-IR, the spectra of thermally-pulsating asymptotic giant
+branch (TP-AGB) stars may dominate the flux in certain intermediate
+age populations (0.5--2 Gyr).  \citet{maraston05a} show that including
+TP-AGB spectra based on nearby TP-AGB stars, rather than on the models
+of \citet{vassiliadis93a} used by \citet{bruzual03a}, produces
+significantly higher flux in the NIR at intermediate ages. Thus, the
+choices for AGB and post-AGB evolution we make here by using the
+models \cite{bruzual03a} are not definitive, but we do not address
+these controversies here and they remain an uncertainty in the
+physical interpretation of galaxy fluxes.  We do note that for low
+redshift galaxies our results from BC03 do explain most galaxies'
+colors (though see below for our discussion of the high redshift GOODS
+sample).
 
@@
 
-\begin{thebibliography}{43}
+\begin{thebibliography}{44}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 Strauss, M.~A. {et~al.} 2002, \aj, 124, 1810
+
+\bibitem[{{Vassiliadis} \& {Wood}(1993)}]{vassiliadis93a}
+{Vassiliadis}, E. \& {Wood}, P.~R. 1993, \apj, 413, 641
 

### 2006-11-30 src/k_qromo.c
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@569 75dff457-8751-0410-ab1a-cd701654ce6b

@@ -9,11 +9,13 @@
 #define K 5
 
 float k_qromo(float (*func)(float), float a, float b,
-							 float (*choose)(float(*)(float), float, float, IDL_LONG))
+							 float (*choose)(float(*)(float), float, float, IDL_LONG), 
+							int *good)
 {
 	IDL_LONG j;
 	float ss,dss,h[JMAXP+1],s[JMAXP+1];
 
+	*good=1;
 	h[1]=1.0;
 	for (j=1;j<=JMAX;j++) {
 		s[j]=(*choose)(func,a,b,j);
@@ -26,7 +28,6 @@ float k_qromo(float (*func)(float), float a, float b,
 	}
 	printf("%le %le\n",dss,ss);
 	printf("Too many steps in routing qromo");
-	exit(1);
 	return 0.0;
 }
 #undef EPS
